### 2010-12-11 paper/paper.tex
Adding source files and initial images for the paper

\documentclass[11pt]{article}
\usepackage{amsmath,amsfonts,latexsym,graphicx}
\usepackage{fullpage}
\usepackage{url,hyperref}
\usepackage{subfig}
%\usepackage{ijcai09}

\pagestyle{empty}

\setlength{\oddsidemargin}{0in}
\setlength{\topmargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{8.5in}

\newtheorem{fact}{Fact}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{theorem}[lemma]{Theorem}
\newtheorem{assumption}[lemma]{Assumption}
\newtheorem{definition}[lemma]{Definition}
\newtheorem{corollary}[lemma]{Corollary}
\newtheorem{prop}[lemma]{Proposition}
\newtheorem{claim}[lemma]{Claim}
\newtheorem{remark}[lemma]{Remark}
\newtheorem{prob}{Problem}
\newtheorem{conjecture}{Conjecture}

\newenvironment{proof}{\vspace{-0.15in}\noindent{\bf Proof:}}%
        {\hspace*{\fill}$\Box$\par}
\newenvironment{proofsketch}{\noindent{\bf Proof Sketch.}}%
        {\hspace*{\fill}$\Box$\par\vspace{4mm}}
\newenvironment{proofof}[1]{\smallskip\noindent{\bf Proof of #1.}}%
        {\hspace*{\fill}$\Box$\par}

\newcommand{\etal}{{\em et al.}\ }
\newcommand{\assign}{\leftarrow}
\newcommand{\eps}{\epsilon}

%\newtheorem{lemma}{Lemma}
%\newtheorem{theorem}[lemma]{Theorem}
\begin{document}

\title{Information Set Generation in Kriegspiel}
\author{Abhishek Gupta	Mark Richards	Osman Sarood\\University of Illinois at Urbana-Champaign\\CS598LVK: Parallel Combinatorial Search\\Fall 2010}
\maketitle

\begin{abstract}
Information Set Generation is the identification of the set of paths in an imperfect information game tree that are consistent with a player's observations.  The ability to reason about the possible state of the world is crucial to the performance of game playing agent's.  In this work, we discuss the problem of information set generation in the context of kriegspiel (partially observable chess).  We implement the algorithm on top of a general purpose combinatorial search engine and discuss its performance under different load balancing strategies.  We discuss some variations in the search strategy and their effect on performance.
\end{abstract}

\section{Introduction}
In imperfect inormation games, players do not have access to full knowledge of the world. Examples of imperfect information include hidden cards in Poker or Bridge, hidden tiles in Scrabble, or hidden pieces in Stratego or Kriegspiel (partially observable chess). The game tree nodes that are indistinguishable to a player because they differ only in the information that is hidden to the player by rule are called that player's information set. The ability to estimate the value of the possible states and to reason about the probability distribution over those states is crucial to playing imperfect information games well. While it is often not possible to consider all nodes in an information set, it can be helpful to sample from them. Solving the problem would have the flavor of an all-solutions search. Good results on Kriegspiel specifically or an approach that works well on a variety of games should be publishable.

For many games, solving the information set generation problem is trivial.  For example, in a poker game, it is easy to see that the unseen cards held by a player's opponents may be any permutation of the cards not seen by that player (i.e., hole cards and any revealed community cards).  After the betting rounds, it would not be reasonable to assume that each of these permutations of unseen cards is equally probable, as betting decisions made by the players up to that point would be affected by the quality of those players' cards.  But the set of {\em possible} hands for all of the opponents is easy to conceive and enumerate.

In the game of kriegspiel, information set generation is non-trivial.  The player knows the opponent's position with certainty when the game begins, but after the initial move there may be varying levels of uncertainty about the location of the opponent's pieces.  Unlike poker, it is not possible to simply permute all of the opponent's possible pieces over all of the squares not occupied by the player's own pieces.  A configuration of pieces for the opponent is only valid.
Information set generation is not the same problem as solving a game tree, although it is arguably a necessary subroutine for quality game play in imperfect information games. The standard game tree problem is to find an equilibrium (i.e., minimax) strategy for the game. This can be done using alpha-beta pruning for perfect information games or linear programming for imperfect information games. Information set generation, by contrast, is the problem of finding sequences of moves in the game tree that are consistent with a player's observations.

And now I make a reference to Figure~\ref{fig:init}.
\begin{figure}
\begin{minipage}{\textwidth}
\centering
\subfloat[1. d4]{\label{fig:1W}\includegraphics[width=0.2\textwidth]{images/1W.png}}
\qquad
\subfloat[1. ... a5]{\label{fig:1B}\includegraphics[width=0.2\textwidth]{images/1B.png}}
\qquad
\subfloat[2. Bg5]{\label{fig:2W}\includegraphics[width=0.2\textwidth]{images/2W.png}}
\qquad
\subfloat[2. ... b6]{\label{fig:2B}\includegraphics[width=0.2\textwidth]{images/2B.png}}
\end{minipage}
\begin{minipage}{\textwidth}
\centering
\subfloat[3. Nc3]{\label{fig:3W}\includegraphics[width=0.2\textwidth]{images/3W.png}}
\qquad
\subfloat[3. ... c5]{\label{fig:3B}\includegraphics[width=0.2\textwidth]{images/3B.png}}
\qquad
\subfloat[4. d5]{\label{fig:4W}\includegraphics[width=0.2\textwidth]{images/4W.png}}
\qquad
\subfloat[4. ... Na6]{\label{fig:4B}\includegraphics[width=0.2\textwidth]{images/4B.png}}
\end{minipage}
\begin{minipage}{\textwidth}
\centering
\subfloat[5. d6!]{\label{fig:5W}\includegraphics[width=0.2\textwidth]{images/5W.png}}
\qquad
\subfloat[5. ... f6!]{\label{fig:5B}\includegraphics[width=0.2\textwidth]{images/5B.png}}
\qquad
\subfloat[6. e4]{\label{fig:6W}\includegraphics[width=0.2\textwidth]{images/6W.png}}
\qquad
\subfloat[6. ... fg]{\label{fig:6B}\includegraphics[width=0.2\textwidth]{images/6B.png}}
\end{minipage}
\begin{minipage}{\textwidth}
\centering
\subfloat[7. Bc4]{\label{fig:7W}\includegraphics[width=0.2\textwidth]{images/7W.png}}
\qquad
\subfloat[7. ... ed]{\label{fig:7B}\includegraphics[width=0.2\textwidth]{images/7B.png}}
\qquad
\subfloat[8 Qd5!]{\label{fig:8W}\includegraphics[width=0.2\textwidth]{images/8W.png}}
\qquad
\subfloat[8 ... Qc7?]{\label{fig:8B}\includegraphics[width=0.2\textwidth]{images/8B.png}}
\end{minipage}
\begin{minipage}{\textwidth}
\centering
\subfloat[9 Qf7+]{\label{fig:9W}\includegraphics[width=0.2\textwidth]{images/9W.png}}
\qquad
\subfloat[9 ... d8]{\label{fig:9B}\includegraphics[width=0.2\textwidth]{images/9B.png}}
\qquad
\subfloat[10 Qxf8++]{\label{fig:10W}\includegraphics[width=0.2\textwidth]{images/8W.png}}
\end{minipage}
\caption{Sequence of moves}
\label{fig:sequence}
\end{figure}
%\nocite{*}
\nocite{richards09information}
\nocite{russell05efficient}
\nocite{kuhn03lectures}
\nocite{kuhn97classics}
\nocite{parker05sampling}
\nocite{nance06reasoning}
\nocite{li94chess}

\bibliographystyle{acm}
\bibliography{paper}
%\begin{thebibliography}{99}
%\end{thebibliography}

\end{document}

### 2010-12-11 serial.C
Resolved (I think) changes from Osman and me (Mark)

@@ -2076,7 +2076,7 @@ int main(int argc, char* argv[])
 	int nExecutedMoves = generateCannedMoves(state,true,moveHistory,failedMoves);
 	// Display the actual sequence of moves (for testing/debugging purposes)
         cout << nExecutedMoves << endl;
-        bool whitePerspective = false; // Black perspective by default
+        //bool whitePerspective = false; // Black perspective by default
         if (argc >= 2) {
           int truncationLevel = atoi(argv[1]);
           assert (truncationLevel <= nExecutedMoves);
@@ -2100,10 +2100,11 @@ int main(int argc, char* argv[])
         //generateInformationSet(false, state, state, true, moveHistory, possHistory, failedMoves, moveList, 0, nExecutedMoves);
 	
 	//Osman's function
+        maxdepth = nExecutedMoves;
 	generateInformationSet(state, state, true, possHistory,  moveList, 0);
     //Mark's new function  
 	//TODO
-	generateInformationSet(whitePerspective, state, state, true, moveHistory, possHistory, failedMoves, moveList, 0, nExecutedMoves);
+	//generateInformationSet(whitePerspective, state, state, true, moveHistory, possHistory, failedMoves, moveList, 0, nExecutedMoves);
 	cout << "Solutions founds: " << nSolutions << endl;
 	return 0;
 }

### 2010-12-12 paper/paper.tex
Adding more stuff to the paper -- timing results for illustrative example, discussion of some related work

@@
 \usepackage{subfig}
+\usepackage{tikz}
+\usetikzlibrary{arrows,decorations.pathmorphing,backgrounds,fit,shapes.geometric}
+\usepgflibrary{shapes.geometric}
 %\usepackage{ijcai09}
@@
 \title{Information Set Generation in Kriegspiel}
-\author{Abhishek Gupta	Mark Richards	Osman Sarood\\University of Illinois at Urbana-Champaign\\CS598LVK: Parallel Combinatorial Search\\Fall 2010}
+\author{Abhishek Gupta\qquad Mark Richards\qquad Osman Sarood\\University of Illinois at Urbana-Champaign\\CS598LVK: Parallel Combinatorial Search}
 \maketitle
@@
 Information Set Generation is the identification of the set of paths in an imperfect information game tree that are consistent with a player's observations.
-The ability to reason about the possible state of the world is crucial to the performance of game playing agent's.
+The ability to reason about the possible state of the world is crucial to the performance of game playing agents.
 In this work, we discuss the problem of information set generation in the context of kriegspiel (partially observable chess).
-We implement the algorithm on top of a general purpose combinatorial search engine and discuss its performance under different load balancing strategies.
+We implement the algorithm on top of a general purpose combinatorial search engine and discuss its performance under different load balancing strategies and with different grainsize parameters.
 We discuss some variations in the search strategy and their effect on performance.
@@
 \section{Introduction}
-In imperfect inormation games, players do not have access to full knowledge of the world.
-Examples of imperfect information include hidden cards in Poker or Bridge, hidden tiles in Scrabble, or hidden pieces in Stratego or Kriegspiel (partially observable chess).
-The game tree nodes that are indistinguishable to a player because they differ only in the information that is hidden to the player by rule are called that player's information set.
-The ability to estimate the value of the possible states and to reason about the probability distribution over those states is crucial to playing imperfect information games well.
-While it is often not possible to consider all nodes in an information set, it can be helpful to sample from them.
-Solving the problem would have the flavor of an all-solutions search.
-Good results on Kriegspiel specifically or an approach that works well on a variety of games should be publishable.
+In imperfect information games, players do not have access to full knowledge of the world.
+Examples of imperfect information include hidden cards in Poker~\cite{billings02challenge} or Bridge~\cite{ginsberg96partition}, hidden tiles in Scrabble~\cite{richards07opponent}, or hidden pieces in Stratego or Kriegspiel (partially observable chess)~\cite{li94chess}.
+The game tree nodes that are indistinguishable to a player because they differ only in the information that is hidden to the player by rule are called that player's {\em information set.}  The ability to estimate the value of the possible states and to reason about the probability distribution over those states is crucial to playing imperfect information games well.
+
+The term {\em belief state} is sometimes used interchangeably with information set to refer to a probability distribution over possible worlds.
+The latter term comes from the game theory community and is preferred here.
+A node in a game tree denotes not only the current state of the game, but also uniquely defines a path from the initial state or root node.
+Thus, a game tree node implicitly encodes not only the current state of the game, but also all the history of all decisions made by all players up to that point in the game.
+Knowing one's information set means knowing all possible game histories.
 
@@
 
-And now I make a reference to Figure~\ref{fig:init}.
+\section{Related Work}
+In this work, we have argued for an information set generation approach to kriegspiel, but other authors have approached the problem differently.
+Parker {\em et al.} consider the problem of sampling from belief states in a variety of large game trees, including kriegspiel ~\cite{parker05game}.
+They show that some performance gains can be achieved by only approximate sampling from the belief state.
+Instead of generating the full information set, or even sampling from it, they sample from possible positions for the current state (i.e., sampling from the belief state) by matching only the most recent observations.
+This motivation for this approach seems to be that it would not be feasible to sample from the full information set, because of the computational expense involved.
+And indeed, our experiments bear this out-- it can indeed be expensive to produce the full information set.
+For the experiments shown in Table XX, generating the full information set took over 25 minutes on a single processor.
+However, by utilizing the Charm++ search engine on a parallel machine (1024 processors), we were able to generate the full set in just 2.9 seconds.
+(Note that this set had more than 4 million members.)  Our results suggest that the problem is highly scalable and our approach might therefore be preferred.
+
+Nance {\em et al.}
+
 \begin{figure}
+%\includegraphics[width=0.2\textwidth]{images/1W.png}
 \begin{minipage}{\textwidth}
@@
 \caption{Sequence of moves}
-\label{fig:sequence}
+\label{sequence}
 \end{figure}
+
+And now I make a reference to Figure \ref{sequence}. and Figure \ref{fulltree}.
+
+\section{Appendix}
+Figure~\ref{fulltree} shows the full game tree for a simple variant of poker known as Kuhn poker.
+There are three cards: a king, a queen, and a jack.
+The dealer deals one card to each of two players.
+Each player antes one unit, and there is one simple round of betting, in which the size of the pot may be doubled.
+Left branches denote check/fold; right branches denote bet/call.
+For this zero-sum game, terminal nodes are labeled with payoffs to player 1, whose decision points are shown in triangles.
+Connections between nodes in the same information set are shown with dotted lines.
+
+\begin{figure}
+\scalebox{.72}{
+\begin{tikzpicture}
+  [chance/.style={circle,draw=blue!50,fill=blue!20,thick,minimum size = 10mm},
+   terminal/.style={rectangle,draw=black!50,fill=black!20,thick, minimum size = 7mm},
+   maxer/.style={shape=regular polygon, regular polygon sides=3,draw=red!50,fill=red!30,thick,minimum size = 10mm,inner sep = 0pt},
+   miner/.style={shape=diamond,draw=green!50,fill=green!30,thick,minimum size = 10mm, inner sep = 0pt},
+   edge from parent/.style={red,thick,draw}, 
+   parent anchor=south,child anchor=north,
+   level 1/.style={sibling distance=4cm,level distance=1.4cm,
+       		growth parent anchor=south},
+   level 2/.style={sibling distance=2cm},
+   level 3/.style={sibling distance=1cm},
+   level 4/.style={sibling distance=1.0cm}]
+		\node  [chance] {}
+		    child {node (M1) [maxer] {\scalebox{.75}{K}}
+			child {node (M11) [miner] {\scalebox{.75}{Q-}}
+				child {node (M111) [terminal] {1}}
+				child {node (M112) [maxer] {\scalebox{.75}{K+}}
+					child {node (M1121) [terminal] {-1}}
+					child {node (M1122) [terminal] {2}}}
+			}
+			child {node (M12) [miner] {\scalebox{.75}{Q+}}
+				child {node (M121) [terminal] {1}}
+				child {node (M122) [terminal] {2}}
+			}
+		    }
+		    child {node (M2) [maxer] {\scalebox{.75}{J}}
+			child {node (M21) [miner] {\scalebox{.75}{Q-}}
+				child {node (M211) [terminal] {-1}}
+				child {node (M212) [maxer] {\scalebox{.75}{J+}}
+					child {node (M2121) [terminal] {-1}}
+					child {node (M2122) [terminal] {-2}}}
+			}
+			child {node (M22) [miner] {\scalebox{.75}{Q+}}
+				child {node (M221) [terminal] {1}}
+				child {node (M222) [terminal] {-2}}
+			}
+		    }
+		    child {node (M3) [maxer] {\scalebox{.75}{K}}
+			child {node (M31) [miner] {\scalebox{.75}{J-}}
+				child {node (M311) [terminal] {1}}
+				child {node (M312) [maxer] {\scalebox{.75}{K+}}
+					child {node (M3121) [terminal] {-1}}
+					child {node (M3122) [terminal] {2}}}
+			}
+			child {node (M32) [miner] {\scalebox{.75}{J+}}
+				child {node (M321) [terminal] {1}}
+				child {node (M322) [terminal] {2}}
+			}
+		    }
+		    child {node (M4) [maxer] {\scalebox{.75}{Q}}
+			child {node (M41) [miner] {\scalebox{.75}{J-}}
+				child {node (M411) [terminal] {1}}
+				child {node (M412) [maxer] {\scalebox{.75}{Q+}}
+					child {node (M4121) [terminal] {-1}}
+					child {node (M4122) [terminal] {2}}}
+			}
+			child {node (M42) [miner] {\scalebox{.75}{J+}}
+				child {node (M421) [terminal] {1}}
+				child {node (M422) [terminal] {2}}
+			}
+		    }
+		    child {node (M5) [maxer] {\scalebox{.75}{J}}
+			child {node (M51) [miner] {\scalebox{.75}{K-}}
+				child {node (M511) [terminal] {-1}}
+				child {node (M512) [maxer] {\scalebox{.75}{J+}}
+					child {node (M5121) [terminal] {-1}}
+					child {node (M5122) [terminal] {-2}}}
+			}
+			child {node (M52) [miner] {\scalebox{.75}{K+}}
+				child {node (M521) [terminal] {1}}
+				child {node (M522) [terminal] {-2}}
+			}
+		    }
+		    child {node (M6) [maxer] {\scalebox{.75}{Q}}
+			child {node (M61) [miner] {\scalebox{.75}{K-}}
+				child {node (M611) [terminal] {-1}}
+				child {node (M612) [maxer] {\scalebox{.75}{Q+}}
+					child {node (M6121) [terminal] {-1}}
+					child {node (M6122) [terminal] {-2}}}
+			}
+			child {node (M62) [miner] {\scalebox{.75}{K+}}
+				child {node (M621) [terminal] {1}}
+				child {node (M622) [terminal] {-2}}
+			}
+		    };
+		\draw  (M1.north) to [dashed,black,out=15,in=165] (M3.north);
+		\draw  (M2.north) to [dashed,black,out=15,in=165] (M5.north);
+		\draw  (M4.north) to [dashed,black,out=15,in=165] (M6.north);
+		\draw  (M112.north) to [dashed,black,out=15,in=165] (M312.north);
+		\draw  (M212.north) to [dashed,black,out=15,in=165] (M512.north);
+		\draw  (M412.north) to [dashed,black,out=15,in=165] (M612.north);
+		\draw  (M11.north) to [dashed,black,out=15,in=165] (M21.north);
+		\draw  (M12.north) to [dashed,black,out=15,in=165] (M22.north);
+		\draw  (M31.north) to [dashed,black,out=15,in=165] (M41.north);
+		\draw  (M32.north) to [dashed,black,out=15,in=165] (M42.north);
+		\draw  (M51.north) to [dashed,black,out=15,in=165] (M61.north);
+		\draw  (M52.north) to [dashed,black,out=15,in=165] (M62.north);
+\end{tikzpicture}
+} %scalebox
+\caption{}
+\label{fulltree}
+\end{figure}
+
+Table ~\ref{sampletimes}
+\begin{table}
+\begin{tabular}{ccc}
+1 & 18 & 0.000
+2 & 18 & 0.000
+3 & 18 & 0.000
+4 & 279 & 0.008
+5 & 242 & 0.028
+6 & 216 & 0.052
+7 & 176 & 0.064
+8 & 3406 & 0.092
+9 & 2191 & 0.288
+10 & 1423 & 0.508
+11 & 1363 & 0.596
+12 & 1416 & 0.744
+13 & 1416 & 0.836
+14 & 3440 & 0.97
+15 & 3428 & 1.25
+16 & 50521 & 1.82
+17 & 43192 & 5.91
+18 & 26128 & 7.94
+19 & 19061 & 9.92
+\end{tabular}
+\caption{Solution counts and running times for a sample kriegspiel game}
+\label{white times}
+\end{table}
+
+Table ~\ref{sampletimes}
+\begin{table}
+\begin{tabular}{ccc}
+1 & 20 & .000\\
+2 & 19 & .000\\
+3 & 404 & .008\\
+4 & 401 & .024\\
+5 & 1472 & .044\\
+6 & 155 & .096\\
+7 & 293 & .116\\
+8 & 158 & .128\\
+9 & 118 & .148\\
+10 & 25 & .152\\
+11 & 798 & .164\\
+12 & 518 & .192\\
+13 & 13564 & .304\\
+14 & 12394 & .776\\
+15 & 343652 & 3.75\\
+16 & 320704 & 17.345\\
+17 & 490162 & 99.9\\
+18 & 3792 & 119\\
+19 & 14836 & 121 \\
+\end{tabular}
+\caption{Solution counts and running times for a sample kriegspiel game}
+\label{blacktimes}
+\end{table}
+
 %\nocite{*}

### 2010-12-13 paper/paper.tex
Continuing to progress on the paper.  Added some stuff to introductory and related work sections.

@@
 \title{Information Set Generation in Kriegspiel}
-\author{Abhishek Gupta\qquad Mark Richards\qquad Osman Sarood\\University of Illinois at Urbana-Champaign\\CS598LVK: Parallel Combinatorial Search}
-\maketitle
+\author{Abhishek Gupta\qquad Mark Richards\qquad Osman Sarood\\University of Illinois at Urbana-Champaign\\CS598LVK:
+Parallel Combinatorial Search} \maketitle
 
 \begin{abstract}
-Information Set Generation is the identification of the set of paths in an imperfect information game tree that are consistent with a player's observations.
-The ability to reason about the possible state of the world is crucial to the performance of game playing agents.
-In this work, we discuss the problem of information set generation in the context of kriegspiel (partially observable chess).
-We implement the algorithm on top of a general purpose combinatorial search engine and discuss its performance under different load balancing strategies and with different grainsize parameters.
-We discuss some variations in the search strategy and their effect on performance.
-\end{abstract}
+Information Set Generation is the identification of the set of paths in an imperfect information game tree that are
+consistent with a player's observations.  The ability to reason about the possible state of the world is crucial to the
+performance of game playing agents.  In this work, we discuss the problem of information set generation in the context
+of kriegspiel (partially observable chess).  We implement the algorithm on top of a general purpose combinatorial search
+engine and discuss its performance under different load balancing strategies and with different grainsize parameters.
+We discuss some variations in the search strategy and their effect on performance.  \end{abstract}
 
 \section{Introduction}
-In imperfect information games, players do not have access to full knowledge of the world.
-Examples of imperfect information include hidden cards in Poker~\cite{billings02challenge} or Bridge~\cite{ginsberg96partition}, hidden tiles in Scrabble~\cite{richards07opponent}, or hidden pieces in Stratego or Kriegspiel (partially observable chess)~\cite{li94chess}.
-The game tree nodes that are indistinguishable to a player because they differ only in the information that is hidden to the player by rule are called that player's {\em information set.}  The ability to estimate the value of the possible states and to reason about the probability distribution over those states is crucial to playing imperfect information games well.
+In imperfect information games, players do not have access to full knowledge of the world. Examples of imperfect
+information include hidden cards in Poker~\cite{billings02challenge} or Bridge~\cite{ginsberg96partition}, hidden tiles
+in Scrabble~\cite{richards07opponent}, or hidden pieces in Stratego or Kriegspiel (partially observable
+chess)~\cite{li94chess}. The game tree nodes that are indistinguishable to a player because they differ only in the
+information that is hidden to the player by rule are called that player's {\em information set.}  The ability to
+estimate the value of the possible states and to reason about the probability distribution over those states is crucial
+to playing imperfect information games well.
 
-The term {\em belief state} is sometimes used interchangeably with information set to refer to a probability distribution over possible worlds.
-The latter term comes from the game theory community and is preferred here.
-A node in a game tree denotes not only the current state of the game, but also uniquely defines a path from the initial state or root node.
-Thus, a game tree node implicitly encodes not only the current state of the game, but also all the history of all decisions made by all players up to that point in the game.
-Knowing one's information set means knowing all possible game histories.
+The term {\em belief state} is sometimes used interchangeably with information set to refer to a probability
+distribution over possible worlds.  The latter term comes from the game theory community and is preferred here.  A node
+in a game tree denotes not only the current state of the game, but also uniquely defines a path from the initial state
+or root node.  Thus, a game tree node implicitly encodes not only the current state of the game, but also all the
+history of all decisions made by all players up to that point in the game.  Knowing one's information set means knowing
+all possible game histories.
 
-For many games, solving the information set generation problem is trivial.
-For example, in a poker game, it is easy to see that the unseen cards held by a player's opponents may be any permutation of the cards not seen by that player (i.e., hole cards and any revealed community cards).
-After the betting rounds, it would not be reasonable to assume that each of these permutations of unseen cards is equally probable, as betting decisions made by the players up to that point would be affected by the quality of those players' cards.
-But the set of {\em possible} hands for all of the opponents is easy to conceive and enumerate.
+For many games, solving the information set generation problem is trivial.  For example, in a poker game, it is easy to
+see that the unseen cards held by a player's opponents may be any permutation of the cards not seen by that player
+(i.e., hole cards and any revealed community cards).  After the betting rounds, it would not be reasonable to assume
+that each of these permutations of unseen cards is equally probable, as betting decisions made by the players up to that
+point would be affected by the quality of those players' cards.  But the set of {\em possible} hands for all of the
+opponents is easy to conceive and enumerate.
 
-In the game of kriegspiel, information set generation is non-trivial.
-The player knows the opponent's position with certainty when the game begins, but after the initial move there may be varying levels of uncertainty about the location of the opponent's pieces.
-Unlike poker, it is not possible to simply permute all of the opponent's possible pieces over all of the squares not occupied by the player's own pieces.
-A configuration of pieces for the opponent is only valid.
-Information set generation is not the same problem as solving a game tree, although it is arguably a necessary subroutine for quality game play in imperfect information games.
-The standard game tree problem is to find an equilibrium (i.e., minimax) strategy for the game.
-This can be done using alpha-beta pruning for perfect information games or linear programming for imperfect information games.
-Information set generation, by contrast, is the problem of finding sequences of moves in the game tree that are consistent with a player's observations.
+In the game of kriegspiel, information set generation is non-trivial.  The player knows the opponent's position with
+certainty when the game begins, but after the initial move there may be varying levels of uncertainty about the location
+of the opponent's pieces.  Unlike poker, it is not possible to simply permute all of the opponent's possible pieces over
+all of the squares not occupied by the player's own pieces.  A configuration of pieces for the opponent is only valid.
+Information set generation is not the same problem as solving a game tree, although it is arguably a necessary
+subroutine for quality game play in imperfect information games. The standard game tree problem is to find an
+equilibrium (i.e., minimax) strategy for the game. This can be done using alpha-beta pruning for perfect information
+games or linear programming for imperfect information games. Information set generation, by contrast, is the problem of
+finding sequences of moves in the game tree that are consistent with a player's observations.
+
+\section{Background}
+In this section, we describe the game of kriegspiel, and the particular variation of it that we have implemented.
+
+Kriegspiel was invented in the 1800s.  Historically, the game has required three parties: two competing players and one
+referee.  Each player sits with a chessboard that is partitioned off so as not be seen by the others.  One player
+controls the white pieces; the other controls the black pieces.  The referee keeps track of both sets of pieces.  Each
+player has a full set of chess pieces.  He may position the opponent's pieces however he wishes on his own board (if at
+all).  The players alternate moves as in a regular chess game, but instead of announcing their moves out loud, they
+write the source and destination square of their desired play son a piece of paper and pass them to the referee, so that
+the move is not heard by the other player. The referee, who knows the location of all pieces, checks the legality of the
+move.  If the move is legal, the referee makes the move on his own board and announces that a legal move has been made
+and that it is the other player's turn to move.  If the move is illegal because it is blocked by an opponent's piece or
+would place or leave the active player in check, the referee announces, "No."  There is no penalty for attempting an
+illegal move.  The player continues to attempt moves until one finally succeeds.  If a player has no legal moves
+(because of stalemate or checkmate), the referee announces the game's result.  With the advent of computers, it is
+convenient to play the game over a network, with a computer acting as the referee.
+
+The referee makes other announcements besides declaring move illegal, and there are several variations of the game that
+differ only in the nature of these announcements.  All declarations by the referee are heard by both players.  Thus, for
+example, a player will hear and know if her opponent has attempted an illegal move, and will thus know that her
+opponent's pieces are configured in such a way as to allow at least one attemptable move that blocked or would leave him
+in check.
+
+The referee announces when a player is in check.  (And again, this announcement is heard by both players.)  There is
+some variation in what additional information is supplied in this case.  In one popular variant, the referee declares
+that the player is in check "by rank," "by file," "by diagonal," or "by knight."  In the case of a diagonal attack, the
+referee may announce whether the checking piece is along the long or short diagonal (from the king's perspective).  It
+is possible to be in check by two pieces at the same time, in which case the direction of both attacks would be
+declared.  In our implementation, for simplicity, the referee declares only that the player is in check, without
+specifying the nature of the attack.
+
+In the case of the capture, the referee announces the location of the piece that was captured, so that both players can
+remove the piece from the board.  In some variations, the referee also announces whether the captured piece was a pawn
+or non-pawn.  In our implementation, information about the captured piece is not explicitly provided to the player who
+makes the capture.
+
+Because there is no penalty for making illegal moves, it is often to a player's advantage to attempt to make moves which
+are likely to be illegal, as hearing such declarations from the referee can provide important information to a player
+about the location of the opponent's pieces.  In particular, since a diagonal move by a pawn is legal only the case
+where such a move would capture an opponent's piece, it could potentially be profitable to attempt all or many pawn
+captures on every turn.  In order to speed the game up, popular variants of the game require the referee to make some
+kind of declaration at the beginning of a player's turn with respect to the pawn captures that are available.  This may
+be as simple as declaring "Try" to indicate that at least one pawn capture is possible.  In our implementation, the
+referee declares all possible pawn captures.
+
+In the case of a pawn promotion, the player making the pawn advancement would secretly notify the referee as to which
+piece (queen, rook, bishop, or knight) he would like.  To reduce the number of bits needed to specify a move, our
+implementation assumes and allows only promotion to queen.
+
+In general, any move that is legal in a regular chess match may be at least attempted by a player in kriegspiel.  And
+any kriegspiel move that is ultimately allowed by the referee would also be a legal move in chess.
+
+We assume that a player may not attempt the same illegal move more than once per turn.  This means that if a player
+hears multiple declarations from the referee that the opponent has attempted an illegal move, the player may assume that
+the opponent has at least that many possibilities available to him.
+
+\subsubsection{Special Moves}
+Our implementation supports the capability for pawns to move two spaces on their first turn.  However, we did not
+implement support for castling or {\em en passant} pawn capture.
+
+\subsection{Example}
+The concept of information sets is perhaps best explained through an example.  Figure~\ref{listing} shows a full listing of an
+actual kriegspiel game from ~\cite{li94chess}.  We have adapted the notation used by Wolfe for Berkeley Kriegspiel.  The
+moves are numbered in pairs: a move by white on one line followed by black's move on the next line.  The actual move is
+shown first followed by a list of referee announcements.  Captures announcements are prefixed with an X and give the
+location of the captured piece.  Similarly, pawn tries announcements are prefixed with T.  Attempted moves that were
+ultimately declared illegal are shown in a list following the ':'.  For example at move 9, (Figure~\ref{fig:9W}--\ref{fig:9B}), black
+attempts to block a potential threat along the diagonal by moving a pawn to f6, then attempts to escape or capture a
+threat at f7.  When both of these attempts fail, black can infer that his king is being threatened by a protected bishop or
+queen at f7.  The attempt to move to e7 is black's way of finding out whether the threatening piece is a queen or
+bishop.  When this move fails, black knows it is white's queen and ultimately retreats to d8.  Note that the failed
+moves by white on turns 3 and 4 are a result of being blocked by the pawn at e2.
+
+\begin{figure}
+\begin{verbatim}
+1. d4 {(:)}
+   a5 {(:)}
+2. Bg5 {(:)}
+   b6 {(:)}
+3. Nc3 {(:Bd8)}
+   c5 {(:)}
+4. d5 {(:Bd8;Tc5)}
+   Na6 {(:)}
+5. d6 {(:)}
+   f6 {(:Td6)}
+6. e4 {(:)}
+   fxg5 {(Xg5:Td6;Tg5)}
+7. Bc4 {(:)}
+   exd6 {(Xd6:Td6)}
+8. Qd5 {(:)}
+   Qc7 {(:)}
+9. Qf7+ {(:)} 
+   d8 {:g6,f7,e7}
+10. Qxf8++ {(1-0:)}
+\end{verbatim}
+\caption{Listing of a kriegspiel game resulting in checkmate by white after 10 moves.}
+\label{listing}
+\end{figure}
+
+Note that the listing in Figure~\ref{listing} corresponds to the referee's view of the game and would be useful for
+post-game analysis or commentary.  During the game, players would not have access to the full information.  The
+transcript from white's perspective is shown in Figure~\ref{filteredlisting}.  Note that black's actual moves have been
+replaced with ``??'' and lists of black's illegal moves have been replaced with the {\em number} of illegal attempted
+moves.  For example, on move 9, black attempted three illegal moves.
+\begin{figure}
+\begin{verbatim}
+1. d4 {(:)}
+   ?? {(:0)}
+2. Bg5 {(:)}
+   ?? {(:0)}
+3. Nc3 {(:Bd8)}
+   ??  {(:0)}
+4. d5 {(:Bd8;Tc5)}
+   ??  {(:0)}
+5. d6 {(:)}
+   ?? {(:0,Td6)}
+6. e4 {(:)}
+   fxg5 {(Xg5:Td6;Tg5)}
+7. Bc4 {(:)}
+   ?? {(Xd6:Td6)}
+8. Qd5 {(:)}
+   ?? {(:0)}
+9. Qf7+ {(:)} 
+   ?? {:0}
+10. Qxf8++ {(1-0:)}
+\end{verbatim}
+\caption{Listing of a kriegspiel game resulting in checkmate by white after 10 moves.}
+\label{filteredlisting}
+\end{figure}
 
 \section{Related Work}
-In this work, we have argued for an information set generation approach to kriegspiel, but other authors have approached the problem differently.
-Parker {\em et al.} consider the problem of sampling from belief states in a variety of large game trees, including kriegspiel ~\cite{parker05game}.
-They show that some performance gains can be achieved by only approximate sampling from the belief state.
-Instead of generating the full information set, or even sampling from it, they sample from possible positions for the current state (i.e., sampling from the belief state) by matching only the most recent observations.
-This motivation for this approach seems to be that it would not be feasible to sample from the full information set, because of the computational expense involved.
-And indeed, our experiments bear this out-- it can indeed be expensive to produce the full information set.
-For the experiments shown in Table XX, generating the full information set took over 25 minutes on a single processor.
-However, by utilizing the Charm++ search engine on a parallel machine (1024 processors), we were able to generate the full set in just 2.9 seconds.
-(Note that this set had more than 4 million members.)  Our results suggest that the problem is highly scalable and our approach might therefore be preferred.
+In this work, we have argued for an information set generation approach to kriegspiel, but other authors have approached
+the problem differently.  Parker \etal, consider the problem of sampling from belief states in a variety of large
+game trees, including kriegspiel~\cite{parker05game}.  They show that some performance gains can be achieved by only
+approximate sampling from the belief state.  Instead of generating the full information set, or even sampling from it,
+they sample from possible positions for the current state (i.e., sampling from the belief state) by matching only the
+most recent observations.  This motivation for this approach seems to be that it would not be feasible to sample from
+the full information set, because of the computational expense involved.  And indeed, our experiments bear this out-- it
+can indeed be expensive to produce the full information set.  For the experiments shown in Table XX, generating the full
+information set took over 25 minutes on a single processor.  However, by utilizing the Charm++ search engine on a
+parallel machine (1024 processors), we were able to generate the full set in just 2.9 seconds.  (Note that this set had
+more than 4 million members.)  Our results suggest that the problem is highly scalable and our approach might therefore
+be preferred.
 
-Nance {\em et al.}
+Li gives an overview of kriegspiel and discusses strategy from the perspective of human players~\cite{li94chess}.  He
+walks through the analysis of several actual games between two human players and discusses the inferences that each
+player can and should make at each position.  As in chess, human players are able to reason about the game at an
+abstract level; they tend to focus on a small number of critical possibilities and to emphasize the most important
+pieces and squares in each position.  Computer agents, lacking such sophistication, nevertheless have some advantages
+when it comes to brute force search strategies. Programs such Deep Blue have shown that a brute force approach to game
+tree search can be effective, even against humans with vastly superior abstract reasoning capabilities, provided that
+significant computational resources are available and wisely used~\cite{campbell02deep}.
+
+Russell and Wolfe have performed some analyses on kriegspiel positions for which it possible to prove that one player
+has a forced win (that is, regardless of the configuration of the opponent's unseen pieces, the player has a sequence of
+moves that is guaranteed to produce a checkmate)~\cite{russell05efficient, wolfe07exploiting}.  The authors assume
+knowledge of the player's belief state.  They claim that under certain "aggressive" styles of play, it is uncommon for
+information sets to exceed 10,000 nodes in size.  They throw out (i.e., do not analyze) any positions that exceed this
+threshold.
+
+In our experiments, we have found positions in which the size of the information set is many tens of millions.  Our goal
+is to develop strategies that are amenable to use in this more general setting.   In other words, rather than being able
+to analyze positions only at the beginning or end of the game (when the size of the information set tends to be
+smaller), we want to be able to develop a strategy that can be used in the middle of the game as well.  game).
+
+Nance \etal, approach the game from the perspective of {\em logical filtering.}~\cite{nance06reasoning}  The goal of
+logical filtering is to maintain a compact logical representation of an agent's belief state.  Recall that a belief
+state is a probability distribution over the possible current states of the world.  A belief state for kriegspiel might
+take the form of $(at(whitepawn,a2) \wedge at whitepawn(b2) \ldots \wedge at(whiterook,h1)) \vee (at(whitepawn,a3)
+\wedge \ldots \wedge at(whiterook,h1)) \ldots$  For even moderately sizes information sets, such
+representations can become unwieldy in terms of storage and the cost of updating.  Furthermore, reasoning with such
+representations often requires the use of some theorem-proving mechanism to run logical queries, such
+as $at(blackking,a3?)$. Nance et al. develop some strategies to maintain compact representations under
+sequences of actions and observations.  They give a theoretical analysis of the kinds of problems that can be treated in
+this way.  Unfortunately, kriegspiel in its full complexity is beyond the scope of their algorithm.  In order to reason
+about the belief states in the game, they must assume knowledge of the particular kind of piece that move in each turn.
+
+The information set generation approach that we have presented is certainly also computationally expensive.  And
+representing a full information set (e.g., as a list of sequences of moves that encode a path from the start state to
+the current node) would be similarly expensive in terms of storage space.  However, we have shown that our algorithm is
+also highly scalable.  Furthermore, contrary to the approach in~\cite{nance06reasoning}, our approach is readily
+adaptable to sampling algorithms.  Given a sample size that matches available memory, any single sequence of moves is
+readily extractable without the computational expense of a theorem prover or SAT solver.
 
@@
 
-And now I make a reference to Figure \ref{sequence}. and Figure \ref{fulltree}.
+And now I make a reference to Figure\ref{sequence}. and Figure\ref{fulltree}.
 
 \section{Appendix}
-Figure~\ref{fulltree} shows the full game tree for a simple variant of poker known as Kuhn poker.
-There are three cards: a king, a queen, and a jack.
-The dealer deals one card to each of two players.
-Each player antes one unit, and there is one simple round of betting, in which the size of the pot may be doubled.
-Left branches denote check/fold; right branches denote bet/call.
-For this zero-sum game, terminal nodes are labeled with payoffs to player 1, whose decision points are shown in triangles.
-Connections between nodes in the same information set are shown with dotted lines.
+Figure~\ref{fulltree} shows the full game tree for a simple variant of poker known as Kuhn poker.  There are three
+cards: a king, a queen, and a jack.  The dealer deals one card to each of two players.  Each player antes one unit, and
+there is one simple round of betting, in which the size of the pot may be doubled.  Left branches denote check/fold;
+right branches denote bet/call.  For this zero-sum game, terminal nodes are labeled with payoffs to player 1, whose
+decision points are shown in triangles.  Connections between nodes in the same information set are shown with dotted
+lines.
 
@@
 
-Table ~\ref{sampletimes}
+Table~\ref{sampletimes}
 \begin{table}
 \begin{tabular}{ccc}
-1 & 18 & 0.000
-2 & 18 & 0.000
-3 & 18 & 0.000
-4 & 279 & 0.008
-5 & 242 & 0.028
-6 & 216 & 0.052
-7 & 176 & 0.064
-8 & 3406 & 0.092
-9 & 2191 & 0.288
-10 & 1423 & 0.508
-11 & 1363 & 0.596
-12 & 1416 & 0.744
-13 & 1416 & 0.836
-14 & 3440 & 0.97
-15 & 3428 & 1.25
-16 & 50521 & 1.82
-17 & 43192 & 5.91
-18 & 26128 & 7.94
-19 & 19061 & 9.92
+1 & 18 & 0.000\\
+2 & 18 & 0.000\\
+3 & 18 & 0.000\\
+4 & 279 & 0.008\\
+5 & 242 & 0.028\\
+6 & 216 & 0.052\\
+7 & 176 & 0.064\\
+8 & 3406 & 0.092\\
+9 & 2191 & 0.288\\
+10 & 1423 & 0.508\\
+11 & 1363 & 0.596\\
+12 & 1416 & 0.744\\
+13 & 1416 & 0.836\\
+14 & 3440 & 0.97\\
+15 & 3428 & 1.25\\
+16 & 50521 & 1.82\\
+17 & 43192 & 5.91\\
+18 & 26128 & 7.94\\
+19 & 19061 & 9.92\\
 \end{tabular}
@@
 
-Table ~\ref{sampletimes}
+Table~\ref{sampletimes}
 \begin{table}
@@
 15 & 343652 & 3.75\\
-16 & 320704 & 17.345\\
+16 & 320704 & 17.3\\
 17 & 490162 & 99.9\\

### 2010-12-13 paper/paper.tex
More progress on paper: added plots for results

@@
 of the opponent's pieces.  Unlike poker, it is not possible to simply permute all of the opponent's possible pieces over
-all of the squares not occupied by the player's own pieces.  A configuration of pieces for the opponent is only valid.
-Information set generation is not the same problem as solving a game tree, although it is arguably a necessary
-subroutine for quality game play in imperfect information games. The standard game tree problem is to find an
-equilibrium (i.e., minimax) strategy for the game. This can be done using alpha-beta pruning for perfect information
-games or linear programming for imperfect information games. Information set generation, by contrast, is the problem of
-finding sequences of moves in the game tree that are consistent with a player's observations.
+all of the squares not occupied by the player's own pieces.  A configuration of pieces for the opponent is only valid if
+it can be reached by a legal sequence of moves.  In the general case, finding the nodes in an information set is a
+combinatorial search problem.
+
+Information set generation is not the same problem as solving a game
+tree, although it is arguably a necessary subroutine for quality game play in imperfect information games. The standard
+game tree problem is to find an equilibrium (i.e., minimax) strategy for the game. This can be done using alpha-beta
+pruning for perfect information games or linear programming for imperfect information games. Information set generation,
+by contrast, is the problem of finding sequences of moves in the game tree that are consistent with a player's
+observations.
 
@@
 
-Kriegspiel was invented in the 1800s.  Historically, the game has required three parties: two competing players and one
-referee.  Each player sits with a chessboard that is partitioned off so as not be seen by the others.  One player
-controls the white pieces; the other controls the black pieces.  The referee keeps track of both sets of pieces.  Each
-player has a full set of chess pieces.  He may position the opponent's pieces however he wishes on his own board (if at
-all).  The players alternate moves as in a regular chess game, but instead of announcing their moves out loud, they
-write the source and destination square of their desired play son a piece of paper and pass them to the referee, so that
-the move is not heard by the other player. The referee, who knows the location of all pieces, checks the legality of the
-move.  If the move is legal, the referee makes the move on his own board and announces that a legal move has been made
-and that it is the other player's turn to move.  If the move is illegal because it is blocked by an opponent's piece or
-would place or leave the active player in check, the referee announces, "No."  There is no penalty for attempting an
-illegal move.  The player continues to attempt moves until one finally succeeds.  If a player has no legal moves
-(because of stalemate or checkmate), the referee announces the game's result.  With the advent of computers, it is
-convenient to play the game over a network, with a computer acting as the referee.
+Kriegspiel was invented by Henry Michael Temple in 1899.  Historically, the game has required three parties: two
+competing players and one referee.  Each player sits with a chessboard that is partitioned off so as not be seen by the
+others.  One player controls the white pieces; the other controls the black pieces.  The referee keeps track of both
+sets of pieces.  Each player has a full set of chess pieces.  He may position the opponent's pieces however he wishes on
+his own board (if at all).  The players alternate moves as in a regular chess game, but instead of announcing their
+moves out loud, they write the source and destination square of their desired play son a piece of paper and pass them to
+the referee, so that the move is not heard by the other player. The referee, who knows the location of all pieces,
+checks the legality of the move.  If the move is legal, the referee makes the move on his own board and announces that a
+legal move has been made and that it is the other player's turn to move.  If the move is illegal because it is blocked
+by an opponent's piece or would place or leave the active player in check, the referee announces, "No."  There is no
+penalty for attempting an illegal move.  The player continues to attempt moves until one finally succeeds.  If a player
+has no legal moves (because of stalemate or checkmate), the referee announces the game's result.  With the advent of
+computers, it is convenient to play the game over a network, with a computer acting as the referee.
 
@@
 Our implementation supports the capability for pawns to move two spaces on their first turn.  However, we did not
-implement support for castling or {\em en passant} pawn capture.
+implement support for castling or {\em en passant} pawn capture.  (These moves require maintaining additional bits of
+information at each state that keep track of whether pieces have moved and/or whether pieces moved on the previous
+turn.)
 
 \subsection{Example}
-The concept of information sets is perhaps best explained through an example.  Figure~\ref{listing} shows a full listing of an
-actual kriegspiel game from ~\cite{li94chess}.  We have adapted the notation used by Wolfe for Berkeley Kriegspiel.  The
-moves are numbered in pairs: a move by white on one line followed by black's move on the next line.  The actual move is
-shown first followed by a list of referee announcements.  Captures announcements are prefixed with an X and give the
-location of the captured piece.  Similarly, pawn tries announcements are prefixed with T.  Attempted moves that were
-ultimately declared illegal are shown in a list following the ':'.  For example at move 9, (Figure~\ref{fig:9W}--\ref{fig:9B}), black
-attempts to block a potential threat along the diagonal by moving a pawn to f6, then attempts to escape or capture a
-threat at f7.  When both of these attempts fail, black can infer that his king is being threatened by a protected bishop or
-queen at f7.  The attempt to move to e7 is black's way of finding out whether the threatening piece is a queen or
-bishop.  When this move fails, black knows it is white's queen and ultimately retreats to d8.  Note that the failed
-moves by white on turns 3 and 4 are a result of being blocked by the pawn at e2.
-
-\begin{figure}
-\begin{verbatim}
-1. d4 {(:)}
-   a5 {(:)}
-2. Bg5 {(:)}
-   b6 {(:)}
-3. Nc3 {(:Bd8)}
-   c5 {(:)}
-4. d5 {(:Bd8;Tc5)}
-   Na6 {(:)}
-5. d6 {(:)}
-   f6 {(:Td6)}
-6. e4 {(:)}
-   fxg5 {(Xg5:Td6;Tg5)}
-7. Bc4 {(:)}
-   exd6 {(Xd6:Td6)}
-8. Qd5 {(:)}
-   Qc7 {(:)}
-9. Qf7+ {(:)} 
-   d8 {:g6,f7,e7}
-10. Qxf8++ {(1-0:)}
-\end{verbatim}
-\caption{Listing of a kriegspiel game resulting in checkmate by white after 10 moves.}
-\label{listing}
-\end{figure}
-
-Note that the listing in Figure~\ref{listing} corresponds to the referee's view of the game and would be useful for
-post-game analysis or commentary.  During the game, players would not have access to the full information.  The
-transcript from white's perspective is shown in Figure~\ref{filteredlisting}.  Note that black's actual moves have been
-replaced with ``??'' and lists of black's illegal moves have been replaced with the {\em number} of illegal attempted
-moves.  For example, on move 9, black attempted three illegal moves.
-\begin{figure}
-\begin{verbatim}
-1. d4 {(:)}
-   ?? {(:0)}
-2. Bg5 {(:)}
-   ?? {(:0)}
-3. Nc3 {(:Bd8)}
-   ??  {(:0)}
-4. d5 {(:Bd8;Tc5)}
-   ??  {(:0)}
-5. d6 {(:)}
-   ?? {(:0,Td6)}
-6. e4 {(:)}
-   fxg5 {(Xg5:Td6;Tg5)}
-7. Bc4 {(:)}
-   ?? {(Xd6:Td6)}
-8. Qd5 {(:)}
-   ?? {(:0)}
-9. Qf7+ {(:)} 
-   ?? {:0}
-10. Qxf8++ {(1-0:)}
-\end{verbatim}
-\caption{Listing of a kriegspiel game resulting in checkmate by white after 10 moves.}
-\label{filteredlisting}
-\end{figure}
-
-\section{Related Work}
-In this work, we have argued for an information set generation approach to kriegspiel, but other authors have approached
-the problem differently.  Parker \etal, consider the problem of sampling from belief states in a variety of large
-game trees, including kriegspiel~\cite{parker05game}.  They show that some performance gains can be achieved by only
-approximate sampling from the belief state.  Instead of generating the full information set, or even sampling from it,
-they sample from possible positions for the current state (i.e., sampling from the belief state) by matching only the
-most recent observations.  This motivation for this approach seems to be that it would not be feasible to sample from
-the full information set, because of the computational expense involved.  And indeed, our experiments bear this out-- it
-can indeed be expensive to produce the full information set.  For the experiments shown in Table XX, generating the full
-information set took over 25 minutes on a single processor.  However, by utilizing the Charm++ search engine on a
-parallel machine (1024 processors), we were able to generate the full set in just 2.9 seconds.  (Note that this set had
-more than 4 million members.)  Our results suggest that the problem is highly scalable and our approach might therefore
-be preferred.
-
-Li gives an overview of kriegspiel and discusses strategy from the perspective of human players~\cite{li94chess}.  He
-walks through the analysis of several actual games between two human players and discusses the inferences that each
-player can and should make at each position.  As in chess, human players are able to reason about the game at an
-abstract level; they tend to focus on a small number of critical possibilities and to emphasize the most important
-pieces and squares in each position.  Computer agents, lacking such sophistication, nevertheless have some advantages
-when it comes to brute force search strategies. Programs such Deep Blue have shown that a brute force approach to game
-tree search can be effective, even against humans with vastly superior abstract reasoning capabilities, provided that
-significant computational resources are available and wisely used~\cite{campbell02deep}.
-
-Russell and Wolfe have performed some analyses on kriegspiel positions for which it possible to prove that one player
-has a forced win (that is, regardless of the configuration of the opponent's unseen pieces, the player has a sequence of
-moves that is guaranteed to produce a checkmate)~\cite{russell05efficient, wolfe07exploiting}.  The authors assume
-knowledge of the player's belief state.  They claim that under certain "aggressive" styles of play, it is uncommon for
-information sets to exceed 10,000 nodes in size.  They throw out (i.e., do not analyze) any positions that exceed this
-threshold.
-
-In our experiments, we have found positions in which the size of the information set is many tens of millions.  Our goal
-is to develop strategies that are amenable to use in this more general setting.   In other words, rather than being able
-to analyze positions only at the beginning or end of the game (when the size of the information set tends to be
-smaller), we want to be able to develop a strategy that can be used in the middle of the game as well.  game).
-
-Nance \etal, approach the game from the perspective of {\em logical filtering.}~\cite{nance06reasoning}  The goal of
-logical filtering is to maintain a compact logical representation of an agent's belief state.  Recall that a belief
-state is a probability distribution over the possible current states of the world.  A belief state for kriegspiel might
-take the form of $(at(whitepawn,a2) \wedge at whitepawn(b2) \ldots \wedge at(whiterook,h1)) \vee (at(whitepawn,a3)
-\wedge \ldots \wedge at(whiterook,h1)) \ldots$  For even moderately sizes information sets, such
-representations can become unwieldy in terms of storage and the cost of updating.  Furthermore, reasoning with such
-representations often requires the use of some theorem-proving mechanism to run logical queries, such
-as $at(blackking,a3?)$. Nance et al. develop some strategies to maintain compact representations under
-sequences of actions and observations.  They give a theoretical analysis of the kinds of problems that can be treated in
-this way.  Unfortunately, kriegspiel in its full complexity is beyond the scope of their algorithm.  In order to reason
-about the belief states in the game, they must assume knowledge of the particular kind of piece that move in each turn.
-
-The information set generation approach that we have presented is certainly also computationally expensive.  And
-representing a full information set (e.g., as a list of sequences of moves that encode a path from the start state to
-the current node) would be similarly expensive in terms of storage space.  However, we have shown that our algorithm is
-also highly scalable.  Furthermore, contrary to the approach in~\cite{nance06reasoning}, our approach is readily
-adaptable to sampling algorithms.  Given a sample size that matches available memory, any single sequence of moves is
-readily extractable without the computational expense of a theorem prover or SAT solver.
-
 \begin{figure}
@@
 
-And now I make a reference to Figure\ref{sequence}. and Figure\ref{fulltree}.
+\begin{figure}
+\begin{verbatim}
+1. d4 {(:)}
+   a5 {(:)}
+2. Bg5 {(:)}
+   b6 {(:)}
+3. Nc3 {(:Bd8)}
+   c5 {(:)}
+4. d5 {(:Bd8;Tc5)}
+   Na6 {(:)}
+5. d6 {(:)}
+   f6 {(:Td6)}
+6. e4 {(:)}
+   fxg5 {(Xg5:Td6;Tg5)}
+7. Bc4 {(:)}
+   exd6 {(Xd6:Td6)}
+8. Qd5 {(:)}
+   Qc7 {(:)}
+9. Qf7+ {(:)} 
+   d8 {:g6,f7,e7}
+10. Qxf8++ {(1-0:)}
+\end{verbatim}
+\caption{Listing of a kriegspiel game resulting in checkmate by white after 10 moves.}
+\label{listing}
+\end{figure}
+
+The concept of information sets is perhaps best explained through an example.  Figures~\ref{listing} and~\ref{sequence}
+show a full listing and the associated pictorial simulation of an actual kriegspiel game from ~\cite{li94chess}.  We
+have adapted the notation used by Wolfe for Berkeley Kriegspiel.  The moves are numbered in pairs: a move by white on
+one line followed by black's move on the next line.  The actual move is shown first followed by a list of referee
+announcements.  Captures announcements are prefixed with an X and give the location of the captured piece.  Similarly,
+pawn tries announcements are prefixed with T.  Attempted moves that were ultimately declared illegal are shown in a list
+following the ':'.  For example at move 9, (Figure~\ref{fig:9W}--\ref{fig:9B}), black attempts to block a potential
+threat along the diagonal by moving a pawn to f6, then attempts to escape or capture a threat at f7.  When both of these
+attempts fail, black can infer that his king is being threatened by a protected bishop or queen at f7.  The attempt to
+move to e7 is black's way of finding out whether the threatening piece is a queen or bishop.  When this move fails,
+black knows it is white's queen and ultimately retreats to d8.  Note that the failed moves by white on turns 3 and 4 are
+a result of being blocked by the pawn at e2.
+
+Note that the listing in Figure~\ref{listing} corresponds to the referee's view of the game and would be useful for
+post-game analysis or commentary.  During the game, players would not have access to the full information.  The
+transcript from white's perspective is shown in Figure~\ref{filteredlisting}.  Note that black's actual moves have been
+replaced with ``??'' and lists of black's illegal moves have been replaced with the {\em number} of illegal attempted
+moves.  For example, on move 9, black attempted three illegal moves.
+\begin{figure}
+\begin{verbatim}
+1. d4 {(:)}
+   ?? {(:0)}
+2. Bg5 {(:)}
+   ?? {(:0)}
+3. Nc3 {(:Bd8)}
+   ??  {(:0)}
+4. d5 {(:Bd8;Tc5)}
+   ??  {(:0)}
+5. d6 {(:)}
+   ?? {(:0,Td6)}
+6. e4 {(:)}
+   fxg5 {(Xg5:Td6;Tg5)}
+7. Bc4 {(:)}
+   ?? {(Xd6:Td6)}
+8. Qd5 {(:)}
+   ?? {(:0)}
+9. Qf7+ {(:)} 
+   ?? {:0}
+10. Qxf8++ {(1-0:)}
+\end{verbatim}
+\caption{Listing of a kriegspiel game resulting in checkmate by white after 10 moves.}
+\label{filteredlisting}
+\end{figure}
+
+The listing of a sequence of moves from the perspective of one player lists all of that player's moves exactly and
+includes the declarations from the referee.  We refer to such a list as a player's observations.  Given a list of
+observations for a sequence of moves, an information set is the set of all possible {\em sequences of moves for both
+players}, that are consistent with those observations.  Figure~\ref{abbrevoutput} shows the information set for black
+after both player's have made five moves.  From this list of possibilities, black can infer that white must have a
+bishop at g5 and a pawn at d6.  There are 25 possible sequences of moves that are consistent with black's observations
+up to that point. Tables~\ref{whitetimes} and ~\ref{blacktimes} show the size of the information sets for each player
+for each move in the game and the amount of time to find the full set on a single processor.  Figure~\ref{verboseoutput}
+shows the full output of the program for the same point in the game (five moves for each player) and illustrates all of
+the possible positions in the belief state.
+
+\section{Performance Results}
+
+Table~\ref{speedups} shows the performance results for our algorithm on a randomly selected problem instance in which
+the size of the information set was about 4.5 million.  (This instance is not necessarily representative of all games or
+states of the game, but it was not cherry-picked in any way.)
+
+\begin{table}
+\centering
+\begin{tabular}{rrr}
+{\bf Procs}	&	{\bf Time (s)} 	&	{\bf Speedup}\\
+1	&	1524	&	1\\
+2	&	812	&	1.88\\
+4	&	407	&	3.74\\
+8	&	208	&	7.32\\
+16	&	102	&	14.8\\
+32	&	53.2	&	28.7\\
+64	&	28.0	&	54.5\\
+128	&	14.9	&	102\\
+256	&	7.70	&	197\\
+512	&	4.59	&	331\\
+1024	&	2.64	&	576\\
+\end{tabular}
+\caption{Speedups for a randomly selected problem instance on 1--1024 processors.  The size of the information set in
+this instance is 4,495,121.}
+\label{speedups}
+\end{table}
+
+Figure~\ref{3schemes}
+\begin{figure}
+\centering
+\includegraphics[width=0.7\textwidth]{plots/3schemes.pdf}
+\caption{3schemes}
+\label{3schemes}
+\end{figure}
+
+Figure~\ref{3depths}
+\begin{figure}
+\centering
+%\begin{minipage}{.4\textwidth}
+\includegraphics[width=0.7\textwidth]{plots/3depths.pdf}
+%\end{minipage}
+%\begin{minipage}{.4\textwidth}
+\includegraphics[width=0.7\textwidth]{plots/3depths_nolog.pdf}
+%\end{minipage}
+\caption{3depths}
+\label{3depths}
+\end{figure}
+
+Figure~\ref{dep8}
+\begin{figure}
+\centering
+\includegraphics[width=0.7\textwidth]{plots/dep8.pdf}
+\caption{dep8}
+\label{dep8}
+\end{figure}
+
+Figure~\ref{dep10}
+\begin{figure}
+\centering
+\includegraphics[width=0.7\textwidth]{plots/dep10.pdf}
+\caption{dep10}
+\label{dep10}
+\end{figure}
+
+Figure~\ref{dep12}
+\begin{figure}
+\centering
+\includegraphics[width=0.7\textwidth]{plots/dep12.pdf}
+\caption{dep12}
+\label{dep12}
+\end{figure}
+
+\section{Related Work}
+An overview of the theory of imperfect information games and other basic topics in game theory can be found in
+~\cite{kuhn03lectures} and~\cite{kuhn97classics}.  A technical explanation of the concept of information sets can be
+found in~\cite{gilpin07algorithms}.  Algorithms for ``solving'' imperfect information games---in the sense of finding a
+Nash equilibrium---can be found in~\cite{koller94fast} and~\cite{koller97representations}.  Koller \etal, showed that an
+equilibrium can be found in time that is polynomial in the size of the game tree.  And in fact, they found that it took
+about as much time to generate the full game tree as it did to actually solve it.  Unfortunately, this is feasible only
+for small games.  In larger games where theoretically optimal strategies cannot be computed efficiently, computer
+systems tend to utilize the concept of infomration sets in some form or another.  A common strategy is to estimate the
+value of a move by averaging the estimated value of the associated descendant for each node in the current information
+set (or a random sample from the set).  While this sort of ``averaging over clairvoyance'' {\em can} lead to very poor
+decisions, it is nevertheless a reasonable strategy in some games (e.g., Scrabble~\cite{sheppard02world}).
+
+In this work, we have argued for an information set generation approach to kriegspiel, but other authors have approached
+the problem differently.  Parker \etal, consider the problem of sampling from belief states in a variety of large
+game trees, including kriegspiel~\cite{parker05game}.  They show that some performance gains can be achieved by only
+approximate sampling from the belief state.  Instead of generating the full information set, or even sampling from it,
+they sample from possible positions for the current state (i.e., sampling from the belief state) by matching only the
+most recent observations.  This motivation for this approach seems to be that it would not be feasible to sample from
+the full information set, because of the computational expense involved.  And indeed, our experiments bear this out-- it
+can indeed be expensive to produce the full information set.  For the experiments shown in Table XX, generating the full
+information set took over 25 minutes on a single processor.  However, by utilizing the Charm++ search engine on a
+parallel machine (1024 processors), we were able to generate the full set in just 2.9 seconds.  (Note that this set had
+more than 4 million members.)  Our results suggest that the problem is highly scalable and our approach might therefore
+be preferred.
+
+Li gives an overview of kriegspiel and discusses strategy from the perspective of human players~\cite{li94chess}.  He
+walks through the analysis of several actual games between two human players and discusses the inferences that each
+player can and should make at each position.  As in chess, human players are able to reason about the game at an
+abstract level; they tend to focus on a small number of critical possibilities and to emphasize the most important
+pieces and squares in each position.  Computer agents, lacking such sophistication, nevertheless have some advantages
+when it comes to brute force search strategies. Programs such Deep Blue have shown that a brute force approach to game
+tree search can be effective, even against humans with vastly superior abstract reasoning capabilities, provided that
+significant computational resources are available and wisely used~\cite{campbell02deep}.
+
+Russell and Wolfe have performed some analyses on kriegspiel positions for which it possible to prove that one player
+has a forced win (that is, regardless of the configuration of the opponent's unseen pieces, the player has a sequence of
+moves that is guaranteed to produce a checkmate)~\cite{russell05efficient, wolfe07exploiting}.  The authors assume
+knowledge of the player's belief state.  They claim that under certain "aggressive" styles of play, it is uncommon for
+information sets to exceed 10,000 nodes in size.  They throw out (i.e., do not analyze) any positions that exceed this
+threshold.
+
+In our experiments, we have found positions in which the size of the information set is many tens of millions.  Our goal
+is to develop strategies that are amenable to use in this more general setting.   In other words, rather than being able
+to analyze positions only at the beginning or end of the game (when the size of the information set tends to be
+smaller), we want to be able to develop a strategy that can be used in the middle of the game as well.  game).
+
+Nance \etal, approach the game from the perspective of {\em logical filtering.}~\cite{nance06reasoning}  The goal of
+logical filtering is to maintain a compact logical representation of an agent's belief state.  Recall that a belief
+state is a probability distribution over the possible current states of the world.  A belief state for kriegspiel might
+take the form of $(at(whitepawn,a2) \wedge at whitepawn(b2) \ldots \wedge at(whiterook,h1)) \vee (at(whitepawn,a3)
+\wedge \ldots \wedge at(whiterook,h1)) \ldots$  For even moderately sizes information sets, such
+representations can become unwieldy in terms of storage and the cost of updating.  Furthermore, reasoning with such
+representations often requires the use of some theorem-proving mechanism to run logical queries, such
+as $at(blackking,a3?)$. Nance et al. develop some strategies to maintain compact representations under
+sequences of actions and observations.  They give a theoretical analysis of the kinds of problems that can be treated in
+this way.  Unfortunately, kriegspiel in its full complexity is beyond the scope of their algorithm.  In order to reason
+about the belief states in the game, they must assume knowledge of the particular kind of piece that move in each turn.
+
+The information set generation approach that we have presented is certainly also computationally expensive.  And
+representing a full information set (e.g., as a list of sequences of moves that encode a path from the start state to
+the current node) would be similarly expensive in terms of storage space.  However, we have shown that our algorithm is
+also highly scalable.  Furthermore, contrary to the approach in~\cite{nance06reasoning}, our approach is readily
+adaptable to sampling algorithms.  Given a sample size that matches available memory, any single sequence of moves is
+readily extractable without the computational expense of a theorem prover or SAT solver.
+
+\section{Future Work}
+
+First and foremost, it would be interesting to see how our information set generation algorithm would perform as a
+subroutine to an actual kriegspiel playing program.  The utility of generating an information set is ultimately
+determined by how well that information improves a player's decision-making capabilities.  We noted earlier that a
+common use of information sets is to estimate the value of a move by averaging the estimated value of the associated
+descendant for each node in the current information set (or a random sample from the set).  This kind of strategy can be
+applied naively using a belief state sampling algorithm (i.e., an algorithm that samples from the distribution of
+possible worlds in the current state without respect to the overall history mof moves).  However, it has been shown that
+significant improvement in play can be achieved by estimating the value of future moves from the perspective of the
+opponent, based on the knowledge that the opponent would have had at prior positions in the move
+hisotry~\cite{richards07opponent}.  Such a strategy would require the use of an information set generation algorithm of the
+kind that we have described here.\footnote{This work is underway.}
+
+Ideally, we would run our algorithm on some deeper instances from real kriegspiel games.  We are not aware of a
+repository of such games.  Unfortunately, notation for kriegspiel is not standardized and there are several minor
+variations in rules (i.e., with respect to the nature of referee announcements) that are unlikely to significantly to
+impact the computability and scalability results but which nevertheless make it difficult to write an input reader that
+converts the transcript of a game into a sequence of moves readable by our algorithm.  An alternative would be to play
+some games ourselves.  Currently, the manual encoding of games is not difficult but is tedious.  Still, it would be nice
+to have some games that go on for 40 or 50 moves and have positions where the information sets are on the order of a few
+hundred thousand or a few million in size.
+
+The information set generation algorithm that we have implemented here is one the most straightforward one and is
+alluded to in~\cite{parker05game} and~\cite{russell05efficient}.  It is analogous to the ``expand-at-the-tip'' strategy
+for the Hamiltonian Circuit problem.  An alternative information set generation algorithm is described in
+~\cite{richards09information} and could potentially improve the search efficiency greatly by exploiting variable
+ordering.  This would be analgous to the disconnected edge-pairs\footnote{what we referred to as ``algorithm 2'' in
+class} heuristic for the Hamiltonian Circuit problem.  Rather than search from the root node going forward, this
+algorithm would seek opportunities for pruning based on the propagation of logical consequences from each move both
+forward and backward in time.  This, incidentally, is more consistent with the way human players would analyze the game.
+For example, if a player's proposed bishop move is rejected and the bishop is sufficiently removed from its own king,
+then the player may infer that one of his opponent's pieces is along that rejected path.  At least one of those spaces
+must have been the destination of a prior move by the opponent.  If a recent move by that bishop crossed the same path,
+then that would be hard evidence that those squares were empty at the time.  And therefore, the arrival of the
+opponent's piece must have been between the previous (successful) bishop move and the current (unsuccessful) bishop
+move.  To implement this algorithm, we envision the use of planning graph type data structures ~\cite{GRAPHPLAN}.
+Planning graphs have a sequence of alternating levels of state constraints and action constraints.  State constraints
+would be of the form $at(bishop,d4) \vee at(bishop,e6)$ or $!occupied(e4)$.  Action constraints would be of the form
+$move(knight,a4,c5) \vee move(knight,e6)$.  At a minimum, each level would require a data structure with a bit for each
+possible action/state.  Propagating the constraints would require much of the same machinery used in STRIPS-like
+planning algorithms~\cite{STRIPS}, and one of the many knobs that would need to be tuned would be the depth of the
+propagation for each constraint.  (Should the consequences of a known action or non-action be propagated one move into
+the future and one move into the past?  More?  Less?)  Additionally, there would be a significant and interesting
+tradeoff in the effectiveness of pruning and the and the size of the data structures that would have to be encoded in
+each chare.  The best algorithm might be a hybrid strategy that does forward propagation from the root at the parallel
+level and constraint propagation at the sequential level.\footnote{This work is also in progress.}
+
+There are certainly several places where the efficiency of the details of each node expansion could be improved.  Many
+of these would reduce the overall running time but would not necessarily be interesting from the standpoint of
+evaluating the efficiency and isoefficiency of parallelization.  For example, in checking to see if a position leaves a
+player in check, we currently utilize existing helper functions that generate possible moves for the opponent.  It would
+be possible to write a more specialized function that focuses only on the position of the king in question and looks
+only at the squares along its rank, file, diagonals, and knight-edges for threats.  There are also some memory
+allocation issues that could be improved.
 
@@
 \caption{Solution counts and running times for a sample kriegspiel game}
-\label{white times}
+\label{whitetimes}
 \end{table}
@@
 
+\begin{figure}
+\begin{verbatim}
+1. Pa2:a4 Pa7:a5 2. Pd2:d4 Pb7:b6 3. Bc1:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+1. Pd2:d3 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pd3:d4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+1. Pd2:d4 Pa7:a5 2. Pa2:a4 Pb7:b6 3. Bc1:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+1. Pd2:d4 Pa7:a5 2. Bc1:f4 Pb7:b6 3. Bf4:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pb2:b3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pc2:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:c1 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ke1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+\end{verbatim}
+\caption{Abbreviated output from example game.  Generating information sets for black after five turns by each player.
+From this, black can infer that there is definitely a white bishop at g5 and a white pawn at d6.}
+\label{abbrevoutput}
+\end{figure}
+
+\begin{figure}
+\begin{verbatim}
+19
+Moves:
+
+**********************************************
+R N B Q K B N R 
+P P P P P P P P 
+- - - - - - - - 
+- - - - - - - - 
+- - - - - - - - 
+- - - - - - - - 
+p p p p p p p p 
+r n b q k b n r 
+**********************************************
+Failed moves: 0
+from: 51 to: 35 blocked: 0 checked: 0
+
+**********************************************
+R N B Q K B N R 
+P P P P P P P P 
+- - - - - - - - 
+- - - - - - - - 
+- - - p - - - - 
+- - - - - - - - 
+p p p - p p p p 
+r n b q k b n r 
+**********************************************
+Failed moves: 0
+from: 8 to: 24 blocked: 0 checked: 0
+
+**********************************************
+R N B Q K B N R 
+- P P P P P P P 
+- - - - - - - - 
+P - - - - - - - 
+- - - p - - - - 
+- - - - - - - - 
+p p p - p p p p 
+r n b q k b n r 
+**********************************************
+Failed moves: 0
+from: 58 to: 30 blocked: 0 checked: 0
+
+**********************************************
+R N B Q K B N R 
+- P P P P P P P 
+- - - - - - - - 
+P - - - - - b - 
+- - - p - - - - 
+- - - - - - - - 
+p p p - p p p p 
+r n - q k b n r 
+**********************************************
+Failed moves: 0
+from: 9 to: 17 blocked: 0 checked: 0
+
+**********************************************
+R N B Q K B N R 
+- - P P P P P P 
+- P - - - - - - 
+P - - - - - b - 
+- - - p - - - - 
+- - - - - - - - 
+p p p - p p p p 
+r n - q k b n r 
+**********************************************
+Failed moves: 1
+from: 57 to: 42 blocked: 0 checked: 0
+
+**********************************************
+R N B Q K B N R 
+- - P P P P P P 
+- P - - - - - - 
+P - - - - - b - 
+- - - p - - - - 
+- - n - - - - - 
+p p p - p p p p 
+r - - q k b n r 
+**********************************************
+Failed moves: 0
+from: 10 to: 26 blocked: 0 checked: 0
+
+**********************************************
+R N B Q K B N R 
+- - - P P P P P 
+- P - - - - - - 
+P - P - - - b - 
+- - - p - - - - 
+- - n - - - - - 
+p p p - p p p p 
+r - - q k b n r 
+**********************************************
+Failed moves: 1
+from: 35 to: 27 blocked: 0 checked: 0
+
+**********************************************
+R N B Q K B N R 
+- - - P P P P P 
+- P - - - - - - 
+P - P p - - b - 
+- - - - - - - - 
+- - n - - - - - 
+p p p - p p p p 
+r - - q k b n r 
+**********************************************
+Failed moves: 0
+from: 1 to: 16 blocked: 0 checked: 0
+
+**********************************************
+R - B Q K B N R 
+- - - P P P P P 
+N P - - - - - - 
+P - P p - - b - 
+- - - - - - - - 
+- - n - - - - - 
+p p p - p p p p 
+r - - q k b n r 
+**********************************************
+Failed moves: 0
+from: 27 to: 19 blocked: 0 checked: 0
+
+**********************************************
+R - B Q K B N R 
+- - - P P P P P 
+N P - p - - - - 
+P - P - - - b - 
+- - - - - - - - 
+- - n - - - - - 
+p p p - p p p p 
+r - - q k b n r 
+**********************************************
+Failed moves: 0
+from: 13 to: 21 blocked: 0 checked: 0
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - n - - - - - 
+p p p - p p p p 
+r - - q k b n r 
+**********************************************
+BEGINNING INFORMATION SET GENERATION
+1. Pa2:a4 Pa7:a5 2. Pd2:d4 Pb7:b6 3. Bc1:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+p - - - - - - - 
+- - - - - - - - 
+- p p - p p p p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d3 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pd3:d4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - - - - - - - 
+p p p - p p p p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Pa2:a4 Pb7:b6 3. Bc1:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+p - - - - - - - 
+- - - - - - - - 
+- p p - p p p p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:f4 Pb7:b6 3. Bf4:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - - - - - - - 
+p p p - p p p p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+p - - - - - - - 
+- p p - p p p p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+p - - - - - - - 
+- - - - - - - - 
+- p p - p p p p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pb2:b3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- p - - - - - - 
+p - p - p p p p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pc2:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - p - - - - - 
+p p - - p p p p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - - - p - - - 
+p p p - - p p p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - p - - - 
+- - - - - - - - 
+p p p - - p p p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - - - - p - - 
+p p p - p - p p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - p - - 
+- - - - - - - - 
+p p p - p - p p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - - - - - p - 
+p p p - p p - p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - p - 
+- - - - - - - - 
+p p p - p p - p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - - - - - - p 
+p p p - p p p - 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - p 
+- - - - - - - - 
+p p p - p p p - 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+n - - - - - - - 
+p p p - p p p p 
+r - - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - n - - - - - 
+p p p - p p p p 
+r - - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - - - - - - - 
+p p p n p p p p 
+r - - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - - - - - - - 
+p p p q p p p p 
+r n - - k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - - q - - - - 
+p p p - p p p p 
+r n - - k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:c1 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - - - - - - - 
+p p p - p p p p 
+r n q - k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ke1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - - - - - - - 
+p p p k p p p p 
+r n - q - b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - - - - n - - 
+p p p - p p p p 
+r n - q k b - r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - - - - - - n 
+p p p - p p p p 
+r n - q k b - r 
+**********************************************
+Solutions founds: 25
+\end{verbatim}
+\caption{Verbose output from example game.  Generating information sets for black after five turns by each player.
+From this, black can infer that there is definitely a white bishop at g5 and a white pawn at d6.}
+\label{verboseoutput}
+\end{figure}
 \end{document}

### 2010-12-14 paper/paper.tex
Added code listing; saving work

@@
 \usepackage{tikz}
+\usepackage{examplep}
+\usepackage{moreverb}
+\usepackage{listings}
+\lstset{language=C,numbers=left,basicstyle=\footnotesize,numberstyle=\footnotesize}
 \usetikzlibrary{arrows,decorations.pathmorphing,backgrounds,fit,shapes.geometric}
@@
 %\usepackage{ijcai09}
-
-\pagestyle{empty}
 
@@
 
+\newenvironment{mylisting}
+{\begin{list}{}{\setlength{\leftmargin}{1em}}\item\scriptsize}
+{\end{list}}
+
+\newenvironment{mytinylisting}
+{\begin{list}{}{\setlength{\leftmargin}{1em}}\item\scriptsize\texttt}
+{\end{list}}
 \newenvironment{proof}{\vspace{-0.15in}\noindent{\bf Proof:}}%
@@
 \begin{figure}
+\begin{minipage}{0.5\linewidth}
 \centering
-%\begin{minipage}{.4\textwidth}
-\includegraphics[width=0.7\textwidth]{plots/3depths.pdf}
-%\end{minipage}
-%\begin{minipage}{.4\textwidth}
-\includegraphics[width=0.7\textwidth]{plots/3depths_nolog.pdf}
-%\end{minipage}
+\includegraphics[width=\textwidth]{plots/3depths.pdf}
+\end{minipage}
+\begin{minipage}{0.5\linewidth}
+\centering
+\includegraphics[width=\textwidth]{plots/3depthsnolog.pdf}
+\end{minipage}
 \caption{3depths}
@@
 \begin{figure}
-\begin{verbatim}
-1. Pa2:a4 Pa7:a5 2. Pd2:d4 Pb7:b6 3. Bc1:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
-1. Pd2:d3 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pd3:d4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
-1. Pd2:d4 Pa7:a5 2. Pa2:a4 Pb7:b6 3. Bc1:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
-1. Pd2:d4 Pa7:a5 2. Bc1:f4 Pb7:b6 3. Bf4:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pb2:b3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pc2:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:c1 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ke1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
-\end{verbatim}
+\begin{mytinylisting}
+1. Pa2:a4 Pa7:a5 2. Pd2:d4 Pb7:b6 3. Bc1:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+1. Pd2:d3 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pd3:d4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+1. Pd2:d4 Pa7:a5 2. Pa2:a4 Pb7:b6 3. Bc1:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+1. Pd2:d4 Pa7:a5 2. Bc1:f4 Pb7:b6 3. Bf4:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pb2:b3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pc2:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:c1 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ke1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+\end{mytinylisting}
 \caption{Abbreviated output from example game.  Generating information sets for black after five turns by each player.
@@
 
-\begin{figure}
+%\begin{figure}
 \begin{verbatim}
@@
 \end{verbatim}
-\caption{Verbose output from example game.  Generating information sets for black after five turns by each player.
-From this, black can infer that there is definitely a white bishop at g5 and a white pawn at d6.}
-\label{verboseoutput}
+%\caption{Verbose output from example game.  Generating information sets for black after five turns by each player.
+%From this, black can infer that there is definitely a white bishop at g5 and a white pawn at d6.}
+%\label{verboseoutput}
+%\end{figure}
+
+\begin{figure}
+%\begin{listing}
+\begin{lstlisting}
+void generateInformationSet(uint16_t* trueState, uint16_t* possState, bool whiteMove, 
+	uint16_t* possHistory, uint16_t** levels, int depth)
+{
+  if (!samePawnTries(trueState, possState, whiteMove)) return;  
+  if (!sameCheckStatus(trueState, possState, whiteMove)) return; 
+  if (depth == maxdepth) { // Then we have found a solution
+	nSolutions++; return; // Report solution
+  }
+  uint16_t newPossState[16]; 
+  int nMoves = 0;
+  generateAttemptableMoves(possState, whiteMove, levels[depth], nMoves,false);
+  checkForCheck(possState, whiteMove, levels[depth], nMoves); // prunes if not equal
+  assert (nMoves < NMOVES); //
+
+  if (whitePerspective == whiteMove) { // move from perspective plyaer 
+    SetMove& failures = failedMoves[depth];
+    for (SetMove::const_iterator itr = failures.begin(); itr != failures.end(); ++itr) {
+      uint16_t move = *itr ;
+      if (!foundMatchingMove(move,levels[depth],nMoves)) {
+        return; // Prune: one of my own failed moves is not attemptable in this state
+      }
+    }
+    // We know exactly what the actual move was; make it
+    uint16_t& actualMove = moveHistory[depth];
+    if (!foundMatchingMove(actualMove,levels[depth],nMoves)) {
+      return; // Prune (because the move that we know we made at this depth is not legal)
+    }
+    // Otherwise, recurse 
+    applyMove(possState,newPossState,actualMove);
+    possHistory[depth] = actualMove;
+    generateInformationSet(newTrueState, newPossState, !whiteMove, 
+      possHistory, levels, depth+1);
+  } else { 
+    // We are at a level in the search tree where 
+    // we are considering the possible moves for the opponent.
+    // We know the number of attempted illegal moves, but not which ones
+
+    // Must be enough attemptable moves in the state to match the number of failed attempts
+    unsigned nIllegalMoves = countIllegalMoves(levels[depth], nMoves);
+    if (nIllegalMoves < failedMoves[depth].size()) {
+	return; // Prune if not enough attemptable moves
+    }
+    // Now we want to try each possible move that is legally executable (not just attemptable) 
+    for (int i = 0; i < nMoves; i++) {
+      uint16_t& move = levels[depth][i];
+      if (isLegal(move)) { // Obviously, we can only execute the moves that are actually legal from this state
+        applyMove(possState,newPossState,move);
+        possHistory[depth] = move;
+        generateInformationSet(newTrueState, newPossState, !whiteMove, 
+          possHistory, levels, depth+1);
+      }
+    }
+  }
+}
+\end{lstlisting}
+\caption{The heart of the information set generation algorithm.}
+\label{codelisting}
 \end{figure}
+
+Figure~\ref{codelisting}
 \end{document}

### 2010-12-14 paper/paper.tex
Reordered several elements and combined a table

@@
 bishop at g5 and a pawn at d6.  There are 25 possible sequences of moves that are consistent with black's observations
-up to that point. Tables~\ref{whitetimes} and ~\ref{blacktimes} show the size of the information sets for each player
-for each move in the game and the amount of time to find the full set on a single processor.  Figure~\ref{verboseoutput}
-shows the full output of the program for the same point in the game (five moves for each player) and illustrates all of
-the possible positions in the belief state.
-
-\section{Performance Results}
-
-Table~\ref{speedups} shows the performance results for our algorithm on a randomly selected problem instance in which
-the size of the information set was about 4.5 million.  (This instance is not necessarily representative of all games or
-states of the game, but it was not cherry-picked in any way.)
-
-\begin{table}
-\centering
-\begin{tabular}{rrr}
-{\bf Procs}	&	{\bf Time (s)} 	&	{\bf Speedup}\\
-1	&	1524	&	1\\
-2	&	812	&	1.88\\
-4	&	407	&	3.74\\
-8	&	208	&	7.32\\
-16	&	102	&	14.8\\
-32	&	53.2	&	28.7\\
-64	&	28.0	&	54.5\\
-128	&	14.9	&	102\\
-256	&	7.70	&	197\\
-512	&	4.59	&	331\\
-1024	&	2.64	&	576\\
-\end{tabular}
-\caption{Speedups for a randomly selected problem instance on 1--1024 processors.  The size of the information set in
-this instance is 4,495,121.}
-\label{speedups}
-\end{table}
-
-Figure~\ref{3schemes}
-\begin{figure}
-\centering
-\includegraphics[width=0.7\textwidth]{plots/3schemes.pdf}
-\caption{3schemes}
-\label{3schemes}
-\end{figure}
-
-Figure~\ref{3depths}
-\begin{figure}
-\begin{minipage}{0.5\linewidth}
-\centering
-\includegraphics[width=\textwidth]{plots/3depths.pdf}
-\end{minipage}
-\begin{minipage}{0.5\linewidth}
-\centering
-\includegraphics[width=\textwidth]{plots/3depthsnolog.pdf}
-\end{minipage}
-\caption{3depths}
-\label{3depths}
-\end{figure}
-
-Figure~\ref{dep8}
-\begin{figure}
-\centering
-\includegraphics[width=0.7\textwidth]{plots/dep8.pdf}
-\caption{dep8}
-\label{dep8}
-\end{figure}
-
-Figure~\ref{dep10}
-\begin{figure}
-\centering
-\includegraphics[width=0.7\textwidth]{plots/dep10.pdf}
-\caption{dep10}
-\label{dep10}
-\end{figure}
-
-Figure~\ref{dep12}
-\begin{figure}
-\centering
-\includegraphics[width=0.7\textwidth]{plots/dep12.pdf}
-\caption{dep12}
-\label{dep12}
-\end{figure}
-
-\section{Related Work}
-An overview of the theory of imperfect information games and other basic topics in game theory can be found in
-~\cite{kuhn03lectures} and~\cite{kuhn97classics}.  A technical explanation of the concept of information sets can be
-found in~\cite{gilpin07algorithms}.  Algorithms for ``solving'' imperfect information games---in the sense of finding a
-Nash equilibrium---can be found in~\cite{koller94fast} and~\cite{koller97representations}.  Koller \etal, showed that an
-equilibrium can be found in time that is polynomial in the size of the game tree.  And in fact, they found that it took
-about as much time to generate the full game tree as it did to actually solve it.  Unfortunately, this is feasible only
-for small games.  In larger games where theoretically optimal strategies cannot be computed efficiently, computer
-systems tend to utilize the concept of infomration sets in some form or another.  A common strategy is to estimate the
-value of a move by averaging the estimated value of the associated descendant for each node in the current information
-set (or a random sample from the set).  While this sort of ``averaging over clairvoyance'' {\em can} lead to very poor
-decisions, it is nevertheless a reasonable strategy in some games (e.g., Scrabble~\cite{sheppard02world}).
-
-In this work, we have argued for an information set generation approach to kriegspiel, but other authors have approached
-the problem differently.  Parker \etal, consider the problem of sampling from belief states in a variety of large
-game trees, including kriegspiel~\cite{parker05game}.  They show that some performance gains can be achieved by only
-approximate sampling from the belief state.  Instead of generating the full information set, or even sampling from it,
-they sample from possible positions for the current state (i.e., sampling from the belief state) by matching only the
-most recent observations.  This motivation for this approach seems to be that it would not be feasible to sample from
-the full information set, because of the computational expense involved.  And indeed, our experiments bear this out-- it
-can indeed be expensive to produce the full information set.  For the experiments shown in Table XX, generating the full
-information set took over 25 minutes on a single processor.  However, by utilizing the Charm++ search engine on a
-parallel machine (1024 processors), we were able to generate the full set in just 2.9 seconds.  (Note that this set had
-more than 4 million members.)  Our results suggest that the problem is highly scalable and our approach might therefore
-be preferred.
-
-Li gives an overview of kriegspiel and discusses strategy from the perspective of human players~\cite{li94chess}.  He
-walks through the analysis of several actual games between two human players and discusses the inferences that each
-player can and should make at each position.  As in chess, human players are able to reason about the game at an
-abstract level; they tend to focus on a small number of critical possibilities and to emphasize the most important
-pieces and squares in each position.  Computer agents, lacking such sophistication, nevertheless have some advantages
-when it comes to brute force search strategies. Programs such Deep Blue have shown that a brute force approach to game
-tree search can be effective, even against humans with vastly superior abstract reasoning capabilities, provided that
-significant computational resources are available and wisely used~\cite{campbell02deep}.
-
-Russell and Wolfe have performed some analyses on kriegspiel positions for which it possible to prove that one player
-has a forced win (that is, regardless of the configuration of the opponent's unseen pieces, the player has a sequence of
-moves that is guaranteed to produce a checkmate)~\cite{russell05efficient, wolfe07exploiting}.  The authors assume
-knowledge of the player's belief state.  They claim that under certain "aggressive" styles of play, it is uncommon for
-information sets to exceed 10,000 nodes in size.  They throw out (i.e., do not analyze) any positions that exceed this
-threshold.
-
-In our experiments, we have found positions in which the size of the information set is many tens of millions.  Our goal
-is to develop strategies that are amenable to use in this more general setting.   In other words, rather than being able
-to analyze positions only at the beginning or end of the game (when the size of the information set tends to be
-smaller), we want to be able to develop a strategy that can be used in the middle of the game as well.  game).
-
-Nance \etal, approach the game from the perspective of {\em logical filtering.}~\cite{nance06reasoning}  The goal of
-logical filtering is to maintain a compact logical representation of an agent's belief state.  Recall that a belief
-state is a probability distribution over the possible current states of the world.  A belief state for kriegspiel might
-take the form of $(at(whitepawn,a2) \wedge at whitepawn(b2) \ldots \wedge at(whiterook,h1)) \vee (at(whitepawn,a3)
-\wedge \ldots \wedge at(whiterook,h1)) \ldots$  For even moderately sizes information sets, such
-representations can become unwieldy in terms of storage and the cost of updating.  Furthermore, reasoning with such
-representations often requires the use of some theorem-proving mechanism to run logical queries, such
-as $at(blackking,a3?)$. Nance et al. develop some strategies to maintain compact representations under
-sequences of actions and observations.  They give a theoretical analysis of the kinds of problems that can be treated in
-this way.  Unfortunately, kriegspiel in its full complexity is beyond the scope of their algorithm.  In order to reason
-about the belief states in the game, they must assume knowledge of the particular kind of piece that move in each turn.
-
-The information set generation approach that we have presented is certainly also computationally expensive.  And
-representing a full information set (e.g., as a list of sequences of moves that encode a path from the start state to
-the current node) would be similarly expensive in terms of storage space.  However, we have shown that our algorithm is
-also highly scalable.  Furthermore, contrary to the approach in~\cite{nance06reasoning}, our approach is readily
-adaptable to sampling algorithms.  Given a sample size that matches available memory, any single sequence of moves is
-readily extractable without the computational expense of a theorem prover or SAT solver.
-
-\section{Future Work}
-
-First and foremost, it would be interesting to see how our information set generation algorithm would perform as a
-subroutine to an actual kriegspiel playing program.  The utility of generating an information set is ultimately
-determined by how well that information improves a player's decision-making capabilities.  We noted earlier that a
-common use of information sets is to estimate the value of a move by averaging the estimated value of the associated
-descendant for each node in the current information set (or a random sample from the set).  This kind of strategy can be
-applied naively using a belief state sampling algorithm (i.e., an algorithm that samples from the distribution of
-possible worlds in the current state without respect to the overall history mof moves).  However, it has been shown that
-significant improvement in play can be achieved by estimating the value of future moves from the perspective of the
-opponent, based on the knowledge that the opponent would have had at prior positions in the move
-hisotry~\cite{richards07opponent}.  Such a strategy would require the use of an information set generation algorithm of the
-kind that we have described here.\footnote{This work is underway.}
-
-Ideally, we would run our algorithm on some deeper instances from real kriegspiel games.  We are not aware of a
-repository of such games.  Unfortunately, notation for kriegspiel is not standardized and there are several minor
-variations in rules (i.e., with respect to the nature of referee announcements) that are unlikely to significantly to
-impact the computability and scalability results but which nevertheless make it difficult to write an input reader that
-converts the transcript of a game into a sequence of moves readable by our algorithm.  An alternative would be to play
-some games ourselves.  Currently, the manual encoding of games is not difficult but is tedious.  Still, it would be nice
-to have some games that go on for 40 or 50 moves and have positions where the information sets are on the order of a few
-hundred thousand or a few million in size.
-
-The information set generation algorithm that we have implemented here is one the most straightforward one and is
-alluded to in~\cite{parker05game} and~\cite{russell05efficient}.  It is analogous to the ``expand-at-the-tip'' strategy
-for the Hamiltonian Circuit problem.  An alternative information set generation algorithm is described in
-~\cite{richards09information} and could potentially improve the search efficiency greatly by exploiting variable
-ordering.  This would be analgous to the disconnected edge-pairs\footnote{what we referred to as ``algorithm 2'' in
-class} heuristic for the Hamiltonian Circuit problem.  Rather than search from the root node going forward, this
-algorithm would seek opportunities for pruning based on the propagation of logical consequences from each move both
-forward and backward in time.  This, incidentally, is more consistent with the way human players would analyze the game.
-For example, if a player's proposed bishop move is rejected and the bishop is sufficiently removed from its own king,
-then the player may infer that one of his opponent's pieces is along that rejected path.  At least one of those spaces
-must have been the destination of a prior move by the opponent.  If a recent move by that bishop crossed the same path,
-then that would be hard evidence that those squares were empty at the time.  And therefore, the arrival of the
-opponent's piece must have been between the previous (successful) bishop move and the current (unsuccessful) bishop
-move.  To implement this algorithm, we envision the use of planning graph type data structures ~\cite{GRAPHPLAN}.
-Planning graphs have a sequence of alternating levels of state constraints and action constraints.  State constraints
-would be of the form $at(bishop,d4) \vee at(bishop,e6)$ or $!occupied(e4)$.  Action constraints would be of the form
-$move(knight,a4,c5) \vee move(knight,e6)$.  At a minimum, each level would require a data structure with a bit for each
-possible action/state.  Propagating the constraints would require much of the same machinery used in STRIPS-like
-planning algorithms~\cite{STRIPS}, and one of the many knobs that would need to be tuned would be the depth of the
-propagation for each constraint.  (Should the consequences of a known action or non-action be propagated one move into
-the future and one move into the past?  More?  Less?)  Additionally, there would be a significant and interesting
-tradeoff in the effectiveness of pruning and the and the size of the data structures that would have to be encoded in
-each chare.  The best algorithm might be a hybrid strategy that does forward propagation from the root at the parallel
-level and constraint propagation at the sequential level.\footnote{This work is also in progress.}
-
-There are certainly several places where the efficiency of the details of each node expansion could be improved.  Many
-of these would reduce the overall running time but would not necessarily be interesting from the standpoint of
-evaluating the efficiency and isoefficiency of parallelization.  For example, in checking to see if a position leaves a
-player in check, we currently utilize existing helper functions that generate possible moves for the opponent.  It would
-be possible to write a more specialized function that focuses only on the position of the king in question and looks
-only at the squares along its rank, file, diagonals, and knight-edges for threats.  There are also some memory
-allocation issues that could be improved.
-
-\section{Appendix}
-Figure~\ref{fulltree} shows the full game tree for a simple variant of poker known as Kuhn poker.  There are three
-cards: a king, a queen, and a jack.  The dealer deals one card to each of two players.  Each player antes one unit, and
-there is one simple round of betting, in which the size of the pot may be doubled.  Left branches denote check/fold;
-right branches denote bet/call.  For this zero-sum game, terminal nodes are labeled with payoffs to player 1, whose
-decision points are shown in triangles.  Connections between nodes in the same information set are shown with dotted
-lines.
-
-\begin{figure}
-\scalebox{.72}{
-\begin{tikzpicture}
-  [chance/.style={circle,draw=blue!50,fill=blue!20,thick,minimum size = 10mm},
-   terminal/.style={rectangle,draw=black!50,fill=black!20,thick, minimum size = 7mm},
-   maxer/.style={shape=regular polygon, regular polygon sides=3,draw=red!50,fill=red!30,thick,minimum size = 10mm,inner sep = 0pt},
-   miner/.style={shape=diamond,draw=green!50,fill=green!30,thick,minimum size = 10mm, inner sep = 0pt},
-   edge from parent/.style={red,thick,draw}, 
-   parent anchor=south,child anchor=north,
-   level 1/.style={sibling distance=4cm,level distance=1.4cm,
-       		growth parent anchor=south},
-   level 2/.style={sibling distance=2cm},
-   level 3/.style={sibling distance=1cm},
-   level 4/.style={sibling distance=1.0cm}]
-		\node  [chance] {}
-		    child {node (M1) [maxer] {\scalebox{.75}{K}}
-			child {node (M11) [miner] {\scalebox{.75}{Q-}}
-				child {node (M111) [terminal] {1}}
-				child {node (M112) [maxer] {\scalebox{.75}{K+}}
-					child {node (M1121) [terminal] {-1}}
-					child {node (M1122) [terminal] {2}}}
-			}
-			child {node (M12) [miner] {\scalebox{.75}{Q+}}
-				child {node (M121) [terminal] {1}}
-				child {node (M122) [terminal] {2}}
-			}
-		    }
-		    child {node (M2) [maxer] {\scalebox{.75}{J}}
-			child {node (M21) [miner] {\scalebox{.75}{Q-}}
-				child {node (M211) [terminal] {-1}}
-				child {node (M212) [maxer] {\scalebox{.75}{J+}}
-					child {node (M2121) [terminal] {-1}}
-					child {node (M2122) [terminal] {-2}}}
-			}
-			child {node (M22) [miner] {\scalebox{.75}{Q+}}
-				child {node (M221) [terminal] {1}}
-				child {node (M222) [terminal] {-2}}
-			}
-		    }
-		    child {node (M3) [maxer] {\scalebox{.75}{K}}
-			child {node (M31) [miner] {\scalebox{.75}{J-}}
-				child {node (M311) [terminal] {1}}
-				child {node (M312) [maxer] {\scalebox{.75}{K+}}
-					child {node (M3121) [terminal] {-1}}
-					child {node (M3122) [terminal] {2}}}
-			}
-			child {node (M32) [miner] {\scalebox{.75}{J+}}
-				child {node (M321) [terminal] {1}}
-				child {node (M322) [terminal] {2}}
-			}
-		    }
-		    child {node (M4) [maxer] {\scalebox{.75}{Q}}
-			child {node (M41) [miner] {\scalebox{.75}{J-}}
-				child {node (M411) [terminal] {1}}
-				child {node (M412) [maxer] {\scalebox{.75}{Q+}}
-					child {node (M4121) [terminal] {-1}}
-					child {node (M4122) [terminal] {2}}}
-			}
-			child {node (M42) [miner] {\scalebox{.75}{J+}}
-				child {node (M421) [terminal] {1}}
-				child {node (M422) [terminal] {2}}
-			}
-		    }
-		    child {node (M5) [maxer] {\scalebox{.75}{J}}
-			child {node (M51) [miner] {\scalebox{.75}{K-}}
-				child {node (M511) [terminal] {-1}}
-				child {node (M512) [maxer] {\scalebox{.75}{J+}}
-					child {node (M5121) [terminal] {-1}}
-					child {node (M5122) [terminal] {-2}}}
-			}
-			child {node (M52) [miner] {\scalebox{.75}{K+}}
-				child {node (M521) [terminal] {1}}
-				child {node (M522) [terminal] {-2}}
-			}
-		    }
-		    child {node (M6) [maxer] {\scalebox{.75}{Q}}
-			child {node (M61) [miner] {\scalebox{.75}{K-}}
-				child {node (M611) [terminal] {-1}}
-				child {node (M612) [maxer] {\scalebox{.75}{Q+}}
-					child {node (M6121) [terminal] {-1}}
-					child {node (M6122) [terminal] {-2}}}
-			}
-			child {node (M62) [miner] {\scalebox{.75}{K+}}
-				child {node (M621) [terminal] {1}}
-				child {node (M622) [terminal] {-2}}
-			}
-		    };
-		\draw  (M1.north) to [dashed,black,out=15,in=165] (M3.north);
-		\draw  (M2.north) to [dashed,black,out=15,in=165] (M5.north);
-		\draw  (M4.north) to [dashed,black,out=15,in=165] (M6.north);
-		\draw  (M112.north) to [dashed,black,out=15,in=165] (M312.north);
-		\draw  (M212.north) to [dashed,black,out=15,in=165] (M512.north);
-		\draw  (M412.north) to [dashed,black,out=15,in=165] (M612.north);
-		\draw  (M11.north) to [dashed,black,out=15,in=165] (M21.north);
-		\draw  (M12.north) to [dashed,black,out=15,in=165] (M22.north);
-		\draw  (M31.north) to [dashed,black,out=15,in=165] (M41.north);
-		\draw  (M32.north) to [dashed,black,out=15,in=165] (M42.north);
-		\draw  (M51.north) to [dashed,black,out=15,in=165] (M61.north);
-		\draw  (M52.north) to [dashed,black,out=15,in=165] (M62.north);
-\end{tikzpicture}
-} %scalebox
-\caption{}
-\label{fulltree}
-\end{figure}
-
-Table~\ref{sampletimes}
-\begin{table}
-\begin{tabular}{ccc}
-1 & 18 & 0.000\\
-2 & 18 & 0.000\\
-3 & 18 & 0.000\\
-4 & 279 & 0.008\\
-5 & 242 & 0.028\\
-6 & 216 & 0.052\\
-7 & 176 & 0.064\\
-8 & 3406 & 0.092\\
-9 & 2191 & 0.288\\
-10 & 1423 & 0.508\\
-11 & 1363 & 0.596\\
-12 & 1416 & 0.744\\
-13 & 1416 & 0.836\\
-14 & 3440 & 0.97\\
-15 & 3428 & 1.25\\
-16 & 50521 & 1.82\\
-17 & 43192 & 5.91\\
-18 & 26128 & 7.94\\
-19 & 19061 & 9.92\\
-\end{tabular}
-\caption{Solution counts and running times for a sample kriegspiel game}
-\label{whitetimes}
-\end{table}
-
-Table~\ref{sampletimes}
-\begin{table}
-\begin{tabular}{ccc}
-1 & 20 & .000\\
-2 & 19 & .000\\
-3 & 404 & .008\\
-4 & 401 & .024\\
-5 & 1472 & .044\\
-6 & 155 & .096\\
-7 & 293 & .116\\
-8 & 158 & .128\\
-9 & 118 & .148\\
-10 & 25 & .152\\
-11 & 798 & .164\\
-12 & 518 & .192\\
-13 & 13564 & .304\\
-14 & 12394 & .776\\
-15 & 343652 & 3.75\\
-16 & 320704 & 17.3\\
-17 & 490162 & 99.9\\
-18 & 3792 & 119\\
-19 & 14836 & 121 \\
-\end{tabular}
-\caption{Solution counts and running times for a sample kriegspiel game}
-\label{blacktimes}
-\end{table}
-
-%\nocite{*}
-\nocite{richards09information}
-\nocite{russell05efficient}
-\nocite{kuhn03lectures}
-\nocite{kuhn97classics}
-\nocite{parker05sampling}
-\nocite{nance06reasoning}
-\nocite{li94chess}
-
-\bibliographystyle{acm}
-\bibliography{paper}
-%\begin{thebibliography}{99}
-%\end{thebibliography}
+up to that point. Table~\ref{bothtimes} shows the size of the information sets for each player for each move in the game
+and the amount of time to find the full set on a single processor.  Appendix B shows the full output of the program for
+the same point in the game (five moves for each player) and illustrates all of the possible positions in the belief
+state.
 
@@
 
-%\begin{figure}
-\begin{verbatim}
-19
-Moves:
-
-**********************************************
-R N B Q K B N R 
-P P P P P P P P 
-- - - - - - - - 
-- - - - - - - - 
-- - - - - - - - 
-- - - - - - - - 
-p p p p p p p p 
-r n b q k b n r 
-**********************************************
-Failed moves: 0
-from: 51 to: 35 blocked: 0 checked: 0
-
-**********************************************
-R N B Q K B N R 
-P P P P P P P P 
-- - - - - - - - 
-- - - - - - - - 
-- - - p - - - - 
-- - - - - - - - 
-p p p - p p p p 
-r n b q k b n r 
-**********************************************
-Failed moves: 0
-from: 8 to: 24 blocked: 0 checked: 0
-
-**********************************************
-R N B Q K B N R 
-- P P P P P P P 
-- - - - - - - - 
-P - - - - - - - 
-- - - p - - - - 
-- - - - - - - - 
-p p p - p p p p 
-r n b q k b n r 
-**********************************************
-Failed moves: 0
-from: 58 to: 30 blocked: 0 checked: 0
-
-**********************************************
-R N B Q K B N R 
-- P P P P P P P 
-- - - - - - - - 
-P - - - - - b - 
-- - - p - - - - 
-- - - - - - - - 
-p p p - p p p p 
-r n - q k b n r 
-**********************************************
-Failed moves: 0
-from: 9 to: 17 blocked: 0 checked: 0
-
-**********************************************
-R N B Q K B N R 
-- - P P P P P P 
-- P - - - - - - 
-P - - - - - b - 
-- - - p - - - - 
-- - - - - - - - 
-p p p - p p p p 
-r n - q k b n r 
-**********************************************
-Failed moves: 1
-from: 57 to: 42 blocked: 0 checked: 0
-
-**********************************************
-R N B Q K B N R 
-- - P P P P P P 
-- P - - - - - - 
-P - - - - - b - 
-- - - p - - - - 
-- - n - - - - - 
-p p p - p p p p 
-r - - q k b n r 
-**********************************************
-Failed moves: 0
-from: 10 to: 26 blocked: 0 checked: 0
-
-**********************************************
-R N B Q K B N R 
-- - - P P P P P 
-- P - - - - - - 
-P - P - - - b - 
-- - - p - - - - 
-- - n - - - - - 
-p p p - p p p p 
-r - - q k b n r 
-**********************************************
-Failed moves: 1
-from: 35 to: 27 blocked: 0 checked: 0
-
-**********************************************
-R N B Q K B N R 
-- - - P P P P P 
-- P - - - - - - 
-P - P p - - b - 
-- - - - - - - - 
-- - n - - - - - 
-p p p - p p p p 
-r - - q k b n r 
-**********************************************
-Failed moves: 0
-from: 1 to: 16 blocked: 0 checked: 0
-
-**********************************************
-R - B Q K B N R 
-- - - P P P P P 
-N P - - - - - - 
-P - P p - - b - 
-- - - - - - - - 
-- - n - - - - - 
-p p p - p p p p 
-r - - q k b n r 
-**********************************************
-Failed moves: 0
-from: 27 to: 19 blocked: 0 checked: 0
-
-**********************************************
-R - B Q K B N R 
-- - - P P P P P 
-N P - p - - - - 
-P - P - - - b - 
-- - - - - - - - 
-- - n - - - - - 
-p p p - p p p p 
-r - - q k b n r 
-**********************************************
-Failed moves: 0
-from: 13 to: 21 blocked: 0 checked: 0
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - n - - - - - 
-p p p - p p p p 
-r - - q k b n r 
-**********************************************
-BEGINNING INFORMATION SET GENERATION
-1. Pa2:a4 Pa7:a5 2. Pd2:d4 Pb7:b6 3. Bc1:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-p - - - - - - - 
-- - - - - - - - 
-- p p - p p p p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d3 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pd3:d4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - - - - - - - 
-p p p - p p p p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Pa2:a4 Pb7:b6 3. Bc1:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-p - - - - - - - 
-- - - - - - - - 
-- p p - p p p p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:f4 Pb7:b6 3. Bf4:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - - - - - - - 
-p p p - p p p p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-p - - - - - - - 
-- p p - p p p p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-p - - - - - - - 
-- - - - - - - - 
-- p p - p p p p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pb2:b3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- p - - - - - - 
-p - p - p p p p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pc2:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - p - - - - - 
-p p - - p p p p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - - - p - - - 
-p p p - - p p p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - p - - - 
-- - - - - - - - 
-p p p - - p p p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - - - - p - - 
-p p p - p - p p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - p - - 
-- - - - - - - - 
-p p p - p - p p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - - - - - p - 
-p p p - p p - p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - p - 
-- - - - - - - - 
-p p p - p p - p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - - - - - - p 
-p p p - p p p - 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - p 
-- - - - - - - - 
-p p p - p p p - 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-n - - - - - - - 
-p p p - p p p p 
-r - - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - n - - - - - 
-p p p - p p p p 
-r - - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - - - - - - - 
-p p p n p p p p 
-r - - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - - - - - - - 
-p p p q p p p p 
-r n - - k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - - q - - - - 
-p p p - p p p p 
-r n - - k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:c1 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - - - - - - - 
-p p p - p p p p 
-r n q - k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ke1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - - - - - - - 
-p p p k p p p p 
-r n - q - b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - - - - n - - 
-p p p - p p p p 
-r n - q k b - r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - - - - - - n 
-p p p - p p p p 
-r n - q k b - r 
-**********************************************
-Solutions founds: 25
-\end{verbatim}
-%\caption{Verbose output from example game.  Generating information sets for black after five turns by each player.
-%From this, black can infer that there is definitely a white bishop at g5 and a white pawn at d6.}
-%\label{verboseoutput}
-%\end{figure}
-
+\begin{table}
+\centering
+\begin{tabular}{rrrrr}
+ & \multicolumn{2}{c}{\bf White} & \multicolumn{2}{c}{\bf Black} \\
+{\bf Ply} & {\bf Size} & {\bf Time} & {\bf Size} & {\bf Time} \\
+1 & 18 & 0.000 & 20 & .000 \\
+2 & 18 & 0.000 & 19 & .000 \\
+3 & 18 & 0.000 & 404 & .008\\
+4 & 279 & 0.008 & 401 & .024\\
+5 & 242 & 0.028 & 1472 & .044\\
+6 & 216 & 0.052 & 155 & .096\\
+7 & 176 & 0.064 & 293 & .116\\
+8 & 3406 & 0.092 & 158 & .128\\
+9 & 2191 & 0.288 & 118 & .148\\
+10 & 1423 & 0.508 & 25 & .152\\
+11 & 1363 & 0.596 & 798 & .164\\
+12 & 1416 & 0.744 & 518 & .192\\
+13 & 1416 & 0.836 & 13564 & .304\\
+14 & 3440 & 0.97 & 12394 & .776\\
+15 & 3428 & 1.25 & 343652 & 3.75\\
+16 & 50521 & 1.82 & 320704 & 17.3\\
+17 & 43192 & 5.91 & 490162 & 99.9\\
+18 & 26128 & 7.94 & 3792 & 119\\
+19 & 19061 & 9.92 & 14836 & 121\\
+\end{tabular}
+\caption{Solution counts and running times for a sample kriegspiel game}
+\label{bothtimes}
+\end{table}
+
+%Table~\ref{sampletimes}
+%\begin{table}
+%\begin{tabular}{ccc}
+%1 & 20 & .000\\
+%2 & 19 & .000\\
+%3 & 404 & .008\\
+%4 & 401 & .024\\
+%5 & 1472 & .044\\
+%6 & 155 & .096\\
+%7 & 293 & .116\\
+%8 & 158 & .128\\
+%9 & 118 & .148\\
+%10 & 25 & .152\\
+%11 & 798 & .164\\
+%12 & 518 & .192\\
+%13 & 13564 & .304\\
+%14 & 12394 & .776\\
+%15 & 343652 & 3.75\\
+%16 & 320704 & 17.3\\
+%17 & 490162 & 99.9\\
+%18 & 3792 & 119\\
+%19 & 14836 & 121 \\
+%\end{tabular}
+%\caption{Solution counts and running times for a sample kriegspiel game}
+%\label{blacktimes}
+%\end{table}
+
+\section{The Information Set Generation Algorithm}
 \begin{figure}
@@
 Figure~\ref{codelisting}
+
+\section{Performance Results}
+
+Table~\ref{speedups} shows the performance results for our algorithm on a randomly selected problem instance in which
+the size of the information set was about 4.5 million.  (This instance is not necessarily representative of all games or
+states of the game, but it was not cherry-picked in any way.)
+
+\begin{table}
+\centering
+\begin{tabular}{rrr}
+{\bf Procs}	&	{\bf Time (s)} 	&	{\bf Speedup}\\
+1	&	1524	&	1\\
+2	&	812	&	1.88\\
+4	&	407	&	3.74\\
+8	&	208	&	7.32\\
+16	&	102	&	14.8\\
+32	&	53.2	&	28.7\\
+64	&	28.0	&	54.5\\
+128	&	14.9	&	102\\
+256	&	7.70	&	197\\
+512	&	4.59	&	331\\
+1024	&	2.64	&	576\\
+\end{tabular}
+\caption{Speedups for a randomly selected problem instance on 1--1024 processors.  The size of the information set in
+this instance is 4,495,121.}
+\label{speedups}
+\end{table}
+
+Figure~\ref{3schemes}
+\begin{figure}
+\centering
+\includegraphics[width=0.7\textwidth]{plots/3schemes.pdf}
+\caption{3schemes}
+\label{3schemes}
+\end{figure}
+
+Figure~\ref{3depths}
+\begin{figure}
+\begin{minipage}{0.5\linewidth}
+\centering
+\includegraphics[width=\textwidth]{plots/3depths.pdf}
+\end{minipage}
+\begin{minipage}{0.5\linewidth}
+\centering
+\includegraphics[width=\textwidth]{plots/3depthsnolog.pdf}
+\end{minipage}
+\caption{3depths}
+\label{3depths}
+\end{figure}
+
+Figure~\ref{dep8}
+\begin{figure}
+\centering
+\includegraphics[width=0.7\textwidth]{plots/dep8.pdf}
+\caption{dep8}
+\label{dep8}
+\end{figure}
+
+Figure~\ref{dep10}
+\begin{figure}
+\centering
+\includegraphics[width=0.7\textwidth]{plots/dep10.pdf}
+\caption{dep10}
+\label{dep10}
+\end{figure}
+
+Figure~\ref{dep12}
+\begin{figure}
+\centering
+\includegraphics[width=0.7\textwidth]{plots/dep12.pdf}
+\caption{dep12}
+\label{dep12}
+\end{figure}
+
+\section{Related Work}
+An overview of the theory of imperfect information games and other basic topics in game theory can be found in
+~\cite{kuhn03lectures} and~\cite{kuhn97classics}.  A technical explanation of the concept of information sets can be
+found in~\cite{gilpin07algorithms}.  Algorithms for ``solving'' imperfect information games---in the sense of finding a
+Nash equilibrium---can be found in~\cite{koller94fast} and~\cite{koller97representations}.  Koller \etal, showed that an
+equilibrium can be found in time that is polynomial in the size of the game tree.  And in fact, they found that it took
+about as much time to generate the full game tree as it did to actually solve it.  Unfortunately, this is feasible only
+for small games.  In larger games where theoretically optimal strategies cannot be computed efficiently, computer
+systems tend to utilize the concept of infomration sets in some form or another.  A common strategy is to estimate the
+value of a move by averaging the estimated value of the associated descendant for each node in the current information
+set (or a random sample from the set).  While this sort of ``averaging over clairvoyance'' {\em can} lead to very poor
+decisions, it is nevertheless a reasonable strategy in some games (e.g., Scrabble~\cite{sheppard02world}).
+
+In this work, we have argued for an information set generation approach to kriegspiel, but other authors have approached
+the problem differently.  Parker \etal, consider the problem of sampling from belief states in a variety of large
+game trees, including kriegspiel~\cite{parker05game}.  They show that some performance gains can be achieved by only
+approximate sampling from the belief state.  Instead of generating the full information set, or even sampling from it,
+they sample from possible positions for the current state (i.e., sampling from the belief state) by matching only the
+most recent observations.  This motivation for this approach seems to be that it would not be feasible to sample from
+the full information set, because of the computational expense involved.  And indeed, our experiments bear this out-- it
+can indeed be expensive to produce the full information set.  For the experiments shown in Table XX, generating the full
+information set took over 25 minutes on a single processor.  However, by utilizing the Charm++ search engine on a
+parallel machine (1024 processors), we were able to generate the full set in just 2.9 seconds.  (Note that this set had
+more than 4 million members.)  Our results suggest that the problem is highly scalable and our approach might therefore
+be preferred.
+
+Li gives an overview of kriegspiel and discusses strategy from the perspective of human players~\cite{li94chess}.  He
+walks through the analysis of several actual games between two human players and discusses the inferences that each
+player can and should make at each position.  As in chess, human players are able to reason about the game at an
+abstract level; they tend to focus on a small number of critical possibilities and to emphasize the most important
+pieces and squares in each position.  Computer agents, lacking such sophistication, nevertheless have some advantages
+when it comes to brute force search strategies. Programs such Deep Blue have shown that a brute force approach to game
+tree search can be effective, even against humans with vastly superior abstract reasoning capabilities, provided that
+significant computational resources are available and wisely used~\cite{campbell02deep}.
+
+Russell and Wolfe have performed some analyses on kriegspiel positions for which it possible to prove that one player
+has a forced win (that is, regardless of the configuration of the opponent's unseen pieces, the player has a sequence of
+moves that is guaranteed to produce a checkmate)~\cite{russell05efficient, wolfe07exploiting}.  The authors assume
+knowledge of the player's belief state.  They claim that under certain "aggressive" styles of play, it is uncommon for
+information sets to exceed 10,000 nodes in size.  They throw out (i.e., do not analyze) any positions that exceed this
+threshold.
+
+In our experiments, we have found positions in which the size of the information set is many tens of millions.  Our goal
+is to develop strategies that are amenable to use in this more general setting.   In other words, rather than being able
+to analyze positions only at the beginning or end of the game (when the size of the information set tends to be
+smaller), we want to be able to develop a strategy that can be used in the middle of the game as well.  game).
+
+Nance \etal, approach the game from the perspective of {\em logical filtering.}~\cite{nance06reasoning}  The goal of
+logical filtering is to maintain a compact logical representation of an agent's belief state.  Recall that a belief
+state is a probability distribution over the possible current states of the world.  A belief state for kriegspiel might
+take the form of $(at(whitepawn,a2) \wedge at whitepawn(b2) \ldots \wedge at(whiterook,h1)) \vee (at(whitepawn,a3)
+\wedge \ldots \wedge at(whiterook,h1)) \ldots$  For even moderately sizes information sets, such
+representations can become unwieldy in terms of storage and the cost of updating.  Furthermore, reasoning with such
+representations often requires the use of some theorem-proving mechanism to run logical queries, such
+as $at(blackking,a3?)$. Nance et al. develop some strategies to maintain compact representations under
+sequences of actions and observations.  They give a theoretical analysis of the kinds of problems that can be treated in
+this way.  Unfortunately, kriegspiel in its full complexity is beyond the scope of their algorithm.  In order to reason
+about the belief states in the game, they must assume knowledge of the particular kind of piece that move in each turn.
+
+The information set generation approach that we have presented is certainly also computationally expensive.  And
+representing a full information set (e.g., as a list of sequences of moves that encode a path from the start state to
+the current node) would be similarly expensive in terms of storage space.  However, we have shown that our algorithm is
+also highly scalable.  Furthermore, contrary to the approach in~\cite{nance06reasoning}, our approach is readily
+adaptable to sampling algorithms.  Given a sample size that matches available memory, any single sequence of moves is
+readily extractable without the computational expense of a theorem prover or SAT solver.
+
+\section{Future Work}
+
+First and foremost, it would be interesting to see how our information set generation algorithm would perform as a
+subroutine to an actual kriegspiel playing program.  The utility of generating an information set is ultimately
+determined by how well that information improves a player's decision-making capabilities.  We noted earlier that a
+common use of information sets is to estimate the value of a move by averaging the estimated value of the associated
+descendant for each node in the current information set (or a random sample from the set).  This kind of strategy can be
+applied naively using a belief state sampling algorithm (i.e., an algorithm that samples from the distribution of
+possible worlds in the current state without respect to the overall history mof moves).  However, it has been shown that
+significant improvement in play can be achieved by estimating the value of future moves from the perspective of the
+opponent, based on the knowledge that the opponent would have had at prior positions in the move
+hisotry~\cite{richards07opponent}.  Such a strategy would require the use of an information set generation algorithm of the
+kind that we have described here.\footnote{This work is underway.}
+
+Ideally, we would run our algorithm on some deeper instances from real kriegspiel games.  We are not aware of a
+repository of such games.  Unfortunately, notation for kriegspiel is not standardized and there are several minor
+variations in rules (i.e., with respect to the nature of referee announcements) that are unlikely to significantly to
+impact the computability and scalability results but which nevertheless make it difficult to write an input reader that
+converts the transcript of a game into a sequence of moves readable by our algorithm.  An alternative would be to play
+some games ourselves.  Currently, the manual encoding of games is not difficult but is tedious.  Still, it would be nice
+to have some games that go on for 40 or 50 moves and have positions where the information sets are on the order of a few
+hundred thousand or a few million in size.
+
+The information set generation algorithm that we have implemented here is one the most straightforward one and is
+alluded to in~\cite{parker05game} and~\cite{russell05efficient}.  It is analogous to the ``expand-at-the-tip'' strategy
+for the Hamiltonian Circuit problem.  An alternative information set generation algorithm is described in
+~\cite{richards09information} and could potentially improve the search efficiency greatly by exploiting variable
+ordering.  This would be analgous to the disconnected edge-pairs\footnote{what we referred to as ``algorithm 2'' in
+class} heuristic for the Hamiltonian Circuit problem.  Rather than search from the root node going forward, this
+algorithm would seek opportunities for pruning based on the propagation of logical consequences from each move both
+forward and backward in time.  This, incidentally, is more consistent with the way human players would analyze the game.
+For example, if a player's proposed bishop move is rejected and the bishop is sufficiently removed from its own king,
+then the player may infer that one of his opponent's pieces is along that rejected path.  At least one of those spaces
+must have been the destination of a prior move by the opponent.  If a recent move by that bishop crossed the same path,
+then that would be hard evidence that those squares were empty at the time.  And therefore, the arrival of the
+opponent's piece must have been between the previous (successful) bishop move and the current (unsuccessful) bishop
+move.  To implement this algorithm, we envision the use of planning graph type data structures ~\cite{GRAPHPLAN}.
+Planning graphs have a sequence of alternating levels of state constraints and action constraints.  State constraints
+would be of the form $at(bishop,d4) \vee at(bishop,e6)$ or $!occupied(e4)$.  Action constraints would be of the form
+$move(knight,a4,c5) \vee move(knight,e6)$.  At a minimum, each level would require a data structure with a bit for each
+possible action/state.  Propagating the constraints would require much of the same machinery used in STRIPS-like
+planning algorithms~\cite{STRIPS}, and one of the many knobs that would need to be tuned would be the depth of the
+propagation for each constraint.  (Should the consequences of a known action or non-action be propagated one move into
+the future and one move into the past?  More?  Less?)  Additionally, there would be a significant and interesting
+tradeoff in the effectiveness of pruning and the and the size of the data structures that would have to be encoded in
+each chare.  The best algorithm might be a hybrid strategy that does forward propagation from the root at the parallel
+level and constraint propagation at the sequential level.\footnote{This work is also in progress.}
+
+There are certainly several places where the efficiency of the details of each node expansion could be improved.  Many
+of these would reduce the overall running time but would not necessarily be interesting from the standpoint of
+evaluating the efficiency and isoefficiency of parallelization.  For example, in checking to see if a position leaves a
+player in check, we currently utilize existing helper functions that generate possible moves for the opponent.  It would
+be possible to write a more specialized function that focuses only on the position of the king in question and looks
+only at the squares along its rank, file, diagonals, and knight-edges for threats.  There are also some memory
+allocation issues that could be improved.
+
+\section{Appendix A}
+Figure~\ref{fulltree} shows the full game tree for a simple variant of poker known as Kuhn poker.  There are three
+cards: a king, a queen, and a jack.  The dealer deals one card to each of two players.  Each player antes one unit, and
+there is one simple round of betting, in which the size of the pot may be doubled.  Left branches denote check/fold;
+right branches denote bet/call.  For this zero-sum game, terminal nodes are labeled with payoffs to player 1, whose
+decision points are shown in triangles.  Connections between nodes in the same information set are shown with dotted
+lines.
+
+\begin{figure}
+\scalebox{.72}{
+\begin{tikzpicture}
+  [chance/.style={circle,draw=blue!50,fill=blue!20,thick,minimum size = 10mm},
+   terminal/.style={rectangle,draw=black!50,fill=black!20,thick, minimum size = 7mm},
+   maxer/.style={shape=regular polygon, regular polygon sides=3,draw=red!50,fill=red!30,thick,minimum size = 10mm,inner sep = 0pt},
+   miner/.style={shape=diamond,draw=green!50,fill=green!30,thick,minimum size = 10mm, inner sep = 0pt},
+   edge from parent/.style={red,thick,draw}, 
+   parent anchor=south,child anchor=north,
+   level 1/.style={sibling distance=4cm,level distance=1.4cm,
+       		growth parent anchor=south},
+   level 2/.style={sibling distance=2cm},
+   level 3/.style={sibling distance=1cm},
+   level 4/.style={sibling distance=1.0cm}]
+		\node  [chance] {}
+		    child {node (M1) [maxer] {\scalebox{.75}{K}}
+			child {node (M11) [miner] {\scalebox{.75}{Q-}}
+				child {node (M111) [terminal] {1}}
+				child {node (M112) [maxer] {\scalebox{.75}{K+}}
+					child {node (M1121) [terminal] {-1}}
+					child {node (M1122) [terminal] {2}}}
+			}
+			child {node (M12) [miner] {\scalebox{.75}{Q+}}
+				child {node (M121) [terminal] {1}}
+				child {node (M122) [terminal] {2}}
+			}
+		    }
+		    child {node (M2) [maxer] {\scalebox{.75}{J}}
+			child {node (M21) [miner] {\scalebox{.75}{Q-}}
+				child {node (M211) [terminal] {-1}}
+				child {node (M212) [maxer] {\scalebox{.75}{J+}}
+					child {node (M2121) [terminal] {-1}}
+					child {node (M2122) [terminal] {-2}}}
+			}
+			child {node (M22) [miner] {\scalebox{.75}{Q+}}
+				child {node (M221) [terminal] {1}}
+				child {node (M222) [terminal] {-2}}
+			}
+		    }
+		    child {node (M3) [maxer] {\scalebox{.75}{K}}
+			child {node (M31) [miner] {\scalebox{.75}{J-}}
+				child {node (M311) [terminal] {1}}
+				child {node (M312) [maxer] {\scalebox{.75}{K+}}
+					child {node (M3121) [terminal] {-1}}
+					child {node (M3122) [terminal] {2}}}
+			}
+			child {node (M32) [miner] {\scalebox{.75}{J+}}
+				child {node (M321) [terminal] {1}}
+				child {node (M322) [terminal] {2}}
+			}
+		    }
+		    child {node (M4) [maxer] {\scalebox{.75}{Q}}
+			child {node (M41) [miner] {\scalebox{.75}{J-}}
+				child {node (M411) [terminal] {1}}
+				child {node (M412) [maxer] {\scalebox{.75}{Q+}}
+					child {node (M4121) [terminal] {-1}}
+					child {node (M4122) [terminal] {2}}}
+			}
+			child {node (M42) [miner] {\scalebox{.75}{J+}}
+				child {node (M421) [terminal] {1}}
+				child {node (M422) [terminal] {2}}
+			}
+		    }
+		    child {node (M5) [maxer] {\scalebox{.75}{J}}
+			child {node (M51) [miner] {\scalebox{.75}{K-}}
+				child {node (M511) [terminal] {-1}}
+				child {node (M512) [maxer] {\scalebox{.75}{J+}}
+					child {node (M5121) [terminal] {-1}}
+					child {node (M5122) [terminal] {-2}}}
+			}
+			child {node (M52) [miner] {\scalebox{.75}{K+}}
+				child {node (M521) [terminal] {1}}
+				child {node (M522) [terminal] {-2}}
+			}
+		    }
+		    child {node (M6) [maxer] {\scalebox{.75}{Q}}
+			child {node (M61) [miner] {\scalebox{.75}{K-}}
+				child {node (M611) [terminal] {-1}}
+				child {node (M612) [maxer] {\scalebox{.75}{Q+}}
+					child {node (M6121) [terminal] {-1}}
+					child {node (M6122) [terminal] {-2}}}
+			}
+			child {node (M62) [miner] {\scalebox{.75}{K+}}
+				child {node (M621) [terminal] {1}}
+				child {node (M622) [terminal] {-2}}
+			}
+		    };
+		\draw  (M1.north) to [dashed,black,out=15,in=165] (M3.north);
+		\draw  (M2.north) to [dashed,black,out=15,in=165] (M5.north);
+		\draw  (M4.north) to [dashed,black,out=15,in=165] (M6.north);
+		\draw  (M112.north) to [dashed,black,out=15,in=165] (M312.north);
+		\draw  (M212.north) to [dashed,black,out=15,in=165] (M512.north);
+		\draw  (M412.north) to [dashed,black,out=15,in=165] (M612.north);
+		\draw  (M11.north) to [dashed,black,out=15,in=165] (M21.north);
+		\draw  (M12.north) to [dashed,black,out=15,in=165] (M22.north);
+		\draw  (M31.north) to [dashed,black,out=15,in=165] (M41.north);
+		\draw  (M32.north) to [dashed,black,out=15,in=165] (M42.north);
+		\draw  (M51.north) to [dashed,black,out=15,in=165] (M61.north);
+		\draw  (M52.north) to [dashed,black,out=15,in=165] (M62.north);
+\end{tikzpicture}
+} %scalebox
+\caption{}
+\label{fulltree}
+\end{figure}
+
+%\nocite{*}
+%\nocite{richards09information}
+%\nocite{russell05efficient}
+%\nocite{kuhn03lectures}
+%\nocite{kuhn97classics}
+%\nocite{parker05sampling}
+%\nocite{nance06reasoning}
+%\nocite{li94chess}
+
+%\begin{thebibliography}{99}
+%\end{thebibliography}
+
+%\begin{figure}
+
+\section{Appendix B}
+\begin{verbatim}
+1. Pa2:a4 Pa7:a5 2. Pd2:d4 Pb7:b6 3. Bc1:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+p - - - - - - - 
+- - - - - - - - 
+- p p - p p p p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d3 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pd3:d4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - - - - - - - 
+p p p - p p p p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Pa2:a4 Pb7:b6 3. Bc1:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+p - - - - - - - 
+- - - - - - - - 
+- p p - p p p p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:f4 Pb7:b6 3. Bf4:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - - - - - - - 
+p p p - p p p p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+p - - - - - - - 
+- p p - p p p p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+p - - - - - - - 
+- - - - - - - - 
+- p p - p p p p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pb2:b3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- p - - - - - - 
+p - p - p p p p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pc2:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - p - - - - - 
+p p - - p p p p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - - - p - - - 
+p p p - - p p p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - p - - - 
+- - - - - - - - 
+p p p - - p p p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - - - - p - - 
+p p p - p - p p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - p - - 
+- - - - - - - - 
+p p p - p - p p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - - - - - p - 
+p p p - p p - p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - p - 
+- - - - - - - - 
+p p p - p p - p 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - - - - - - p 
+p p p - p p p - 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - p 
+- - - - - - - - 
+p p p - p p p - 
+r n - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+n - - - - - - - 
+p p p - p p p p 
+r - - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - n - - - - - 
+p p p - p p p p 
+r - - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - - - - - - - 
+p p p n p p p p 
+r - - q k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - - - - - - - 
+p p p q p p p p 
+r n - - k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - - q - - - - 
+p p p - p p p p 
+r n - - k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:c1 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - - - - - - - 
+p p p - p p p p 
+r n q - k b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ke1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - - - - - - - 
+p p p k p p p p 
+r n - q - b n r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - - - - n - - 
+p p p - p p p p 
+r n - q k b - r 
+**********************************************
+1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
+
+**********************************************
+R - B Q K B N R 
+- - - P P - P P 
+N P - p - P - - 
+P - P - - - b - 
+- - - - - - - - 
+- - - - - - - n 
+p p p - p p p p 
+r n - q k b - r 
+**********************************************
+Solutions founds: 25
+\end{verbatim}
+%\caption{Verbose output from example game.  Generating information sets for black after five turns by each player.
+%From this, black can infer that there is definitely a white bishop at g5 and a white pawn at d6.}
+%\label{verboseoutput}
+%\end{figure}
+
+\bibliographystyle{acm}
+\bibliography{paper}
+
 \end{document}

### 2010-12-14 paper/paper.tex
One hour to go--shifting to some formatting issues and rewrites.

@@
 \section{The Information Set Generation Algorithm}
+We have formulated information set generation as a combinatorial search problem.  The input to the algorithm is a list
+of observations for one player in a kriegspiel game.  We assume that the agent has perfect recall (i.e., remembers all
+of its own decisions), which is a reasonable assumption for kriegspiel.  (See~\cite{gilpin07algorithms} for a further
+discussion of perfect recall.)
+
+The nodes in the search tree at depth $n$ correspond possible to possible sequences of $n$ plies from the start state.
+There is a search space variable for each ply in the game, and the values that a variable can take on are all possible
+legal moves.  A kriegspiel game may last for on the order of tens of plies.  Moves are parameterized by the piece that
+is relocating (both black and white have six different types of pieces: king, queen, rook, bishop, knight, pawn), the
+source square, and the destination square.  Only a few hundred of the $64 \times 64 \times 6$ combinations represent
+moves that are legitimate in at least some scenarios.  For example, no piece could ever move from a3 to h2\footnote{We
+observe the convention that on an $8 \times 8$ chessboard the rows are numbered 1 to 8 and the columns are labeled a to
+h.} in one move.
+
+Information set generation, then, is an all-solutions search through the space.  Whenever a node in the search tree is
+found to be inconsistent with the actual observation given in the input, the subtree rooted at that node is pruned.
+There are types of observations that must match in order for the state to be found: check status, pawn tries, and 
+illegal moves available/attempted.
+
+Figure~\ref{codelisting} shows the function at the heart of our information set generation algorithm.  The actual
+observations can be viewed as global, read-only variables.  All solutions (i.e., all nodes in the information set) will
+be located at the same depth in the tree: the depth that is equal to the number of plies in the observation list.
+Referee announcements regarding pawn tries or a player being in check are made prior to a move.  The algorithm checks
+for the consistency of these observations in lines $5--6$.  If the search reaches the appropriate depth and passes these
+tests, then a solution has been found and is reported (line $7$).
+
+Otherwise, we consider the possible moves that can be made from that state.  If the player whose turn it is to move at
+the current level of the search tree is the same player from whose perspective we have received observations, then the
+next step is to make sure that every legal move that was attempted at the player at that ply is also at least
+attemptable at the game position that corresponds to the game tree node (lines $15--22$).  We know exactly the legal
+move that was ultimately made at this depth (because we are assuming perfect recall), and we check to make sure that
+this move is indeed legal (pruning otherwise).  If so, we apply that move and recurse (lines $24--32$).  This
+corresponds to a ``forced move'' in the general search paradigm.
+
+On the other hand, if the player whose turn it is to move at the current depth of the search tree is the opponent, then
+must consider all possible legal moves to branch on.  We first check to make sure that the number of legal moves from
+this position is at least as large as the number of illegal moves that the opponent attempted (i.e., the number of times
+the referee said ``No'' before finally accepting the opponent's move).  This is shown in lines $39--42$.  At this point
+we generate a child for each move that would be accepted by the referee and recurse (lines $44--52$).
+
+One of our goals was to assess the varying utilities of the different pruning criteria.  We tried some different
+orderings of the tests and were somewhat surprised (and disappointed) to see that there was actually a fairly minimal
+impact on performance.  This could be due to the fact that because of the tedious nature of encoding problem instances
+from actual games, we were not able to run experiments of a wide variety of types of positions (i.e., different styles
+of play).
+
 \begin{figure}
@@
 
-Figure~\ref{codelisting}
-
 \section{Performance Results}
+We built our search on top of the same general search engine that we used in class.  The performance results reported in
+this section were obtained by running our program on Bleuprint with up to 1024 processors.
 
@@
 the size of the information set was about 4.5 million.  (This instance is not necessarily representative of all games or
-states of the game, but it was not cherry-picked in any way.)
+states of the game, but it was not cherry-picked in any way.)  The table shows that the problem is indeed highly
+parallizable.  The speedups are nearly linear for up  to 32 processors, and even on 1024 processors, we get a good
+speedup.  In this particular instance, the sequential algorithm takes almost 25 minutes to run.  But when parallelized
+on 1024 processors---perfectly reasonable resource usage on modern hardware---the search takes less than three seconds.
+This is a speedup of 576 times!  It might not be reasonable to use the algorithm at all if it takes 25 or more to run
+for some turns.  But if the cost is only a few seconds, it becomes entirely feasible.
 
@@
 \end{tabular}
-\caption{Speedups for a randomly selected problem instance on 1--1024 processors.  The size of the information set in
-this instance is 4,495,121.}
+\caption{Speedups for a randomly selected problem instance on 1--1024 processors.  The input observations were from the
+10th ply of the game tree and the size of the information set in this instance is 4,495,121.}
 \label{speedups}
@@
 \includegraphics[width=0.7\textwidth]{plots/3schemes.pdf}
-\caption{3schemes}
+\caption{Experiments using different load balancing strategies.}
 \label{3schemes}
 \end{figure}
+
+In our next round of experiments, we set out to compare a few different load-balancing strategies for the problem.  We
+used the random, work stealing, and neighborhood balancing strategies that are available in the general purpose search
+engine.  These results are shown in Figure~\ref{3schemes}.  Our results show that there was actually very little
+variation in performance across the different strategies.  We did notice that across all of our experiments, the work
+stealing approach appeared to be consistently---if only marginally---better than the others.  Our hypothesis here is
+that perhaps the nodes are rarely starved for work.  In other words, perhaps the very nature of the problem means that
+the load is naturally well-balanced and that the load-balancer rarely needs to do a significant amount of
+redistribution.  This is probably also related to the fact that we are doing an all-solutions search.
 
@@
 
+We next attempted to measure the isoefficiency of the problem---how is efficiency of the parallelization of the
+algorithm affected by {\em size} of the problem?  To test this we varied the number of plies in the test instance from
+8 to 12. We used the work stealing load balancing strategy.  The results are shown in Table~\ref{3depths}.  The data are
+the same in both tables, but the one on the left uses a log-log scale.  In the easiest case (depth 8), there isn't
+enough additional parallel work to justify the jump from 256 to 1024 processors, and performance actually degrades.
+Other than that, however, we see consistent speedups as we increase the number of processors.  The plot on the right helps to
+illustrate the good isoefficiency of the problem.  As the size of the problem increaes, the amount of parallelizable
+work available also increases, and therefore the steepness of the speed curve is higher for larger problem sizes, as the
+number of processors increases.
+
 Figure~\ref{dep8}
@@
 \centering
-\includegraphics[width=0.7\textwidth]{plots/dep8.pdf}
+\includegraphics[width=0.6\textwidth]{plots/dep8.pdf}
 \caption{dep8}
-\label{dep8}
+\label{Grainsize control data for depth 8 problems.}
 \end{figure}
@@
 \centering
-\includegraphics[width=0.7\textwidth]{plots/dep10.pdf}
+\includegraphics[width=0.6\textwidth]{plots/dep10.pdf}
 \caption{dep10}
-\label{dep10}
+\label{Grainsize control data for depth 10 problems.}
 \end{figure}
@@
 \centering
-\includegraphics[width=0.7\textwidth]{plots/dep12.pdf}
+\includegraphics[width=0.6\textwidth]{plots/dep12.pdf}
 \caption{dep12}
-\label{dep12}
+\label{Grainsize control data for depth 12 problems.}
 \end{figure}
+
+Finally, we ran some experiments to measure the effect of grainsize on the problem.  The results are shown in
+Tables~\ref{dep8}--~\ref{dep12}.  It is clear that there is a ``sweet spot'' for each problem size.  (At least it is
+clear for this particular instance.  It is possible that there would be different behavior on different problem
+instances.)  When the grainsize is too small, the overall computation is hampered by excessive parallel overhead
+(communication, queueing of nodes to expand, load balancing operations).  On the other hand, when the grain size is too
+large, it becomes too difficult to ensure that all processors are busy and processors are not idling with no work to do
+while a small number of processors are busy doing large-sized chunks of work.
+
+In our experiments, the optimal grainsize is clearly related to the size of the problem, the depth of the search tree.
+The optimal sequential threshold seems to be $2--4$ less than the problem depth.  Since the depth of all solutions is
+the same for each problem instance, this means that the number of levels that should be searched sequentially is fairly
+consistent across problem sizes.  We noted that the number of solutions found---in other words, the size of the
+information sets---is typically fairly large with respect to the total number of nodes expanded.  (We don't have hard
+figures, but we estimate that it was often $30-40\%$).  This high density of solutions limits the variability that can
+be present in the amount of work to be done at the nodes representing the last few levels of the tree.  This means that
+the best grainsize is also likely to be fairly predictable.
 
@@
 ~\cite{richards09information} and could potentially improve the search efficiency greatly by exploiting variable
-ordering.  This would be analgous to the disconnected edge-pairs\footnote{what we referred to as ``algorithm 2'' in
-class} heuristic for the Hamiltonian Circuit problem.  Rather than search from the root node going forward, this
+ordering.  This would be analgous to the disconnected edge-pairs\footnote{This is what we referred to as ``algorithm 2'' in
+class.} heuristic for the Hamiltonian Circuit problem.  Rather than search from the root node going forward, this
 algorithm would seek opportunities for pruning based on the propagation of logical consequences from each move both
@@
 opponent's piece must have been between the previous (successful) bishop move and the current (unsuccessful) bishop
-move.  To implement this algorithm, we envision the use of planning graph type data structures ~\cite{GRAPHPLAN}.
+move.  To implement this algorithm, we envision the use of planning graph type data structures ~\cite{blum97fast}.
 Planning graphs have a sequence of alternating levels of state constraints and action constraints.  State constraints
@@
 possible action/state.  Propagating the constraints would require much of the same machinery used in STRIPS-like
-planning algorithms~\cite{STRIPS}, and one of the many knobs that would need to be tuned would be the depth of the
+planning algorithms~\cite{chen05solving}, and one of the many knobs that would need to be tuned would be the depth of the
 propagation for each constraint.  (Should the consequences of a known action or non-action be propagated one move into
@@
 \section{Appendix A}
-Figure~\ref{fulltree} shows the full game tree for a simple variant of poker known as Kuhn poker.  There are three
-cards: a king, a queen, and a jack.  The dealer deals one card to each of two players.  Each player antes one unit, and
-there is one simple round of betting, in which the size of the pot may be doubled.  Left branches denote check/fold;
-right branches denote bet/call.  For this zero-sum game, terminal nodes are labeled with payoffs to player 1, whose
-decision points are shown in triangles.  Connections between nodes in the same information set are shown with dotted
-lines.
+The purpose of this figure is simply to help visualize the concept of information sets in game trees.  The tree for
+kriegspiel is very difficult to visualize/depict (even an interesting part of the tree).  Figure~\ref{fulltree} shows
+the full game tree for a simple variant of poker known as Kuhn poker, which we hope will serve to at least make the
+concept clear, even if the poker game itself is not immediately relevant.  There are three cards: a king, a queen, and a
+jack.  The dealer deals one card to each of two players.  Each player antes one unit, and there is one simple round of
+betting, in which the size of the pot may be doubled.  Left branches denote check/fold; right branches denote bet/call.
+For this zero-sum game, terminal nodes are labeled with payoffs to player 1, whose decision points are shown in
+triangles.  Connections between nodes in the same information set are shown with dotted lines.  The top level shows
+three different information sets for the first player, one for each possible card that could be dealt to him.  There are
+two nodes in each set, because no matter what the deal is, there are two possibilities for the card that the other
+player holds.  The leftmost information set on the bottom level corresponds to player 1's observation that he was dealt
+a king and checked, and then the opponent wagered.  The second player has six information sets as well, but they are all
+at the same level of the tree.  Like the first player, the second player has a separate information set for each
+possible card that he could be dealt.  And the two nodes in each set are there because there are two possibilities for
+the opponent card.  The second player also gets to know whether the first player checked of bet, and that is why there
+are twice as many information sets.
 

### 2012-06-08 sbac/paper.tex

--- 2010-12-14 paper/paper.tex
@@
-\documentclass[11pt]{article}
-\usepackage{amsmath,amsfonts,latexsym,graphicx}
-\usepackage{fullpage}
+\documentclass[times, 10pt,twocolumn]{article} 
+\usepackage{latex8}
+\usepackage{times}
 \usepackage{url,hyperref}
@@
 \usepackage{moreverb}
+\usepackage{algorithm}
+\usepackage{boxedminipage}
+\usepackage[noend]{algpseudocode}
 \usepackage{listings}
+\usepackage{amsmath,amsfonts,latexsym,graphicx}
 \lstset{language=C,numbers=left,basicstyle=\footnotesize,numberstyle=\footnotesize}
@@
 \usepgflibrary{shapes.geometric}
+
+%\documentclass[11pt]{article}
+%\usepackage{fullpage}
 %\usepackage{ijcai09}
 
-\setlength{\oddsidemargin}{0in}
-\setlength{\topmargin}{0in}
-\setlength{\textwidth}{6.5in}
-\setlength{\textheight}{8.5in}
+%\setlength{\oddsidemargin}{0in}
+%\setlength{\topmargin}{0in}
+%\setlength{\textwidth}{6.5in}
+%\setlength{\textheight}{8.5in}
 
@@
 \newcommand{\eps}{\epsilon}
+%\documentstyle[times,art10,twocolumn,latex8]{article}
+
+%------------------------------------------------------------------------- 
+% take the % away on next line to produce the final camera-ready version 
+\pagestyle{empty}
+
+%------------------------------------------------------------------------- 
+\begin{document}
+
+\title{Parallel Information Set Generation for Kriegspiel 
+       }
+
+\author{Mark Richards \and Abhishek Gupta \and Osman Sarood \\
+%Department of Computer Science\\ 
+%University of Illinois\\ 
+%Urbana, IL\\ mdrichar@illinois.edu\\
+% For a paper whose authors are all at the same institution, 
+% omit the following lines up until the closing ``}''.
+% Additional authors and addresses can be added with ``\and'', 
+% just like the second author.
+%\and
+%Second Author\\
+%Institution2\\
+%First line of institution2 address\\ Second line of institution2 address\\ 
+%SecondAuthor@institution2.com\\
+}
+
+\maketitle
+\thispagestyle{empty}
 
@@
 %\newtheorem{theorem}[lemma]{Theorem}
-\begin{document}
-
-\title{Information Set Generation in Kriegspiel}
-\author{Abhishek Gupta\qquad Mark Richards\qquad Osman Sarood\\University of Illinois at Urbana-Champaign\\CS598LVK:
-Parallel Combinatorial Search} \maketitle
+%\begin{document}
+
+%\title{Information Set Generation in Kriegspiel}
+%\author{Abhishek Gupta\qquad Mark Richards\qquad Osman Sarood\\University of Illinois at Urbana-Champaign\\CS598LVK:
+%Parallel Combinatorial Search} \maketitle
 
@@
 Information Set Generation is the identification of the set of paths in an imperfect information game tree that are
-consistent with a player's observations.  The ability to reason about the possible state of the world is crucial to the
+consistent with a player's observations.  The ability to reason about the possible game history is crucial to the
 performance of game playing agents.  In this work, we discuss the problem of information set generation in the context
@@
 engine and discuss its performance under different load balancing strategies and with different grainsize parameters.
-We discuss some variations in the search strategy and their effect on performance.  \end{abstract}
-
-\section{Introduction}
+Our results suggest that our algorithm can be efficiently parallelized.  
+\end{abstract}
+
+\Section{Introduction}
 In imperfect information games, players do not have access to full knowledge of the world. Examples of imperfect
@@
 in a game tree denotes not only the current state of the game, but also uniquely defines a path from the initial state
-or root node.  Thus, a game tree node implicitly encodes not only the current state of the game, but also all the
-history of all decisions made by all players up to that point in the game.  Knowing one's information set means knowing
-all possible game histories.
-
-For many games, solving the information set generation problem is trivial.  For example, in a poker game, it is easy to
+or root node.  Thus, a game tree node implicitly encodes not only the current state of the game but also the complete
+history of all decisions made by all players up to that point in the game, including the outcome of any chance moves
+such as dice rolls or card shuffling.  Knowing one's information set means knowing all possible game histories.
+
+For many specific games, solving the information set generation problem is trivial.  For example, in a poker game, it is easy to
 see that the unseen cards held by a player's opponents may be any permutation of the cards not seen by that player
@@
 
-In the game of kriegspiel, information set generation is non-trivial.  The player knows the opponent's position with
+In the game of kriegspiel, information set generation is much more complicated.  The player knows the opponent's position with
 certainty when the game begins, but after the initial move there may be varying levels of uncertainty about the location
@@
 
-Information set generation is not the same problem as solving a game
-tree, although it is arguably a necessary subroutine for quality game play in imperfect information games. The standard
-game tree problem is to find an equilibrium (i.e., minimax) strategy for the game. This can be done using alpha-beta
-pruning for perfect information games or linear programming for imperfect information games. Information set generation,
-by contrast, is the problem of finding sequences of moves in the game tree that are consistent with a player's
-observations.
-
-\section{Background}
+Information set generation is not the same problem as solving a game tree, although it is arguably a necessary
+subroutine for quality game play in imperfect information games.  In this work, we describe an approach to generating
+information sets for the game of kriegspiel.
+
+\Section{Background}
 In this section, we describe the game of kriegspiel, and the particular variation of it that we have implemented.
@@
 competing players and one referee.  Each player sits with a chessboard that is partitioned off so as not be seen by the
-others.  One player controls the white pieces; the other controls the black pieces.  The referee keeps track of both
-sets of pieces.  Each player has a full set of chess pieces.  He may position the opponent's pieces however he wishes on
-his own board (if at all).  The players alternate moves as in a regular chess game, but instead of announcing their
-moves out loud, they write the source and destination square of their desired play son a piece of paper and pass them to
-the referee, so that the move is not heard by the other player. The referee, who knows the location of all pieces,
-checks the legality of the move.  If the move is legal, the referee makes the move on his own board and announces that a
-legal move has been made and that it is the other player's turn to move.  If the move is illegal because it is blocked
-by an opponent's piece or would place or leave the active player in check, the referee announces, "No."  There is no
-penalty for attempting an illegal move.  The player continues to attempt moves until one finally succeeds.  If a player
-has no legal moves (because of stalemate or checkmate), the referee announces the game's result.  With the advent of
-computers, it is convenient to play the game over a network, with a computer acting as the referee.
+other.  One player controls the white pieces; the other controls the black pieces.  The referee has a board on which she
+keeps track of both sets of pieces.  Each player has a full set of chess pieces.  He may position the opponent's pieces
+however he wishes on his own board (if at all).  The players alternate moves as in a regular chess game, but instead of
+announcing their moves out loud, they write the source and destination square of their desired plays on a piece of paper
+and pass them to the referee, so that the requested move is not known to the other player. The referee, who knows the location of
+all pieces, checks the legality of the move.  If the move is legal, the referee makes the move on his own board and
+announces that a legal move has been made and that it is the other player's turn to move.  If the move is illegal
+because it is blocked by an opponent's piece or would place or leave the active player in check, the referee announces,
+``No.''  There is no penalty for attempting an illegal move.  The player continues to attempt moves until one finally
+succeeds.  If a player has no legal moves (because of stalemate or checkmate), the referee announces the game's result.
+With the advent of computers, it is convenient to play the game over a network, with a computer acting as the referee.
 
@@
 some variation in what additional information is supplied in this case.  In one popular variant, the referee declares
-that the player is in check "by rank," "by file," "by diagonal," or "by knight."  In the case of a diagonal attack, the
+that the player is in check ``by rank,'' ``by file,'' ``by diagonal,'' or ``by knight.''  In the case of a diagonal attack, the
 referee may announce whether the checking piece is along the long or short diagonal (from the king's perspective).  It
@@
 kind of declaration at the beginning of a player's turn with respect to the pawn captures that are available.  This may
-be as simple as declaring "Try" to indicate that at least one pawn capture is possible.  In our implementation, the
-referee declares all possible pawn captures.
+be as simple as declaring ``Try'' to indicate that at least one pawn capture is possible.  In our implementation, the
+referee declares all potential pawn captures.
 
@@
 piece (queen, rook, bishop, or knight) he would like.  To reduce the number of bits needed to specify a move, our
-implementation assumes and allows only promotion to queen.
+implementation allows only promotion to queen.
 
@@
 
-\subsubsection{Special Moves}
+\SubSection{Special Moves}
 Our implementation supports the capability for pawns to move two spaces on their first turn.  However, we did not
@@
 
-\subsection{Example}
-\begin{figure}
+\SubSection{Example}
 %\includegraphics[width=0.2\textwidth]{images/1W.png}
-\begin{minipage}{\textwidth}
+%\begin{minipage}{\textwidth}
+%\centering
+%\subfloat[1. d4]{\label{fig:1W}\includegraphics[width=0.2\textwidth]{images/1W.png}}
+%\qquad
+%\subfloat[1. ... a5]{\label{fig:1B}\includegraphics[width=0.2\textwidth]{images/1B.png}}
+%\qquad
+%\subfloat[2. Bg5]{\label{fig:2W}\includegraphics[width=0.2\textwidth]{images/2W.png}}
+%\qquad
+%\subfloat[2. ... b6]{\label{fig:2B}\includegraphics[width=0.2\textwidth]{images/2B.png}}
+%\end{minipage}
+%\begin{minipage}{\textwidth}
+%\centering
+%\subfloat[3. Nc3]{\label{fig:3W}\includegraphics[width=0.2\textwidth]{images/3W.png}}
+%\qquad
+%\subfloat[3. ... c5]{\label{fig:3B}\includegraphics[width=0.2\textwidth]{images/3B.png}}
+%\qquad
+%\subfloat[4. d5]{\label{fig:4W}\includegraphics[width=0.2\textwidth]{images/4W.png}}
+%\qquad
+%\subfloat[4. ... Na6]{\label{fig:4B}\includegraphics[width=0.2\textwidth]{images/4B.png}}
+%\end{minipage}
+%\begin{minipage}{\textwidth}
+%\centering
+%\subfloat[5. d6!]{\label{fig:5W}\includegraphics[width=0.2\textwidth]{images/5W.png}}
+%\qquad
+%\subfloat[5. ... f6!]{\label{fig:5B}\includegraphics[width=0.2\textwidth]{images/5B.png}}
+%\qquad
+%\subfloat[6. e4]{\label{fig:6W}\includegraphics[width=0.2\textwidth]{images/6W.png}}
+%\qquad
+%\subfloat[6. ... fg]{\label{fig:6B}\includegraphics[width=0.2\textwidth]{images/6B.png}}
+%\end{minipage}
+%\begin{minipage}{\textwidth}
+%\centering
+%\subfloat[7. Bc4]{\label{fig:7W}\includegraphics[width=0.2\textwidth]{images/7W.png}}
+%\qquad
+%\subfloat[7. ... ed]{\label{fig:7B}\includegraphics[width=0.2\textwidth]{images/7B.png}}
+%\qquad
+%\subfloat[8 Qd5!]{\label{fig:8W}\includegraphics[width=0.2\textwidth]{images/8W.png}}
+%\qquad
+%\subfloat[8 ... Qc7?]{\label{fig:8B}\includegraphics[width=0.2\textwidth]{images/8B.png}}
+%\end{minipage}
+%\begin{minipage}{\textwidth}
+%\centering
+%\subfloat[9 Qf7+]{\label{fig:9W}\includegraphics[width=0.2\textwidth]{images/9W.png}}
+%\qquad
+%\subfloat[9 ... d8]{\label{fig:9B}\includegraphics[width=0.2\textwidth]{images/9B.png}}
+%\qquad
+%\subfloat[10 Qxf8++]{\label{fig:10W}\includegraphics[width=0.2\textwidth]{images/8W.png}}
+%\end{minipage}
+\begin{figure}[ht]
 \centering
-\subfloat[1. d4]{\label{fig:1W}\includegraphics[width=0.2\textwidth]{images/1W.png}}
-\qquad
-\subfloat[1. ... a5]{\label{fig:1B}\includegraphics[width=0.2\textwidth]{images/1B.png}}
-\qquad
-\subfloat[2. Bg5]{\label{fig:2W}\includegraphics[width=0.2\textwidth]{images/2W.png}}
-\qquad
-\subfloat[2. ... b6]{\label{fig:2B}\includegraphics[width=0.2\textwidth]{images/2B.png}}
+\begin{minipage}[b]{0.5\linewidth}
+\includegraphics[width=0.7\textwidth]{images/9B.png}
+\caption{Actual Position}
+\label{fig:figure1}
 \end{minipage}
-\begin{minipage}{\textwidth}
+\hspace{0.5cm}
 \centering
-\subfloat[3. Nc3]{\label{fig:3W}\includegraphics[width=0.2\textwidth]{images/3W.png}}
-\qquad
-\subfloat[3. ... c5]{\label{fig:3B}\includegraphics[width=0.2\textwidth]{images/3B.png}}
-\qquad
-\subfloat[4. d5]{\label{fig:4W}\includegraphics[width=0.2\textwidth]{images/4W.png}}
-\qquad
-\subfloat[4. ... Na6]{\label{fig:4B}\includegraphics[width=0.2\textwidth]{images/4B.png}}
+\begin{minipage}[b]{0.5\linewidth}
+\includegraphics[width=0.7\textwidth]{images/OneView.png}
+\caption{White's view}
+\label{fig:figure2}
 \end{minipage}
-\begin{minipage}{\textwidth}
-\centering
-\subfloat[5. d6!]{\label{fig:5W}\includegraphics[width=0.2\textwidth]{images/5W.png}}
-\qquad
-\subfloat[5. ... f6!]{\label{fig:5B}\includegraphics[width=0.2\textwidth]{images/5B.png}}
-\qquad
-\subfloat[6. e4]{\label{fig:6W}\includegraphics[width=0.2\textwidth]{images/6W.png}}
-\qquad
-\subfloat[6. ... fg]{\label{fig:6B}\includegraphics[width=0.2\textwidth]{images/6B.png}}
-\end{minipage}
-\begin{minipage}{\textwidth}
-\centering
-\subfloat[7. Bc4]{\label{fig:7W}\includegraphics[width=0.2\textwidth]{images/7W.png}}
-\qquad
-\subfloat[7. ... ed]{\label{fig:7B}\includegraphics[width=0.2\textwidth]{images/7B.png}}
-\qquad
-\subfloat[8 Qd5!]{\label{fig:8W}\includegraphics[width=0.2\textwidth]{images/8W.png}}
-\qquad
-\subfloat[8 ... Qc7?]{\label{fig:8B}\includegraphics[width=0.2\textwidth]{images/8B.png}}
-\end{minipage}
-\begin{minipage}{\textwidth}
-\centering
-\subfloat[9 Qf7+]{\label{fig:9W}\includegraphics[width=0.2\textwidth]{images/9W.png}}
-\qquad
-\subfloat[9 ... d8]{\label{fig:9B}\includegraphics[width=0.2\textwidth]{images/9B.png}}
-\qquad
-\subfloat[10 Qxf8++]{\label{fig:10W}\includegraphics[width=0.2\textwidth]{images/8W.png}}
-\end{minipage}
-\caption{Sequence of moves}
-\label{sequence}
 \end{figure}
 
-\begin{figure}
-\begin{verbatim}
-1. d4 {(:)}
-   a5 {(:)}
-2. Bg5 {(:)}
-   b6 {(:)}
-3. Nc3 {(:Bd8)}
-   c5 {(:)}
-4. d5 {(:Bd8;Tc5)}
-   Na6 {(:)}
-5. d6 {(:)}
-   f6 {(:Td6)}
-6. e4 {(:)}
-   fxg5 {(Xg5:Td6;Tg5)}
-7. Bc4 {(:)}
-   exd6 {(Xd6:Td6)}
-8. Qd5 {(:)}
-   Qc7 {(:)}
-9. Qf7+ {(:)} 
-   d8 {:g6,f7,e7}
-10. Qxf8++ {(1-0:)}
-\end{verbatim}
-\caption{Listing of a kriegspiel game resulting in checkmate by white after 10 moves.}
-\label{listing}
-\end{figure}
-
-The concept of information sets is perhaps best explained through an example.  Figures~\ref{listing} and~\ref{sequence}
-show a full listing and the associated pictorial simulation of an actual kriegspiel game from ~\cite{li94chess}.  We
-have adapted the notation used by Wolfe for Berkeley Kriegspiel.  The moves are numbered in pairs: a move by white on
-one line followed by black's move on the next line.  The actual move is shown first followed by a list of referee
-announcements.  Captures announcements are prefixed with an X and give the location of the captured piece.  Similarly,
-pawn tries announcements are prefixed with T.  Attempted moves that were ultimately declared illegal are shown in a list
-following the ':'.  For example at move 9, (Figure~\ref{fig:9W}--\ref{fig:9B}), black attempts to block a potential
-threat along the diagonal by moving a pawn to f6, then attempts to escape or capture a threat at f7.  When both of these
-attempts fail, black can infer that his king is being threatened by a protected bishop or queen at f7.  The attempt to
-move to e7 is black's way of finding out whether the threatening piece is a queen or bishop.  When this move fails,
-black knows it is white's queen and ultimately retreats to d8.  Note that the failed moves by white on turns 3 and 4 are
-a result of being blocked by the pawn at e2.
-
-Note that the listing in Figure~\ref{listing} corresponds to the referee's view of the game and would be useful for
+%\begin{figure}
+%\begin{verbatim}
+%1. d4 {(:)}
+%   a5 {(:)}
+%2. Bg5 {(:)}
+%   b6 {(:)}
+%3. Nc3 {(:Bd8)}
+%   c5 {(:)}
+%4. d5 {(:Bd8;Tc5)}
+%   Na6 {(:)}
+%5. d6 {(:)}
+%   f6 {(:Td6)}
+%6. e4 {(:)}
+%   fxg5 {(Xg5:Td6;Tg5)}
+%7. Bc4 {(:)}
+%   exd6 {(Xd6:Td6)}
+%8. Qd5 {(:)}
+%   Qc7 {(:)}
+%9. Qf7+ {(:)} 
+%   d8 {:g6,f7,e7}
+%10. Qxf8++ {(1-0:)}
+%\end{verbatim}
+%\label{listing}
+%\caption{Listing of a kriegspiel game resulting in checkmate by white after 10 moves.}
+%\end{figure}
+
+The concept of information sets is perhaps best explained through an example.  (See also Appendix A.)
+Figures~\ref{listing} and~\ref{sequence} show a full listing and the associated pictorial progression of an actual
+kriegspiel game from ~\cite{li94chess}.  We have adapted the notation used by Wolfe for Berkeley Kriegspiel.  The moves
+are numbered in pairs: a move by white on one line followed by black's move on the next line.  The actual move is shown
+first followed by a list of referee announcements.  Captures announcements are prefixed with an X and give the location
+of the captured piece.  Similarly, pawn tries announcements are prefixed with T.  Attempted moves that were ultimately
+declared illegal are shown in a list following the `:'.  For example at move 9, (Figure~\ref{fig:9W}--\ref{fig:9B}),
+black attempts to block a potential threat along the diagonal by moving a pawn to f6, then attempts to escape or capture
+a threat at f7.  When both of these attempts fail, black can infer that his king is being threatened by a protected
+bishop or queen at f7.  The attempt to move to e7 is black's way of finding out whether the threatening piece is a queen
+or bishop.  When this move fails, black knows it is white's queen and ultimately retreats to d8.  Note that the failed
+moves by white on turns 3 and 4 are a result of being blocked by the pawn at e2.
+
+The listing in Figure~\ref{listing} corresponds to the referee's view of the game and would be useful for
 post-game analysis or commentary.  During the game, players would not have access to the full information.  The
@@
 observations for a sequence of moves, an information set is the set of all possible {\em sequences of moves for both
-players}, that are consistent with those observations.  Figure~\ref{abbrevoutput} shows the information set for black
-after both player's have made five moves.  From this list of possibilities, black can infer that white must have a
-bishop at g5 and a pawn at d6.  There are 25 possible sequences of moves that are consistent with black's observations
-up to that point. Table~\ref{bothtimes} shows the size of the information sets for each player for each move in the game
-and the amount of time to find the full set on a single processor.  Appendix B shows the full output of the program for
-the same point in the game (five moves for each player) and illustrates all of the possible positions in the belief
-state.
+players}, that are consistent with those observations.  In other words, any sequence of moves that would have generated
+the same sequence of observations for black is in black's information set.  Figure~\ref{abbrevoutput} shows the
+information set for black after both player's have made five moves.  From this list of possibilities, black can infer
+that white must have a bishop at g5 and a pawn at d6.  There are 25 possible sequences of moves that are consistent with
+black's observations up to that point. Table~\ref{bothtimes} shows the size of the information sets for each player for
+each move in the game and the amount of time to find the full set on a single processor.  Appendix B shows the full
+output of the program for the same point in the game (five moves for each player) and illustrates all of the possible
+positions in the belief state.
 
@@
 \begin{mytinylisting}
-1. Pa2:a4 Pa7:a5 2. Pd2:d4 Pb7:b6 3. Bc1:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
-1. Pd2:d3 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pd3:d4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
-1. Pd2:d4 Pa7:a5 2. Pa2:a4 Pb7:b6 3. Bc1:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
-1. Pd2:d4 Pa7:a5 2. Bc1:f4 Pb7:b6 3. Bf4:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pb2:b3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pc2:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:c1 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ke1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+1. Pa2:a4 Pa7:a5 2. Pd2:d4 Pb7:b6 3. Bc1:g5 Pc7:c5 \\
+4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+\\
+1. Pd2:d3 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pd3:d4 Pc7:c5 \\
+4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+\\
+1. Pd2:d4 Pa7:a5 2. Pa2:a4 Pb7:b6 3. Bc1:g5 Pc7:c5 \\
+4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+
+%1. Pd2:d4 Pa7:a5 2. Bc1:f4 Pb7:b6 3. Bf4:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pb2:b3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pc2:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:c1 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ke1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
 \end{mytinylisting}
@@
 
-\section{The Information Set Generation Algorithm}
+\Section{The Information Set Generation Algorithm}
 We have formulated information set generation as a combinatorial search problem.  The input to the algorithm is a list
@@
 
-\begin{figure}
-%\begin{listing}
-\begin{lstlisting}
-void generateInformationSet(uint16_t* trueState, uint16_t* possState, bool whiteMove, 
-	uint16_t* possHistory, uint16_t** levels, int depth)
-{
-  if (!samePawnTries(trueState, possState, whiteMove)) return;  
-  if (!sameCheckStatus(trueState, possState, whiteMove)) return; 
-  if (depth == maxdepth) { // Then we have found a solution
-	nSolutions++; return; // Report solution
-  }
-  uint16_t newPossState[16]; 
-  int nMoves = 0;
-  generateAttemptableMoves(possState, whiteMove, levels[depth], nMoves,false);
-  checkForCheck(possState, whiteMove, levels[depth], nMoves); // prunes if not equal
-  assert (nMoves < NMOVES); //
-
-  if (whitePerspective == whiteMove) { // move from perspective plyaer 
-    SetMove& failures = failedMoves[depth];
-    for (SetMove::const_iterator itr = failures.begin(); itr != failures.end(); ++itr) {
-      uint16_t move = *itr ;
-      if (!foundMatchingMove(move,levels[depth],nMoves)) {
-        return; // Prune: one of my own failed moves is not attemptable in this state
-      }
-    }
-    // We know exactly what the actual move was; make it
-    uint16_t& actualMove = moveHistory[depth];
-    if (!foundMatchingMove(actualMove,levels[depth],nMoves)) {
-      return; // Prune (because the move that we know we made at this depth is not legal)
-    }
-    // Otherwise, recurse 
-    applyMove(possState,newPossState,actualMove);
-    possHistory[depth] = actualMove;
-    generateInformationSet(newTrueState, newPossState, !whiteMove, 
-      possHistory, levels, depth+1);
-  } else { 
-    // We are at a level in the search tree where 
-    // we are considering the possible moves for the opponent.
-    // We know the number of attempted illegal moves, but not which ones
-
-    // Must be enough attemptable moves in the state to match the number of failed attempts
-    unsigned nIllegalMoves = countIllegalMoves(levels[depth], nMoves);
-    if (nIllegalMoves < failedMoves[depth].size()) {
-	return; // Prune if not enough attemptable moves
-    }
-    // Now we want to try each possible move that is legally executable (not just attemptable) 
-    for (int i = 0; i < nMoves; i++) {
-      uint16_t& move = levels[depth][i];
-      if (isLegal(move)) { // Obviously, we can only execute the moves that are actually legal from this state
-        applyMove(possState,newPossState,move);
-        possHistory[depth] = move;
-        generateInformationSet(newTrueState, newPossState, !whiteMove, 
-          possHistory, levels, depth+1);
-      }
-    }
-  }
-}
-\end{lstlisting}
-\caption{The heart of the information set generation algorithm.}
+%\begin{figure}
+%%\begin{listing}
+%\begin{lstlisting}
+%void generateInformationSet(uint16_t* trueState, uint16_t* possState, bool whiteMove, 
+%	uint16_t* possHistory, uint16_t** levels, int depth)
+%{
+%  if (!samePawnTries(trueState, possState, whiteMove)) return;  
+%  if (!sameCheckStatus(trueState, possState, whiteMove)) return; 
+%  if (depth == maxdepth) { // Then we have found a solution
+%	nSolutions++; return; // Report solution
+%  }
+%  uint16_t newPossState[16]; 
+%  int nMoves = 0;
+%  generateAttemptableMoves(possState, whiteMove, levels[depth], nMoves,false);
+%  checkForCheck(possState, whiteMove, levels[depth], nMoves); // prunes if not equal
+%  assert (nMoves < NMOVES); // 
+%
+%  if (whitePerspective == whiteMove) { // move from perspective plyaer 
+%    SetMove& failures = failedMoves[depth];
+%    for (SetMove::const_iterator itr = failures.begin(); itr != failures.end(); ++itr) {
+%      uint16_t move = *itr ;
+%      if (!foundMatchingMove(move,levels[depth],nMoves)) {
+%        return; // Prune: one of my own failed moves is not attemptable in this state
+%      }
+%    }
+%    // We know exactly what the actual move was; make it
+%    uint16_t& actualMove = moveHistory[depth];
+%    if (!foundMatchingMove(actualMove,levels[depth],nMoves)) {
+%      return; // Prune (because the move that we know we made at this depth is not legal)
+%    }
+%    // Otherwise, recurse 
+%    applyMove(possState,newPossState,actualMove);
+%    possHistory[depth] = actualMove;
+%    generateInformationSet(newTrueState, newPossState, !whiteMove, 
+%      possHistory, levels, depth+1);
+%  } else { 
+%    // We are at a level in the search tree where 
+%    // we are considering the possible moves for the opponent.
+%    // We know the number of attempted illegal moves, but not which ones
+%
+%    // Must be enough attemptable moves in the state to match the number of failed attempts
+%    unsigned nIllegalMoves = countIllegalMoves(levels[depth], nMoves);
+%    if (nIllegalMoves < failedMoves[depth].size()) {
+%	return; // Prune if not enough attemptable moves
+%    }
+%    // Now we want to try each possible move that is legally executable (not just attemptable) 
+%    for (int i = 0; i < nMoves; i++) {
+%      uint16_t& move = levels[depth][i];
+%      if (isLegal(move)) { // Obviously, we can only execute the moves that are actually legal from this state
+%        applyMove(possState,newPossState,move);
+%        possHistory[depth] = move;
+%        generateInformationSet(newTrueState, newPossState, !whiteMove, 
+%          possHistory, levels, depth+1);
+%      }
+%    }
+%  }
+%}
+%\end{lstlisting}
+%\caption{The heart of the information set generation algorithm.}
+%\label{codelisting}
+%\end{figure}
+
+\begin{figure}[htpb]
+\small
+%\begin{algorithm}
+\begin{boxedminipage}{\columnwidth}
+\begin{algorithmic}[1]
+\Function{DFS-Infoset}{$a_{1:t}$, $o_{1:T}^{(k)}$, $I$, $s$, $t$}
+  \If {$t > T$}
+    \State {$I \leftarrow I \cup a_{1:t}$} %\Comment{$a_{1:t}$ leads to $o_{1:T}$ for $P_k$; add it to the information set.}
+  \Else {
+    \For{\textbf{each} legal action $a$ at $s$}
+      \State {$s'$ = {\sc GetNextState}($s$,$a$)}\label{line:update}
+      \If {$s'.o_{t+1}^{(k)}=o_{t+1}^{(k)}$}
+        \State {{\sc DFS-Infoset}($a_{1:t}a$, $o_{1:T}^{(k)}$, $I$, $s'$, $t+1$)} 
+          %\Comment{ Recall that $N(a_{1:t}a) = T(T(a_{1:t}),a)$}
+      \EndIf
+    \EndFor
+  }\EndIf
+\EndFunction
 \label{codelisting}
+\end{algorithmic}
+\end{boxedminipage}
 \end{figure}
 
-\section{Performance Results}
+\Section{Charm++}
+We built our search engine on top of the {\sc Charm++} run-time
+system for parallel processing~\cite{kale93charm,kale09charm}.  {\sc Charm++}
+is an extension to the C++ programming language that provides
+machine-independent infrastructure for parallel computation.  The language and
+its accompanying runtime system have been ported to many shared-memory and
+distributed-memory platforms.
+
+In contrast to other popular parallel programming frameworks such as {\sc MPI}
+and {\sc OpenMP}, the {\sc Charm++} language is based on object-oriented
+programming principles.  The programmer is responsible to decompose its problem
+into a collection of objects that represent elements of work to be done.
+Typically the number of work objects far exceeds the number of processing
+elements (cores) available.  These objects are defined so as to be able to be
+migrated between processors {\em during execution}, in order to improve the
+overall load balancing.
+
+Work objects, known as {\em chares}, communicate with each other via
+asynchronous message-passing.  Objects can invoke {\em entry methods} on other
+objects, which are executed according to a schedule set by the {\sc Charm++}
+runtime system.  The runtime is responsible to interleave computation and
+communication tasks in order to maximize efficiency.  To facilitate this, each
+processor maintains an {\em incoming queue} of messages targeted to chares that
+it currently holds.  A corresponding {\em outgoing queue} collects messages
+generated by objects on the processor that need to be sent to objects assigned
+to other processors.
+
+Some work tasks may inherently require more computation than others, but the
+amount of computation associated with each piece of work is not always known
+before execution.  In tree search, for example, a piece of work may be defined
+as the computation of a minimax value on a node in the tree using alpha-beta
+pruning.  A natural decomposition is to assign different nodes to different
+processors for evaluation.  Because some nodes will be pruned at shallower
+levels of the tree than others, the amount of work performed at each subtree can
+vary considerably.  In a traditional parallel framework, this would result in
+many processors idling while a small number of processors execute the larger
+workloads. By allowing the {\sc Charm++} runtime system to schedule a large
+number of small pieces of work, better adaptive load balancing is achieved.
+
+\Section{The {\sc Charm++} Parallel State Space Search Engine}
+Sun \etal have developed the Parallel State Space Search Engine ({\sc ParSSSE})
+within the {\sc Charm++} framework, which exposes an API that is useful for a
+variety of graph-based and tree-based search problems~\cite{sun11adaptive}. In
+{\sc ParSSE} the fundamental unit of work is the processing of a node.
+Generally, this means identifying whether the node corresponds to a solution
+and/or generating descendant nodes and spawning work tasks related to them. A
+task is {\em terminal} if it spawns no children.
+
+A {\em grainsize control} parameter specifies a bound on the size of a piece of
+work that can be passed along to the runtime system for scheduling.  Once a
+task has been subdivided into an appropriate number of subtasks, those tasks
+run to completion on a single processor without preemption and without generating
+any additional parallel tasks.  If the grainsize is too large, then the number
+of pieces of work to be done may be lower than the number of processors
+available, forcing some processors to idle.  If the grainsize is too small,
+then the runtime system is flooded with a large number of trivial tasks, and
+the system gets bogged down by the overhead of creating and scheduling tasks
+instead of actually doing them.  {\sc ParSSE} supports an adaptive grainsize
+management, so that the threshold for sequential processing can be changed
+during the execution of the program, depending on the properties of the problem
+and/or the realtime performance analysis of the runtime system.
+
+{\sc ParSSE} currently supports two different load balancing strategies.  The
+first is the randomized strategy.  In this approach, each newly created object
+is assigned to a random processor.  Generally, if a large enough number of 
+chares is created, the system will achieve an approximately uniform
+distribution of work across processors.  The approach is also appealing because
+of its simplicity. However, randomization can sometimes lead to inefficiencies
+from parallel overhead.  For a program running on $p$ processors, the
+probability that a piece of work is assigned to a processor other than the
+one that created it is $1-1/p$.  As the number of processors gets large, this
+probability approaches 1, which means that the parallel overhead related to the
+creation and scheduling of a new task is incurred for almost every piece of new
+work.
+
+A second load balancing strategy is known as {\em work stealing}.  In this
+approach, all new tasks are inserted in the message queue of the processor that
+spawns them.  Whenever a processor is idle it becomes a {\em thief}.  A thief
+randomly selects another processor as its {\em victim} and sends a message to
+that processor requesting work.  When the victim receives the message, it
+removes the work with the highest priority from its own queue and sends it to
+the thief.  If it has no work available, it sends a negative acknowledgment
+message to the thief, which causes the thief to look for another victim.
+
+The advantage of the work stealing strategy is that the message-passing
+overhead is only incurred if at least one processor is idling.  Furthermore,
+because victims respond by sending the highest priority work, the most
+important tasks are more likely to get done sooner.
+
+In addition to these strategies, a user may define its own load balancing
+strategies by overriding {\sc Charm++} task dispersal functions.
+
+\Section{Performance Results}
 We built our search on top of the same general search engine that we used in class.  The performance results reported in
@@
 
-Figure~\ref{3schemes}
-\begin{figure}
+\begin{figure}[h]
 \centering
-\includegraphics[width=0.7\textwidth]{plots/3schemes.pdf}
+\includegraphics[width=0.7\columnwidth]{plots/3schemes.pdf}
 \caption{Experiments using different load balancing strategies.}
@@
 
-Figure~\ref{3depths}
-\begin{figure}
+\begin{figure}[h]
+\centering
 \begin{minipage}{0.5\linewidth}
+\includegraphics[width=\columnwidth]{plots/3depths.pdf}
+\end{minipage}
 \centering
-\includegraphics[width=\textwidth]{plots/3depths.pdf}
+\begin{minipage}{0.5\linewidth}
+\includegraphics[width=\columnwidth]{plots/3depthsnolog.pdf}
 \end{minipage}
-\begin{minipage}{0.5\linewidth}
-\centering
-\includegraphics[width=\textwidth]{plots/3depthsnolog.pdf}
-\end{minipage}
-\caption{3depths}
+\caption{(Left) Log-log plot of running time for three different sizes of problem as the number of cores used increases.
+(Right) Same data with a linear y-axis for the running time (to help see the isoefficiency.)}
 \label{3depths}
@@
 Other than that, however, we see consistent speedups as we increase the number of processors.  The plot on the right helps to
-illustrate the good isoefficiency of the problem.  As the size of the problem increaes, the amount of parallelizable
+illustrate the good isoefficiency of the problem.  As the size of the problem increases, the amount of parallelizable
 work available also increases, and therefore the steepness of the speed curve is higher for larger problem sizes, as the
@@
 
-Figure~\ref{dep8}
-\begin{figure}
+\begin{figure}[h]
 \centering
-\includegraphics[width=0.6\textwidth]{plots/dep8.pdf}
-\caption{dep8}
-\label{Grainsize control data for depth 8 problems.}
+\includegraphics[width=0.6\columnwidth]{plots/dep8.pdf}
+\caption{Grainsize control data for depth 8 problems.}
+\label{dep8}
 \end{figure}
 
-Figure~\ref{dep10}
-\begin{figure}
+\begin{figure}[h]
 \centering
-\includegraphics[width=0.6\textwidth]{plots/dep10.pdf}
-\caption{dep10}
-\label{Grainsize control data for depth 10 problems.}
+\includegraphics[width=0.6\columnwidth]{plots/dep10.pdf}
+\caption{Grainsize control data for depth 10 problems.}
+\label{dep10}
 \end{figure}
 
-Figure~\ref{dep12}
-\begin{figure}
+\begin{figure}[h]
 \centering
-\includegraphics[width=0.6\textwidth]{plots/dep12.pdf}
-\caption{dep12}
-\label{Grainsize control data for depth 12 problems.}
+\includegraphics[width=0.6\columnwidth]{plots/dep12.pdf}
+\caption{Grainsize control data for depth 12 problems.}
+\label{dep12}
 \end{figure}
@@
 
-\section{Related Work}
-An overview of the theory of imperfect information games and other basic topics in game theory can be found in
-~\cite{kuhn03lectures} and~\cite{kuhn97classics}.  A technical explanation of the concept of information sets can be
+\Section{New Results}
+In addition to the random instances, we also evaluated our information set generation routine on 40 positions from actual kriespiel games, including human vs.
+human and computer vs.
+computer.
+Many of these were trivial to solve in a few seconds or less on a desktop computer.
+The running times and speedups for two of the more challenging instances are shown in Tables~\ref{prob4}--\ref{tab:prob5speedups}.
+These tables show the results for runs on one to $1024$ cores.
+Each column corresponds to a different parameter.
+In this case, the grainsize is the number of levels of the search tree that are searched in parallel.
+Up to that threshold, each node that is generated spawns a new piece of work, which the {\sc Charm++} runtime schedules to execute on one of the processing elements according to its load balancing strategy.
+Above the threshold the entire subtree is searched to completion on a single processor, with no additional parallel overhead.
+
+Note that the optimal grainsize varies for the different numbers of cores and the pattern is relatively consistent across problem instances.
+The optimal grainsize increases as the number of cores increases.
+This makes sense, because as this parameter grows, the number of distinct pieces of work increases.
+When the number of pieces of work per processor is high, much of the overhead of parallelization is wasted.
+But as the number of cores increases, the work can be efficiently broken up into smaller pieces and doled out to all the cores.
+With more pieces of work, it becomes easier to balance the load.
+
+\begin{figure}[htpb]
+ \centering \includegraphics[scale=0.4]{images/KriegspielProblem4.pdf} %
+\caption{Kriegspiel position from Virgil-David 1969 with information set size 764,209.}
+\label{prob4} 
+\end{figure}
+
+\begin{figure}[htpb]
+ \centering \includegraphics[scale=0.4]{images/KriegspielProblem5.pdf} %
+\caption{Kriegspiel position from Game 6 of Kbot-Darkboard at the 2006 Computer Chess championship.  Information set size is 5,099,257.}
+\label{prob5} 
+\end{figure}
+
+\begin{table}[tphb]
+\centering
+\begin{tabular}{rrrrr}
+PE/G & 2 & 4 & 6 & 8 \\
+1 & {\bf 177} & 178 & 184 & 493 \\
+4 & 71 & 53 & {\bf 48} & 75 \\
+16 & 34 & 19 & {\bf 12} & 16 \\
+64 & 24 & 7 & {\bf 3.7} & 4.1 \\
+256 & 24 & 4.4 & 1.2 & {\bf 1.0} \\
+1024 & 26 & 3.8 & 0.64 & {\bf 0.40}
+\end{tabular}
+\caption{Running times, in seconds, to find full information set of size 764,209 at position 10 of Virgil-David 1969. Optimal grainsize (G) for each number of cores (PE) is shown in bold}
+\label{tab:prob4}
+\end{table}
+
+\begin{table}[tphb]
+\centering
+\begin{tabular}{rrrrr}
+PE/G & 2 & 4 & 6 & 8 \\
+1 & 617 & {\bf 612} & 650 & 3312 \\
+4 & 213 & 170 & {\bf 158} & 367 \\
+16 & 83 & 49 & {\bf 40} & 64 \\
+64 & 54 & 19 & {\bf 11} & 14.9 \\
+256 & 54 & 9.8 & 3.5 & {\bf 3.5} \\
+1024 & 58 & 5.5 & 1.3 & {\bf 1.2} \\
+\end{tabular}
+\caption{Running times, in seconds, to find full information set of size 5,099,257 at position 10 of Game 6 Kbot-Darkboard 2006. Optimal grainsize for each number of cores is shown in bold.}
+\label{tab:prob5}
+\end{table}
+
+\begin{table}[tphb]
+\centering
+\begin{tabular}{rrrrr}
+PE/G & 2 & 4 & 6 & 8 \\
+1 & 1 & 0.99 & 0.96 & 0.36 \\ 
+4 & 2.48 & 3.34 & 3.63 & 2.35 \\
+16 & 5.19 & 9.27 & 13.8 & 10.7 \\
+64 & 7.25 & 23.6 & 47.39 & 43.1 \\
+256 & 7.1 & 40.1 & 145 & 172 \\
+1024 & 6.61 & 46.5 & 276 & {\bf 443}
+\end{tabular}
+\caption{Speedups for Virgil-David 1969.}
+\label{tab:prob4speedups}
+\end{table}
+
+\begin{table}[tphb]
+\centering
+\begin{tabular}{rrrrr}
+PE/G & 2 & 4 & 6 & 8 \\
+1 & 1 & 1 & 0.95 & 0.19 \\
+4 & 2.89 & 3.63 & 3.89 & 1.68 \\
+16 & 7.41 & 12.4 & 15.1 & 9.57 \\
+64 & 11.3 & 32.3 & 52.9 & 41.4 \\
+256 & 11.3 & 62.9 & 172 & 174 \\
+1024 & 10.6 & 111 & 464 & {\bf 536}
+\end{tabular}
+\caption{Speedups for Kbot-Darkboard 2006.}
+\label{tab:prob5speedups}
+\end{table}
+
+Note that both problems scale well even up to 1024 processors.
+We achieve a speedup of $443$ in the game between humans and up to $536$ in the game between computers.
+Figures~\ref{prob4} and~\ref{prob5} show the plots of the running times for the best grainsize for each core setting.
+These plots show that the running time continues to decrease as the number of cores increases.
+
+Our results compare favorably to previous work in parallel game tree search, shown in Table~\ref{bestspeedups}.
+The previous best parallelization results were for parallelizing alpha-beta pruning.
+Ferguson and Korf achieved a speedup of 12 on 32 cores in the game of Othello~\cite{ferguson88distributed}.
+In the game of chess, Felton and Otto showed a speedup of 100 on 256 cores~\cite{felten88highly}.
+This limited scalability is due to speculative loss.
+Parts of the game tree that are searched during the parallel execution would have been pruned had the algorithm run sequentially.
+This is a fundamental limitation in alpha-beta search.
+
+For partially observable games, information set generation represents a different variety of tree search that does not suffer from speculative loss.
+All of the nodes that are visited in parallel would have also been visited in a sequential run.
+The key challenge for parallelizing information set generation is load balancing: some parts of the tree will be pruned at a higher level than others, resulting in a wide variation in cost for each subtree.
+The {\sc Charm++/ParSSSE} framework is perfectly suited to handle the load balancing issues that this problem naturally presents.
+
+Note also that information set generation is just one subroutine needed to reason about partially observable games.
+Once plausible game histories have been found, an agent may simulate some sequences of moves up to previous moves made by opponents and then recursively call the information set generation procedure from the perspective of the opponent.
+This results in many parallel applications of information set generation.
+Furthermore, given a node in an informaion set, it is common for an agent to perform forward search on that node (e.g., using MCTS).
+With these two additional levels of parallelism on top of the information set generation problem, it is conceivable that the problem of playing partially observable games could scale efficiently to millions of processors or more.
+\begin{table}[tphb]
+\centering
+\begin{tabular}{lrrc}
+{\bf Authors} & {\bf Game} & {\bf Year} & {\bf Best Speedup (cores)} \\
+Ferguson \& Korf & Othello & 1988 & 12(32) \\
+Felton \& Otto & Chess & 1988 & 100(256) \\
+Richards \etal & Kriegspiel & 2012 & 178(256) \\
+Richards \etal & Kriegspiel & 2012 & {\bf 536(1024)} \\
+\end{tabular}
+\caption{Best Speedups reported for game tree applications.}
+\label{bestspeedups}
+\end{table}
+
+\Section{Related Work}
+An overview of the theory of imperfect information games and other basic topics in game theory can be found
+in~\cite{kuhn03lectures} and~\cite{kuhn97classics}.  A technical explanation of the concept of information sets can be
 found in~\cite{gilpin07algorithms}.  Algorithms for ``solving'' imperfect information games---in the sense of finding a
 Nash equilibrium---can be found in~\cite{koller94fast} and~\cite{koller97representations}.  Koller \etal, showed that an
-equilibrium can be found in time that is polynomial in the size of the game tree.  And in fact, they found that it took
-about as much time to generate the full game tree as it did to actually solve it.  Unfortunately, this is feasible only
-for small games.  In larger games where theoretically optimal strategies cannot be computed efficiently, computer
-systems tend to utilize the concept of infomration sets in some form or another.  A common strategy is to estimate the
-value of a move by averaging the estimated value of the associated descendant for each node in the current information
-set (or a random sample from the set).  While this sort of ``averaging over clairvoyance'' {\em can} lead to very poor
-decisions, it is nevertheless a reasonable strategy in some games (e.g., Scrabble~\cite{sheppard02world}).
+equilibrium can be found in time that is polynomial in the number of nodes in the game tree.  And in fact, they found
+that it took about as much time to generate the full game tree as it did to actually solve it.  Unfortunately, this is
+feasible only for small games.  In larger games where theoretically optimal strategies cannot be computed efficiently,
+computer systems tend to utilize the concept of information sets or belief states in some form or another.  A common
+strategy is to estimate the value of a move by averaging the estimated value of the associated descendant for each node
+in the current information set (or a random sample from the set).  While this sort of ``averaging over clairvoyance''
+{\em can} lead to very poor decisions, it is nevertheless a reasonable strategy in some games (e.g.,
+Scrabble~\cite{sheppard02world}).
 
 In this work, we have argued for an information set generation approach to kriegspiel, but other authors have approached
-the problem differently.  Parker \etal, consider the problem of sampling from belief states in a variety of large
-game trees, including kriegspiel~\cite{parker05game}.  They show that some performance gains can be achieved by only
+the problem differently.  Parker \etal, consider the problem of sampling from belief states in a variety of large game
+trees, including kriegspiel~\cite{parker05game}.  They show that some performance gains can be achieved by only
 approximate sampling from the belief state.  Instead of generating the full information set, or even sampling from it,
@@
 the full information set, because of the computational expense involved.  And indeed, our experiments bear this out-- it
-can indeed be expensive to produce the full information set.  For the experiments shown in Table XX, generating the full
-information set took over 25 minutes on a single processor.  However, by utilizing the Charm++ search engine on a
-parallel machine (1024 processors), we were able to generate the full set in just 2.9 seconds.  (Note that this set had
-more than 4 million members.)  Our results suggest that the problem is highly scalable and our approach might therefore
-be preferred.
+can indeed be expensive to produce the full information set.  For the experiments shown in Table~\ref{speedups},
+generating the full information set took over 25 minutes on a single processor.  However, by utilizing the Charm++
+search engine on a parallel machine (1024 processors), we were able to generate the full set in just 2.9 seconds.  (Note
+that this set had more than 4 million members.)  Our results suggest that the problem is highly scalable and our
+approach might therefore be preferred.
 
@@
 
-\section{Future Work}
+\Section{Future Work}
 
@@
 applied naively using a belief state sampling algorithm (i.e., an algorithm that samples from the distribution of
-possible worlds in the current state without respect to the overall history mof moves).  However, it has been shown that
+possible worlds in the current state without respect to the overall history of moves).  However, it has been shown that
 significant improvement in play can be achieved by estimating the value of future moves from the perspective of the
 opponent, based on the knowledge that the opponent would have had at prior positions in the move
-hisotry~\cite{richards07opponent}.  Such a strategy would require the use of an information set generation algorithm of the
+history~\cite{richards07opponent}.  Such a strategy would require the use of an information set generation algorithm of the
 kind that we have described here.\footnote{This work is underway.}
@@
 some games ourselves.  Currently, the manual encoding of games is not difficult but is tedious.  Still, it would be nice
-to have some games that go on for 40 or 50 moves and have positions where the information sets are on the order of a few
-hundred thousand or a few million in size.
-
-The information set generation algorithm that we have implemented here is one the most straightforward one and is
-alluded to in~\cite{parker05game} and~\cite{russell05efficient}.  It is analogous to the ``expand-at-the-tip'' strategy
-for the Hamiltonian Circuit problem.  An alternative information set generation algorithm is described in
-~\cite{richards09information} and could potentially improve the search efficiency greatly by exploiting variable
-ordering.  This would be analgous to the disconnected edge-pairs\footnote{This is what we referred to as ``algorithm 2'' in
-class.} heuristic for the Hamiltonian Circuit problem.  Rather than search from the root node going forward, this
-algorithm would seek opportunities for pruning based on the propagation of logical consequences from each move both
-forward and backward in time.  This, incidentally, is more consistent with the way human players would analyze the game.
-For example, if a player's proposed bishop move is rejected and the bishop is sufficiently removed from its own king,
-then the player may infer that one of his opponent's pieces is along that rejected path.  At least one of those spaces
-must have been the destination of a prior move by the opponent.  If a recent move by that bishop crossed the same path,
-then that would be hard evidence that those squares were empty at the time.  And therefore, the arrival of the
-opponent's piece must have been between the previous (successful) bishop move and the current (unsuccessful) bishop
-move.  To implement this algorithm, we envision the use of planning graph type data structures ~\cite{blum97fast}.
-Planning graphs have a sequence of alternating levels of state constraints and action constraints.  State constraints
-would be of the form $at(bishop,d4) \vee at(bishop,e6)$ or $!occupied(e4)$.  Action constraints would be of the form
-$move(knight,a4,c5) \vee move(knight,e6)$.  At a minimum, each level would require a data structure with a bit for each
-possible action/state.  Propagating the constraints would require much of the same machinery used in STRIPS-like
-planning algorithms~\cite{chen05solving}, and one of the many knobs that would need to be tuned would be the depth of the
-propagation for each constraint.  (Should the consequences of a known action or non-action be propagated one move into
-the future and one move into the past?  More?  Less?)  Additionally, there would be a significant and interesting
-tradeoff in the effectiveness of pruning and the and the size of the data structures that would have to be encoded in
-each chare.  The best algorithm might be a hybrid strategy that does forward propagation from the root at the parallel
-level and constraint propagation at the sequential level.\footnote{This work is also in progress.}
+to have some games that go on for 40 or 50 moves and have positions where the information sets are of moderate size (on
+the order of a few hundred thousand to a few million).
+
+The information set generation algorithm that we have implemented here is the most straightforward one and is alluded to
+in~\cite{parker05game} and~\cite{russell05efficient}.  It is analogous to the ``expand-at-the-tip'' strategy for the
+Hamiltonian Circuit problem.  An alternative information set generation algorithm is described
+in~\cite{richards09information} (but not implemented in parallel) and could potentially improve the search efficiency
+greatly by exploiting variable ordering.  This would be analogous to the disconnected edge-pairs\footnote{This is what we
+referred to as ``algorithm 2'' in class.} heuristic for the Hamiltonian Circuit problem.  Rather than search from the
+root node going forward, this algorithm would seek opportunities for pruning based on the propagation of logical
+consequences from each move both forward and backward in time.  This, incidentally, is more consistent with the way
+human players would analyze the game.  For example, if a player's proposed bishop move is rejected and the bishop is
+sufficiently removed from its own king, then the player may infer that one of his opponent's pieces is along that
+rejected path.  At least one of those spaces must have been the destination of a prior move by the opponent.  If a
+recent move by that bishop crossed the same path, then that would be hard evidence that those squares were empty at the
+time.  And therefore, the arrival of the opponent's piece must have been between the previous (successful) bishop move
+and the current (unsuccessful) bishop move.  To implement this algorithm, we envision the use of planning graph type
+data structures~\cite{blum97fast}.  Planning graphs have a sequence of alternating levels of state constraints and
+action constraints.  State constraints would be of the form $at(bishop,d4) \vee at(bishop,e6)$ or $!occupied(e4)$.
+Action constraints would be of the form $move(knight,a4,c5) \vee move(knight,e6)$.  At a minimum, each level would
+require a data structure with a bit for each possible action/state.  Propagating the constraints would require much of
+the same machinery used in STRIPS-like planning algorithms~\cite{chen05solving}, and one of the many knobs that would
+need to be tuned would be the depth of the propagation for each constraint.  (Should the consequences of a known action
+or non-action be propagated one move into the future and one move into the past?  More?  Less?)  Additionally, there
+would be a significant and interesting tradeoff in the effectiveness of pruning and the size of the data
+structures that would have to be encoded in each chare.  The best algorithm might be a hybrid strategy that does forward
+propagation from the root at the parallel level and constraint propagation at the sequential level.\footnote{This work
+is also in progress.}
 
@@
 
-\section{Appendix A}
-The purpose of this figure is simply to help visualize the concept of information sets in game trees.  The tree for
-kriegspiel is very difficult to visualize/depict (even an interesting part of the tree).  Figure~\ref{fulltree} shows
-the full game tree for a simple variant of poker known as Kuhn poker, which we hope will serve to at least make the
-concept clear, even if the poker game itself is not immediately relevant.  There are three cards: a king, a queen, and a
-jack.  The dealer deals one card to each of two players.  Each player antes one unit, and there is one simple round of
-betting, in which the size of the pot may be doubled.  Left branches denote check/fold; right branches denote bet/call.
-For this zero-sum game, terminal nodes are labeled with payoffs to player 1, whose decision points are shown in
-triangles.  Connections between nodes in the same information set are shown with dotted lines.  The top level shows
-three different information sets for the first player, one for each possible card that could be dealt to him.  There are
-two nodes in each set, because no matter what the deal is, there are two possibilities for the card that the other
-player holds.  The leftmost information set on the bottom level corresponds to player 1's observation that he was dealt
-a king and checked, and then the opponent wagered.  The second player has six information sets as well, but they are all
-at the same level of the tree.  Like the first player, the second player has a separate information set for each
-possible card that he could be dealt.  And the two nodes in each set are there because there are two possibilities for
-the opponent card.  The second player also gets to know whether the first player checked of bet, and that is why there
-are twice as many information sets.
-
-\begin{figure}
-\scalebox{.72}{
-\begin{tikzpicture}
-  [chance/.style={circle,draw=blue!50,fill=blue!20,thick,minimum size = 10mm},
-   terminal/.style={rectangle,draw=black!50,fill=black!20,thick, minimum size = 7mm},
-   maxer/.style={shape=regular polygon, regular polygon sides=3,draw=red!50,fill=red!30,thick,minimum size = 10mm,inner sep = 0pt},
-   miner/.style={shape=diamond,draw=green!50,fill=green!30,thick,minimum size = 10mm, inner sep = 0pt},
-   edge from parent/.style={red,thick,draw}, 
-   parent anchor=south,child anchor=north,
-   level 1/.style={sibling distance=4cm,level distance=1.4cm,
-       		growth parent anchor=south},
-   level 2/.style={sibling distance=2cm},
-   level 3/.style={sibling distance=1cm},
-   level 4/.style={sibling distance=1.0cm}]
-		\node  [chance] {}
-		    child {node (M1) [maxer] {\scalebox{.75}{K}}
-			child {node (M11) [miner] {\scalebox{.75}{Q-}}
-				child {node (M111) [terminal] {1}}
-				child {node (M112) [maxer] {\scalebox{.75}{K+}}
-					child {node (M1121) [terminal] {-1}}
-					child {node (M1122) [terminal] {2}}}
-			}
-			child {node (M12) [miner] {\scalebox{.75}{Q+}}
-				child {node (M121) [terminal] {1}}
-				child {node (M122) [terminal] {2}}
-			}
-		    }
-		    child {node (M2) [maxer] {\scalebox{.75}{J}}
-			child {node (M21) [miner] {\scalebox{.75}{Q-}}
-				child {node (M211) [terminal] {-1}}
-				child {node (M212) [maxer] {\scalebox{.75}{J+}}
-					child {node (M2121) [terminal] {-1}}
-					child {node (M2122) [terminal] {-2}}}
-			}
-			child {node (M22) [miner] {\scalebox{.75}{Q+}}
-				child {node (M221) [terminal] {1}}
-				child {node (M222) [terminal] {-2}}
-			}
-		    }
-		    child {node (M3) [maxer] {\scalebox{.75}{K}}
-			child {node (M31) [miner] {\scalebox{.75}{J-}}
-				child {node (M311) [terminal] {1}}
-				child {node (M312) [maxer] {\scalebox{.75}{K+}}
-					child {node (M3121) [terminal] {-1}}
-					child {node (M3122) [terminal] {2}}}
-			}
-			child {node (M32) [miner] {\scalebox{.75}{J+}}
-				child {node (M321) [terminal] {1}}
-				child {node (M322) [terminal] {2}}
-			}
-		    }
-		    child {node (M4) [maxer] {\scalebox{.75}{Q}}
-			child {node (M41) [miner] {\scalebox{.75}{J-}}
-				child {node (M411) [terminal] {1}}
-				child {node (M412) [maxer] {\scalebox{.75}{Q+}}
-					child {node (M4121) [terminal] {-1}}
-					child {node (M4122) [terminal] {2}}}
-			}
-			child {node (M42) [miner] {\scalebox{.75}{J+}}
-				child {node (M421) [terminal] {1}}
-				child {node (M422) [terminal] {2}}
-			}
-		    }
-		    child {node (M5) [maxer] {\scalebox{.75}{J}}
-			child {node (M51) [miner] {\scalebox{.75}{K-}}
-				child {node (M511) [terminal] {-1}}
-				child {node (M512) [maxer] {\scalebox{.75}{J+}}
-					child {node (M5121) [terminal] {-1}}
-					child {node (M5122) [terminal] {-2}}}
-			}
-			child {node (M52) [miner] {\scalebox{.75}{K+}}
-				child {node (M521) [terminal] {1}}
-				child {node (M522) [terminal] {-2}}
-			}
-		    }
-		    child {node (M6) [maxer] {\scalebox{.75}{Q}}
-			child {node (M61) [miner] {\scalebox{.75}{K-}}
-				child {node (M611) [terminal] {-1}}
-				child {node (M612) [maxer] {\scalebox{.75}{Q+}}
-					child {node (M6121) [terminal] {-1}}
-					child {node (M6122) [terminal] {-2}}}
-			}
-			child {node (M62) [miner] {\scalebox{.75}{K+}}
-				child {node (M621) [terminal] {1}}
-				child {node (M622) [terminal] {-2}}
-			}
-		    };
-		\draw  (M1.north) to [dashed,black,out=15,in=165] (M3.north);
-		\draw  (M2.north) to [dashed,black,out=15,in=165] (M5.north);
-		\draw  (M4.north) to [dashed,black,out=15,in=165] (M6.north);
-		\draw  (M112.north) to [dashed,black,out=15,in=165] (M312.north);
-		\draw  (M212.north) to [dashed,black,out=15,in=165] (M512.north);
-		\draw  (M412.north) to [dashed,black,out=15,in=165] (M612.north);
-		\draw  (M11.north) to [dashed,black,out=15,in=165] (M21.north);
-		\draw  (M12.north) to [dashed,black,out=15,in=165] (M22.north);
-		\draw  (M31.north) to [dashed,black,out=15,in=165] (M41.north);
-		\draw  (M32.north) to [dashed,black,out=15,in=165] (M42.north);
-		\draw  (M51.north) to [dashed,black,out=15,in=165] (M61.north);
-		\draw  (M52.north) to [dashed,black,out=15,in=165] (M62.north);
-\end{tikzpicture}
-} %scalebox
-\caption{}
-\label{fulltree}
-\end{figure}
+\Section{Conclusion}
+We have shown that information set generation for kriegspiel can be efficiently parallelized.  The opportunities for
+parallelism increase as the problem size increases.  Our algorithm has performed well on problems with information sets
+with as many as 65 million nodes---much larger than any other treatment of the problem that we are aware of.  Because
+the algorithm can be efficiently parallelized, it would be reasonable to explore using it in a full kriegspiel player.
+It remains to be seen whether these scalability properties hold over a wide variety of playing styles and problem
+instances.  We suspect that fruitful research remains with respect to alternative algorithms that implement variable
+ordering heuristics.
+
+%\Section{Appendix A}
+%The purpose of this figure is simply to help visualize the concept of information sets in game trees.  The tree for
+%kriegspiel is very difficult to visualize/depict (even an interesting part of the tree).  Figure~\ref{fulltree} shows
+%the full game tree for a simple variant of poker known as Kuhn poker, which we hope will serve to at least make the
+%concept clear, even if the poker game itself is not immediately relevant.  There are three cards: a king, a queen, and a
+%jack.  The dealer deals one card to each of two players.  Each player antes one unit, and there is one simple round of
+%betting, in which the size of the pot may be doubled.  Left branches denote check/fold; right branches denote bet/call.
+%For this zero-sum game, terminal nodes are labeled with payoffs to player 1, whose decision points are shown in
+%triangles.  Connections between nodes in the same information set are shown with dotted lines.  The top level shows
+%three different information sets for the first player, one for each possible card that could be dealt to him.  There are
+%two nodes in each set, because no matter what the deal is, there are two possibilities for the card that the other
+%player holds.  The leftmost information set on the bottom level corresponds to player 1's observation that he was dealt
+%a king and checked, and then the opponent wagered.  The second player has six information sets as well, but they are all
+%at the same level of the tree.  Like the first player, the second player has a separate information set for each
+%possible card that he could be dealt.  And the two nodes in each set are there because there are two possibilities for
+%the opponent card.  The second player also gets to know whether the first player checked of bet, and that is why there
+%are twice as many information sets.
+%
+%\begin{figure}
+%\scalebox{.72}{
+%\begin{tikzpicture}
+%  [chance/.style={circle,draw=blue!50,fill=blue!20,thick,minimum size = 10mm},
+%   terminal/.style={rectangle,draw=black!50,fill=black!20,thick, minimum size = 7mm},
+%   maxer/.style={shape=regular polygon, regular polygon sides=3,draw=red!50,fill=red!30,thick,minimum size = 10mm,inner sep = 0pt},
+%   miner/.style={shape=diamond,draw=green!50,fill=green!30,thick,minimum size = 10mm, inner sep = 0pt},
+%   edge from parent/.style={red,thick,draw}, 
+%   parent anchor=south,child anchor=north,
+%   level 1/.style={sibling distance=4cm,level distance=1.4cm,
+%       		growth parent anchor=south},
+%   level 2/.style={sibling distance=2cm},
+%   level 3/.style={sibling distance=1cm},
+%   level 4/.style={sibling distance=1.0cm}]
+%		\node  [chance] {}
+%		    child {node (M1) [maxer] {\scalebox{.75}{K}}
+%			child {node (M11) [miner] {\scalebox{.75}{Q-}}
+%				child {node (M111) [terminal] {1}}
+%				child {node (M112) [maxer] {\scalebox{.75}{K+}}
+%					child {node (M1121) [terminal] {-1}}
+%					child {node (M1122) [terminal] {2}}}
+%			}
+%			child {node (M12) [miner] {\scalebox{.75}{Q+}}
+%				child {node (M121) [terminal] {1}}
+%				child {node (M122) [terminal] {2}}
+%			}
+%		    }
+%		    child {node (M2) [maxer] {\scalebox{.75}{J}}
+%			child {node (M21) [miner] {\scalebox{.75}{Q-}}
+%				child {node (M211) [terminal] {-1}}
+%				child {node (M212) [maxer] {\scalebox{.75}{J+}}
+%					child {node (M2121) [terminal] {-1}}
+%					child {node (M2122) [terminal] {-2}}}
+%			}
+%			child {node (M22) [miner] {\scalebox{.75}{Q+}}
+%				child {node (M221) [terminal] {1}}
+%				child {node (M222) [terminal] {-2}}
+%			}
+%		    }
+%		    child {node (M3) [maxer] {\scalebox{.75}{K}}
+%			child {node (M31) [miner] {\scalebox{.75}{J-}}
+%				child {node (M311) [terminal] {1}}
+%				child {node (M312) [maxer] {\scalebox{.75}{K+}}
+%					child {node (M3121) [terminal] {-1}}
+%					child {node (M3122) [terminal] {2}}}
+%			}
+%			child {node (M32) [miner] {\scalebox{.75}{J+}}
+%				child {node (M321) [terminal] {1}}
+%				child {node (M322) [terminal] {2}}
+%			}
+%		    }
+%		    child {node (M4) [maxer] {\scalebox{.75}{Q}}
+%			child {node (M41) [miner] {\scalebox{.75}{J-}}
+%				child {node (M411) [terminal] {1}}
+%				child {node (M412) [maxer] {\scalebox{.75}{Q+}}
+%					child {node (M4121) [terminal] {-1}}
+%					child {node (M4122) [terminal] {2}}}
+%			}
+%			child {node (M42) [miner] {\scalebox{.75}{J+}}
+%				child {node (M421) [terminal] {1}}
+%				child {node (M422) [terminal] {2}}
+%			}
+%		    }
+%		    child {node (M5) [maxer] {\scalebox{.75}{J}}
+%			child {node (M51) [miner] {\scalebox{.75}{K-}}
+%				child {node (M511) [terminal] {-1}}
+%				child {node (M512) [maxer] {\scalebox{.75}{J+}}
+%					child {node (M5121) [terminal] {-1}}
+%					child {node (M5122) [terminal] {-2}}}
+%			}
+%			child {node (M52) [miner] {\scalebox{.75}{K+}}
+%				child {node (M521) [terminal] {1}}
+%				child {node (M522) [terminal] {-2}}
+%			}
+%		    }
+%		    child {node (M6) [maxer] {\scalebox{.75}{Q}}
+%			child {node (M61) [miner] {\scalebox{.75}{K-}}
+%				child {node (M611) [terminal] {-1}}
+%				child {node (M612) [maxer] {\scalebox{.75}{Q+}}
+%					child {node (M6121) [terminal] {-1}}
+%					child {node (M6122) [terminal] {-2}}}
+%			}
+%			child {node (M62) [miner] {\scalebox{.75}{K+}}
+%				child {node (M621) [terminal] {1}}
+%				child {node (M622) [terminal] {-2}}
+%			}
+%		    };
+%		\draw  (M1.north) to [dashed,black,out=15,in=165] (M3.north);
+%		\draw  (M2.north) to [dashed,black,out=15,in=165] (M5.north);
+%		\draw  (M4.north) to [dashed,black,out=15,in=165] (M6.north);
+%		\draw  (M112.north) to [dashed,black,out=15,in=165] (M312.north);
+%		\draw  (M212.north) to [dashed,black,out=15,in=165] (M512.north);
+%		\draw  (M412.north) to [dashed,black,out=15,in=165] (M612.north);
+%		\draw  (M11.north) to [dashed,black,out=15,in=165] (M21.north);
+%		\draw  (M12.north) to [dashed,black,out=15,in=165] (M22.north);
+%		\draw  (M31.north) to [dashed,black,out=15,in=165] (M41.north);
+%		\draw  (M32.north) to [dashed,black,out=15,in=165] (M42.north);
+%		\draw  (M51.north) to [dashed,black,out=15,in=165] (M61.north);
+%		\draw  (M52.north) to [dashed,black,out=15,in=165] (M62.north);
+%\end{tikzpicture}
+%} %scalebox
+%\caption{}
+%\label{fulltree}
+%\end{figure}
 
@@
 
-\section{Appendix B}
-\begin{verbatim}
-1. Pa2:a4 Pa7:a5 2. Pd2:d4 Pb7:b6 3. Bc1:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-p - - - - - - - 
-- - - - - - - - 
-- p p - p p p p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d3 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pd3:d4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - - - - - - - 
-p p p - p p p p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Pa2:a4 Pb7:b6 3. Bc1:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-p - - - - - - - 
-- - - - - - - - 
-- p p - p p p p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:f4 Pb7:b6 3. Bf4:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - - - - - - - 
-p p p - p p p p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-p - - - - - - - 
-- p p - p p p p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-p - - - - - - - 
-- - - - - - - - 
-- p p - p p p p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pb2:b3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- p - - - - - - 
-p - p - p p p p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pc2:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - p - - - - - 
-p p - - p p p p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - - - p - - - 
-p p p - - p p p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - p - - - 
-- - - - - - - - 
-p p p - - p p p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - - - - p - - 
-p p p - p - p p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - p - - 
-- - - - - - - - 
-p p p - p - p p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - - - - - p - 
-p p p - p p - p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - p - 
-- - - - - - - - 
-p p p - p p - p 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - - - - - - p 
-p p p - p p p - 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - p 
-- - - - - - - - 
-p p p - p p p - 
-r n - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-n - - - - - - - 
-p p p - p p p p 
-r - - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - n - - - - - 
-p p p - p p p p 
-r - - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - - - - - - - 
-p p p n p p p p 
-r - - q k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - - - - - - - 
-p p p q p p p p 
-r n - - k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - - q - - - - 
-p p p - p p p p 
-r n - - k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:c1 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - - - - - - - 
-p p p - p p p p 
-r n q - k b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ke1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - - - - - - - 
-p p p k p p p p 
-r n - q - b n r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - - - - n - - 
-p p p - p p p p 
-r n - q k b - r 
-**********************************************
-1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6
-
-**********************************************
-R - B Q K B N R 
-- - - P P - P P 
-N P - p - P - - 
-P - P - - - b - 
-- - - - - - - - 
-- - - - - - - n 
-p p p - p p p p 
-r n - q k b - r 
-**********************************************
-Solutions founds: 25
-\end{verbatim}
-%\caption{Verbose output from example game.  Generating information sets for black after five turns by each player.
-%From this, black can infer that there is definitely a white bishop at g5 and a white pawn at d6.}
-%\label{verboseoutput}
-%\end{figure}
-
-\bibliographystyle{acm}
+%\Section{Appendix B}
+%Verbose output from the information set generation algorithm.  The listing shows possible sequences of moves that are
+%consistent with the black's observations after ten plies.  The resulting position after each consistent sequence of
+%moves is shown below.  Based on these results, black can know with certainty the location of white's pawn at d6 and
+%white's bishop at g5.
+%
+%\begin{verbatim}
+%1. Pa2:a4 Pa7:a5 2. Pd2:d4 Pb7:b6 3. Bc1:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+%
+%**********************************************
+%R - B Q K B N R 
+%- - - P P - P P 
+%N P - p - P - - 
+%P - P - - - b - 
+%p - - - - - - - 
+%- - - - - - - - 
+%- p p - p p p p 
+%r n - q k b n r 
+%**********************************************
+%1. Pd2:d3 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pd3:d4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+%
+%**********************************************
+%R - B Q K B N R 
+%- - - P P - P P 
+%N P - p - P - - 
+%P - P - - - b - 
+%- - - - - - - - 
+%- - - - - - - - 
+%p p p - p p p p 
+%r n - q k b n r 
+%**********************************************
+%1. Pd2:d4 Pa7:a5 2. Pa2:a4 Pb7:b6 3. Bc1:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+%
+%**********************************************
+%R - B Q K B N R 
+%- - - P P - P P 
+%N P - p - P - - 
+%P - P - - - b - 
+%p - - - - - - - 
+%- - - - - - - - 
+%- p p - p p p p 
+%r n - q k b n r 
+%**********************************************
+%1. Pd2:d4 Pa7:a5 2. Bc1:f4 Pb7:b6 3. Bf4:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+%
+%**********************************************
+%R - B Q K B N R 
+%- - - P P - P P 
+%N P - p - P - - 
+%P - P - - - b - 
+%- - - - - - - - 
+%- - - - - - - - 
+%p p p - p p p p 
+%r n - q k b n r 
+%**********************************************
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+%
+%**********************************************
+%R - B Q K B N R 
+%- - - P P - P P 
+%N P - p - P - - 
+%P - P - - - b - 
+%- - - - - - - - 
+%p - - - - - - - 
+%- p p - p p p p 
+%r n - q k b n r 
+%**********************************************
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+%
+%**********************************************
+%R - B Q K B N R 
+%- - - P P - P P 
+%N P - p - P - - 
+%P - P - - - b - 
+%p - - - - - - - 
+%- - - - - - - - 
+%- p p - p p p p 
+%r n - q k b n r 
+%**********************************************
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pb2:b3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+%
+%**********************************************
+%R - B Q K B N R 
+%- - - P P - P P 
+%N P - p - P - - 
+%P - P - - - b - 
+%- - - - - - - - 
+%- p - - - - - - 
+%p - p - p p p p 
+%r n - q k b n r 
+%**********************************************
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pc2:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+%
+%**********************************************
+%R - B Q K B N R 
+%- - - P P - P P 
+%N P - p - P - - 
+%P - P - - - b - 
+%- - - - - - - - 
+%- - p - - - - - 
+%p p - - p p p p 
+%r n - q k b n r 
+%**********************************************
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+%
+%**********************************************
+%R - B Q K B N R 
+%- - - P P - P P 
+%N P - p - P - - 
+%P - P - - - b - 
+%- - - - - - - - 
+%- - - - p - - - 
+%p p p - - p p p 
+%r n - q k b n r 
+%**********************************************
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+%
+%**********************************************
+%R - B Q K B N R 
+%- - - P P - P P 
+%N P - p - P - - 
+%P - P - - - b - 
+%- - - - p - - - 
+%- - - - - - - - 
+%p p p - - p p p 
+%r n - q k b n r 
+%**********************************************
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+%
+%**********************************************
+%R - B Q K B N R 
+%- - - P P - P P 
+%N P - p - P - - 
+%P - P - - - b - 
+%- - - - - - - - 
+%- - - - - p - - 
+%p p p - p - p p 
+%r n - q k b n r 
+%**********************************************
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+%
+%**********************************************
+%R - B Q K B N R 
+%- - - P P - P P 
+%N P - p - P - - 
+%P - P - - - b - 
+%- - - - - p - - 
+%- - - - - - - - 
+%p p p - p - p p 
+%r n - q k b n r 
+%**********************************************
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+%
+%**********************************************
+%R - B Q K B N R 
+%- - - P P - P P 
+%N P - p - P - - 
+%P - P - - - b - 
+%- - - - - - - - 
+%- - - - - - p - 
+%p p p - p p - p 
+%r n - q k b n r 
+%**********************************************
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+%
+%**********************************************
+%R - B Q K B N R 
+%- - - P P - P P 
+%N P - p - P - - 
+%P - P - - - b - 
+%- - - - - - p - 
+%- - - - - - - - 
+%p p p - p p - p 
+%r n - q k b n r 
+%**********************************************
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+%
+%**********************************************
+%R - B Q K B N R 
+%- - - P P - P P 
+%N P - p - P - - 
+%P - P - - - b - 
+%- - - - - - - - 
+%- - - - - - - p 
+%p p p - p p p - 
+%r n - q k b n r 
+%**********************************************
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+%
+%**********************************************
+%R - B Q K B N R 
+%- - - P P - P P 
+%N P - p - P - - 
+%P - P - - - b - 
+%- - - - - - - p 
+%- - - - - - - - 
+%p p p - p p p - 
+%r n - q k b n r 
+%**********************************************
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+%
+%**********************************************
+%R - B Q K B N R 
+%- - - P P - P P 
+%N P - p - P - - 
+%P - P - - - b - 
+%- - - - - - - - 
+%n - - - - - - - 
+%p p p - p p p p 
+%r - - q k b n r 
+%**********************************************
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+%
+%**********************************************
+%R - B Q K B N R 
+%- - - P P - P P 
+%N P - p - P - - 
+%P - P - - - b - 
+%- - - - - - - - 
+%- - n - - - - - 
+%p p p - p p p p 
+%r - - q k b n r 
+%**********************************************
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+%
+%**********************************************
+%R - B Q K B N R 
+%- - - P P - P P 
+%N P - p - P - - 
+%P - P - - - b - 
+%- - - - - - - - 
+%- - - - - - - - 
+%p p p n p p p p 
+%r - - q k b n r 
+%**********************************************
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+%
+%**********************************************
+%R - B Q K B N R 
+%- - - P P - P P 
+%N P - p - P - - 
+%P - P - - - b - 
+%- - - - - - - - 
+%- - - - - - - - 
+%p p p q p p p p 
+%r n - - k b n r 
+%**********************************************
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+%
+%**********************************************
+%R - B Q K B N R 
+%- - - P P - P P 
+%N P - p - P - - 
+%P - P - - - b - 
+%- - - - - - - - 
+%- - - q - - - - 
+%p p p - p p p p 
+%r n - - k b n r 
+%**********************************************
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:c1 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+%
+%**********************************************
+%R - B Q K B N R 
+%- - - P P - P P 
+%N P - p - P - - 
+%P - P - - - b - 
+%- - - - - - - - 
+%- - - - - - - - 
+%p p p - p p p p 
+%r n q - k b n r 
+%**********************************************
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ke1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+%
+%**********************************************
+%R - B Q K B N R 
+%- - - P P - P P 
+%N P - p - P - - 
+%P - P - - - b - 
+%- - - - - - - - 
+%- - - - - - - - 
+%p p p k p p p p 
+%r n - q - b n r 
+%**********************************************
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+%
+%**********************************************
+%R - B Q K B N R 
+%- - - P P - P P 
+%N P - p - P - - 
+%P - P - - - b - 
+%- - - - - - - - 
+%- - - - - n - - 
+%p p p - p p p p 
+%r n - q k b - r 
+%**********************************************
+%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 
+%
+%**********************************************
+%R - B Q K B N R 
+%- - - P P - P P 
+%N P - p - P - - 
+%P - P - - - b - 
+%- - - - - - - - 
+%- - - - - - - n 
+%p p p - p p p p 
+%r n - q k b - r 
+%**********************************************
+%Solutions founds: 25
+%\end{verbatim}
+%%\caption{Verbose output from example game.  Generating information sets for black after five turns by each player.
+%%From this, black can infer that there is definitely a white bishop at g5 and a white pawn at d6.}
+%%\label{verboseoutput}
+%%\end{figure}
+
+\bibliographystyle{latex8}
 \bibliography{paper}

### 2012-06-25 sbac/paper.tex
Pared doc down to eight pages

@@
 Information Set Generation is the identification of the set of paths in an imperfect information game tree that are
-consistent with a player's observations.  The ability to reason about the possible game history is crucial to the
-performance of game playing agents.  In this work, we discuss the problem of information set generation in the context
+consistent with a player's observations.  The ability to reason about the possible game history is critical to the
+performance of game playing agents.  In this work, we address the parallelization of information set generation in the context
 of kriegspiel (partially observable chess).  We implement the algorithm on top of a general purpose combinatorial search
 engine and discuss its performance under different load balancing strategies and with different grainsize parameters.
-Our results suggest that our algorithm can be efficiently parallelized.  
+We achieve speedups of over 500 on 1024 processors, far exceeding previous scalability results for game tree search applications.
 \end{abstract}
@@
 \Section{Introduction}
-In imperfect information games, players do not have access to full knowledge of the world. Examples of imperfect
-information include hidden cards in Poker~\cite{billings02challenge} or Bridge~\cite{ginsberg96partition}, hidden tiles
-in Scrabble~\cite{richards07opponent}, or hidden pieces in Stratego or Kriegspiel (partially observable
-chess)~\cite{li94chess}. The game tree nodes that are indistinguishable to a player because they differ only in the
-information that is hidden to the player by rule are called that player's {\em information set.}  The ability to
-estimate the value of the possible states and to reason about the probability distribution over those states is crucial
-to playing imperfect information games well.
+In imperfect information games, players do not have access to full knowledge of
+the world. Examples of imperfect information include hidden cards in
+Poker~\cite{billings02challenge} or Bridge~\cite{ginsberg96partition}, hidden
+tiles in Scrabble~\cite{richards07opponent}, or hidden pieces 
+Kriegspiel (partially observable chess)~\cite{li94chess}. The game tree nodes
+that are indistinguishable to a player because they differ only in the
+information that is hidden to the player by rule compose that player's {\em
+information set.}  The ability to estimate the value of the possible states and
+to reason about the probability distribution over those states is crucial to
+playing imperfect information games well~\cite{thielscher12hyperplay}.
+
+Information Set Generation (ISG) is an operation that game tree search agents
+can use to identify nodes in the game tree that are consistent with their
+observations~\cite{richards12reasoning}.  Because the operation is relatively
+expensive and may be repeatedly applied by a decision-making agent, its
+efficiency is a key factor in an agent's overall performance.  In this work, we
+show that ISG for kriegspiel can be efficiently parallelized.
 
@@
 
-\Section{Background}
-In this section, we describe the game of kriegspiel, and the particular variation of it that we have implemented.
-
-Kriegspiel was invented by Henry Michael Temple in 1899.  Historically, the game has required three parties: two
-competing players and one referee.  Each player sits with a chessboard that is partitioned off so as not be seen by the
-other.  One player controls the white pieces; the other controls the black pieces.  The referee has a board on which she
-keeps track of both sets of pieces.  Each player has a full set of chess pieces.  He may position the opponent's pieces
-however he wishes on his own board (if at all).  The players alternate moves as in a regular chess game, but instead of
-announcing their moves out loud, they write the source and destination square of their desired plays on a piece of paper
-and pass them to the referee, so that the requested move is not known to the other player. The referee, who knows the location of
-all pieces, checks the legality of the move.  If the move is legal, the referee makes the move on his own board and
-announces that a legal move has been made and that it is the other player's turn to move.  If the move is illegal
-because it is blocked by an opponent's piece or would place or leave the active player in check, the referee announces,
-``No.''  There is no penalty for attempting an illegal move.  The player continues to attempt moves until one finally
-succeeds.  If a player has no legal moves (because of stalemate or checkmate), the referee announces the game's result.
-With the advent of computers, it is convenient to play the game over a network, with a computer acting as the referee.
-
-The referee makes other announcements besides declaring move illegal, and there are several variations of the game that
-differ only in the nature of these announcements.  All declarations by the referee are heard by both players.  Thus, for
-example, a player will hear and know if her opponent has attempted an illegal move, and will thus know that her
-opponent's pieces are configured in such a way as to allow at least one attemptable move that blocked or would leave him
-in check.
-
-The referee announces when a player is in check.  (And again, this announcement is heard by both players.)  There is
-some variation in what additional information is supplied in this case.  In one popular variant, the referee declares
-that the player is in check ``by rank,'' ``by file,'' ``by diagonal,'' or ``by knight.''  In the case of a diagonal attack, the
-referee may announce whether the checking piece is along the long or short diagonal (from the king's perspective).  It
-is possible to be in check by two pieces at the same time, in which case the direction of both attacks would be
-declared.  In our implementation, for simplicity, the referee declares only that the player is in check, without
-specifying the nature of the attack.
-
-In the case of the capture, the referee announces the location of the piece that was captured, so that both players can
-remove the piece from the board.  In some variations, the referee also announces whether the captured piece was a pawn
-or non-pawn.  In our implementation, information about the captured piece is not explicitly provided to the player who
-makes the capture.
-
-Because there is no penalty for making illegal moves, it is often to a player's advantage to attempt to make moves which
-are likely to be illegal, as hearing such declarations from the referee can provide important information to a player
-about the location of the opponent's pieces.  In particular, since a diagonal move by a pawn is legal only the case
-where such a move would capture an opponent's piece, it could potentially be profitable to attempt all or many pawn
-captures on every turn.  In order to speed the game up, popular variants of the game require the referee to make some
-kind of declaration at the beginning of a player's turn with respect to the pawn captures that are available.  This may
-be as simple as declaring ``Try'' to indicate that at least one pawn capture is possible.  In our implementation, the
-referee declares all potential pawn captures.
-
-In the case of a pawn promotion, the player making the pawn advancement would secretly notify the referee as to which
-piece (queen, rook, bishop, or knight) he would like.  To reduce the number of bits needed to specify a move, our
-implementation allows only promotion to queen.
-
-In general, any move that is legal in a regular chess match may be at least attempted by a player in kriegspiel.  And
-any kriegspiel move that is ultimately allowed by the referee would also be a legal move in chess.
-
-We assume that a player may not attempt the same illegal move more than once per turn.  This means that if a player
-hears multiple declarations from the referee that the opponent has attempted an illegal move, the player may assume that
-the opponent has at least that many possibilities available to him.
-
-\SubSection{Special Moves}
-Our implementation supports the capability for pawns to move two spaces on their first turn.  However, we did not
-implement support for castling or {\em en passant} pawn capture.  (These moves require maintaining additional bits of
-information at each state that keep track of whether pieces have moved and/or whether pieces moved on the previous
-turn.)
+The paper is organized as follows.
+In Section~\ref{Background}, we review necessary background information, including the rules of kriegspiel and the nature of the information set generation problem.
+In Section ~\ref{ParSSSE}, we describe the {\sc Charm++} Parallel State Space Search Engine (ParSSSE) platform.
+We then provide our experimental results in Section~\ref{Results}.
+Related work is discussed in Section~\ref{Related}.
+We conclude with a discussion of future work.
+
+\Section{Background}\label{Background}
+\subsection{Kriegspiel}
+Kriegspiel, or partially observable chess,  was invented by Henry Michael Temple in 1899.  Historically, the
+game has required three parties: two competing players and one referee.  Today,
+players often compete over a computer network, with the role of the referee
+being automated.  One player sees and controls only the white pieces; the
+other, black.  The referee track of both sets of pieces.  The players alternate
+moves as in a regular chess game, but instead of openly declaring their moves,
+they secretly communicate move requests to the referee.  The referee, who knows
+the location of all pieces, checks the legality of the move.
+
+If the move is legal, the referee makes executes the move and announces that a
+legal move has been made and that it is the other player's turn to move.  If
+the move is illegal because it is blocked by an opponent's piece or would place
+or leave the active player in check, the referee announces, ``No.''  The player
+then continues, without penalty, to attempt moves until one finally succeeds.
+If a player has no legal moves because of stalemate or checkmate, the referee
+announces the game's result.
+
+The referee makes other announcements besides declaring move illegal, and there
+are several variations of the game that differ only in the nature of these
+announcements.  All declarations by the referee are made to both players.  In
+particular, a player will hear and know if her opponent has attempted an
+illegal move, and will thus know that her opponent's pieces are configured in
+such a way as to allow at least one attemptable move that is blocked or would
+leave him in check.
+
+The referee announces when a player is in check. In the case of the capture,
+the referee announces the location of the piece that was captured but not its
+identity.  Because there is no penalty for making illegal moves, it is often to
+a player's advantage to attempt to make moves which are likely to be illegal,
+as hearing such declarations from the referee can provide important information
+to a player about the location of the opponent's pieces.  In particular, since
+a diagonal move by a pawn is legal only the case where such a move would
+capture an opponent's piece, it could potentially be profitable to attempt all
+or many pawn captures on every turn.  In order to speed the game up, the
+referee declares, at the beginning of a player's turn, all locations where a
+pawn capture is possible.
+
+In general, any move that is legal in a regular chess match may be at least
+attempted by a player in Kriegspiel.  And any Kriegspiel move that is
+ultimately allowed by the referee would also be a legal move in chess.
 
 \SubSection{Example}
-%\includegraphics[width=0.2\textwidth]{images/1W.png}
-%\begin{minipage}{\textwidth}
-%\centering
-%\subfloat[1. d4]{\label{fig:1W}\includegraphics[width=0.2\textwidth]{images/1W.png}}
-%\qquad
-%\subfloat[1. ... a5]{\label{fig:1B}\includegraphics[width=0.2\textwidth]{images/1B.png}}
-%\qquad
-%\subfloat[2. Bg5]{\label{fig:2W}\includegraphics[width=0.2\textwidth]{images/2W.png}}
-%\qquad
-%\subfloat[2. ... b6]{\label{fig:2B}\includegraphics[width=0.2\textwidth]{images/2B.png}}
-%\end{minipage}
-%\begin{minipage}{\textwidth}
-%\centering
-%\subfloat[3. Nc3]{\label{fig:3W}\includegraphics[width=0.2\textwidth]{images/3W.png}}
-%\qquad
-%\subfloat[3. ... c5]{\label{fig:3B}\includegraphics[width=0.2\textwidth]{images/3B.png}}
-%\qquad
-%\subfloat[4. d5]{\label{fig:4W}\includegraphics[width=0.2\textwidth]{images/4W.png}}
-%\qquad
-%\subfloat[4. ... Na6]{\label{fig:4B}\includegraphics[width=0.2\textwidth]{images/4B.png}}
-%\end{minipage}
-%\begin{minipage}{\textwidth}
-%\centering
-%\subfloat[5. d6!]{\label{fig:5W}\includegraphics[width=0.2\textwidth]{images/5W.png}}
-%\qquad
-%\subfloat[5. ... f6!]{\label{fig:5B}\includegraphics[width=0.2\textwidth]{images/5B.png}}
-%\qquad
-%\subfloat[6. e4]{\label{fig:6W}\includegraphics[width=0.2\textwidth]{images/6W.png}}
-%\qquad
-%\subfloat[6. ... fg]{\label{fig:6B}\includegraphics[width=0.2\textwidth]{images/6B.png}}
-%\end{minipage}
-%\begin{minipage}{\textwidth}
-%\centering
-%\subfloat[7. Bc4]{\label{fig:7W}\includegraphics[width=0.2\textwidth]{images/7W.png}}
-%\qquad
-%\subfloat[7. ... ed]{\label{fig:7B}\includegraphics[width=0.2\textwidth]{images/7B.png}}
-%\qquad
-%\subfloat[8 Qd5!]{\label{fig:8W}\includegraphics[width=0.2\textwidth]{images/8W.png}}
-%\qquad
-%\subfloat[8 ... Qc7?]{\label{fig:8B}\includegraphics[width=0.2\textwidth]{images/8B.png}}
-%\end{minipage}
-%\begin{minipage}{\textwidth}
-%\centering
-%\subfloat[9 Qf7+]{\label{fig:9W}\includegraphics[width=0.2\textwidth]{images/9W.png}}
-%\qquad
-%\subfloat[9 ... d8]{\label{fig:9B}\includegraphics[width=0.2\textwidth]{images/9B.png}}
-%\qquad
-%\subfloat[10 Qxf8++]{\label{fig:10W}\includegraphics[width=0.2\textwidth]{images/8W.png}}
-%\end{minipage}
 \begin{figure}[ht]
@@
 
-The concept of information sets is perhaps best explained through an example.  (See also Appendix A.)
-Figures~\ref{listing} and~\ref{sequence} show a full listing and the associated pictorial progression of an actual
-kriegspiel game from ~\cite{li94chess}.  We have adapted the notation used by Wolfe for Berkeley Kriegspiel.  The moves
-are numbered in pairs: a move by white on one line followed by black's move on the next line.  The actual move is shown
+\subsection{Information Set Generation}
+The concept of information sets is perhaps best explained through an example. 
+Figure~\ref{listing} and show a full listing and the associated pictorial progression of an actual
+kriegspiel game~\cite{li94chess}.  We have adapted the notation used by Wolfe for Berkeley Kriegspiel.  Odd (resp. even) moves are for white (resp. black).  The actual move is shown
 first followed by a list of referee announcements.  Captures announcements are prefixed with an X and give the location
@@
 \begin{figure}
+\begin{minipage}[b]{0.35\linewidth}
+\centering
+\small
 \begin{verbatim}
@@
    ?? {(:0,Td6)}
+\end{verbatim}
+\end{minipage}
+\begin{minipage}[b]{0.25\linewidth}
+\small
+\centering
+\begin{verbatim}
 6. e4 {(:)}
@@
 \end{verbatim}
+\end{minipage}
 \caption{Listing of a kriegspiel game resulting in checkmate by white after 10 moves.}
@@
 that white must have a bishop at g5 and a pawn at d6.  There are 25 possible sequences of moves that are consistent with
-black's observations up to that point. Table~\ref{bothtimes} shows the size of the information sets for each player for
-each move in the game and the amount of time to find the full set on a single processor.  Appendix B shows the full
-output of the program for the same point in the game (five moves for each player) and illustrates all of the possible
-positions in the belief state.
+black's observations up to that point.
+
+%Table~\ref{bothtimes} shows the size of the information sets for each player for
+%each move in the game and the amount of time to find the full set on a single processor.  Appendix B shows the full
+%output of the program for the same point in the game (five moves for each player) and illustrates all of the possible
+%positions in the belief state.
 
@@
 
-\begin{table}
-\centering
-\begin{tabular}{rrrrr}
- & \multicolumn{2}{c}{\bf White} & \multicolumn{2}{c}{\bf Black} \\
-{\bf Ply} & {\bf Size} & {\bf Time} & {\bf Size} & {\bf Time} \\
-1 & 18 & 0.000 & 20 & .000 \\
-2 & 18 & 0.000 & 19 & .000 \\
-3 & 18 & 0.000 & 404 & .008\\
-4 & 279 & 0.008 & 401 & .024\\
-5 & 242 & 0.028 & 1472 & .044\\
-6 & 216 & 0.052 & 155 & .096\\
-7 & 176 & 0.064 & 293 & .116\\
-8 & 3406 & 0.092 & 158 & .128\\
-9 & 2191 & 0.288 & 118 & .148\\
-10 & 1423 & 0.508 & 25 & .152\\
-11 & 1363 & 0.596 & 798 & .164\\
-12 & 1416 & 0.744 & 518 & .192\\
-13 & 1416 & 0.836 & 13564 & .304\\
-14 & 3440 & 0.97 & 12394 & .776\\
-15 & 3428 & 1.25 & 343652 & 3.75\\
-16 & 50521 & 1.82 & 320704 & 17.3\\
-17 & 43192 & 5.91 & 490162 & 99.9\\
-18 & 26128 & 7.94 & 3792 & 119\\
-19 & 19061 & 9.92 & 14836 & 121\\
-\end{tabular}
-\caption{Solution counts and running times for a sample kriegspiel game}
-\label{bothtimes}
-\end{table}
+%\begin{table}
+%\centering
+%\begin{tabular}{rrrrr}
+% & \multicolumn{2}{c}{\bf White} & \multicolumn{2}{c}{\bf Black} \\
+%{\bf Ply} & {\bf Size} & {\bf Time} & {\bf Size} & {\bf Time} \\
+%1 & 18 & 0.000 & 20 & .000 \\
+%2 & 18 & 0.000 & 19 & .000 \\
+%3 & 18 & 0.000 & 404 & .008\\
+%4 & 279 & 0.008 & 401 & .024\\
+%5 & 242 & 0.028 & 1472 & .044\\
+%6 & 216 & 0.052 & 155 & .096\\
+%7 & 176 & 0.064 & 293 & .116\\
+%8 & 3406 & 0.092 & 158 & .128\\
+%9 & 2191 & 0.288 & 118 & .148\\
+%10 & 1423 & 0.508 & 25 & .152\\
+%11 & 1363 & 0.596 & 798 & .164\\
+%12 & 1416 & 0.744 & 518 & .192\\
+%13 & 1416 & 0.836 & 13564 & .304\\
+%14 & 3440 & 0.97 & 12394 & .776\\
+%15 & 3428 & 1.25 & 343652 & 3.75\\
+%16 & 50521 & 1.82 & 320704 & 17.3\\
+%17 & 43192 & 5.91 & 490162 & 99.9\\
+%18 & 26128 & 7.94 & 3792 & 119\\
+%19 & 19061 & 9.92 & 14836 & 121\\
+%\end{tabular}
+%\caption{Solution counts and running times for a sample kriegspiel game}
+%\label{bothtimes}
+%\end{table}
 
@@
 
-\Section{The Information Set Generation Algorithm}
-We have formulated information set generation as a combinatorial search problem.  The input to the algorithm is a list
-of observations for one player in a kriegspiel game.  We assume that the agent has perfect recall (i.e., remembers all
-of its own decisions), which is a reasonable assumption for kriegspiel.  (See~\cite{gilpin07algorithms} for a further
-discussion of perfect recall.)
+{The Information Set Generation Algorithm}
+ISG can be formulated as a combinatorial search problem.  The input to the algorithm is a list
+of observations for one player in a kriegspiel game.
 
@@
 There is a search space variable for each ply in the game, and the values that a variable can take on are all possible
-legal moves.  A kriegspiel game may last for on the order of tens of plies.  Moves are parameterized by the piece that
-is relocating (both black and white have six different types of pieces: king, queen, rook, bishop, knight, pawn), the
-source square, and the destination square.  Only a few hundred of the $64 \times 64 \times 6$ combinations represent
-moves that are legitimate in at least some scenarios.  For example, no piece could ever move from a3 to h2\footnote{We
-observe the convention that on an $8 \times 8$ chessboard the rows are numbered 1 to 8 and the columns are labeled a to
-h.} in one move.
-
-Information set generation, then, is an all-solutions search through the space.  Whenever a node in the search tree is
-found to be inconsistent with the actual observation given in the input, the subtree rooted at that node is pruned.
-There are types of observations that must match in order for the state to be found: check status, pawn tries, and 
-illegal moves available/attempted.
+legal moves.  Information set generation, then, is an all-solutions search through the space.  Whenever a node in the search tree is
+found to be inconsistent with the actual observation given in the input, the
+subtree rooted at that node is pruned (i.e., search is discontinued along that
+path).  There are types of observations that must match in order for the state
+to be found: check status, pawn tries, and illegal moves available/attempted.
 
@@
 Referee announcements regarding pawn tries or a player being in check are made prior to a move.  The algorithm checks
-for the consistency of these observations in lines $5--6$.  If the search reaches the appropriate depth and passes these
-tests, then a solution has been found and is reported (line $7$).
-
-Otherwise, we consider the possible moves that can be made from that state.  If the player whose turn it is to move at
-the current level of the search tree is the same player from whose perspective we have received observations, then the
-next step is to make sure that every legal move that was attempted at the player at that ply is also at least
-attemptable at the game position that corresponds to the game tree node (lines $15--22$).  We know exactly the legal
-move that was ultimately made at this depth (because we are assuming perfect recall), and we check to make sure that
-this move is indeed legal (pruning otherwise).  If so, we apply that move and recurse (lines $24--32$).  This
-corresponds to a ``forced move'' in the general search paradigm.
-
-On the other hand, if the player whose turn it is to move at the current depth of the search tree is the opponent, then
-must consider all possible legal moves to branch on.  We first check to make sure that the number of legal moves from
-this position is at least as large as the number of illegal moves that the opponent attempted (i.e., the number of times
-the referee said ``No'' before finally accepting the opponent's move).  This is shown in lines $39--42$.  At this point
-we generate a child for each move that would be accepted by the referee and recurse (lines $44--52$).
-
-One of our goals was to assess the varying utilities of the different pruning criteria.  We tried some different
-orderings of the tests and were somewhat surprised (and disappointed) to see that there was actually a fairly minimal
-impact on performance.  This could be due to the fact that because of the tedious nature of encoding problem instances
-from actual games, we were not able to run experiments of a wide variety of types of positions (i.e., different styles
-of play).
+for the consistency of these observations at line $7$.  If the search reaches the appropriate depth and passes these
+tests, then a solution has been found and is added to the information set (line $3$).
+
+%Otherwise, we consider the possible moves that can be made from that state.  If the player whose turn it is to move at
+%the current level of the search tree is the same player from whose perspective we have received observations, then the
+%next step is to make sure that every legal move that was attempted at the player at that ply is also at least
+%attemptable at the game position that corresponds to the game tree node (lines $15--22$).  We know exactly the legal
+%move that was ultimately made at this depth (because we are assuming perfect recall), and we check to make sure that
+%this move is indeed legal (pruning otherwise).  If so, we apply that move and recurse (lines $24--32$).  This
+%corresponds to a ``forced move'' in the general search paradigm.
+%
+%On the other hand, if the player whose turn it is to move at the current depth of the search tree is the opponent, then
+%must consider all possible legal moves to branch on.  We first check to make sure that the number of legal moves from
+%this position is at least as large as the number of illegal moves that the opponent attempted (i.e., the number of times
+%the referee said ``No'' before finally accepting the opponent's move).  This is shown in lines $39--42$.  At this point
+%we generate a child for each move that would be accepted by the referee and recurse (lines $44--52$).
+%
+%One of our goals was to assess the varying utilities of the different pruning criteria.  We tried some different
+%orderings of the tests and were somewhat surprised (and disappointed) to see that there was actually a fairly minimal
+%impact on performance.  This could be due to the fact that because of the tedious nature of encoding problem instances
+%from actual games, we were not able to run experiments of a wide variety of types of positions (i.e., different styles
+%of play).
 
@@
 
-\Section{Charm++}
+\Section{Charm++ and ParSSSE}\label{ParSSSE}
 We built our search engine on top of the {\sc Charm++} run-time
@@
 
-\Section{The {\sc Charm++} Parallel State Space Search Engine}
+%\subsection{The {\sc Charm++} Parallel State Space Search Engine}
 Sun \etal have developed the Parallel State Space Search Engine ({\sc ParSSSE})
@@
 
-\Section{Performance Results}
-We built our search on top of the same general search engine that we used in class.  The performance results reported in
-this section were obtained by running our program on Bleuprint with up to 1024 processors.
-
-Table~\ref{speedups} shows the performance results for our algorithm on a randomly selected problem instance in which
-the size of the information set was about 4.5 million.  (This instance is not necessarily representative of all games or
-states of the game, but it was not cherry-picked in any way.)  The table shows that the problem is indeed highly
-parallizable.  The speedups are nearly linear for up  to 32 processors, and even on 1024 processors, we get a good
-speedup.  In this particular instance, the sequential algorithm takes almost 25 minutes to run.  But when parallelized
-on 1024 processors---perfectly reasonable resource usage on modern hardware---the search takes less than three seconds.
-This is a speedup of 576 times!  It might not be reasonable to use the algorithm at all if it takes 25 or more to run
-for some turns.  But if the cost is only a few seconds, it becomes entirely feasible.
-
-\begin{table}
-\centering
-\begin{tabular}{rrr}
-{\bf Procs}	&	{\bf Time (s)} 	&	{\bf Speedup}\\
-1	&	1524	&	1\\
-2	&	812	&	1.88\\
-4	&	407	&	3.74\\
-8	&	208	&	7.32\\
-16	&	102	&	14.8\\
-32	&	53.2	&	28.7\\
-64	&	28.0	&	54.5\\
-128	&	14.9	&	102\\
-256	&	7.70	&	197\\
-512	&	4.59	&	331\\
-1024	&	2.64	&	576\\
-\end{tabular}
-\caption{Speedups for a randomly selected problem instance on 1--1024 processors.  The input observations were from the
-10th ply of the game tree and the size of the information set in this instance is 4,495,121.}
-\label{speedups}
-\end{table}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.7\columnwidth]{plots/3schemes.pdf}
-\caption{Experiments using different load balancing strategies.}
-\label{3schemes}
-\end{figure}
-
-In our next round of experiments, we set out to compare a few different load-balancing strategies for the problem.  We
-used the random, work stealing, and neighborhood balancing strategies that are available in the general purpose search
-engine.  These results are shown in Figure~\ref{3schemes}.  Our results show that there was actually very little
-variation in performance across the different strategies.  We did notice that across all of our experiments, the work
-stealing approach appeared to be consistently---if only marginally---better than the others.  Our hypothesis here is
-that perhaps the nodes are rarely starved for work.  In other words, perhaps the very nature of the problem means that
-the load is naturally well-balanced and that the load-balancer rarely needs to do a significant amount of
-redistribution.  This is probably also related to the fact that we are doing an all-solutions search.
-
-\begin{figure}[h]
-\centering
-\begin{minipage}{0.5\linewidth}
-\includegraphics[width=\columnwidth]{plots/3depths.pdf}
-\end{minipage}
-\centering
-\begin{minipage}{0.5\linewidth}
-\includegraphics[width=\columnwidth]{plots/3depthsnolog.pdf}
-\end{minipage}
-\caption{(Left) Log-log plot of running time for three different sizes of problem as the number of cores used increases.
-(Right) Same data with a linear y-axis for the running time (to help see the isoefficiency.)}
-\label{3depths}
-\end{figure}
-
-We next attempted to measure the isoefficiency of the problem---how is efficiency of the parallelization of the
-algorithm affected by {\em size} of the problem?  To test this we varied the number of plies in the test instance from
-8 to 12. We used the work stealing load balancing strategy.  The results are shown in Table~\ref{3depths}.  The data are
-the same in both tables, but the one on the left uses a log-log scale.  In the easiest case (depth 8), there isn't
-enough additional parallel work to justify the jump from 256 to 1024 processors, and performance actually degrades.
-Other than that, however, we see consistent speedups as we increase the number of processors.  The plot on the right helps to
-illustrate the good isoefficiency of the problem.  As the size of the problem increases, the amount of parallelizable
-work available also increases, and therefore the steepness of the speed curve is higher for larger problem sizes, as the
-number of processors increases.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.6\columnwidth]{plots/dep8.pdf}
-\caption{Grainsize control data for depth 8 problems.}
-\label{dep8}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.6\columnwidth]{plots/dep10.pdf}
-\caption{Grainsize control data for depth 10 problems.}
-\label{dep10}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.6\columnwidth]{plots/dep12.pdf}
-\caption{Grainsize control data for depth 12 problems.}
-\label{dep12}
-\end{figure}
-
-Finally, we ran some experiments to measure the effect of grainsize on the problem.  The results are shown in
-Tables~\ref{dep8}--~\ref{dep12}.  It is clear that there is a ``sweet spot'' for each problem size.  (At least it is
-clear for this particular instance.  It is possible that there would be different behavior on different problem
-instances.)  When the grainsize is too small, the overall computation is hampered by excessive parallel overhead
-(communication, queueing of nodes to expand, load balancing operations).  On the other hand, when the grain size is too
-large, it becomes too difficult to ensure that all processors are busy and processors are not idling with no work to do
-while a small number of processors are busy doing large-sized chunks of work.
-
-In our experiments, the optimal grainsize is clearly related to the size of the problem, the depth of the search tree.
-The optimal sequential threshold seems to be $2--4$ less than the problem depth.  Since the depth of all solutions is
-the same for each problem instance, this means that the number of levels that should be searched sequentially is fairly
-consistent across problem sizes.  We noted that the number of solutions found---in other words, the size of the
-information sets---is typically fairly large with respect to the total number of nodes expanded.  (We don't have hard
-figures, but we estimate that it was often $30-40\%$).  This high density of solutions limits the variability that can
-be present in the amount of work to be done at the nodes representing the last few levels of the tree.  This means that
-the best grainsize is also likely to be fairly predictable.
-
-\Section{New Results}
-In addition to the random instances, we also evaluated our information set generation routine on 40 positions from actual kriespiel games, including human vs.
+%%We built our search on top of the same general search engine that we used in class.  The performance results reported in
+%%this section were obtained by running our program on Bleuprint with up to 1024 processors.
+%%
+%%Table~\ref{speedups} shows the performance results for our algorithm on a randomly selected problem instance in which
+%%the size of the information set was about 4.5 million.  (This instance is not necessarily representative of all games or
+%%states of the game, but it was not cherry-picked in any way.)  The table shows that the problem is indeed highly
+%%parallizable.  The speedups are nearly linear for up  to 32 processors, and even on 1024 processors, we get a good
+%%speedup.  In this particular instance, the sequential algorithm takes almost 25 minutes to run.  But when parallelized
+%%on 1024 processors---perfectly reasonable resource usage on modern hardware---the search takes less than three seconds.
+%%This is a speedup of 576 times!  It might not be reasonable to use the algorithm at all if it takes 25 or more to run
+%%for some turns.  But if the cost is only a few seconds, it becomes entirely feasible.
+%%
+%%\begin{table}
+%%\centering
+%%\begin{tabular}{rrr}
+%%{\bf Procs}	&	{\bf Time (s)} 	&	{\bf Speedup}\\
+%%1	&	1524	&	1\\
+%%2	&	812	&	1.88\\
+%%4	&	407	&	3.74\\
+%%8	&	208	&	7.32\\
+%%16	&	102	&	14.8\\
+%%32	&	53.2	&	28.7\\
+%%64	&	28.0	&	54.5\\
+%%128	&	14.9	&	102\\
+%%256	&	7.70	&	197\\
+%%512	&	4.59	&	331\\
+%%1024	&	2.64	&	576\\
+%%\end{tabular}
+%%\caption{Speedups for a randomly selected problem instance on 1--1024 processors.  The input observations were from the
+%%10th ply of the game tree and the size of the information set in this instance is 4,495,121.}
+%%\label{speedups}
+%%\end{table}
+%%
+%%\begin{figure}[h]
+%%\centering
+%%\includegraphics[width=0.7\columnwidth]{plots/3schemes.pdf}
+%%\caption{Experiments using different load balancing strategies.}
+%%\label{3schemes}
+%%\end{figure}
+%%
+%%In our next round of experiments, we set out to compare a few different load-balancing strategies for the problem.  We
+%%used the random, work stealing, and neighborhood balancing strategies that are available in the general purpose search
+%%engine.  These results are shown in Figure~\ref{3schemes}.  Our results show that there was actually very little
+%%variation in performance across the different strategies.  We did notice that across all of our experiments, the work
+%%stealing approach appeared to be consistently---if only marginally---better than the others.  Our hypothesis here is
+%%that perhaps the nodes are rarely starved for work.  In other words, perhaps the very nature of the problem means that
+%%the load is naturally well-balanced and that the load-balancer rarely needs to do a significant amount of
+%%redistribution.  This is probably also related to the fact that we are doing an all-solutions search. 
+%%
+%% 
+%%\begin{figure}[h]
+%%\centering
+%%\begin{minipage}{0.5\linewidth}
+%%\includegraphics[width=\columnwidth]{plots/3depths.pdf}
+%%\end{minipage}
+%%\centering
+%%\begin{minipage}{0.5\linewidth}
+%%\includegraphics[width=\columnwidth]{plots/3depthsnolog.pdf}
+%%\end{minipage}
+%%\caption{(Left) Log-log plot of running time for three different sizes of problem as the number of cores used increases.
+%%(Right) Same data with a linear y-axis for the running time (to help see the isoefficiency.)}
+%%\label{3depths}
+%%\end{figure}
+%%
+%%We next attempted to measure the isoefficiency of the problem---how is efficiency of the parallelization of the
+%%algorithm affected by {\em size} of the problem?  To test this we varied the number of plies in the test instance from
+%%8 to 12. We used the work stealing load balancing strategy.  The results are shown in Table~\ref{3depths}.  The data are
+%%the same in both tables, but the one on the left uses a log-log scale.  In the easiest case (depth 8), there isn't
+%%enough additional parallel work to justify the jump from 256 to 1024 processors, and performance actually degrades.
+%%Other than that, however, we see consistent speedups as we increase the number of processors.  The plot on the right helps to
+%%illustrate the good isoefficiency of the problem.  As the size of the problem increases, the amount of parallelizable
+%%work available also increases, and therefore the steepness of the speed curve is higher for larger problem sizes, as the
+%%number of processors increases.
+%% 
+%%\begin{figure}[h]
+%%\centering
+%%\includegraphics[width=0.6\columnwidth]{plots/dep8.pdf}
+%%\caption{Grainsize control data for depth 8 problems.}
+%%\label{dep8}
+%%\end{figure}
+%% 
+%%\begin{figure}[h]
+%%\centering
+%%\includegraphics[width=0.6\columnwidth]{plots/dep10.pdf}
+%%\caption{Grainsize control data for depth 10 problems.}
+%%\label{dep10}
+%%\end{figure}
+%% 
+%%\begin{figure}[h]
+%%\centering
+%%\includegraphics[width=0.6\columnwidth]{plots/dep12.pdf}
+%%\caption{Grainsize control data for depth 12 problems.}
+%%\label{dep12}
+%%\end{figure}
+%%
+%%Finally, we ran some experiments to measure the effect of grainsize on the problem.  The results are shown in
+%%Tables~\ref{dep8}--~\ref{dep12}.  It is clear that there is a ``sweet spot'' for each problem size.  (At least it is
+%%clear for this particular instance.  It is possible that there would be different behavior on different problem
+%%instances.)  When the grainsize is too small, the overall computation is hampered by excessive parallel overhead
+%%(communication, queueing of nodes to expand, load balancing operations).  On the other hand, when the grain size is too
+%%large, it becomes too difficult to ensure that all processors are busy and processors are not idling with no work to do
+%%while a small number of processors are busy doing large-sized chunks of work.  
+%%
+%%In our experiments, the optimal grainsize is clearly related to the size of the problem, the depth of the search tree.
+%%The optimal sequential threshold seems to be $2--4$ less than the problem depth.  Since the depth of all solutions is
+%%the same for each problem instance, this means that the number of levels that should be searched sequentially is fairly
+%%consistent across problem sizes.  We noted that the number of solutions found---in other words, the size of the
+%%information sets---is typically fairly large with respect to the total number of nodes expanded.  (We don't have hard
+%%figures, but we estimate that it was often $30-40\%$).  This high density of solutions limits the variability that can
+%%be present in the amount of work to be done at the nodes representing the last few levels of the tree.  This means that
+%%the best grainsize is also likely to be fairly predictable.
+
+\Section{Performance Results}\label{Results}
+We also evaluated our information set generation routine on 40 positions from actual kriespiel games, including human vs.
 human and computer vs.
@@
 in~\cite{kuhn03lectures} and~\cite{kuhn97classics}.  A technical explanation of the concept of information sets can be
-found in~\cite{gilpin07algorithms}.  Algorithms for ``solving'' imperfect information games---in the sense of finding a
-Nash equilibrium---can be found in~\cite{koller94fast} and~\cite{koller97representations}.  Koller \etal, showed that an
-equilibrium can be found in time that is polynomial in the number of nodes in the game tree.  And in fact, they found
+found in~\cite{gilpin07algorithms}.  Koller \etal, showed that a Nash
+equilibrium can be found in time that is polynomial in the number of nodes in the game tree~\cite{koller94fast}.  In fact, they found
 that it took about as much time to generate the full game tree as it did to actually solve it.  Unfortunately, this is
@@
 they sample from possible positions for the current state (i.e., sampling from the belief state) by matching only the
-most recent observations.  This motivation for this approach seems to be that it would not be feasible to sample from
+most recent observations.  The motivation for this approach seems to be that it would not be feasible to sample from
 the full information set, because of the computational expense involved.  And indeed, our experiments bear this out-- it
@@
 search engine on a parallel machine (1024 processors), we were able to generate the full set in just 2.9 seconds.  (Note
-that this set had more than 4 million members.)  Our results suggest that the problem is highly scalable and our
-approach might therefore be preferred.
+that this set had more than 4 million members.)
 
@@
 
-In our experiments, we have found positions in which the size of the information set is many tens of millions.  Our goal
-is to develop strategies that are amenable to use in this more general setting.   In other words, rather than being able
-to analyze positions only at the beginning or end of the game (when the size of the information set tends to be
-smaller), we want to be able to develop a strategy that can be used in the middle of the game as well.  game).
+%In our experiments, we have found positions in which the size of the information set is many tens of millions.  Our goal
+%is to develop strategies that are amenable to use in this more general setting.   In other words, rather than being able
+%to analyze positions only at the beginning or end of the game (when the size of the information set tends to be
+%smaller), we want to be able to develop a strategy that can be used in the middle of the game as well.  game).
 
@@
 
-\Section{Future Work}
-
-First and foremost, it would be interesting to see how our information set generation algorithm would perform as a
-subroutine to an actual kriegspiel playing program.  The utility of generating an information set is ultimately
-determined by how well that information improves a player's decision-making capabilities.  We noted earlier that a
-common use of information sets is to estimate the value of a move by averaging the estimated value of the associated
-descendant for each node in the current information set (or a random sample from the set).  This kind of strategy can be
-applied naively using a belief state sampling algorithm (i.e., an algorithm that samples from the distribution of
-possible worlds in the current state without respect to the overall history of moves).  However, it has been shown that
-significant improvement in play can be achieved by estimating the value of future moves from the perspective of the
-opponent, based on the knowledge that the opponent would have had at prior positions in the move
-history~\cite{richards07opponent}.  Such a strategy would require the use of an information set generation algorithm of the
-kind that we have described here.\footnote{This work is underway.}
-
-Ideally, we would run our algorithm on some deeper instances from real kriegspiel games.  We are not aware of a
-repository of such games.  Unfortunately, notation for kriegspiel is not standardized and there are several minor
-variations in rules (i.e., with respect to the nature of referee announcements) that are unlikely to significantly to
-impact the computability and scalability results but which nevertheless make it difficult to write an input reader that
-converts the transcript of a game into a sequence of moves readable by our algorithm.  An alternative would be to play
-some games ourselves.  Currently, the manual encoding of games is not difficult but is tedious.  Still, it would be nice
-to have some games that go on for 40 or 50 moves and have positions where the information sets are of moderate size (on
-the order of a few hundred thousand to a few million).
-
-The information set generation algorithm that we have implemented here is the most straightforward one and is alluded to
-in~\cite{parker05game} and~\cite{russell05efficient}.  It is analogous to the ``expand-at-the-tip'' strategy for the
-Hamiltonian Circuit problem.  An alternative information set generation algorithm is described
-in~\cite{richards09information} (but not implemented in parallel) and could potentially improve the search efficiency
-greatly by exploiting variable ordering.  This would be analogous to the disconnected edge-pairs\footnote{This is what we
-referred to as ``algorithm 2'' in class.} heuristic for the Hamiltonian Circuit problem.  Rather than search from the
-root node going forward, this algorithm would seek opportunities for pruning based on the propagation of logical
-consequences from each move both forward and backward in time.  This, incidentally, is more consistent with the way
-human players would analyze the game.  For example, if a player's proposed bishop move is rejected and the bishop is
-sufficiently removed from its own king, then the player may infer that one of his opponent's pieces is along that
-rejected path.  At least one of those spaces must have been the destination of a prior move by the opponent.  If a
-recent move by that bishop crossed the same path, then that would be hard evidence that those squares were empty at the
-time.  And therefore, the arrival of the opponent's piece must have been between the previous (successful) bishop move
-and the current (unsuccessful) bishop move.  To implement this algorithm, we envision the use of planning graph type
-data structures~\cite{blum97fast}.  Planning graphs have a sequence of alternating levels of state constraints and
-action constraints.  State constraints would be of the form $at(bishop,d4) \vee at(bishop,e6)$ or $!occupied(e4)$.
-Action constraints would be of the form $move(knight,a4,c5) \vee move(knight,e6)$.  At a minimum, each level would
-require a data structure with a bit for each possible action/state.  Propagating the constraints would require much of
-the same machinery used in STRIPS-like planning algorithms~\cite{chen05solving}, and one of the many knobs that would
-need to be tuned would be the depth of the propagation for each constraint.  (Should the consequences of a known action
-or non-action be propagated one move into the future and one move into the past?  More?  Less?)  Additionally, there
-would be a significant and interesting tradeoff in the effectiveness of pruning and the size of the data
-structures that would have to be encoded in each chare.  The best algorithm might be a hybrid strategy that does forward
-propagation from the root at the parallel level and constraint propagation at the sequential level.\footnote{This work
-is also in progress.}
-
-There are certainly several places where the efficiency of the details of each node expansion could be improved.  Many
-of these would reduce the overall running time but would not necessarily be interesting from the standpoint of
-evaluating the efficiency and isoefficiency of parallelization.  For example, in checking to see if a position leaves a
-player in check, we currently utilize existing helper functions that generate possible moves for the opponent.  It would
-be possible to write a more specialized function that focuses only on the position of the king in question and looks
-only at the squares along its rank, file, diagonals, and knight-edges for threats.  There are also some memory
-allocation issues that could be improved.
+\Section{Conclusion Future Work}
+We have shown that information set generation for kriegspiel can be efficiently parallelized in a way that scales up to at least 1024 processors.
+
+A key line of future research is to use our ISG routine as part of a full Kriegspiel play.
+The ISG method itself could be updated using constraint-satisfaction techniques outlined in ~\cite{richards12information}, which improve pruning performance in the search space by incorporating information from all observations in the list at each search step, instead of performing depth-first search.
+The use of ISG-based opponent modeling to improve overall decision making is described in ~\cite{richards12reasoning}.
+
+%determined by how well that information improves a player's decision-making capabilities.  We noted earlier that a
+%common use of information sets is to estimate the value of a move by averaging the estimated value of the associated
+%descendant for each node in the current information set (or a random sample from the set).  This kind of strategy can be
+%applied naively using a belief state sampling algorithm (i.e., an algorithm that samples from the distribution of
+%possible worlds in the current state without respect to the overall history of moves).  However, it has been shown that
+%significant improvement in play can be achieved by estimating the value of future moves from the perspective of the
+%opponent, based on the knowledge that the opponent would have had at prior positions in the move
+%history~\cite{richards07opponent}.  Such a strategy would require the use of an information set generation algorithm of the
+%kind that we have described here.\footnote{This work is underway.}
+
+%Ideally, we would run our algorithm on some deeper instances from real kriegspiel games.  We are not aware of a
+%repository of such games.  Unfortunately, notation for kriegspiel is not standardized and there are several minor
+%variations in rules (i.e., with respect to the nature of referee announcements) that are unlikely to significantly to
+%impact the computability and scalability results but which nevertheless make it difficult to write an input reader that
+%converts the transcript of a game into a sequence of moves readable by our algorithm.  An alternative would be to play
+%some games ourselves.  Currently, the manual encoding of games is not difficult but is tedious.  Still, it would be nice
+%to have some games that go on for 40 or 50 moves and have positions where the information sets are of moderate size (on
+%the order of a few hundred thousand to a few million).
+
+%The information set generation algorithm that we have implemented here is the most straightforward one and is alluded to
+%in~\cite{parker05game} and~\cite{russell05efficient}.  It is analogous to the ``expand-at-the-tip'' strategy for the
+%Hamiltonian Circuit problem.  An alternative information set generation algorithm is described
+%in~\cite{richards09information} (but not implemented in parallel) and could potentially improve the search efficiency
+%greatly by exploiting variable ordering.  This would be analogous to the disconnected edge-pairs\footnote{This is what we
+%referred to as ``algorithm 2'' in class.} heuristic for the Hamiltonian Circuit problem.  Rather than search from the
+%root node going forward, this algorithm would seek opportunities for pruning based on the propagation of logical
+%consequences from each move both forward and backward in time.  This, incidentally, is more consistent with the way
+%human players would analyze the game.  For example, if a player's proposed bishop move is rejected and the bishop is
+%sufficiently removed from its own king, then the player may infer that one of his opponent's pieces is along that
+%rejected path.  At least one of those spaces must have been the destination of a prior move by the opponent.  If a
+%recent move by that bishop crossed the same path, then that would be hard evidence that those squares were empty at the
+%time.  And therefore, the arrival of the opponent's piece must have been between the previous (successful) bishop move
+%and the current (unsuccessful) bishop move.  To implement this algorithm, we envision the use of planning graph type
+%data structures~\cite{blum97fast}.  Planning graphs have a sequence of alternating levels of state constraints and
+%action constraints.  State constraints would be of the form $at(bishop,d4) \vee at(bishop,e6)$ or $!occupied(e4)$.
+%Action constraints would be of the form $move(knight,a4,c5) \vee move(knight,e6)$.  At a minimum, each level would
+%require a data structure with a bit for each possible action/state.  Propagating the constraints would require much of
+%the same machinery used in STRIPS-like planning algorithms~\cite{chen05solving}, and one of the many knobs that would
+%need to be tuned would be the depth of the propagation for each constraint.  (Should the consequences of a known action
+%or non-action be propagated one move into the future and one move into the past?  More?  Less?)  Additionally, there
+%would be a significant and interesting tradeoff in the effectiveness of pruning and the size of the data
+%structures that would have to be encoded in each chare.  The best algorithm might be a hybrid strategy that does forward
+%propagation from the root at the parallel level and constraint propagation at the sequential level.\footnote{This work
+%is also in progress.}
+%
+%There are certainly several places where the efficiency of the details of each node expansion could be improved.  Many
+%of these would reduce the overall running time but would not necessarily be interesting from the standpoint of
+%evaluating the efficiency and isoefficiency of parallelization.  For example, in checking to see if a position leaves a
+%player in check, we currently utilize existing helper functions that generate possible moves for the opponent.  It would
+%be possible to write a more specialized function that focuses only on the position of the king in question and looks
+%only at the squares along its rank, file, diagonals, and knight-edges for threats.  There are also some memory
+%allocation issues that could be improved.
 

### 2012-06-27 sbac/paper.tex
Add Sanjay as author
remoived dumplicatecoinclusion.. some reorganiuzation

@@
 
-\author{Mark Richards \and Abhishek Gupta \and Osman Sarood \\
-%Department of Computer Science\\ 
-%University of Illinois\\ 
-%Urbana, IL\\ mdrichar@illinois.edu\\
+\author{Mark Richards, Abhishek Gupta, Osman Sarood and Laxmikant Kale\\
+Department of Computer Science\\ 
+University of Illinois\\ 
+Urbana, IL\\ {mdrichar, gupta59, sarood1, kale}@illinois.edu\\
 % For a paper whose authors are all at the same institution, 
@@
 
-\Section{Background}\label{Background}
-\subsection{Kriegspiel}
+%\Section{Information Set Generation}
+\Section{Background: Kriegspiel}
+Information Set Generation is a widely adopted operation used in game trees search.
+Here, we present our techniques and results in the context of a particular game - Kriegspiel.
+However, the techniques and framework presented in this paper is general and can be applied to other imperfect games for generating information set.
+%\Section{Background: Kriegspiel}
 Kriegspiel, or partially observable chess,  was invented by Henry Michael Temple in 1899.  Historically, the
@@
 
-\SubSection{Example}
+%\SubSection{Example}
 \begin{figure}[ht]
@@
 
-\subsection{Information Set Generation}
+\Section{Information Set Generation}
+%\subsection{Information Set Generation}
 The concept of information sets is perhaps best explained through an example. 
@@
 
-{The Information Set Generation Algorithm}
-ISG can be formulated as a combinatorial search problem.  The input to the algorithm is a list
+\textbf{The Information Set Generation Algorithm}
+(ISG) can be formulated as a combinatorial search problem.  The input to the algorithm is a list
 of observations for one player in a kriegspiel game.
@@
 \end{boxedminipage}
+\caption{Algorithm}
 \end{figure}
 
-\Section{Charm++ and ParSSSE}\label{ParSSSE}
+\Section{Parallel Implementation using Charm++ and ParSSSE}\label{ParSSSE}
 We built our search engine on top of the {\sc Charm++} run-time
@@
 
-\Section{Conclusion Future Work}
-We have shown that information set generation for kriegspiel can be efficiently parallelized in a way that scales up to at least 1024 processors.
-
-A key line of future research is to use our ISG routine as part of a full Kriegspiel play.
+%\Section{Conclusion Future Work}
+%We have shown that information set generation for kriegspiel can be efficiently parallelized in a way that scales up to at least 1024 processors.
+
+%A key line of future research is to use our ISG routine as part of a full Kriegspiel play.
 The ISG method itself could be updated using constraint-satisfaction techniques outlined in ~\cite{richards12information}, which improve pruning performance in the search space by incorporating information from all observations in the list at each search step, instead of performing depth-first search.
@@
 
-\Section{Conclusion}
+\Section{Conclusion and Future Work}
 We have shown that information set generation for kriegspiel can be efficiently parallelized.  The opportunities for

### 2012-06-27 sbac/paper.tex
Many changes

@@
 \usepackage{url,hyperref}
-\usepackage{subfig}
+\usepackage{subfigure}
 \usepackage{tikz}
@@
 
-\author{Mark Richards, Abhishek Gupta, Osman Sarood and Laxmikant Kale\\
+\author{Mark Richards, Abhishek Gupta, Osman Sarood and Laxmikant V. Kale\\
 Department of Computer Science\\ 
@@
 \centering
-\begin{minipage}[b]{0.5\linewidth}
+\begin{minipage}[b]{0.45\linewidth}
+\centering
 \includegraphics[width=0.7\textwidth]{images/9B.png}
@@
 \centering
-\begin{minipage}[b]{0.5\linewidth}
+\begin{minipage}[b]{0.45\linewidth}
+\centering
 \includegraphics[width=0.7\textwidth]{images/OneView.png}
@@
 \end{minipage}
+\vspace{-0.2in}
 \end{figure}
@@
 \end{minipage}
+\vspace{-0.1in}
 \caption{Listing of a kriegspiel game resulting in checkmate by white after 10 moves.}
 \label{filteredlisting}
+\vspace{-0.2in}
 \end{figure}
@@
 \end{mytinylisting}
+\vspace{-0.2in}
 \caption{Abbreviated output from example game.  Generating information sets for black after five turns by each player.
@@
 \label{abbrevoutput}
+\vspace{-0.3in}
 \end{figure}
@@
 
-%%We built our search on top of the same general search engine that we used in class.  The performance results reported in
-%%this section were obtained by running our program on Bleuprint with up to 1024 processors.
-%%
-%%Table~\ref{speedups} shows the performance results for our algorithm on a randomly selected problem instance in which
-%%the size of the information set was about 4.5 million.  (This instance is not necessarily representative of all games or
-%%states of the game, but it was not cherry-picked in any way.)  The table shows that the problem is indeed highly
-%%parallizable.  The speedups are nearly linear for up  to 32 processors, and even on 1024 processors, we get a good
-%%speedup.  In this particular instance, the sequential algorithm takes almost 25 minutes to run.  But when parallelized
-%%on 1024 processors---perfectly reasonable resource usage on modern hardware---the search takes less than three seconds.
-%%This is a speedup of 576 times!  It might not be reasonable to use the algorithm at all if it takes 25 or more to run
-%%for some turns.  But if the cost is only a few seconds, it becomes entirely feasible.
-%%
-%%\begin{table}
-%%\centering
-%%\begin{tabular}{rrr}
-%%{\bf Procs}	&	{\bf Time (s)} 	&	{\bf Speedup}\\
-%%1	&	1524	&	1\\
-%%2	&	812	&	1.88\\
-%%4	&	407	&	3.74\\
-%%8	&	208	&	7.32\\
-%%16	&	102	&	14.8\\
-%%32	&	53.2	&	28.7\\
-%%64	&	28.0	&	54.5\\
-%%128	&	14.9	&	102\\
-%%256	&	7.70	&	197\\
-%%512	&	4.59	&	331\\
-%%1024	&	2.64	&	576\\
-%%\end{tabular}
-%%\caption{Speedups for a randomly selected problem instance on 1--1024 processors.  The input observations were from the
-%%10th ply of the game tree and the size of the information set in this instance is 4,495,121.}
-%%\label{speedups}
-%%\end{table}
-%%
-%%\begin{figure}[h]
-%%\centering
-%%\includegraphics[width=0.7\columnwidth]{plots/3schemes.pdf}
-%%\caption{Experiments using different load balancing strategies.}
-%%\label{3schemes}
-%%\end{figure}
-%%
-%%In our next round of experiments, we set out to compare a few different load-balancing strategies for the problem.  We
-%%used the random, work stealing, and neighborhood balancing strategies that are available in the general purpose search
-%%engine.  These results are shown in Figure~\ref{3schemes}.  Our results show that there was actually very little
-%%variation in performance across the different strategies.  We did notice that across all of our experiments, the work
-%%stealing approach appeared to be consistently---if only marginally---better than the others.  Our hypothesis here is
-%%that perhaps the nodes are rarely starved for work.  In other words, perhaps the very nature of the problem means that
-%%the load is naturally well-balanced and that the load-balancer rarely needs to do a significant amount of
-%%redistribution.  This is probably also related to the fact that we are doing an all-solutions search. 
-%%
-%% 
-%%\begin{figure}[h]
-%%\centering
-%%\begin{minipage}{0.5\linewidth}
-%%\includegraphics[width=\columnwidth]{plots/3depths.pdf}
-%%\end{minipage}
-%%\centering
-%%\begin{minipage}{0.5\linewidth}
-%%\includegraphics[width=\columnwidth]{plots/3depthsnolog.pdf}
-%%\end{minipage}
-%%\caption{(Left) Log-log plot of running time for three different sizes of problem as the number of cores used increases.
-%%(Right) Same data with a linear y-axis for the running time (to help see the isoefficiency.)}
-%%\label{3depths}
-%%\end{figure}
-%%
-%%We next attempted to measure the isoefficiency of the problem---how is efficiency of the parallelization of the
-%%algorithm affected by {\em size} of the problem?  To test this we varied the number of plies in the test instance from
-%%8 to 12. We used the work stealing load balancing strategy.  The results are shown in Table~\ref{3depths}.  The data are
-%%the same in both tables, but the one on the left uses a log-log scale.  In the easiest case (depth 8), there isn't
-%%enough additional parallel work to justify the jump from 256 to 1024 processors, and performance actually degrades.
-%%Other than that, however, we see consistent speedups as we increase the number of processors.  The plot on the right helps to
-%%illustrate the good isoefficiency of the problem.  As the size of the problem increases, the amount of parallelizable
-%%work available also increases, and therefore the steepness of the speed curve is higher for larger problem sizes, as the
-%%number of processors increases.
-%% 
-%%\begin{figure}[h]
-%%\centering
-%%\includegraphics[width=0.6\columnwidth]{plots/dep8.pdf}
-%%\caption{Grainsize control data for depth 8 problems.}
-%%\label{dep8}
-%%\end{figure}
-%% 
-%%\begin{figure}[h]
-%%\centering
-%%\includegraphics[width=0.6\columnwidth]{plots/dep10.pdf}
-%%\caption{Grainsize control data for depth 10 problems.}
-%%\label{dep10}
-%%\end{figure}
-%% 
-%%\begin{figure}[h]
-%%\centering
-%%\includegraphics[width=0.6\columnwidth]{plots/dep12.pdf}
-%%\caption{Grainsize control data for depth 12 problems.}
-%%\label{dep12}
-%%\end{figure}
-%%
-%%Finally, we ran some experiments to measure the effect of grainsize on the problem.  The results are shown in
-%%Tables~\ref{dep8}--~\ref{dep12}.  It is clear that there is a ``sweet spot'' for each problem size.  (At least it is
-%%clear for this particular instance.  It is possible that there would be different behavior on different problem
-%%instances.)  When the grainsize is too small, the overall computation is hampered by excessive parallel overhead
-%%(communication, queueing of nodes to expand, load balancing operations).  On the other hand, when the grain size is too
-%%large, it becomes too difficult to ensure that all processors are busy and processors are not idling with no work to do
-%%while a small number of processors are busy doing large-sized chunks of work.  
-%%
-%%In our experiments, the optimal grainsize is clearly related to the size of the problem, the depth of the search tree.
-%%The optimal sequential threshold seems to be $2--4$ less than the problem depth.  Since the depth of all solutions is
-%%the same for each problem instance, this means that the number of levels that should be searched sequentially is fairly
-%%consistent across problem sizes.  We noted that the number of solutions found---in other words, the size of the
-%%information sets---is typically fairly large with respect to the total number of nodes expanded.  (We don't have hard
-%%figures, but we estimate that it was often $30-40\%$).  This high density of solutions limits the variability that can
-%%be present in the amount of work to be done at the nodes representing the last few levels of the tree.  This means that
-%%the best grainsize is also likely to be fairly predictable.
-
 \Section{Performance Results}\label{Results}
+In order to understand the effect of various parameters such as grain size and load balancing, we first evaluated the scalability of our implementation using a randomly selected problem instance.
+Next, we show that our implementation performs equally well for instances from actual games.
+\subsection{Analysis using Random Instance}
+The performance results reported in
+this section were obtained by running our program on Bleuprint with up to 1024 processors. TODO: Blueprint H/w or citation
+
+Table~\ref{speedups} shows the performance results for our algorithm on a randomly selected problem instance in which
+the size of the information set was about 4.5 million.  This instance is not necessarily representative of all games or
+states of the game, but it was not cherry-picked in any way. The table shows that the problem is indeed highly
+parallizable.  The speedups are nearly linear for up  to 32 processors, and even on 1024 processors, we get a good
+speedup.  In this particular instance, the sequential algorithm takes almost 25 minutes to run.  But when parallelized
+on 1024 processors---perfectly reasonable resource usage on modern hardware---the search takes less than three seconds.
+This is a speedup of 576 times!  It might not be reasonable to use the algorithm at all if it takes 25 or more to run
+for some turns.  But if the cost is only a few seconds, it becomes entirely feasible.
+
+\begin{table}
+\centering
+\begin{tabular}{rrr}
+{\bf Procs}	&	{\bf Time (s)} 	&	{\bf Speedup}\\
+1	&	1524	&	1\\
+2	&	812	&	1.88\\
+4	&	407	&	3.74\\
+8	&	208	&	7.32\\
+16	&	102	&	14.8\\
+32	&	53.2	&	28.7\\
+64	&	28.0	&	54.5\\
+128	&	14.9	&	102\\
+256	&	7.70	&	197\\
+512	&	4.59	&	331\\
+1024	&	2.64	&	576\\
+\end{tabular}
+\vspace{-0.1in}
+\caption{Speedups for a randomly selected problem instance on 1--1024 processors.  The input observations were from the
+10th ply of the game tree and the size of the information set in this instance is 4,495,121.}
+\label{speedups}
+\vspace{-0.3in}
+\end{table}
+
+\subsubsection{Effect of Load Balancing}
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.7\columnwidth]{plots/3schemes.pdf}
+\vspace{-0.1in}
+\caption{Effect of load balancing strategy}
+\label{3schemes}
+\vspace{-0.2in}
+\end{figure}
+
+In our next round of experiments, we set out to compare a few different load-balancing strategies for the problem.  We
+used the random, work stealing, and neighborhood balancing strategies that are available in the general purpose search
+engine.  These results are shown in Figure~\ref{3schemes}.  Our results show that there was actually very little
+variation in performance across the different strategies.  We did notice that across all of our experiments, the work
+stealing approach appeared to be consistently---if only marginally---better than the others.  Our hypothesis here is
+that perhaps the nodes are rarely starved for work.  In other words, perhaps the very nature of the problem means that
+the load is naturally well-balanced and that the load-balancer rarely needs to do a significant amount of
+redistribution.  This is probably also related to the fact that we are doing an all-solutions search.
+
+\subsubsection{Iso-efficiency}
+\begin{figure}[h]
+\centering
+\begin{minipage}{0.5\linewidth}
+\includegraphics[width=\columnwidth]{plots/3depths.pdf}
+\end{minipage}
+\centering
+\begin{minipage}{0.5\linewidth}
+\includegraphics[width=\columnwidth]{plots/3depthsnolog.pdf}
+\end{minipage}
+\vspace{-0.1in}
+\caption{(Top) Log-log plot of running time for three different sizes of problem as the number of cores used increases.
+(Bottom) Same data with a linear y-axis for the running time (to help see the isoefficiency.)}
+\vspace{-0.2in}
+\label{3depths}
+\end{figure}
+
+We next attempted to measure the isoefficiency [TODO:cite] of the problem---how is efficiency of the parallelization of the
+algorithm affected by {\em size} of the problem?  To test this, we varied the number of plies in the test instance from
+8 to 12. We used the work stealing load balancing strategy.  The results are shown in Figure~\ref{3depths}.   In the easiest case (depth 8), there isn't
+enough additional parallel work to justify the jump from 256 to 1024 processors, and performance actually degrades.
+Other than that, however, we see consistent speedups as we increase the number of processors.  The plot on the right helps to
+illustrate the good isoefficiency of the problem.  As the size of the problem increases, the amount of parallelizable
+work available also increases, and therefore the steepness of the speed curve is higher for larger problem sizes, as the
+number of processors increases.
+
+\subsubsection{Effect of Grain Size}
+
+\begin{figure*}[!ht]
+\centering
+\subfigure[Depth 8]{
+\includegraphics[width=2in]{plots/dep8.pdf}
+\label{dep8}
+} 
+\subfigure[Depth 10]{
+\includegraphics[width=2in]{plots/dep10.pdf}
+\label{dep10}
+} 
+\subfigure[Depth 12]{
+\includegraphics[width=2in]{plots/dep12.pdf}
+\label{dep12}
+} 
+\vspace{-0.1in}
+\caption{Grain Size Control for different problem sizes}
+%\vspace{-0.1in}
+\label{fig:dep}
+\vspace{-0.1in}
+\end{figure*}
+
+Next, we evaluate the effect of grainsize on the problem execution time.  The results are shown in
+Tables~\ref{dep8}--~\ref{dep12}.  It is clear that there is a ``sweet spot'' for each problem size.  (At least it is
+clear for this particular instance.  It is possible that there would be different behavior on different problem
+instances.)  When the grainsize is too small, the overall computation is hampered by excessive parallel overhead
+(communication, queueing of nodes to expand, load balancing operations).  On the other hand, when the grain size is too
+large, it becomes too difficult to ensure that all processors are busy and processors are not idling with no work to do
+while a small number of processors are busy doing large-sized chunks of work.
+
+In our experiments, the optimal grainsize is clearly related to the size of the problem, the depth of the search tree.
+The optimal sequential threshold seems to be $2--4$ less than the problem depth.  Since the depth of all solutions is
+the same for each problem instance, this means that the number of levels that should be searched sequentially is fairly
+consistent across problem sizes.  We noted that the number of solutions found---in other words, the size of the
+information sets---is typically fairly large with respect to the total number of nodes expanded.  (We don't have hard
+figures, but we estimate that it was often $30-40\%$).  This high density of solutions limits the variability that can
+be present in the amount of work to be done at the nodes representing the last few levels of the tree.  This means that
+the best grainsize is also likely to be fairly predictable.
+
+\subsection{Real Game Instance}
+\begin{figure}[ht]
+ \centering \includegraphics[viewport= 280 200 300 550, scale=0.35]{images/KriegspielProblem4.pdf} %
+\caption{Kriegspiel position from Virgil-David 1969 with information set size 764,209.}
+\label{prob4} 
+\vspace{-0.2in}
+\end{figure}
+
+\begin{figure}[ht]
+ \centering \includegraphics[viewport= 280 200 300 550, scale=0.35]{images/KriegspielProblem5.pdf} %
+\caption{Kriegspiel position from Game 6 of Kbot-Darkboard at the 2006 Computer Chess championship.  Information set size is 5,099,257.}
+\label{prob5} 
+\vspace{-0.2in}
+\end{figure}
+
 We also evaluated our information set generation routine on 40 positions from actual kriespiel games, including human vs.
@@
 
-\begin{figure}[htpb]
- \centering \includegraphics[scale=0.4]{images/KriegspielProblem4.pdf} %
-\caption{Kriegspiel position from Virgil-David 1969 with information set size 764,209.}
-\label{prob4} 
-\end{figure}
-
-\begin{figure}[htpb]
- \centering \includegraphics[scale=0.4]{images/KriegspielProblem5.pdf} %
-\caption{Kriegspiel position from Game 6 of Kbot-Darkboard at the 2006 Computer Chess championship.  Information set size is 5,099,257.}
-\label{prob5} 
-\end{figure}
-
 \begin{table}[tphb]
@@
 \caption{Running times, in seconds, to find full information set of size 764,209 at position 10 of Virgil-David 1969. Optimal grainsize (G) for each number of cores (PE) is shown in bold}
+\vspace{-0.2in}
 \label{tab:prob4}
@@
 \caption{Running times, in seconds, to find full information set of size 5,099,257 at position 10 of Game 6 Kbot-Darkboard 2006. Optimal grainsize for each number of cores is shown in bold.}
+\vspace{-0.2in}
 \label{tab:prob5}
@@
 \caption{Speedups for Virgil-David 1969.}
+\vspace{-0.2in}
 \label{tab:prob4speedups}
@@
 \caption{Speedups for Kbot-Darkboard 2006.}
+\vspace{-0.2in}
 \label{tab:prob5speedups}
@@
 \caption{Best Speedups reported for game tree applications.}
+\vspace{-0.2in}
 \label{bestspeedups}
@@
 
-Li gives an overview of kriegspiel and discusses strategy from the perspective of human players~\cite{li94chess}.  He
-walks through the analysis of several actual games between two human players and discusses the inferences that each
-player can and should make at each position.  As in chess, human players are able to reason about the game at an
-abstract level; they tend to focus on a small number of critical possibilities and to emphasize the most important
-pieces and squares in each position.  Computer agents, lacking such sophistication, nevertheless have some advantages
-when it comes to brute force search strategies. Programs such Deep Blue have shown that a brute force approach to game
-tree search can be effective, even against humans with vastly superior abstract reasoning capabilities, provided that
-significant computational resources are available and wisely used~\cite{campbell02deep}.
-
-Russell and Wolfe have performed some analyses on kriegspiel positions for which it possible to prove that one player
-has a forced win (that is, regardless of the configuration of the opponent's unseen pieces, the player has a sequence of
-moves that is guaranteed to produce a checkmate)~\cite{russell05efficient, wolfe07exploiting}.  The authors assume
-knowledge of the player's belief state.  They claim that under certain "aggressive" styles of play, it is uncommon for
-information sets to exceed 10,000 nodes in size.  They throw out (i.e., do not analyze) any positions that exceed this
-threshold.
+%Li gives an overview of kriegspiel and discusses strategy from the perspective of human players~\cite{li94chess}.  He
+%walks through the analysis of several actual games between two human players and discusses the inferences that each
+%player can and should make at each position.  As in chess, human players are able to reason about the game at an
+%abstract level; they tend to focus on a small number of critical possibilities and to emphasize the most important
+%pieces and squares in each position.  Computer agents, lacking such sophistication, nevertheless have some advantages
+%when it comes to brute force search strategies. Programs such Deep Blue have shown that a brute force approach to game
+%tree search can be effective, even against humans with vastly superior abstract reasoning capabilities, provided that
+%significant computational resources are available and wisely used~\cite{campbell02deep}.
+
+%Russell and Wolfe have performed some analyses on kriegspiel positions for which it possible to prove that one player
+%has a forced win (that is, regardless of the configuration of the opponent's unseen pieces, the player has a sequence of
+%moves that is guaranteed to produce a checkmate)~\cite{russell05efficient, wolfe07exploiting}.  The authors assume
+%knowledge of the player's belief state.  They claim that under certain "aggressive" styles of play, it is uncommon for
+%information sets to exceed 10,000 nodes in size.  They throw out (i.e., do not analyze) any positions that exceed this
+%threshold.
 
@@
 %smaller), we want to be able to develop a strategy that can be used in the middle of the game as well.  game).
-
-Nance \etal, approach the game from the perspective of {\em logical filtering.}~\cite{nance06reasoning}  The goal of
-logical filtering is to maintain a compact logical representation of an agent's belief state.  Recall that a belief
-state is a probability distribution over the possible current states of the world.  A belief state for kriegspiel might
-take the form of $(at(whitepawn,a2) \wedge at whitepawn(b2) \ldots \wedge at(whiterook,h1)) \vee (at(whitepawn,a3)
-\wedge \ldots \wedge at(whiterook,h1)) \ldots$  For even moderately sizes information sets, such
-representations can become unwieldy in terms of storage and the cost of updating.  Furthermore, reasoning with such
-representations often requires the use of some theorem-proving mechanism to run logical queries, such
-as $at(blackking,a3?)$. Nance et al. develop some strategies to maintain compact representations under
-sequences of actions and observations.  They give a theoretical analysis of the kinds of problems that can be treated in
-this way.  Unfortunately, kriegspiel in its full complexity is beyond the scope of their algorithm.  In order to reason
-about the belief states in the game, they must assume knowledge of the particular kind of piece that move in each turn.
 

### 2012-06-27 sbac/paper.tex
spell check

@@
 performance of game playing agents.  In this work, we address the parallelization of information set generation in the context
-of kriegspiel (partially observable chess).  We implement the algorithm on top of a general purpose combinatorial search
-engine and discuss its performance under different load balancing strategies and with different grainsize parameters.
+of Kriegspiel (partially observable chess).  We implement the algorithm on top of a general purpose combinatorial search
+engine and discuss its performance under different load balancing strategies and with different grain-size parameters.
 We achieve speedups of over 500 on 1024 processors, far exceeding previous scalability results for game tree search applications.
@@
 efficiency is a key factor in an agent's overall performance.  In this work, we
-show that ISG for kriegspiel can be efficiently parallelized.
+show that ISG for Kriegspiel can be efficiently parallelized.
 
@@
 
-In the game of kriegspiel, information set generation is much more complicated.  The player knows the opponent's position with
+In the game of Kriegspiel, information set generation is much more complicated.  The player knows the opponent's position with
 certainty when the game begins, but after the initial move there may be varying levels of uncertainty about the location
@@
 subroutine for quality game play in imperfect information games.  In this work, we describe an approach to generating
-information sets for the game of kriegspiel.
+information sets for the game of Kriegspiel.
 
@@
 The performance results reported in
-this section were obtained by running our program on Bleuprint with up to 1024 processors. TODO: Blueprint H/w or citation
+this section were obtained by running our program on Blueprint with up to 1024 processors. TODO: Blueprint H/w or citation
 
@@
 \caption{(Top) Log-log plot of running time for three different sizes of problem as the number of cores used increases.
-(Bottom) Same data with a linear y-axis for the running time (to help see the isoefficiency.)}
+(Bottom) Same data with a linear y-axis for the running time (to help see the Iso-efficiency.)}
 \vspace{-0.2in}
@@
 
-We next attempted to measure the isoefficiency [TODO:cite] of the problem---how is efficiency of the parallelization of the
+We next attempted to measure the Iso-efficiency [TODO:cite] of the problem---how is efficiency of the parallelization of the
 algorithm affected by {\em size} of the problem?  To test this, we varied the number of plies in the test instance from
@@
 This results in many parallel applications of information set generation.
-Furthermore, given a node in an informaion set, it is common for an agent to perform forward search on that node (e.g., using MCTS).
+Furthermore, given a node in an information set, it is common for an agent to perform forward search on that node (e.g., using MCTS).
 With these two additional levels of parallelism on top of the information set generation problem, it is conceivable that the problem of playing partially observable games could scale efficiently to millions of processors or more.
@@
 
-%Russell and Wolfe have performed some analyses on kriegspiel positions for which it possible to prove that one player
+%Russell and Wolfe have performed some analysis on kriegspiel positions for which it possible to prove that one player
 %has a forced win (that is, regardless of the configuration of the opponent's unseen pieces, the player has a sequence of

### 2012-06-27 sbac/paper.tex
Tables merge an, abstract, intro

@@
 consistent with a player's observations.  The ability to reason about the possible game history is critical to the
-performance of game playing agents.  In this work, we address the parallelization of information set generation in the context
+performance of game playing agents.  ISG represent a class of combinatorial search problems which is computationaly intensive but difficult to efficiently parallelize.
+
+In this paper, we address the parallelization of information set generation in the context
 of Kriegspiel (partially observable chess).  We implement the algorithm on top of a general purpose combinatorial search
@@
 expensive and may be repeatedly applied by a decision-making agent, its
-efficiency is a key factor in an agent's overall performance.  In this work, we
-show that ISG for Kriegspiel can be efficiently parallelized.
+efficiency is a key factor in an agent's overall performance.
+Information set generation is not the same problem as solving a game tree, although it is arguably a necessary subroutine for quality game play in imperfect information games.
 
@@
 
-Information set generation is not the same problem as solving a game tree, although it is arguably a necessary
-subroutine for quality game play in imperfect information games.  In this work, we describe an approach to generating
-information sets for the game of Kriegspiel.
+In this work, we show that ISG for Kriegspiel can be efficiently parallelized.
+The primary contributions of this work are the following.
+We describe an approach to efficient parallelization of an application which is not typical of high performance computing domain.
+We present scalability results for our parallel implementation of information set generation algorithm for the game of Kriegspiel using real game problem instances.
+Further, we demonstrate the effect of load balancing and grain size on parallel performance of our solution.
 
@@
 
-\subsubsection{Iso-efficiency}
+\subsubsection{Effect of Problem Size}
 \begin{figure}[h]
@@
 
-We next attempted to measure the Iso-efficiency [TODO:cite] of the problem---how is efficiency of the parallelization of the
-algorithm affected by {\em size} of the problem?  To test this, we varied the number of plies in the test instance from
+We next study the variation in the efficiency of the parallelization of the
+algorithm with {\em size} of the problem?  To test this, we varied the number of plies in the test instance from
 8 to 12. We used the work stealing load balancing strategy.  The results are shown in Figure~\ref{3depths}.   In the easiest case (depth 8), there isn't
@@
 PE/G & 2 & 4 & 6 & 8 \\
-1 & {\bf 177} & 178 & 184 & 493 \\
-4 & 71 & 53 & {\bf 48} & 75 \\
-16 & 34 & 19 & {\bf 12} & 16 \\
-64 & 24 & 7 & {\bf 3.7} & 4.1 \\
-256 & 24 & 4.4 & 1.2 & {\bf 1.0} \\
-1024 & 26 & 3.8 & 0.64 & {\bf 0.40}
+1 & {\bf 177$|$1} & 178$|$0.99 & 1844$|$0.96 & 493$|$0.36 \\
+4 & 71$|$2.48 & 53$|$3.34 & {\bf 48$|$3.63} & 75$|$2.35 \\
+16 & 34$|$5.19 & 19 $|$9.27& {\bf 12$|$13.8} & 16$|$10.7 \\
+64 & 24 $|$7.25 & 7 $|$23.6& {\bf 3.7$|$47.39} & 4.1$|$43.1 \\
+256 & 24 $|$ 7.1& 4.4 $|$40.1 & 1.2 $|$145& {\bf 1.0$|$172} \\
+1024 & 26 $|$6.61 & 3.8 $|$46.5& 0.64 $|$276& {\bf 0.40$|$443}
 \end{tabular}
-\caption{Running times, in seconds, to find full information set of size 764,209 at position 10 of Virgil-David 1969. Optimal grainsize (G) for each number of cores (PE) is shown in bold}
+\caption{Running times in seconds and speedups (after $|$), to find full information set of size 764,209 at position 10 of Virgil-David 1969.
+Optimal grainsize (G) for each number of cores (PE) is shown in bold}
 \vspace{-0.2in}
@@
 PE/G & 2 & 4 & 6 & 8 \\
-1 & 617 & {\bf 612} & 650 & 3312 \\
-4 & 213 & 170 & {\bf 158} & 367 \\
-16 & 83 & 49 & {\bf 40} & 64 \\
-64 & 54 & 19 & {\bf 11} & 14.9 \\
-256 & 54 & 9.8 & 3.5 & {\bf 3.5} \\
-1024 & 58 & 5.5 & 1.3 & {\bf 1.2} \\
+1 & 617$|$1 & {\bf 612$|$1} & 650$|$0.95 & 3312$|$0.19 \\
+4 & 213$|$2.89 & 170$|$3.63 & {\bf 158$|$3.89} & 367$|$1.68 \\
+16 & 83$|$7.41 & 49$|$12.4 & {\bf 40$|$15.1} & 64$|$9.57 \\
+64 & 54$|$11.3 & 19$|$32.3 & {\bf 11$|$52.9} & 14.9$|$41.4 \\
+256 & 54$|$11.3 & 9.8$|$62.9 & 3.5$|$172 & {\bf 3.5$|$174} \\
+1024 & 58$|$10.6 & 5.5$|$111 & 1.3$|$464 & {\bf 1.2$|$536} \\
 \end{tabular}
-\caption{Running times, in seconds, to find full information set of size 5,099,257 at position 10 of Game 6 Kbot-Darkboard 2006. Optimal grainsize for each number of cores is shown in bold.}
+\caption{Running times, in seconds and speedups (after $|$), to find full information set of size 5,099,257 at position 10 of Game 6 Kbot-Darkboard 2006.
+Optimal grainsize for each number of cores is shown in bold.}
 \vspace{-0.2in}
 \label{tab:prob5}
-\end{table}
-
-\begin{table}[tphb]
-\centering
-\begin{tabular}{rrrrr}
-PE/G & 2 & 4 & 6 & 8 \\
-1 & 1 & 0.99 & 0.96 & 0.36 \\ 
-4 & 2.48 & 3.34 & 3.63 & 2.35 \\
-16 & 5.19 & 9.27 & 13.8 & 10.7 \\
-64 & 7.25 & 23.6 & 47.39 & 43.1 \\
-256 & 7.1 & 40.1 & 145 & 172 \\
-1024 & 6.61 & 46.5 & 276 & {\bf 443}
-\end{tabular}
-\caption{Speedups for Virgil-David 1969.}
-\vspace{-0.2in}
-\label{tab:prob4speedups}
-\end{table}
-
-\begin{table}[tphb]
-\centering
-\begin{tabular}{rrrrr}
-PE/G & 2 & 4 & 6 & 8 \\
-1 & 1 & 1 & 0.95 & 0.19 \\
-4 & 2.89 & 3.63 & 3.89 & 1.68 \\
-16 & 7.41 & 12.4 & 15.1 & 9.57 \\
-64 & 11.3 & 32.3 & 52.9 & 41.4 \\
-256 & 11.3 & 62.9 & 172 & 174 \\
-1024 & 10.6 & 111 & 464 & {\bf 536}
-\end{tabular}
-\caption{Speedups for Kbot-Darkboard 2006.}
-\vspace{-0.2in}
-\label{tab:prob5speedups}
 \end{table}

### 2012-06-28 sbac/paper.tex
Increase font size for figures
Added description of Ranger

@@
 \centering
-\includegraphics[width=0.7\columnwidth]{plots/3schemes.pdf}
+\includegraphics[width=0.9\columnwidth]{plots/3schemes.pdf}
 \vspace{-0.1in}
@@
 \centering
-\begin{minipage}{0.5\linewidth}
+\begin{minipage}{0.7\linewidth}
 \includegraphics[width=\columnwidth]{plots/3depths.pdf}
@@
 \centering
-\begin{minipage}{0.5\linewidth}
+\begin{minipage}{0.7\linewidth}
 \includegraphics[width=\columnwidth]{plots/3depthsnolog.pdf}
@@
 \subfigure[Depth 8]{
-\includegraphics[width=2in]{plots/dep8.pdf}
+\includegraphics[width=2.2in]{plots/dep8.pdf}
 \label{dep8}
@@
 \subfigure[Depth 10]{
-\includegraphics[width=2in]{plots/dep10.pdf}
+\includegraphics[width=2.2in]{plots/dep10.pdf}
 \label{dep10}
@@
 \subfigure[Depth 12]{
-\includegraphics[width=2in]{plots/dep12.pdf}
+\includegraphics[width=2.2in]{plots/dep12.pdf}
 \label{dep12}
@@
 Many of these were trivial to solve in a few seconds or less on a desktop computer.
-The running times and speedups for two of the more challenging instances are shown in Tables~\ref{prob4}--\ref{tab:prob5speedups}.
+For these experiments, we used a real supercomputer - Ranger at TACC~\cite{ranger}.
+Ranger has 16-way SMP nodes containing AMD Opteron processors.
+The nodes are interconnected using Infiniband technology in a fully-CLOS topology providing a 1GBps point-to-point bandwidth.
+The running times and speedups for two of the more challenging instances on Ranger are shown in Tables~\ref{prob4}--\ref{tab:prob5speedups}.
 These tables show the results for runs on one to $1024$ cores.

### 2012-07-10 sbac/paper.tex
text for blueprint
Placement etc

@@
 The paper is organized as follows.
-In Section~\ref{Background}, we review necessary background information, including the rules of kriegspiel and the nature of the information set generation problem.
-In Section ~\ref{ParSSSE}, we describe the {\sc Charm++} Parallel State Space Search Engine (ParSSSE) platform.
+In Section~\ref{Background}, we review necessary background information, including the rules of kriegspiel.
+Next, in Section~\ref{info} we describe the nature of the information set generation problem and the algorithm used.
+In Section~\ref{ParSSSE}, we describe our parallel implementation, including the {\sc Charm++} Parallel State Space Search Engine (ParSSSE) platform.
 We then provide our experimental results in Section~\ref{Results}.
@@
 \Section{Background: Kriegspiel}
+\label{Background}
 Information Set Generation is a widely adopted operation used in game trees search.
@@
 \Section{Information Set Generation}
+\label{info}
 %\subsection{Information Set Generation}
 The concept of information sets is perhaps best explained through an example. 
-Figure~\ref{listing} and show a full listing and the associated pictorial progression of an actual
-kriegspiel game~\cite{li94chess}.  We have adapted the notation used by Wolfe for Berkeley Kriegspiel.  Odd (resp. even) moves are for white (resp. black).  The actual move is shown
+Figure~\ref{filteredlisting},~\ref{fig:figure1} and \ref{fig:figure2} show a full listing and the associated pictorial progression of an actual
+kriegspiel game~\cite{li94chess}.  We have adopted the notation used by Wolfe for Berkeley Kriegspiel.  Odd (resp. even) moves are for white (resp. black).  The actual move is shown
 first followed by a list of referee announcements.  Captures announcements are prefixed with an X and give the location
 of the captured piece.  Similarly, pawn tries announcements are prefixed with T.  Attempted moves that were ultimately
-declared illegal are shown in a list following the `:'.  For example at move 9, (Figure~\ref{fig:9W}--\ref{fig:9B}),
+declared illegal are shown in a list following the `:'.  For example at move 9, (Figure~\ref{filteredlisting}),
 black attempts to block a potential threat along the diagonal by moving a pawn to f6, then attempts to escape or capture
@@
 to be found: check status, pawn tries, and illegal moves available/attempted.
-
-Figure~\ref{codelisting} shows the function at the heart of our information set generation algorithm.  The actual
-observations can be viewed as global, read-only variables.  All solutions (i.e., all nodes in the information set) will
-be located at the same depth in the tree: the depth that is equal to the number of plies in the observation list.
-Referee announcements regarding pawn tries or a player being in check are made prior to a move.  The algorithm checks
-for the consistency of these observations at line $7$.  If the search reaches the appropriate depth and passes these
-tests, then a solution has been found and is added to the information set (line $3$).
-
-%Otherwise, we consider the possible moves that can be made from that state.  If the player whose turn it is to move at
-%the current level of the search tree is the same player from whose perspective we have received observations, then the
-%next step is to make sure that every legal move that was attempted at the player at that ply is also at least
-%attemptable at the game position that corresponds to the game tree node (lines $15--22$).  We know exactly the legal
-%move that was ultimately made at this depth (because we are assuming perfect recall), and we check to make sure that
-%this move is indeed legal (pruning otherwise).  If so, we apply that move and recurse (lines $24--32$).  This
-%corresponds to a ``forced move'' in the general search paradigm.
-%
-%On the other hand, if the player whose turn it is to move at the current depth of the search tree is the opponent, then
-%must consider all possible legal moves to branch on.  We first check to make sure that the number of legal moves from
-%this position is at least as large as the number of illegal moves that the opponent attempted (i.e., the number of times
-%the referee said ``No'' before finally accepting the opponent's move).  This is shown in lines $39--42$.  At this point
-%we generate a child for each move that would be accepted by the referee and recurse (lines $44--52$).
-%
-%One of our goals was to assess the varying utilities of the different pruning criteria.  We tried some different
-%orderings of the tests and were somewhat surprised (and disappointed) to see that there was actually a fairly minimal
-%impact on performance.  This could be due to the fact that because of the tedious nature of encoding problem instances
-%from actual games, we were not able to run experiments of a wide variety of types of positions (i.e., different styles
-%of play).
-
-%\begin{figure}
-%%\begin{listing}
-%\begin{lstlisting}
-%void generateInformationSet(uint16_t* trueState, uint16_t* possState, bool whiteMove, 
-%	uint16_t* possHistory, uint16_t** levels, int depth)
-%{
-%  if (!samePawnTries(trueState, possState, whiteMove)) return;  
-%  if (!sameCheckStatus(trueState, possState, whiteMove)) return; 
-%  if (depth == maxdepth) { // Then we have found a solution
-%	nSolutions++; return; // Report solution
-%  }
-%  uint16_t newPossState[16]; 
-%  int nMoves = 0;
-%  generateAttemptableMoves(possState, whiteMove, levels[depth], nMoves,false);
-%  checkForCheck(possState, whiteMove, levels[depth], nMoves); // prunes if not equal
-%  assert (nMoves < NMOVES); // 
-%
-%  if (whitePerspective == whiteMove) { // move from perspective plyaer 
-%    SetMove& failures = failedMoves[depth];
-%    for (SetMove::const_iterator itr = failures.begin(); itr != failures.end(); ++itr) {
-%      uint16_t move = *itr ;
-%      if (!foundMatchingMove(move,levels[depth],nMoves)) {
-%        return; // Prune: one of my own failed moves is not attemptable in this state
-%      }
-%    }
-%    // We know exactly what the actual move was; make it
-%    uint16_t& actualMove = moveHistory[depth];
-%    if (!foundMatchingMove(actualMove,levels[depth],nMoves)) {
-%      return; // Prune (because the move that we know we made at this depth is not legal)
-%    }
-%    // Otherwise, recurse 
-%    applyMove(possState,newPossState,actualMove);
-%    possHistory[depth] = actualMove;
-%    generateInformationSet(newTrueState, newPossState, !whiteMove, 
-%      possHistory, levels, depth+1);
-%  } else { 
-%    // We are at a level in the search tree where 
-%    // we are considering the possible moves for the opponent.
-%    // We know the number of attempted illegal moves, but not which ones
-%
-%    // Must be enough attemptable moves in the state to match the number of failed attempts
-%    unsigned nIllegalMoves = countIllegalMoves(levels[depth], nMoves);
-%    if (nIllegalMoves < failedMoves[depth].size()) {
-%	return; // Prune if not enough attemptable moves
-%    }
-%    // Now we want to try each possible move that is legally executable (not just attemptable) 
-%    for (int i = 0; i < nMoves; i++) {
-%      uint16_t& move = levels[depth][i];
-%      if (isLegal(move)) { // Obviously, we can only execute the moves that are actually legal from this state
-%        applyMove(possState,newPossState,move);
-%        possHistory[depth] = move;
-%        generateInformationSet(newTrueState, newPossState, !whiteMove, 
-%          possHistory, levels, depth+1);
-%      }
-%    }
-%  }
-%}
-%\end{lstlisting}
-%\caption{The heart of the information set generation algorithm.}
-%\label{codelisting}
-%\end{figure}
-
 \begin{figure}[htpb]
@@
 \EndFunction
-\label{codelisting}
+\label{codelisting1}
 \end{algorithmic}
@@
 \end{figure}
+
+Figure~\ref{codelisting1} shows the function at the heart of our information set generation algorithm.  The actual
+observations can be viewed as global, read-only variables.  All solutions (i.e., all nodes in the information set) will
+be located at the same depth in the tree: the depth that is equal to the number of plies in the observation list.
+Referee announcements regarding pawn tries or a player being in check are made prior to a move.  The algorithm checks
+for the consistency of these observations at line $7$.  If the search reaches the appropriate depth and passes these
+tests, then a solution has been found and is added to the information set (line $3$).
+
+%Otherwise, we consider the possible moves that can be made from that state.  If the player whose turn it is to move at
+%the current level of the search tree is the same player from whose perspective we have received observations, then the
+%next step is to make sure that every legal move that was attempted at the player at that ply is also at least
+%attemptable at the game position that corresponds to the game tree node (lines $15--22$).  We know exactly the legal
+%move that was ultimately made at this depth (because we are assuming perfect recall), and we check to make sure that
+%this move is indeed legal (pruning otherwise).  If so, we apply that move and recurse (lines $24--32$).  This
+%corresponds to a ``forced move'' in the general search paradigm.
+%
+%On the other hand, if the player whose turn it is to move at the current depth of the search tree is the opponent, then
+%must consider all possible legal moves to branch on.  We first check to make sure that the number of legal moves from
+%this position is at least as large as the number of illegal moves that the opponent attempted (i.e., the number of times
+%the referee said ``No'' before finally accepting the opponent's move).  This is shown in lines $39--42$.  At this point
+%we generate a child for each move that would be accepted by the referee and recurse (lines $44--52$).
+%
+%One of our goals was to assess the varying utilities of the different pruning criteria.  We tried some different
+%orderings of the tests and were somewhat surprised (and disappointed) to see that there was actually a fairly minimal
+%impact on performance.  This could be due to the fact that because of the tedious nature of encoding problem instances
+%from actual games, we were not able to run experiments of a wide variety of types of positions (i.e., different styles
+%of play).
+
+%\begin{figure}
+%%\begin{listing}
+%\begin{lstlisting}
+%void generateInformationSet(uint16_t* trueState, uint16_t* possState, bool whiteMove, 
+%	uint16_t* possHistory, uint16_t** levels, int depth)
+%{
+%  if (!samePawnTries(trueState, possState, whiteMove)) return;  
+%  if (!sameCheckStatus(trueState, possState, whiteMove)) return; 
+%  if (depth == maxdepth) { // Then we have found a solution
+%	nSolutions++; return; // Report solution
+%  }
+%  uint16_t newPossState[16]; 
+%  int nMoves = 0;
+%  generateAttemptableMoves(possState, whiteMove, levels[depth], nMoves,false);
+%  checkForCheck(possState, whiteMove, levels[depth], nMoves); // prunes if not equal
+%  assert (nMoves < NMOVES); // 
+%
+%  if (whitePerspective == whiteMove) { // move from perspective plyaer 
+%    SetMove& failures = failedMoves[depth];
+%    for (SetMove::const_iterator itr = failures.begin(); itr != failures.end(); ++itr) {
+%      uint16_t move = *itr ;
+%      if (!foundMatchingMove(move,levels[depth],nMoves)) {
+%        return; // Prune: one of my own failed moves is not attemptable in this state
+%      }
+%    }
+%    // We know exactly what the actual move was; make it
+%    uint16_t& actualMove = moveHistory[depth];
+%    if (!foundMatchingMove(actualMove,levels[depth],nMoves)) {
+%      return; // Prune (because the move that we know we made at this depth is not legal)
+%    }
+%    // Otherwise, recurse 
+%    applyMove(possState,newPossState,actualMove);
+%    possHistory[depth] = actualMove;
+%    generateInformationSet(newTrueState, newPossState, !whiteMove, 
+%      possHistory, levels, depth+1);
+%  } else { 
+%    // We are at a level in the search tree where 
+%    // we are considering the possible moves for the opponent.
+%    // We know the number of attempted illegal moves, but not which ones
+%
+%    // Must be enough attemptable moves in the state to match the number of failed attempts
+%    unsigned nIllegalMoves = countIllegalMoves(levels[depth], nMoves);
+%    if (nIllegalMoves < failedMoves[depth].size()) {
+%	return; // Prune if not enough attemptable moves
+%    }
+%    // Now we want to try each possible move that is legally executable (not just attemptable) 
+%    for (int i = 0; i < nMoves; i++) {
+%      uint16_t& move = levels[depth][i];
+%      if (isLegal(move)) { // Obviously, we can only execute the moves that are actually legal from this state
+%        applyMove(possState,newPossState,move);
+%        possHistory[depth] = move;
+%        generateInformationSet(newTrueState, newPossState, !whiteMove, 
+%          possHistory, levels, depth+1);
+%      }
+%    }
+%  }
+%}
+%\end{lstlisting}
+%\caption{The heart of the information set generation algorithm.}
+%\label{codelisting}
+%\end{figure}
 
@@
 \Section{Performance Results}\label{Results}
-In order to understand the effect of various parameters such as grain size and load balancing, we first evaluated the scalability of our implementation using a randomly selected problem instance.
-Next, we show that our implementation performs equally well for instances from actual games.
-\subsection{Analysis using Random Instance}
-The performance results reported in
-this section were obtained by running our program on Blueprint with up to 1024 processors. TODO: Blueprint H/w or citation
-
-Table~\ref{speedups} shows the performance results for our algorithm on a randomly selected problem instance in which
-the size of the information set was about 4.5 million.  This instance is not necessarily representative of all games or
-states of the game, but it was not cherry-picked in any way. The table shows that the problem is indeed highly
-parallizable.  The speedups are nearly linear for up  to 32 processors, and even on 1024 processors, we get a good
-speedup.  In this particular instance, the sequential algorithm takes almost 25 minutes to run.  But when parallelized
-on 1024 processors---perfectly reasonable resource usage on modern hardware---the search takes less than three seconds.
-This is a speedup of 576 times!  It might not be reasonable to use the algorithm at all if it takes 25 or more to run
-for some turns.  But if the cost is only a few seconds, it becomes entirely feasible.
-
-\begin{table}
+
+\begin{table}[ht]
 \centering
@@
 \end{table}
+In order to understand the effect of various parameters such as grain size and load balancing, we first evaluated the scalability of our implementation using a randomly selected problem instance.
+Next, we show that our implementation performs equally well for instances from actual games.
+\subsection{Analysis using Random Instance}
+The performance results reported in
+this section were obtained by running our program on Blueprint with up to 1024 processors. TODO: Blueprint H/w or citation
+
+Table~\ref{speedups} shows the performance results for our algorithm on a randomly selected problem instance in which
+the size of the information set was about 4.5 million.  This instance is not necessarily representative of all games or
+states of the game, but it was not cherry-picked in any way. The table shows that the problem is indeed highly
+parallizable.  The speedups are nearly linear for up  to 32 processors, and even on 1024 processors, we get a good
+speedup.  In this particular instance, the sequential algorithm takes almost 25 minutes to run.  But when parallelized
+on 1024 processors---perfectly reasonable resource usage on modern hardware---the search takes less than three seconds.
+This is a speedup of 576 times!  It might not be reasonable to use the algorithm at all if it takes 25 or more to run
+for some turns.  But if the cost is only a few seconds, it becomes entirely feasible.
 
 \subsubsection{Effect of Load Balancing}
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.9\columnwidth]{plots/3schemes.pdf}
-\vspace{-0.1in}
-\caption{Effect of load balancing strategy}
-\label{3schemes}
-\vspace{-0.2in}
-\end{figure}
 
@@
 redistribution.  This is probably also related to the fact that we are doing an all-solutions search. 
-
-\subsubsection{Effect of Problem Size}
-\begin{figure}[h]
+\begin{figure}[ht]
 \centering
-\begin{minipage}{0.7\linewidth}
-\includegraphics[width=\columnwidth]{plots/3depths.pdf}
-\end{minipage}
-\centering
-\begin{minipage}{0.7\linewidth}
-\includegraphics[width=\columnwidth]{plots/3depthsnolog.pdf}
-\end{minipage}
+\includegraphics[width=0.9\columnwidth]{plots/3schemes.pdf}
 \vspace{-0.1in}
-\caption{(Top) Log-log plot of running time for three different sizes of problem as the number of cores used increases.
-(Bottom) Same data with a linear y-axis for the running time (to help see the Iso-efficiency.)}
+\caption{Effect of load balancing strategy}
+\label{3schemes}
 \vspace{-0.2in}
-\label{3depths}
 \end{figure}
-
-We next study the variation in the efficiency of the parallelization of the
-algorithm with {\em size} of the problem?  To test this, we varied the number of plies in the test instance from
-8 to 12. We used the work stealing load balancing strategy.  The results are shown in Figure~\ref{3depths}.   In the easiest case (depth 8), there isn't
-enough additional parallel work to justify the jump from 256 to 1024 processors, and performance actually degrades.
-Other than that, however, we see consistent speedups as we increase the number of processors.  The plot on the right helps to
-illustrate the good isoefficiency of the problem.  As the size of the problem increases, the amount of parallelizable
-work available also increases, and therefore the steepness of the speed curve is higher for larger problem sizes, as the
-number of processors increases.
 
@@
 
+\begin{figure}[ht]
+\centering
+\begin{minipage}{0.7\linewidth}
+\includegraphics[width=\columnwidth]{plots/3depths.pdf}
+\end{minipage}
+\centering
+\begin{minipage}{0.7\linewidth}
+\includegraphics[width=\columnwidth]{plots/3depthsnolog.pdf}
+\end{minipage}
+\vspace{-0.1in}
+\caption{(Top) Log-log plot of running time for three different sizes of problem as the number of cores used increases.
+(Bottom) Same data with a linear y-axis for the running time (to help see the Iso-efficiency.)}
+\vspace{-0.2in}
+\label{3depths}
+\end{figure}
 Next, we evaluate the effect of grainsize on the problem execution time.  The results are shown in
-Tables~\ref{dep8}--~\ref{dep12}.  It is clear that there is a ``sweet spot'' for each problem size.  (At least it is
+Figures~\ref{dep8}--~\ref{dep12}.  It is clear that there is a ``sweet spot'' for each problem size.  (At least it is
 clear for this particular instance.  It is possible that there would be different behavior on different problem
@@
 
+\subsubsection{Effect of Problem Size}
+
+We next study the variation in the efficiency of the parallelization of the
+algorithm with {\em size} of the problem?  To test this, we varied the number of plies in the test instance from
+8 to 12. We used the work stealing load balancing strategy.  The results are shown in Figure~\ref{3depths}.   In the easiest case (depth 8), there isn't
+enough additional parallel work to justify the jump from 256 to 1024 processors, and performance actually degrades.
+Other than that, however, we see consistent speedups as we increase the number of processors.  The plot on the right helps to
+illustrate the good isoefficiency of the problem.  As the size of the problem increases, the amount of parallelizable
+work available also increases, and therefore the steepness of the speed curve is higher for larger problem sizes, as the
+number of processors increases.
 \subsection{Real Game Instance}
+%\begin{figure}[ht]
+% \centering \includegraphics[viewport= 280 200 300 550, scale=0.35]{images/KriegspielProblem4.pdf} %
+%\caption{Kriegspiel position from Virgil-David 1969 with information set size 764,209.}
+%\label{prob4} 
+%\vspace{-0.2in}
+%\end{figure}
+
+%\begin{figure}[ht]
+% \centering \includegraphics[viewport= 280 200 300 550, scale=0.35]{images/KriegspielProblem5.pdf} %
+%\caption{Kriegspiel position from Game 6 of Kbot-Darkboard at the 2006 Computer Chess championship.  Information set size is 5,099,257.}
+%\label{prob5} 
+%\vspace{-0.2in}
+%\end{figure}
+
 \begin{figure}[ht]
- \centering \includegraphics[viewport= 280 200 300 550, scale=0.35]{images/KriegspielProblem4.pdf} %
-\caption{Kriegspiel position from Virgil-David 1969 with information set size 764,209.}
+ \centering \includegraphics[scale=1]{plots/new.pdf} %
+\caption{Kriegspiel position from 2 games, postiion 10}
 \label{prob4} 
@@
 
-\begin{figure}[ht]
- \centering \includegraphics[viewport= 280 200 300 550, scale=0.35]{images/KriegspielProblem5.pdf} %
-\caption{Kriegspiel position from Game 6 of Kbot-Darkboard at the 2006 Computer Chess championship.  Information set size is 5,099,257.}
-\label{prob5} 
-\vspace{-0.2in}
-\end{figure}
-
-We also evaluated our information set generation routine on 40 positions from actual kriespiel games, including human vs.
-human and computer vs.
-computer.
-Many of these were trivial to solve in a few seconds or less on a desktop computer.
-For these experiments, we used a real supercomputer - Ranger at TACC~\cite{ranger}.
-Ranger has 16-way SMP nodes containing AMD Opteron processors.
-The nodes are interconnected using Infiniband technology in a fully-CLOS topology providing a 1GBps point-to-point bandwidth.
-The running times and speedups for two of the more challenging instances on Ranger are shown in Tables~\ref{prob4}--\ref{tab:prob5speedups}.
-These tables show the results for runs on one to $1024$ cores.
-Each column corresponds to a different parameter.
-In this case, the grainsize is the number of levels of the search tree that are searched in parallel.
-Up to that threshold, each node that is generated spawns a new piece of work, which the {\sc Charm++} runtime schedules to execute on one of the processing elements according to its load balancing strategy.
-Above the threshold the entire subtree is searched to completion on a single processor, with no additional parallel overhead.
-
-Note that the optimal grainsize varies for the different numbers of cores and the pattern is relatively consistent across problem instances.
-The optimal grainsize increases as the number of cores increases.
-This makes sense, because as this parameter grows, the number of distinct pieces of work increases.
-When the number of pieces of work per processor is high, much of the overhead of parallelization is wasted.
-But as the number of cores increases, the work can be efficiently broken up into smaller pieces and doled out to all the cores.
-With more pieces of work, it becomes easier to balance the load.
-
-\begin{table}[tphb]
+\begin{table}[ht]
 \centering
@@
 
-\begin{table}[tphb]
+\begin{table}[th]
 \centering
@@
 
+We also evaluated our information set generation routine on 40 positions from actual kriespiel games, including human vs.
+human and computer vs.
+computer.
+Many of these were trivial to solve in a few seconds or less on a desktop computer.
+For these experiments, we used a real supercomputer - Ranger at TACC~\cite{ranger}.
+Ranger has 16-way SMP nodes containing AMD Opteron processors.
+The nodes are interconnected using Infiniband technology in a fully-CLOS topology providing a 1GBps point-to-point bandwidth.
+The running times and speedups for two of the more challenging instances on Ranger are shown in Tables~\ref{tab:prob4}--\ref{tab:prob5}.
+These tables show the results for runs on one to $1024$ cores.
+Each column corresponds to a different parameter.
+In this case, the grainsize is the number of levels of the search tree that are searched in parallel.
+Up to that threshold, each node that is generated spawns a new piece of work, which the {\sc Charm++} runtime schedules to execute on one of the processing elements according to its load balancing strategy.
+Above the threshold the entire subtree is searched to completion on a single processor, with no additional parallel overhead.
+
+Note that the optimal grainsize varies for the different numbers of cores and the pattern is relatively consistent across problem instances.
+The optimal grainsize increases as the number of cores increases.
+This makes sense, because as this parameter grows, the number of distinct pieces of work increases.
+When the number of pieces of work per processor is high, much of the overhead of parallelization is wasted.
+But as the number of cores increases, the work can be efficiently broken up into smaller pieces and doled out to all the cores.
+With more pieces of work, it becomes easier to balance the load.
 Note that both problems scale well even up to 1024 processors.
 We achieve a speedup of $443$ in the game between humans and up to $536$ in the game between computers.
-Figures~\ref{prob4} and~\ref{prob5} show the plots of the running times for the best grainsize for each core setting.
+Figures~\ref{prob4} shows the plots of the running times for the best grainsize for each core setting.
 These plots show that the running time continues to decrease as the number of cores increases.
@@
 This is a fundamental limitation in alpha-beta search.
-
-For partially observable games, information set generation represents a different variety of tree search that does not suffer from speculative loss.
-All of the nodes that are visited in parallel would have also been visited in a sequential run.
-The key challenge for parallelizing information set generation is load balancing: some parts of the tree will be pruned at a higher level than others, resulting in a wide variation in cost for each subtree.
-The {\sc Charm++/ParSSSE} framework is perfectly suited to handle the load balancing issues that this problem naturally presents.
-
-Note also that information set generation is just one subroutine needed to reason about partially observable games.
-Once plausible game histories have been found, an agent may simulate some sequences of moves up to previous moves made by opponents and then recursively call the information set generation procedure from the perspective of the opponent.
-This results in many parallel applications of information set generation.
-Furthermore, given a node in an information set, it is common for an agent to perform forward search on that node (e.g., using MCTS).
-With these two additional levels of parallelism on top of the information set generation problem, it is conceivable that the problem of playing partially observable games could scale efficiently to millions of processors or more.
-\begin{table}[tphb]
+\begin{table}[ht]
 \centering
@@
 \end{table}
+For partially observable games, information set generation represents a different variety of tree search that does not suffer from speculative loss.
+All of the nodes that are visited in parallel would have also been visited in a sequential run.
+The key challenge for parallelizing information set generation is load balancing: some parts of the tree will be pruned at a higher level than others, resulting in a wide variation in cost for each subtree.
+The {\sc Charm++/ParSSSE} framework is perfectly suited to handle the load balancing issues that this problem naturally presents.
+
+Note also that information set generation is just one subroutine needed to reason about partially observable games.
+Once plausible game histories have been found, an agent may simulate some sequences of moves up to previous moves made by opponents and then recursively call the information set generation procedure from the perspective of the opponent.
+This results in many parallel applications of information set generation.
+Furthermore, given a node in an information set, it is common for an agent to perform forward search on that node (e.g., using MCTS).
+With these two additional levels of parallelism on top of the information set generation problem, it is conceivable that the problem of playing partially observable games could scale efficiently to millions of processors or more.
 
 \Section{Related Work}
+\label{Related}
 An overview of the theory of imperfect information games and other basic topics in game theory can be found

### 2012-07-11 sbac/paper.tex
Changed the style according to the conf web site guidelines.. They must have changed the guidelines recently

@@
-\documentclass[times, 10pt,twocolumn]{article} 
-\usepackage{latex8}
-\usepackage{times}
+%\documentclass[times, 10pt,twocolumn]{article}
+%\documentclass[conference]{IEEEtran}
+\documentclass[10pt, conference, compsocconf]{IEEEtran} 
+%\usepackage{latex8}
+%\usepackage{times}
 \usepackage{url,hyperref}
@@
 
-\Section{Introduction}
+\section{Introduction}
 In imperfect information games, players do not have access to full knowledge of
@@
 %\Section{Information Set Generation}
-\Section{Background: Kriegspiel}
+\section{Background: Kriegspiel}
 \label{Background}
@@
 
-\Section{Information Set Generation}
+\section{Information Set Generation}
 \label{info}
@@
 
-\Section{Parallel Implementation using Charm++ and ParSSSE}\label{ParSSSE}
+\section{Parallel Implementation using Charm++ and ParSSSE}\label{ParSSSE}
 We built our search engine on top of the {\sc Charm++} run-time
@@
 
-\Section{Performance Results}\label{Results}
+\section{Performance Results}\label{Results}
 
@@
 The performance results reported in
-this section were obtained by running our program on Blueprint with up to 1024 processors. TODO: Blueprint H/w or citation
+this section were obtained by running our program on Blueprint with up to 1024 processors.
+This system was composed of 120 POWER5+ nodes, each having 16 cores and 64GB of memory.
+The system was decomissioned in June 2011 and is no longer in operation.
 
@@
 \subfigure[Depth 8]{
-\includegraphics[width=2.2in]{plots/dep8.pdf}
+\includegraphics[width=2.1in]{plots/dep8.pdf}
 \label{dep8}
@@
 \subfigure[Depth 10]{
-\includegraphics[width=2.2in]{plots/dep10.pdf}
+\includegraphics[width=2.1in]{plots/dep10.pdf}
 \label{dep10}
@@
 \subfigure[Depth 12]{
-\includegraphics[width=2.2in]{plots/dep12.pdf}
+\includegraphics[width=2.1in]{plots/dep12.pdf}
 \label{dep12}
@@
 
-\Section{Related Work}
+\section{Related Work}
 \label{Related}
@@
 
-\Section{Conclusion and Future Work}
+\section{Conclusion and Future Work}
 We have shown that information set generation for kriegspiel can be efficiently parallelized.  The opportunities for
@@
 
-\bibliographystyle{latex8}
+\bibliographystyle{IEEEtran}
 \bibliography{paper}

### 2012-07-11 sbac/paper.tex
Several minor changes to the main text of the paper. Still need to work on the section about Information Set Generation.

--- 2012-06-28 sbac/paper.tex
@@
 \begin{abstract}
-Information Set Generation is the identification of the set of paths in an imperfect information game tree that are
-consistent with a player's observations.  The ability to reason about the possible game history is critical to the
-performance of game playing agents.  ISG represent a class of combinatorial search problems which is computationaly intensive but difficult to efficiently parallelize.
-
-In this paper, we address the parallelization of information set generation in the context
-of Kriegspiel (partially observable chess).  We implement the algorithm on top of a general purpose combinatorial search
-engine and discuss its performance under different load balancing strategies and with different grain-size parameters.
-We achieve speedups of over 500 on 1024 processors, far exceeding previous scalability results for game tree search applications.
+Information Set Generation is the identification of the set of paths in an
+imperfect information game tree that are consistent with a player's
+observations.  The ability to reason about the possible game history is
+critical to the performance of game-playing agents.  ISG represent a class of
+combinatorial search problems which is computationaly intensive but difficult
+to efficiently parallelize.
+
+In this paper, we address the parallelization of information set generation in
+the context of Kriegspiel (partially observable chess).  We implement the
+algorithm on top of a general purpose combinatorial search engine and discuss
+its performance under different load balancing strategies and with different
+grainsize parameters.  We achieve speedups of over 500 on 1,024 processors, far
+exceeding previous scalability results for game tree search applications.
 \end{abstract}
@@
 can use to identify nodes in the game tree that are consistent with their
-observations~\cite{richards12reasoning}.  Because the operation is relatively
-expensive and may be repeatedly applied by a decision-making agent, its
-efficiency is a key factor in an agent's overall performance.
-Information set generation is not the same problem as solving a game tree, although it is arguably a necessary subroutine for quality game play in imperfect information games.
-
-The term {\em belief state} is sometimes used interchangeably with information set to refer to a probability
-distribution over possible worlds.  The latter term comes from the game theory community and is preferred here.  A node
-in a game tree denotes not only the current state of the game, but also uniquely defines a path from the initial state
-or root node.  Thus, a game tree node implicitly encodes not only the current state of the game but also the complete
-history of all decisions made by all players up to that point in the game, including the outcome of any chance moves
-such as dice rolls or card shuffling.  Knowing one's information set means knowing all possible game histories.
-
-For many specific games, solving the information set generation problem is trivial.  For example, in a poker game, it is easy to
-see that the unseen cards held by a player's opponents may be any permutation of the cards not seen by that player
-(i.e., hole cards and any revealed community cards).  After the betting rounds, it would not be reasonable to assume
-that each of these permutations of unseen cards is equally probable, as betting decisions made by the players up to that
-point would be affected by the quality of those players' cards.  But the set of {\em possible} hands for all of the
-opponents is easy to conceive and enumerate.
-
-In the game of Kriegspiel, information set generation is much more complicated.  The player knows the opponent's position with
-certainty when the game begins, but after the initial move there may be varying levels of uncertainty about the location
-of the opponent's pieces.  Unlike poker, it is not possible to simply permute all of the opponent's possible pieces over
-all of the squares not occupied by the player's own pieces.  A configuration of pieces for the opponent is only valid if
-it can be reached by a legal sequence of moves.  In the general case, finding the nodes in an information set is a
-combinatorial search problem.
+observations~\cite{richards12information}.  Because the operation is
+computationally intensive and may be repeatedly applied by a decision-making
+agent, its efficiency is a key factor in an agent's overall performance.
+
+The term {\em belief state} is sometimes used interchangeably with information
+set to refer to a probability distribution over possible worlds.  The latter
+term comes from the game theory community and is preferred here.  A node in a
+game tree denotes not only the current state of the game, but also uniquely
+defines a path from the initial state or root node.  Thus, a game tree node
+implicitly encodes not only the current state of the game but also the complete
+history of all decisions made by all players up to that point in the game,
+including the outcome of any chance moves such as dice rolls or card shuffling.
+Knowing one's information set means knowing all possible game histories.
+
+For many specific games, solving the information set generation problem is
+trivial.  For example, in a Poker game, the unseen cards held by a player's
+opponents may be any permutation of the cards not seen by that player (i.e.,
+hole cards and any revealed community cards).  After the betting rounds, it
+would not be reasonable to assume that each of these permutations of unseen
+cards is equally probable, as betting decisions made by the players up to that
+point would be affected by the quality of those players' cards.  But the set of
+{\em possible} hands for all of the opponents is easy to conceive and
+enumerate.
+
+In the game of Kriegspiel, information set generation is much more complicated.
+After the first move by the opponent, a player faces uncertainty about the
+location of the opponent's pieces.  Unlike Poker, it is not possible to simply
+permute all of the opponent's possible pieces over all of the squares not
+occupied by the player's own pieces.  A configuration of pieces for the
+opponent is only valid if it can be reached by a legal sequence of moves.  In
+the general case, finding the nodes in an information set is a combinatorial
+search problem.
 
 In this work, we show that ISG for Kriegspiel can be efficiently parallelized.
-The primary contributions of this work are the following.
-We describe an approach to efficient parallelization of an application which is not typical of high performance computing domain.
-We present scalability results for our parallel implementation of information set generation algorithm for the game of Kriegspiel using real game problem instances.
-Further, we demonstrate the effect of load balancing and grain size on parallel performance of our solution.
-
-The paper is organized as follows.
-In Section~\ref{Background}, we review necessary background information, including the rules of kriegspiel and the nature of the information set generation problem.
-In Section ~\ref{ParSSSE}, we describe the {\sc Charm++} Parallel State Space Search Engine (ParSSSE) platform.
-We then provide our experimental results in Section~\ref{Results}.
-Related work is discussed in Section~\ref{Related}.
-We conclude with a discussion of future work.
+The primary contributions of this work are the following. We describe an
+approach to efficient parallelization of an application which is not typical of
+high performance computing domain. We present scalability results for our
+parallel implementation of an information set generation algorithm for the game
+of Kriegspiel using problem instances from real games. Further, we demonstrate
+the effects of load balancing strategy and grainsize on parallel performance.
+
+The paper is organized as follows.  In Section~\ref{Background}, we review
+necessary background material, including the rules of Kriegspiel and the nature
+of the information set generation problem.  In Section ~\ref{ParSSSE}, we
+describe the {\sc Charm++} Parallel State Space Search Engine (ParSSSE)
+platform.  We then provide our experimental results in Section~\ref{Results}.
+Related work is discussed in Section~\ref{Related}.  We conclude with a
+discussion of future work.
 
 %\Section{Information Set Generation}
-\Section{Background: Kriegspiel}
-Information Set Generation is a widely adopted operation used in game trees search.
-Here, we present our techniques and results in the context of a particular game - Kriegspiel.
-However, the techniques and framework presented in this paper is general and can be applied to other imperfect games for generating information set.
+\Section{Kriegspiel}
+Information Set Generation is the process of identifying the nodes in an
+imperfect information game tree that are consistent with a player's
+observations~\cite{richards12information}. Here, we present our techniques and
+results in the context of a particular game: Kriegspiel~\cite{li94chess}.
+However, the techniques and framework presented in this paper are general and
+can be applied to ISG in other imperfect information games.
+
 %\Section{Background: Kriegspiel}
-Kriegspiel, or partially observable chess,  was invented by Henry Michael Temple in 1899.  Historically, the
-game has required three parties: two competing players and one referee.  Today,
-players often compete over a computer network, with the role of the referee
-being automated.  One player sees and controls only the white pieces; the
-other, black.  The referee track of both sets of pieces.  The players alternate
-moves as in a regular chess game, but instead of openly declaring their moves,
-they secretly communicate move requests to the referee.  The referee, who knows
-the location of all pieces, checks the legality of the move.
-
-If the move is legal, the referee makes executes the move and announces that a
+Kriegspiel, or partially observable chess,  was invented by Henry Michael
+Temple in 1899.  Historically, the game has required three parties: two
+competing players and one referee.  Today, players often compete over a
+computer network, with the role of the referee being automated.  One player
+sees and controls only the white pieces; the other, black.  The referee tracks
+both sets of pieces.  Instead of openly declaring their moves, players secretly
+communicate move requests to the referee.  The referee, who knows the location
+of all pieces, assesses the legality of the move according to the standard
+rules of chess.
+
+If the move is legal, the referee executes the move and announces that a
 legal move has been made and that it is the other player's turn to move.  If
@@
 
-The referee makes other announcements besides declaring move illegal, and there
-are several variations of the game that differ only in the nature of these
-announcements.  All declarations by the referee are made to both players.  In
+All declarations by the referee are made to both players.  In
 particular, a player will hear and know if her opponent has attempted an
@@
 
-The referee announces when a player is in check. In the case of the capture,
-the referee announces the location of the piece that was captured but not its
-identity.  Because there is no penalty for making illegal moves, it is often to
-a player's advantage to attempt to make moves which are likely to be illegal,
-as hearing such declarations from the referee can provide important information
-to a player about the location of the opponent's pieces.  In particular, since
-a diagonal move by a pawn is legal only the case where such a move would
-capture an opponent's piece, it could potentially be profitable to attempt all
-or many pawn captures on every turn.  In order to speed the game up, the
-referee declares, at the beginning of a player's turn, all locations where a
-pawn capture is possible.
+The referee makes other announcements besides declaring move illegal, and there
+are several variations of the game that differ only in the nature of these
+announcements.  In our implementation, the referee announces when a player is
+in check, when a piece has been captured, and whenever a pawn may capture an
+opponent's piece.  In the case of a capture, the referee announces the square
+in which the capture occurs but not the identity of the captured piece.
 
@@
 %\label{listing}
-%\caption{Listing of a kriegspiel game resulting in checkmate by white after 10 moves.}
+%\caption{Listing of a Kriegspiel game resulting in checkmate by white after 10 moves.}
 %\end{figure}
@@
 Figure~\ref{listing} and show a full listing and the associated pictorial progression of an actual
-kriegspiel game~\cite{li94chess}.  We have adapted the notation used by Wolfe for Berkeley Kriegspiel.  Odd (resp. even) moves are for white (resp. black).  The actual move is shown
+Kriegspiel game~\cite{li94chess}.  We have adapted the notation used by Wolfe for Berkeley Kriegspiel.  Odd (resp. even) moves are for white (resp. black).  The actual move is shown
 first followed by a list of referee announcements.  Captures announcements are prefixed with an X and give the location
@@
 \vspace{-0.1in}
-\caption{Listing of a kriegspiel game resulting in checkmate by white after 10 moves.}
+\caption{Listing of a Kriegspiel game resulting in checkmate by white after 10 moves.}
 \label{filteredlisting}
@@
 %\end{tabular}
-%\caption{Solution counts and running times for a sample kriegspiel game}
+%\caption{Solution counts and running times for a sample Kriegspiel game}
 %\label{bothtimes}
@@
 %\end{tabular}
-%\caption{Solution counts and running times for a sample kriegspiel game}
+%\caption{Solution counts and running times for a sample Kriegspiel game}
 %\label{blacktimes}
@@
 (ISG) can be formulated as a combinatorial search problem.  The input to the algorithm is a list
-of observations for one player in a kriegspiel game.
+of observations for one player in a Kriegspiel game.
 
@@
 \Section{Parallel Implementation using Charm++ and ParSSSE}\label{ParSSSE}
-We built our search engine on top of the {\sc Charm++} run-time
+We implemented our search engine on top of the {\sc Charm++} run-time
 system for parallel processing~\cite{kale93charm,kale09charm}.  {\sc Charm++}
@@
 and {\sc OpenMP}, the {\sc Charm++} language is based on object-oriented
-programming principles.  The programmer is responsible to decompose its problem
+programming principles.  The programmer is responsible to decompose a problem
 into a collection of objects that represent elements of work to be done.
-Typically the number of work objects far exceeds the number of processing
-elements (cores) available.  These objects are defined so as to be able to be
-migrated between processors {\em during execution}, in order to improve the
-overall load balancing.
-
-Work objects, known as {\em chares}, communicate with each other via
-asynchronous message-passing.  Objects can invoke {\em entry methods} on other
+Typically the number of these work objects, called {\em chares}, far exceeds
+the number of processing elements (cores) available.  Chares are defined
+so as to be able to be migrated between processors {\em during execution}, in
+order to improve the overall load balancing.
+
+Chares communicate with each other via
+asynchronous message passing.  They can invoke {\em entry methods} on other
 objects, which are executed according to a schedule set by the {\sc Charm++}
@@
 it currently holds.  A corresponding {\em outgoing queue} collects messages
-generated by objects on the processor that need to be sent to objects assigned
+generated by objects on the processor that need to be sent to chares assigned
 to other processors.
 
-Some work tasks may inherently require more computation than others, but the
-amount of computation associated with each piece of work is not always known
-before execution.  In tree search, for example, a piece of work may be defined
-as the computation of a minimax value on a node in the tree using alpha-beta
-pruning.  A natural decomposition is to assign different nodes to different
-processors for evaluation.  Because some nodes will be pruned at shallower
-levels of the tree than others, the amount of work performed at each subtree can
-vary considerably.  In a traditional parallel framework, this would result in
-many processors idling while a small number of processors execute the larger
-workloads. By allowing the {\sc Charm++} runtime system to schedule a large
-number of small pieces of work, better adaptive load balancing is achieved.
+Some chares may inherently require more computation than others, but the amount
+of computation associated with each piece of work is not always known before
+execution.  In tree search, for example, a piece of work may be defined as the
+computation of a minimax value on a node in the tree using alpha-beta
+pruning~\cite{knuth75analysis,baudet78analysis}.  A natural decomposition is to
+assign different nodes to different processors for evaluation.  Because some
+nodes will be pruned at shallower levels of the tree than others, the amount of
+work performed at each subtree can vary considerably.  In a traditional
+parallel framework, this would result in many processors idling while a small
+number of processors execute the larger workloads. By allowing the {\sc
+Charm++} runtime system to schedule a large number of small pieces of work,
+better adaptive load balancing is achieved.
 
@@
 variety of graph-based and tree-based search problems~\cite{sun11adaptive}. In
-{\sc ParSSE} the fundamental unit of work is the processing of a node.
+{\sc ParSSSE} the fundamental unit of work is the processing of a node.
 Generally, this means identifying whether the node corresponds to a solution
@@
 the system gets bogged down by the overhead of creating and scheduling tasks
-instead of actually doing them.  {\sc ParSSE} supports an adaptive grainsize
+instead of actually doing them.  {\sc ParSSSE} supports adaptive grainsize
 management, so that the threshold for sequential processing can be changed
@@
 
-{\sc ParSSE} currently supports two different load balancing strategies.  The
+{\sc ParSSSE} currently supports two load balancing strategies.  The
 first is the randomized strategy.  In this approach, each newly created object
@@
 from parallel overhead.  For a program running on $p$ processors, the
-probability that a piece of work is assigned to a processor other than the
+probability that a chare is assigned to a processor other than the
 one that created it is $1-1/p$.  As the number of processors gets large, this
@@
 
-A second load balancing strategy is known as {\em work stealing}.  In this
+The second load balancing strategy is {\em work stealing}.  In this
 approach, all new tasks are inserted in the message queue of the processor that
@@
 because victims respond by sending the highest priority work, the most
-important tasks are more likely to get done sooner.
+important tasks are scheduled sooner.
 
@@
 \Section{Performance Results}\label{Results}
-In order to understand the effect of various parameters such as grain size and load balancing, we first evaluated the scalability of our implementation using a randomly selected problem instance.
-Next, we show that our implementation performs equally well for instances from actual games.
+In order to understand the effect of various parameters such as grainsize and
+load balancing strategy, we first evaluated the scalability of our
+implementation using a randomly selected problem instance. Next, we show that
+our implementation performs equally well for instances from actual games.
+
 \subsection{Analysis using Random Instance} 
-The performance results reported in
-this section were obtained by running our program on Blueprint with up to 1024 processors. TODO: Blueprint H/w or citation
-
-Table~\ref{speedups} shows the performance results for our algorithm on a randomly selected problem instance in which
-the size of the information set was about 4.5 million.  This instance is not necessarily representative of all games or
-states of the game, but it was not cherry-picked in any way. The table shows that the problem is indeed highly
-parallizable.  The speedups are nearly linear for up  to 32 processors, and even on 1024 processors, we get a good
-speedup.  In this particular instance, the sequential algorithm takes almost 25 minutes to run.  But when parallelized
-on 1024 processors---perfectly reasonable resource usage on modern hardware---the search takes less than three seconds.
-This is a speedup of 576 times!  It might not be reasonable to use the algorithm at all if it takes 25 or more to run
-for some turns.  But if the cost is only a few seconds, it becomes entirely feasible.
+The performance results reported in this section were obtained by running our
+program on Blueprint with up to 1,024 processors. TODO: Blueprint H/w or
+citation
+
+Table~\ref{speedups} shows the performance results for our algorithm on a
+randomly selected problem instance in which the size of the information set was
+about 4.5 million.  This instance is not necessarily representative of all
+games or states of the game, but it was not cherry-picked in any way. The table
+shows that the problem is indeed highly parallizable. The speedups are nearly
+linear for up to 32 processors, and the speedups continue to increase even up
+to 1,024 processors. In this particular instance, the sequential algorithm
+takes almost 25 minutes to run.  But when parallelized on 1,024
+processors---perfectly reasonable resource usage on modern hardware---the
+search takes less than three seconds.  This is a speedup of 576 times! In a
+tournament where each player is allotted an average of five minutes per move,
+an ISG operation that requires 25 minutes or more would be prohibitively
+expensive.  On the other hand, if the operation can be done in under three
+seconds, ISG can be used as a subroutine in the decision-making process
+(see~\cite{parker05game}).
 
@@
 \vspace{-0.1in}
-\caption{Speedups for a randomly selected problem instance on 1--1024 processors.  The input observations were from the
-10th ply of the game tree and the size of the information set in this instance is 4,495,121.}
+
+\caption{Speedups for a randomly selected problem instance on 1--1,024
+processors.  The input observations were from the 10th ply of the game tree and
+the size of the information set in this instance is 4,495,121.}
+
 \label{speedups}
@@
 \vspace{-0.1in}
-\caption{Effect of load balancing strategy}
+\caption{Effect of load-balancing strategy}
 \label{3schemes}
@@
 
-In our next round of experiments, we set out to compare a few different load-balancing strategies for the problem.  We
-used the random, work stealing, and neighborhood balancing strategies that are available in the general purpose search
-engine.  These results are shown in Figure~\ref{3schemes}.  Our results show that there was actually very little
-variation in performance across the different strategies.  We did notice that across all of our experiments, the work
-stealing approach appeared to be consistently---if only marginally---better than the others.  Our hypothesis here is
-that perhaps the nodes are rarely starved for work.  In other words, perhaps the very nature of the problem means that
-the load is naturally well-balanced and that the load-balancer rarely needs to do a significant amount of
-redistribution.  This is probably also related to the fact that we are doing an all-solutions search.
+In our next round of experiments, we set out to compare load-balancing
+strategies for the problem.  We used the random, work stealing, and
+neighborhood balancing strategies that are available in the general purpose
+search engine.  These results are shown in Figure~\ref{3schemes}.  Our results
+show that there was actually little variation in performance across the
+different strategies.  Across all of our experiments, the work stealing
+approach appeared to be consistently---if only marginally---better than the
+others.  Our hypothesis here is that the nodes are rarely starved for work.  In
+other words, perhaps the very nature of the problem means that the load is
+naturally well-balanced and that the load-balancer rarely needs to do a
+significant amount of redistribution.  This is probably also related to the
+fact that we are doing an all-solutions search.
 
@@
 \vspace{-0.1in}
-\caption{(Top) Log-log plot of running time for three different sizes of problem as the number of cores used increases.
-(Bottom) Same data with a linear y-axis for the running time (to help see the Iso-efficiency.)}
+
+\caption{(Top) Log-log plot of running time for three different problem sizes
+as the number of cores used increases.  (Bottom) Same data with a linear y-axis
+for the running time (to help see the isoefficiency.)}
+
 \vspace{-0.2in}
@@
 We next study the variation in the efficiency of the parallelization of the
-algorithm with {\em size} of the problem?  To test this, we varied the number of plies in the test instance from
-8 to 12. We used the work stealing load balancing strategy.  The results are shown in Figure~\ref{3depths}.   In the easiest case (depth 8), there isn't
-enough additional parallel work to justify the jump from 256 to 1024 processors, and performance actually degrades.
-Other than that, however, we see consistent speedups as we increase the number of processors.  The plot on the right helps to
-illustrate the good isoefficiency of the problem.  As the size of the problem increases, the amount of parallelizable
-work available also increases, and therefore the steepness of the speed curve is higher for larger problem sizes, as the
-number of processors increases.
-
-\subsubsection{Effect of Grain Size}
+algorithm as size of the problem increases.  To test this, we varied the number
+of plies in the test instance from 8 to 12. We used the work stealing
+load-balancing strategy.
+
+The results are shown in Figure~\ref{3depths}.  In the easiest case (depth 8),
+there is not enough additional parallel work to justify the jump from 256 to
+1,024 processors, and performance actually degrades.  Aside from this anomaly,
+however, we see consistent speedups as we increase the number of processors.
+The plot on the right helps to illustrate the good isoefficiency of the
+problem.  As the size of the problem increases, the amount of parallelizable
+work available also increases, and therefore the steepness of the speed curve
+is higher for larger problem sizes, as the number of processors increases.
+
+\subsubsection{Effect of Grainsize}
 
@@
 \vspace{-0.1in}
-\caption{Grain Size Control for different problem sizes}
+\caption{Grainsize Control for different problem sizes}
 %\vspace{-0.1in}
@@
 
-Next, we evaluate the effect of grainsize on the problem execution time.  The results are shown in
-Tables~\ref{dep8}--~\ref{dep12}.  It is clear that there is a ``sweet spot'' for each problem size.  (At least it is
-clear for this particular instance.  It is possible that there would be different behavior on different problem
-instances.)  When the grainsize is too small, the overall computation is hampered by excessive parallel overhead
-(communication, queueing of nodes to expand, load balancing operations).  On the other hand, when the grain size is too
-large, it becomes too difficult to ensure that all processors are busy and processors are not idling with no work to do
-while a small number of processors are busy doing large-sized chunks of work.
-
-In our experiments, the optimal grainsize is clearly related to the size of the problem, the depth of the search tree.
-The optimal sequential threshold seems to be $2--4$ less than the problem depth.  Since the depth of all solutions is
-the same for each problem instance, this means that the number of levels that should be searched sequentially is fairly
-consistent across problem sizes.  We noted that the number of solutions found---in other words, the size of the
-information sets---is typically fairly large with respect to the total number of nodes expanded.  (We don't have hard
-figures, but we estimate that it was often $30-40\%$).  This high density of solutions limits the variability that can
-be present in the amount of work to be done at the nodes representing the last few levels of the tree.  This means that
-the best grainsize is also likely to be fairly predictable.
+Next, we evaluate the effect of grainsize on execution time.  The results are
+shown in Tables~\ref{dep8}--~\ref{dep12}.  It is clear that there is a ``sweet
+spot'' for each problem size.  (At least it is clear for this particular
+instance.  It is possible that there would be different behavior on different
+problem instances.)  When the grainsize is too small, the overall computation
+is hampered by excessive parallel overhead (communication, queueing of nodes to
+expand, load balancing operations).  On the other hand, when the grainsize is
+too large, it becomes too difficult to ensure that some processors are not
+idling with no work to do while a small number of processors are busy doing
+large chunks of work.
+
+In our experiments, the optimal grainsize is clearly related to the size of the
+problem, the depth of the search tree.  The optimal sequential threshold seems
+to be $2-4$ less than the problem depth.  Since the depth of all solutions is
+the same for each problem instance, this means that the number of levels that
+should be searched sequentially is fairly consistent across problem sizes.  We
+noted that the number of solutions found---in other words, the size of the
+information sets---is typically fairly large with respect to the total number
+of nodes expanded.  (We estimate that it was often $30-40\%$).  This high
+density of solutions limits the variability that can be present in the amount
+of work to be done at the nodes representing the last few levels of the tree.
+This means that the best grainsize is also likely to be fairly predictable,
+which is encouraging.
 
@@
  \centering \includegraphics[viewport= 280 200 300 550, scale=0.35]{images/KriegspielProblem4.pdf} %
-\caption{Kriegspiel position from Virgil-David 1969 with information set size 764,209.}
+
+\caption{Kriegspiel position from Virgil-David 1969 with information set size
+764,209.}
+
 \label{prob4} 
@@
  \centering \includegraphics[viewport= 280 200 300 550, scale=0.35]{images/KriegspielProblem5.pdf} %
-\caption{Kriegspiel position from Game 6 of Kbot-Darkboard at the 2006 Computer Chess championship.  Information set size is 5,099,257.}
+
+\caption{Kriegspiel position from Game 6 of Kbot-Darkboard at the 2006 Computer
+Chess championship.  Information set size is 5,099,257.}
+
 \label{prob5} 
@@
 
-We also evaluated our information set generation routine on 40 positions from actual kriespiel games, including human vs.
-human and computer vs.
-computer.
-Many of these were trivial to solve in a few seconds or less on a desktop computer.
-For these experiments, we used a real supercomputer - Ranger at TACC~\cite{ranger}.
-Ranger has 16-way SMP nodes containing AMD Opteron processors.
-The nodes are interconnected using Infiniband technology in a fully-CLOS topology providing a 1GBps point-to-point bandwidth.
-The running times and speedups for two of the more challenging instances on Ranger are shown in Tables~\ref{prob4}--\ref{tab:prob5speedups}.
-These tables show the results for runs on one to $1024$ cores.
-Each column corresponds to a different parameter.
-In this case, the grainsize is the number of levels of the search tree that are searched in parallel.
-Up to that threshold, each node that is generated spawns a new piece of work, which the {\sc Charm++} runtime schedules to execute on one of the processing elements according to its load balancing strategy.
-Above the threshold the entire subtree is searched to completion on a single processor, with no additional parallel overhead.
-
-Note that the optimal grainsize varies for the different numbers of cores and the pattern is relatively consistent across problem instances.
-The optimal grainsize increases as the number of cores increases.
-This makes sense, because as this parameter grows, the number of distinct pieces of work increases.
-When the number of pieces of work per processor is high, much of the overhead of parallelization is wasted.
-But as the number of cores increases, the work can be efficiently broken up into smaller pieces and doled out to all the cores.
-With more pieces of work, it becomes easier to balance the load.
+We also evaluated our information set generation routine on 40 positions from
+actual kriespiel games, including human vs. human and computer vs. computer
+instances.  Many of these were trivial to solve in a few seconds or less on a
+desktop computer.  For these experiments, we used the Ranger supercomputer
+at TACC~\cite{ranger}. Ranger has 16-way SMP nodes containing AMD Opteron
+processors. The nodes are interconnected using Infiniband technology in a
+fully-CLOS topology providing a 1GBps point-to-point bandwidth.
+
+The running times and speedups for two of the more challenging instances on
+Ranger are shown in Tables~\ref{prob4}--\ref{tab:prob5speedups}.   These tables
+show the results for runs on one to $1,024$ cores.  Each column corresponds to
+a different grainsize parameter.  In this case, the grainsize is the number of
+levels of the search tree that are searched in parallel.  Up to that threshold,
+each node that is generated spawns a new piece of work, which the {\sc Charm++}
+runtime schedules to execute on one of the processing elements according to its
+load balancing strategy.  Above the threshold the entire subtree is searched to
+completion on a single processor, with no additional parallel overhead.
+
+Note that the optimal grainsize varies for the different numbers of cores and
+the pattern is consistent across problem instances.  The optimal
+grainsize increases as the number of cores increases.  This makes sense,
+because as this parameter grows, the number of distinct pieces of work
+increases.  When the number of chares per processor is high, much of
+the overhead of parallelization is wasted.  But as the number of cores
+increases, the work can be efficiently broken up into smaller pieces and doled
+out to all the cores.  With more pieces of work, it becomes easier to balance
+the load.
 
@@
 \end{tabular}
-\caption{Running times in seconds and speedups (after $|$), to find full information set of size 764,209 at position 10 of Virgil-David 1969.
-Optimal grainsize (G) for each number of cores (PE) is shown in bold}
+
+\caption{Running times in seconds and speedups (after $|$), to find full
+information set of size 764,209 at position 10 of Virgil-David 1969. Optimal
+grainsize (G) for each number of cores (PE) is shown in bold.}
+
 \vspace{-0.2in}
@@
 \end{tabular}
-\caption{Running times, in seconds and speedups (after $|$), to find full information set of size 5,099,257 at position 10 of Game 6 Kbot-Darkboard 2006.
+
+\caption{Running times, in seconds and speedups (after $|$), to find full
+information set of size 5,099,257 at position 10 of Game 6 Kbot-Darkboard 2006.
 Optimal grainsize for each number of cores is shown in bold.}
+
 \vspace{-0.2in}
@@
 
-Note that both problems scale well even up to 1024 processors.
-We achieve a speedup of $443$ in the game between humans and up to $536$ in the game between computers.
-Figures~\ref{prob4} and~\ref{prob5} show the plots of the running times for the best grainsize for each core setting.
-These plots show that the running time continues to decrease as the number of cores increases.
-
-Our results compare favorably to previous work in parallel game tree search, shown in Table~\ref{bestspeedups}.
-The previous best parallelization results were for parallelizing alpha-beta pruning.
-Ferguson and Korf achieved a speedup of 12 on 32 cores in the game of Othello~\cite{ferguson88distributed}.
-In the game of chess, Felton and Otto showed a speedup of 100 on 256 cores~\cite{felten88highly}.
-This limited scalability is due to speculative loss.
-Parts of the game tree that are searched during the parallel execution would have been pruned had the algorithm run sequentially.
-This is a fundamental limitation in alpha-beta search.
-
-For partially observable games, information set generation represents a different variety of tree search that does not suffer from speculative loss.
-All of the nodes that are visited in parallel would have also been visited in a sequential run.
-The key challenge for parallelizing information set generation is load balancing: some parts of the tree will be pruned at a higher level than others, resulting in a wide variation in cost for each subtree.
-The {\sc Charm++/ParSSSE} framework is perfectly suited to handle the load balancing issues that this problem naturally presents.
-
-Note also that information set generation is just one subroutine needed to reason about partially observable games.
-Once plausible game histories have been found, an agent may simulate some sequences of moves up to previous moves made by opponents and then recursively call the information set generation procedure from the perspective of the opponent.
-This results in many parallel applications of information set generation.
-Furthermore, given a node in an information set, it is common for an agent to perform forward search on that node (e.g., using MCTS).
-With these two additional levels of parallelism on top of the information set generation problem, it is conceivable that the problem of playing partially observable games could scale efficiently to millions of processors or more.
+Note that both problems scale well even up to 1,024 processors.  We achieve a
+speedup of $443$ in the game between humans and up to $536$ in the game between
+computers.  Figures~\ref{prob4} and~\ref{prob5} show the plots of the running
+times for the best grainsize for each core setting.  These plots show that the
+running time continues to decrease as the number of cores increases.
+
+Our results compare favorably to previous work in parallel game tree search,
+shown in Table~\ref{bestspeedups}.  The previous best parallelization results
+were for parallelizing alpha-beta pruning.  Ferguson and Korf achieved a
+speedup of 12 on 32 cores in the game of Othello~\cite{ferguson88distributed}.
+In the game of chess, Felton and Otto showed a speedup of 100 on 256
+cores~\cite{felten88highly}.  This limited scalability is due to speculative
+loss.  Parts of the game tree that are searched during the parallel execution
+would have been pruned had the algorithm run sequentially.  This is a
+fundamental limitation in alpha-beta search.
+
+For partially observable games, information set generation represents a
+different variety of tree search that does not suffer from speculative loss.
+All of the nodes that are visited in parallel also would have been visited in a
+sequential run.  The key challenge for parallelizing information set generation
+is load balancing: some parts of the tree will be pruned at a higher level than
+others, resulting in a wide variation in cost for each subtree.  The {\sc
+Charm++/ParSSSE} framework is perfectly suited to handle the load balancing
+issues that this problem naturally presents.
+
+Note also that information set generation is just one subroutine needed to
+reason about partially observable games.  Once plausible game histories have
+been found, an agent may simulate some sequences of moves up to previous moves
+made by opponents and then recursively call the ISG procedure from the
+perspective of the opponent.  This results in many parallel applications of
+ISG. Furthermore, given a node in an information set, it is common for an agent
+to perform forward search on that node (e.g., using MCTS).  All of these
+forward search operation can be performed independently in parallel. With these
+two additional levels of parallelism on top of the information set generation
+problem, it is likely that the problem of playing partially observable
+games could scale efficiently to millions of processors or more.
+
 \begin{table}[tphb]
@@
 \Section{Related Work}
-An overview of the theory of imperfect information games and other basic topics in game theory can be found
-in~\cite{kuhn03lectures} and~\cite{kuhn97classics}.  A technical explanation of the concept of information sets can be
-found in~\cite{gilpin07algorithms}.  Koller \etal, showed that a Nash
-equilibrium can be found in time that is polynomial in the number of nodes in the game tree~\cite{koller94fast}.  In fact, they found
-that it took about as much time to generate the full game tree as it did to actually solve it.  Unfortunately, this is
-feasible only for small games.  In larger games where theoretically optimal strategies cannot be computed efficiently,
-computer systems tend to utilize the concept of information sets or belief states in some form or another.  A common
-strategy is to estimate the value of a move by averaging the estimated value of the associated descendant for each node
-in the current information set (or a random sample from the set).  While this sort of ``averaging over clairvoyance''
-{\em can} lead to very poor decisions, it is nevertheless a reasonable strategy in some games (e.g.,
-Scrabble~\cite{sheppard02world}).
-
-In this work, we have argued for an information set generation approach to kriegspiel, but other authors have approached
-the problem differently.  Parker \etal, consider the problem of sampling from belief states in a variety of large game
-trees, including kriegspiel~\cite{parker05game}.  They show that some performance gains can be achieved by only
-approximate sampling from the belief state.  Instead of generating the full information set, or even sampling from it,
-they sample from possible positions for the current state (i.e., sampling from the belief state) by matching only the
-most recent observations.  The motivation for this approach seems to be that it would not be feasible to sample from
-the full information set, because of the computational expense involved.  And indeed, our experiments bear this out-- it
-can indeed be expensive to produce the full information set.  For the experiments shown in Table~\ref{speedups},
-generating the full information set took over 25 minutes on a single processor.  However, by utilizing the Charm++
-search engine on a parallel machine (1024 processors), we were able to generate the full set in just 2.9 seconds.  (Note
-that this set had more than 4 million members.)
-
-%Li gives an overview of kriegspiel and discusses strategy from the perspective of human players~\cite{li94chess}.  He
+An overview of the theory of imperfect information games and other basic topics
+in game theory can be found in~\cite{kuhn03lectures} and~\cite{kuhn97classics}.
+A technical explanation of the concept of information sets can be found
+in~\cite{gilpin07algorithms}.  Koller \etal, showed that a Nash equilibrium can
+be found in time that is polynomial in the number of nodes in the game
+tree~\cite{koller94fast}.  In fact, they found that it took about as much time
+to generate the full game tree as it did to actually solve it.  Unfortunately,
+this is feasible only for small games.  In larger games where theoretically
+optimal strategies cannot be computed efficiently, computer systems tend to
+utilize the concept of information sets or belief states in some form or
+another.  A common strategy is to estimate the value of a move by averaging the
+estimated value of the associated descendant for each node in the current
+information set (or a random sample from the set).  While this sort of
+``averaging over clairvoyance'' {\em can} lead to very poor decisions, it is
+nevertheless a reasonable strategy in some games, such as
+Scrabble~\cite{sheppard02world}.
+
+In this work, we have argued for an ISG approach to Kriegspiel, but other
+authors have approached the problem differently.  Parker \etal, consider the
+problem of sampling from belief states in a variety of large game trees,
+including Kriegspiel~\cite{parker05game}.  They show that some performance
+gains can be achieved by only approximate sampling from the belief state.
+Instead of generating the full information set, or even sampling from it, they
+sample from possible positions for the current state (i.e., sampling from the
+belief state) by matching only the most recent observations.  The motivation
+for this approach seems to be that it would not be feasible to sample from the
+full information set, because of the computational expense involved. Our
+experiments bear this out-- it can indeed be expensive to produce the full
+information set.  For the experiments shown in Table~\ref{speedups}, generating
+the full information set took over 25 minutes on a single processor.  However,
+by utilizing the {\sc Charm++/ParSSSE} search engine on a parallel machine (1,024
+processors), we were able to generate the full set in just 2.9 seconds.
+
+%Li gives an overview of Kriegspiel and discusses strategy from the perspective of human players~\cite{li94chess}.  He
 %walks through the analysis of several actual games between two human players and discusses the inferences that each
@@
 
-%Russell and Wolfe have performed some analysis on kriegspiel positions for which it possible to prove that one player
+%Russell and Wolfe have performed some analysis on Kriegspiel positions for which it possible to prove that one player
 %has a forced win (that is, regardless of the configuration of the opponent's unseen pieces, the player has a sequence of
@@
 
-The information set generation approach that we have presented is certainly also computationally expensive.  And
-representing a full information set (e.g., as a list of sequences of moves that encode a path from the start state to
-the current node) would be similarly expensive in terms of storage space.  However, we have shown that our algorithm is
-also highly scalable.  Furthermore, contrary to the approach in~\cite{nance06reasoning}, our approach is readily
-adaptable to sampling algorithms.  Given a sample size that matches available memory, any single sequence of moves is
-readily extractable without the computational expense of a theorem prover or SAT solver.
+The information set generation approach that we have presented is certainly
+also computationally expensive.  And representing a full information set (e.g.,
+as a list of sequences of moves that encode a path from the start state to the
+current node) would be similarly expensive in terms of storage space.  However,
+we have shown that our algorithm is also highly scalable.  Furthermore,
+contrary to the approach in~\cite{nance06reasoning}, our approach is readily
+adaptable to sampling algorithms.  Given a sample size that matches available
+memory, any single sequence of moves is readily extractable without the
+computational expense of a theorem prover or SAT solver.
 
 %\Section{Conclusion Future Work}
-%We have shown that information set generation for kriegspiel can be efficiently parallelized in a way that scales up to at least 1024 processors.
+%We have shown that information set generation for Kriegspiel can be efficiently parallelized in a way that scales up to at least 1,024 processors.
 
@@
 
-%Ideally, we would run our algorithm on some deeper instances from real kriegspiel games.  We are not aware of a
-%repository of such games.  Unfortunately, notation for kriegspiel is not standardized and there are several minor
+%Ideally, we would run our algorithm on some deeper instances from real Kriegspiel games.  We are not aware of a
+%repository of such games.  Unfortunately, notation for Kriegspiel is not standardized and there are several minor
 %variations in rules (i.e., with respect to the nature of referee announcements) that are unlikely to significantly to
@@
 \Section{Conclusion and Future Work}
-We have shown that information set generation for kriegspiel can be efficiently parallelized.  The opportunities for
-parallelism increase as the problem size increases.  Our algorithm has performed well on problems with information sets
-with as many as 65 million nodes---much larger than any other treatment of the problem that we are aware of.  Because
-the algorithm can be efficiently parallelized, it would be reasonable to explore using it in a full kriegspiel player.
-It remains to be seen whether these scalability properties hold over a wide variety of playing styles and problem
-instances.  We suspect that fruitful research remains with respect to alternative algorithms that implement variable
+We have shown that information set generation for Kriegspiel can be efficiently
+parallelized.  The opportunities for parallelism increase as the problem size
+increases.  Our algorithm has performed well on problems with information sets
+with as many as 65 million nodes---much larger than any other treatment of the
+problem that we are aware of.  Because the algorithm can be efficiently
+parallelized, it would be reasonable to explore using it in a full Kriegspiel
+player.  It remains to be seen whether these scalability properties hold over a
+wide variety of playing styles and problem instances.  We suspect that fruitful
+research remains with respect to alternative algorithms that implement variable
 ordering heuristics.
@@
 %The purpose of this figure is simply to help visualize the concept of information sets in game trees.  The tree for
-%kriegspiel is very difficult to visualize/depict (even an interesting part of the tree).  Figure~\ref{fulltree} shows
-%the full game tree for a simple variant of poker known as Kuhn poker, which we hope will serve to at least make the
-%concept clear, even if the poker game itself is not immediately relevant.  There are three cards: a king, a queen, and a
+%Kriegspiel is very difficult to visualize/depict (even an interesting part of the tree).  Figure~\ref{fulltree} shows
+%the full game tree for a simple variant of Poker known as Kuhn Poker, which we hope will serve to at least make the
+%concept clear, even if the Poker game itself is not immediately relevant.  There are three cards: a king, a queen, and a
 %jack.  The dealer deals one card to each of two players.  Each player antes one unit, and there is one simple round of

### 2012-07-12 sbac/paper.tex

@@
 
-The paper is organized as follows.  In Section~\ref{Background}, we review
-necessary background material, including the rules of Kriegspiel and the nature
-of the information set generation problem.  In Section ~\ref{ParSSSE}, we
-describe the {\sc Charm++} Parallel State Space Search Engine (ParSSSE)
-platform.  We then provide our experimental results in Section~\ref{Results}.
-Related work is discussed in Section~\ref{Related}.  We conclude with a
-discussion of future work.
-
-%\Section{Information Set Generation}
-\Section{Kriegspiel}
+The paper is organized as follows.  In Section~\ref{Kriegspiel}, we review
+necessary background material, including the rules of Kriegspiel. Next, in
+Section~\ref{info} we describe the nature of the information set generation
+problem and the algorithm used.  In Section~\ref{ParSSSE}, we describe our
+parallel implementation, including the {\sc Charm++} Parallel State Space
+Search Engine (ParSSSE) platform.  We then provide our experimental results in
+Section~\ref{Results}.  Related work is discussed in Section~\ref{Related}.  We
+conclude with a discussion of future work.
+
+\section{Kriegspiel}\label{Kriegspiel}
 Information Set Generation is the process of identifying the nodes in an
@@
 
-%\Section{Background: Kriegspiel}
 Kriegspiel, or partially observable chess,  was invented by Henry Michael
@@
 
-\Section{Information Set Generation}
+\section{Information Set Generation}
+\label{info}
 %\subsection{Information Set Generation}
 The concept of information sets is perhaps best explained through an example. 
-Figure~\ref{listing} and show a full listing and the associated pictorial progression of an actual
-Kriegspiel game~\cite{li94chess}.  We have adapted the notation used by Wolfe for Berkeley Kriegspiel.  Odd (resp. even) moves are for white (resp. black).  The actual move is shown
+Figure~\ref{filteredlisting},~\ref{fig:figure1} and \ref{fig:figure2} show a full listing and the associated pictorial progression of an actual
+kriegspiel game~\cite{li94chess}.  We have adopted the notation used by Wolfe for Berkeley Kriegspiel.  Odd (resp. even) moves are for white (resp. black).  The actual move is shown
 first followed by a list of referee announcements.  Captures announcements are prefixed with an X and give the location
 of the captured piece.  Similarly, pawn tries announcements are prefixed with T.  Attempted moves that were ultimately
-declared illegal are shown in a list following the `:'.  For example at move 9, (Figure~\ref{fig:9W}--\ref{fig:9B}),
+declared illegal are shown in a list following the `:'.  For example at move 9, (Figure~\ref{filteredlisting}),
 black attempts to block a potential threat along the diagonal by moving a pawn to f6, then attempts to escape or capture
@@
 
-\Section{Parallel Implementation using Charm++ and ParSSSE}\label{ParSSSE}
-We implemented our search engine on top of the {\sc Charm++} run-time
+Figure~\ref{codelisting1} shows the function at the heart of our information set generation algorithm.  The actual
+observations can be viewed as global, read-only variables.  All solutions (i.e., all nodes in the information set) will
+be located at the same depth in the tree: the depth that is equal to the number of plies in the observation list.
+Referee announcements regarding pawn tries or a player being in check are made prior to a move.  The algorithm checks
+for the consistency of these observations at line $7$.  If the search reaches the appropriate depth and passes these
+tests, then a solution has been found and is added to the information set (line $3$).
+
+%Otherwise, we consider the possible moves that can be made from that state.  If the player whose turn it is to move at
+%the current level of the search tree is the same player from whose perspective we have received observations, then the
+%next step is to make sure that every legal move that was attempted at the player at that ply is also at least
+%attemptable at the game position that corresponds to the game tree node (lines $15--22$).  We know exactly the legal
+%move that was ultimately made at this depth (because we are assuming perfect recall), and we check to make sure that
+%this move is indeed legal (pruning otherwise).  If so, we apply that move and recurse (lines $24--32$).  This
+%corresponds to a ``forced move'' in the general search paradigm.
+%
+%On the other hand, if the player whose turn it is to move at the current depth of the search tree is the opponent, then
+%must consider all possible legal moves to branch on.  We first check to make sure that the number of legal moves from
+%this position is at least as large as the number of illegal moves that the opponent attempted (i.e., the number of times
+%the referee said ``No'' before finally accepting the opponent's move).  This is shown in lines $39--42$.  At this point
+%we generate a child for each move that would be accepted by the referee and recurse (lines $44--52$).
+%
+%One of our goals was to assess the varying utilities of the different pruning criteria.  We tried some different
+%orderings of the tests and were somewhat surprised (and disappointed) to see that there was actually a fairly minimal
+%impact on performance.  This could be due to the fact that because of the tedious nature of encoding problem instances
+%from actual games, we were not able to run experiments of a wide variety of types of positions (i.e., different styles
+%of play).
+
+%\begin{figure}
+%%\begin{listing}
+%\begin{lstlisting}
+%void generateInformationSet(uint16_t* trueState, uint16_t* possState, bool whiteMove, 
+%	uint16_t* possHistory, uint16_t** levels, int depth)
+%{
+%  if (!samePawnTries(trueState, possState, whiteMove)) return;  
+%  if (!sameCheckStatus(trueState, possState, whiteMove)) return; 
+%  if (depth == maxdepth) { // Then we have found a solution
+%	nSolutions++; return; // Report solution
+%  }
+%  uint16_t newPossState[16]; 
+%  int nMoves = 0;
+%  generateAttemptableMoves(possState, whiteMove, levels[depth], nMoves,false);
+%  checkForCheck(possState, whiteMove, levels[depth], nMoves); // prunes if not equal
+%  assert (nMoves < NMOVES); // 
+%
+%  if (whitePerspective == whiteMove) { // move from perspective plyaer 
+%    SetMove& failures = failedMoves[depth];
+%    for (SetMove::const_iterator itr = failures.begin(); itr != failures.end(); ++itr) {
+%      uint16_t move = *itr ;
+%      if (!foundMatchingMove(move,levels[depth],nMoves)) {
+%        return; // Prune: one of my own failed moves is not attemptable in this state
+%      }
+%    }
+%    // We know exactly what the actual move was; make it
+%    uint16_t& actualMove = moveHistory[depth];
+%    if (!foundMatchingMove(actualMove,levels[depth],nMoves)) {
+%      return; // Prune (because the move that we know we made at this depth is not legal)
+%    }
+%    // Otherwise, recurse 
+%    applyMove(possState,newPossState,actualMove);
+%    possHistory[depth] = actualMove;
+%    generateInformationSet(newTrueState, newPossState, !whiteMove, 
+%      possHistory, levels, depth+1);
+%  } else { 
+%    // We are at a level in the search tree where 
+%    // we are considering the possible moves for the opponent.
+%    // We know the number of attempted illegal moves, but not which ones
+%
+%    // Must be enough attemptable moves in the state to match the number of failed attempts
+%    unsigned nIllegalMoves = countIllegalMoves(levels[depth], nMoves);
+%    if (nIllegalMoves < failedMoves[depth].size()) {
+%	return; // Prune if not enough attemptable moves
+%    }
+%    // Now we want to try each possible move that is legally executable (not just attemptable) 
+%    for (int i = 0; i < nMoves; i++) {
+%      uint16_t& move = levels[depth][i];
+%      if (isLegal(move)) { // Obviously, we can only execute the moves that are actually legal from this state
+%        applyMove(possState,newPossState,move);
+%        possHistory[depth] = move;
+%        generateInformationSet(newTrueState, newPossState, !whiteMove, 
+%          possHistory, levels, depth+1);
+%      }
+%    }
+%  }
+%}
+%\end{lstlisting}
+%\caption{The heart of the information set generation algorithm.}
+%\label{codelisting}
+%\end{figure}
+
+\section{Parallel Implementation using Charm++ and ParSSSE}\label{ParSSSE}
+We built our search engine on top of the {\sc Charm++} run-time
 system for parallel processing~\cite{kale93charm,kale09charm}.  {\sc Charm++}
@@
 
-\Section{Performance Results}\label{Results}
+\section{Performance Results}\label{Results}
 In order to understand the effect of various parameters such as grainsize and
@@
 The performance results reported in this section were obtained by running our
-program on Blueprint with up to 1,024 processors. TODO: Blueprint H/w or
-citation
+program on Blueprint with up to 1,024 processors. This system was composed of
+120 POWER5+ nodes, each having 16 cores and 64GB of memory. The system was
+decomissioned in June 2011 and is no longer in operation.
 
@@
 
-\begin{table}
+\begin{table}[ht]
 \centering
@@
 \end{table}
+In order to understand the effect of various parameters such as grain size and load balancing, we first evaluated the scalability of our implementation using a randomly selected problem instance.
+Next, we show that our implementation performs equally well for instances from actual games.
+\subsection{Analysis using Random Instance}
+The performance results reported in
+this section were obtained by running our program on Blueprint with up to 1024 processors.
+This system was composed of 120 POWER5+ nodes, each having 16 cores and 64GB of memory.
+The system was decomissioned in June 2011 and is no longer in operation.
+
+Table~\ref{speedups} shows the performance results for our algorithm on a randomly selected problem instance in which
+the size of the information set was about 4.5 million.  This instance is not necessarily representative of all games or
+states of the game, but it was not cherry-picked in any way. The table shows that the problem is indeed highly
+parallizable.  The speedups are nearly linear for up  to 32 processors, and even on 1024 processors, we get a good
+speedup.  In this particular instance, the sequential algorithm takes almost 25 minutes to run.  But when parallelized
+on 1024 processors---perfectly reasonable resource usage on modern hardware---the search takes less than three seconds.
+This is a speedup of 576 times!  It might not be reasonable to use the algorithm at all if it takes 25 or more to run
+for some turns.  But if the cost is only a few seconds, it becomes entirely feasible.
 
 \subsubsection{Effect of Load Balancing}
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.9\columnwidth]{plots/3schemes.pdf}
-\vspace{-0.1in}
-\caption{Effect of load-balancing strategy}
-\label{3schemes}
-\vspace{-0.2in}
-\end{figure}
 
@@
 fact that we are doing an all-solutions search. 
-
-\subsubsection{Effect of Problem Size}
-\begin{figure}[h]
+\begin{figure}[ht]
 \centering
-\begin{minipage}{0.7\linewidth}
-\includegraphics[width=\columnwidth]{plots/3depths.pdf}
-\end{minipage}
-\centering
-\begin{minipage}{0.7\linewidth}
-\includegraphics[width=\columnwidth]{plots/3depthsnolog.pdf}
-\end{minipage}
+\includegraphics[width=0.9\columnwidth]{plots/3schemes.pdf}
 \vspace{-0.1in}
@@
 
+\caption{Effect of load balancing strategy}
+\label{3schemes}
 \vspace{-0.2in}
-\label{3depths}
 \end{figure}
-
-We next study the variation in the efficiency of the parallelization of the
-algorithm as size of the problem increases.  To test this, we varied the number
-of plies in the test instance from 8 to 12. We used the work stealing
-load-balancing strategy.
-
-The results are shown in Figure~\ref{3depths}.  In the easiest case (depth 8),
-there is not enough additional parallel work to justify the jump from 256 to
-1,024 processors, and performance actually degrades.  Aside from this anomaly,
-however, we see consistent speedups as we increase the number of processors.
-The plot on the right helps to illustrate the good isoefficiency of the
-problem.  As the size of the problem increases, the amount of parallelizable
-work available also increases, and therefore the steepness of the speed curve
-is higher for larger problem sizes, as the number of processors increases.
 
@@
 
+\subsubsection{Effect of Problem Size}
+
+We next study the variation in the efficiency of the parallelization of the
+algorithm as size of the problem increases.  To test this, we varied the number
+of plies in the test instance from 8 to 12. We used the work stealing
+load-balancing strategy.
+
+The results are shown in Figure~\ref{3depths}.  In the easiest case (depth 8),
+there is not enough additional parallel work to justify the jump from 256 to
+1,024 processors, and performance actually degrades.  Aside from this anomaly,
+however, we see consistent speedups as we increase the number of processors.
+The plot on the right helps to illustrate the good isoefficiency of the
+problem.  As the size of the problem increases, the amount of parallelizable
+work available also increases, and therefore the steepness of the speed curve
+is higher for larger problem sizes, as the number of processors increases.
+
 \subsection{Real Game Instance}
+%\begin{figure}[ht]
+% \centering \includegraphics[viewport= 280 200 300 550, scale=0.35]{images/KriegspielProblem4.pdf} %
+%
+%\caption{Kriegspiel position from Virgil-David 1969 with information set size
+%764,209.}
+%
+%\label{prob4} 
+%\vspace{-0.2in}
+%\end{figure}
+%
+%\begin{figure}[ht]
+% \centering \includegraphics[viewport= 280 200 300 550, scale=0.35]{images/KriegspielProblem5.pdf} %
+%
+%\caption{Kriegspiel position from Game 6 of Kbot-Darkboard at the 2006 Computer
+%Chess championship.  Information set size is 5,099,257.}
+%
+%\label{prob5} 
+%\vspace{-0.2in}
+%\end{figure}
+
 \begin{figure}[ht]
- \centering \includegraphics[viewport= 280 200 300 550, scale=0.35]{images/KriegspielProblem4.pdf} %
-
-\caption{Kriegspiel position from Virgil-David 1969 with information set size
-764,209.}
-
+ \centering \includegraphics[scale=1]{plots/new.pdf} %
+\caption{Kriegspiel position from 2 games, position 10}
 \label{prob4} 
@@
 
-\begin{figure}[ht]
- \centering \includegraphics[viewport= 280 200 300 550, scale=0.35]{images/KriegspielProblem5.pdf} %
-
-\caption{Kriegspiel position from Game 6 of Kbot-Darkboard at the 2006 Computer
-Chess championship.  Information set size is 5,099,257.}
-
-\label{prob5} 
-\vspace{-0.2in}
-\end{figure}
-
-We also evaluated our information set generation routine on 40 positions from
-actual kriespiel games, including human vs. human and computer vs. computer
-instances.  Many of these were trivial to solve in a few seconds or less on a
-desktop computer.  For these experiments, we used the Ranger supercomputer
-at TACC~\cite{ranger}. Ranger has 16-way SMP nodes containing AMD Opteron
-processors. The nodes are interconnected using Infiniband technology in a
-fully-CLOS topology providing a 1GBps point-to-point bandwidth.
-
-The running times and speedups for two of the more challenging instances on
-Ranger are shown in Tables~\ref{prob4}--\ref{tab:prob5speedups}.   These tables
-show the results for runs on one to $1,024$ cores.  Each column corresponds to
-a different grainsize parameter.  In this case, the grainsize is the number of
-levels of the search tree that are searched in parallel.  Up to that threshold,
-each node that is generated spawns a new piece of work, which the {\sc Charm++}
-runtime schedules to execute on one of the processing elements according to its
-load balancing strategy.  Above the threshold the entire subtree is searched to
-completion on a single processor, with no additional parallel overhead.
-
-Note that the optimal grainsize varies for the different numbers of cores and
-the pattern is consistent across problem instances.  The optimal
-grainsize increases as the number of cores increases.  This makes sense,
-because as this parameter grows, the number of distinct pieces of work
-increases.  When the number of chares per processor is high, much of
-the overhead of parallelization is wasted.  But as the number of cores
-increases, the work can be efficiently broken up into smaller pieces and doled
-out to all the cores.  With more pieces of work, it becomes easier to balance
-the load.
-
-\begin{table}[tphb]
+\begin{table}[ht]
 \centering
@@
 
+We also evaluated our information set generation routine on 40 positions from
+actual kriespiel games, including human vs. human and computer vs. computer
+instances.  Many of these were trivial to solve in a few seconds or less on a
+desktop computer.  For these experiments, we used the Ranger supercomputer
+at TACC~\cite{ranger}. Ranger has 16-way SMP nodes containing AMD Opteron
+processors. The nodes are interconnected using Infiniband technology in a
+fully-CLOS topology providing a 1GBps point-to-point bandwidth.
+
+The running times and speedups for two of the more challenging instances on
+Ranger are shown in Tables~\ref{prob4}--\ref{tab:prob5speedups}.   These tables
+show the results for runs on one to $1,024$ cores.  Each column corresponds to
+a different grainsize parameter.  In this case, the grainsize is the number of
+levels of the search tree that are searched in parallel.  Up to that threshold,
+each node that is generated spawns a new piece of work, which the {\sc Charm++}
+runtime schedules to execute on one of the processing elements according to its
+load balancing strategy.  Above the threshold the entire subtree is searched to
+completion on a single processor, with no additional parallel overhead.
+
+Note that the optimal grainsize varies for the different numbers of cores and
+the pattern is consistent across problem instances.  The optimal
+grainsize increases as the number of cores increases.  This makes sense,
+because as this parameter grows, the number of distinct pieces of work
+increases.  When the number of chares per processor is high, much of
+the overhead of parallelization is wasted.  But as the number of cores
+increases, the work can be efficiently broken up into smaller pieces and doled
+out to all the cores.  With more pieces of work, it becomes easier to balance
+the load.
+
 Note that both problems scale well even up to 1,024 processors.  We achieve a
@@
 
-For partially observable games, information set generation represents a
-different variety of tree search that does not suffer from speculative loss.
-All of the nodes that are visited in parallel also would have been visited in a
-sequential run.  The key challenge for parallelizing information set generation
-is load balancing: some parts of the tree will be pruned at a higher level than
-others, resulting in a wide variation in cost for each subtree.  The {\sc
-Charm++/ParSSSE} framework is perfectly suited to handle the load balancing
-issues that this problem naturally presents.
-
-Note also that information set generation is just one subroutine needed to
-reason about partially observable games.  Once plausible game histories have
-been found, an agent may simulate some sequences of moves up to previous moves
-made by opponents and then recursively call the ISG procedure from the
-perspective of the opponent.  This results in many parallel applications of
-ISG. Furthermore, given a node in an information set, it is common for an agent
-to perform forward search on that node (e.g., using MCTS).  All of these
-forward search operation can be performed independently in parallel. With these
-two additional levels of parallelism on top of the information set generation
-problem, it is likely that the problem of playing partially observable
-games could scale efficiently to millions of processors or more.
-
-\begin{table}[tphb]
+\begin{table}[ht]
 \centering
@@
 
-\Section{Related Work}
+For partially observable games, information set generation represents a
+different variety of tree search that does not suffer from speculative loss.
+All of the nodes that are visited in parallel also would have been visited in a
+sequential run.  The key challenge for parallelizing information set generation
+is load balancing: some parts of the tree will be pruned at a higher level than
+others, resulting in a wide variation in cost for each subtree.  The {\sc
+Charm++/ParSSSE} framework is perfectly suited to handle the load balancing
+issues that this problem naturally presents.
+
+Note also that information set generation is just one subroutine needed to
+reason about partially observable games.  Once plausible game histories have
+been found, an agent may simulate some sequences of moves up to previous moves
+made by opponents and then recursively call the ISG procedure from the
+perspective of the opponent.  This results in many parallel applications of
+ISG. Furthermore, given a node in an information set, it is common for an agent
+to perform forward search on that node (e.g., using MCTS).  All of these
+forward search operation can be performed independently in parallel. With these
+two additional levels of parallelism on top of the information set generation
+problem, it is likely that the problem of playing partially observable
+games could scale efficiently to millions of processors or more.
+
+\section{Related Work}
+\label{Related}
 An overview of the theory of imperfect information games and other basic topics
@@
 problem of sampling from belief states in a variety of large game trees,
-including Kriegspiel~\cite{parker05game}.  They show that some performance
-gains can be achieved by only approximate sampling from the belief state.
-Instead of generating the full information set, or even sampling from it, they
-sample from possible positions for the current state (i.e., sampling from the
-belief state) by matching only the most recent observations.  The motivation
-for this approach seems to be that it would not be feasible to sample from the
-full information set, because of the computational expense involved. Our
-experiments bear this out-- it can indeed be expensive to produce the full
-information set.  For the experiments shown in Table~\ref{speedups}, generating
-the full information set took over 25 minutes on a single processor.  However,
-by utilizing the {\sc Charm++/ParSSSE} search engine on a parallel machine (1,024
-processors), we were able to generate the full set in just 2.9 seconds.
+including Kriegspiel~\cite{parker05game}.  They show that some performance gains
+can be achieved by only approximate sampling from the belief state.  Instead of
+generating the full information set, or even sampling from it, they sample from
+possible positions for the current state (i.e., sampling from the belief state)
+by matching only the most recent observations.  The motivation for this approach
+seems to be that it would not be feasible to sample from the full information
+set, because of the computational expense involved. Our experiments bear this
+out-- it can indeed be expensive to produce the full information set.  For the
+experiments shown in Table~\ref{speedups}, generating the full information set
+took over 25 minutes on a single processor.  However, by utilizing the {\sc
+Charm++/ParSSSE} search engine on a parallel machine (1,024 processors), we were
+able to generate the full set in just 2.9 seconds.
 
@@
 
-\Section{Conclusion and Future Work}
+\section{Conclusion and Future Work}
 We have shown that information set generation for Kriegspiel can be efficiently

### 2012-07-12 sbac/paper.tex

@@
 to reason about the probability distribution over those states is crucial to
-playing imperfect information games well~\cite{thielscher12hyperplay}.
+playing imperfect information games well~\cite{schofield12hyperplay}.
 

### 2012-07-12 sbac/paper.tex
Removed some duplicate text

@@
 \end{table}
-In order to understand the effect of various parameters such as grain size and load balancing, we first evaluated the scalability of our implementation using a randomly selected problem instance.
-Next, we show that our implementation performs equally well for instances from actual games.
-\subsection{Analysis using Random Instance}
-The performance results reported in
-this section were obtained by running our program on Blueprint with up to 1024 processors.
-This system was composed of 120 POWER5+ nodes, each having 16 cores and 64GB of memory.
-The system was decomissioned in June 2011 and is no longer in operation.
-
-Table~\ref{speedups} shows the performance results for our algorithm on a randomly selected problem instance in which
-the size of the information set was about 4.5 million.  This instance is not necessarily representative of all games or
-states of the game, but it was not cherry-picked in any way. The table shows that the problem is indeed highly
-parallizable.  The speedups are nearly linear for up  to 32 processors, and even on 1024 processors, we get a good
-speedup.  In this particular instance, the sequential algorithm takes almost 25 minutes to run.  But when parallelized
-on 1024 processors---perfectly reasonable resource usage on modern hardware---the search takes less than three seconds.
-This is a speedup of 576 times!  It might not be reasonable to use the algorithm at all if it takes 25 or more to run
-for some turns.  But if the cost is only a few seconds, it becomes entirely feasible.
 

### 2012-07-15 sbac/paper.tex
removed some vspaces

@@
 \newenvironment{mytinylisting}
-{\begin{list}{}{\setlength{\leftmargin}{1em}}\item\scriptsize\texttt}
+{\begin{list}{}{\setlength{\leftmargin}{1em}}\item\small\texttt}
 {\end{list}}
+
 \newenvironment{proof}{\vspace{-0.15in}\noindent{\bf Proof:}}%
@@
 ultimately allowed by the referee would also be a legal move in chess.
+
+\begin{figure}
+\begin{center}
+\begin{verbatim}
+1. d4 {(:)}
+   a5 {(:)}
+2. Bg5 {(:)}
+   b6 {(:)}
+3. Nc3 {(:Bd8)}
+   c5 {(:)}
+4. d5 {(:Bd8;Tc5)}
+   Na6 {(:)}
+5. d6 {(:)}
+   f6 {(:Td6)}
+6. e4 {(:)}
+   fxg5 {(Xg5:Td6;Tg5)}
+7. Bc4 {(:)}
+   exd6 {(Xd6:Td6)}
+8. Qd5 {(:)}
+   Qc7 {(:)}
+9. Qf7+ {(:)} 
+   Kd8 {:g6,Kxf7,Ke7}
+10. Qxf8++ {(1-0:)}
+\end{verbatim}
+\end{center}
+\caption{Listing of a Kriegspiel game resulting in checkmate by white after 10 moves.
+The final, legal move for each player is shown followed by the list of moderator declarations and all moves that were attempted but disallowed before the legal move was accepted.}
+\label{listing}
+\end{figure}
 
@@
 \end{minipage}
-\vspace{-0.2in}
 \end{figure}
-
-%\begin{figure}
-%\begin{verbatim}
-%1. d4 {(:)}
-%   a5 {(:)}
-%2. Bg5 {(:)}
-%   b6 {(:)}
-%3. Nc3 {(:Bd8)}
-%   c5 {(:)}
-%4. d5 {(:Bd8;Tc5)}
-%   Na6 {(:)}
-%5. d6 {(:)}
-%   f6 {(:Td6)}
-%6. e4 {(:)}
-%   fxg5 {(Xg5:Td6;Tg5)}
-%7. Bc4 {(:)}
-%   exd6 {(Xd6:Td6)}
-%8. Qd5 {(:)}
-%   Qc7 {(:)}
-%9. Qf7+ {(:)} 
-%   d8 {:g6,f7,e7}
-%10. Qxf8++ {(1-0:)}
-%\end{verbatim}
-%\label{listing}
-%\caption{Listing of a Kriegspiel game resulting in checkmate by white after 10 moves.}
-%\end{figure}
 
@@
 The concept of information sets is perhaps best explained through an example. 
-Figure~\ref{filteredlisting},~\ref{fig:figure1} and \ref{fig:figure2} show a full listing and the associated pictorial progression of an actual
-kriegspiel game~\cite{li94chess}.  We have adopted the notation used by Wolfe for Berkeley Kriegspiel.  Odd (resp. even) moves are for white (resp. black).  The actual move is shown
+Figures ~\ref{listing} show a full game listing and the associated pictorial progression of an actual
+Kriegspiel game~\cite{li94chess}.
+We have adopted the notation used by Wolfe for Berkeley Kriegspiel~\cite{wolfe07exploiting}.
+Each numbered entry shows one complete turn for both White and Black.
+The actual move is shown
 first followed by a list of referee announcements.  Captures announcements are prefixed with an X and give the location
-of the captured piece.  Similarly, pawn tries announcements are prefixed with T.  Attempted moves that were ultimately
-declared illegal are shown in a list following the `:'.  For example at move 9, (Figure~\ref{filteredlisting}),
-black attempts to block a potential threat along the diagonal by moving a pawn to f6, then attempts to escape or capture
-a threat at f7.  When both of these attempts fail, black can infer that his king is being threatened by a protected
-bishop or queen at f7.  The attempt to move to e7 is black's way of finding out whether the threatening piece is a queen
-or bishop.  When this move fails, black knows it is white's queen and ultimately retreats to d8.  Note that the failed
-moves by white on turns 3 and 4 are a result of being blocked by the pawn at e2.
+of the captured piece.  Similarly, pawn try announcements are prefixed with T.  Attempted moves that were ultimately
+declared illegal are shown in a list following the `:'.
+
+Figure~\ref{fig:figure1} shows the true state of the game after nine moves by
+each player.  Figure~\ref{fig:figure2} shows White's perspective of the game at
+that poin. At move 9, Black attempts to block a
+potential threat along the diagonal by moving a pawn to g6, then attempts to
+escape or capture a threat at f7.  When both of these attempts fail, Black can
+infer that his king is being threatened by a protected bishop or queen at f7.
+The attempt to move to e7 is Black's way of finding out whether the threatening
+piece is a queen or bishop.  When this move fails, black knows it is White's
+queen and ultimately retreats to d8.  Note that the failed moves by White on
+turns 3 and 4 are a result of being blocked by the pawn at e2.
 
@@
 post-game analysis or commentary.  During the game, players would not have access to the full information.  The
-transcript from white's perspective is shown in Figure~\ref{filteredlisting}.  Note that black's actual moves have been
-replaced with ``??'' and lists of black's illegal moves have been replaced with the {\em number} of illegal attempted
-moves.  For example, on move 9, black attempted three illegal moves.
+transcript from White's perspective is shown in Figure~\ref{filteredlisting}.  Note that Black's actual moves have been
+replaced with ``??'' and lists of Black's illegal moves have been replaced with the {\em number} of illegal attempted
+moves.  For example, on move 9, Black attempted three illegal moves.
 \begin{figure}
-\begin{minipage}[b]{0.35\linewidth}
-\centering
-\small
+%\begin{minipage}[b]{0.35\linewidth}
+%\centering
+%\small
 \begin{verbatim}
@@
    ?? {(:0,Td6)}
-\end{verbatim}
-\end{minipage}
-\begin{minipage}[b]{0.25\linewidth}
-\small
-\centering
-\begin{verbatim}
 6. e4 {(:)}
@@
 \end{verbatim}
-\end{minipage}
-\vspace{-0.1in}
-\caption{Listing of a Kriegspiel game resulting in checkmate by white after 10 moves.}
+%\end{verbatim}
+%\end{minipage}
+%\begin{minipage}[b]{0.25\linewidth}
+%\small
+%\centering
+%\begin{verbatim}
+%\end{minipage}
+%\vspace{-0.1in}
+\caption{Observations received by White over the course of the game.
+White does not know Black's actual moves but hears moderator announcements related to captures, pawn tries, and the number of attempted moves by Black that fail.}
 \label{filteredlisting}
-\vspace{-0.2in}
+%\vspace{-0.2in}
 \end{figure}
@@
 players}, that are consistent with those observations.  In other words, any sequence of moves that would have generated
-the same sequence of observations for black is in black's information set.  Figure~\ref{abbrevoutput} shows the
-information set for black after both player's have made five moves.  From this list of possibilities, black can infer
-that white must have a bishop at g5 and a pawn at d6.  There are 25 possible sequences of moves that are consistent with
-black's observations up to that point.
+the same sequence of observations for Black is in Black's information set.  Figure~\ref{abbrevoutput} shows a portion of the
+information set for Black after both player's have made five moves.  From this list of possibilities, Black can infer
+that White must have a bishop at g5 and a pawn at d6.  There are 25 possible sequences of moves that are consistent with
+Black's observations up to that point.
 
@@
 \end{mytinylisting}
-\vspace{-0.2in}
-\caption{Abbreviated output from example game.  Generating information sets for black after five turns by each player.
-From this, black can infer that there is definitely a white bishop at g5 and a white pawn at d6.}
+\caption{Abbreviated output from example game.
+Three of 25 sequences that are consistent with Black's observations after five moves by each player.
+The 25 sequences (or the states that result from them) compose Black's information set at that point in the game.
+From this, Black can infer that there is definitely a white bishop at g5 and a white pawn at d6.}
 \label{abbrevoutput}
-\vspace{-0.3in}
 \end{figure}
@@
 
-\textbf{The Information Set Generation Algorithm}
-(ISG) can be formulated as a combinatorial search problem.  The input to the algorithm is a list
-of observations for one player in a Kriegspiel game.
-
-The nodes in the search tree at depth $n$ correspond possible to possible sequences of $n$ plies from the start state.
-There is a search space variable for each ply in the game, and the values that a variable can take on are all possible
-legal moves.  Information set generation, then, is an all-solutions search through the space.  Whenever a node in the search tree is
-found to be inconsistent with the actual observation given in the input, the
-subtree rooted at that node is pruned (i.e., search is discontinued along that
-path).  There are types of observations that must match in order for the state
-to be found: check status, pawn tries, and illegal moves available/attempted.
+\textbf{The Information Set Generation}
+(ISG) operation can be formulated as a combinatorial search problem.  The input to the algorithm is a list
+of $t$ observations for one player $k$ (denoted $o_{1:t}^{(k)}$ in a Kriegspiel game.
+
+The nodes in the search tree at depth $n$ correspond possible to possible move 
+sequences of $n$ plies from the start state.  (A sequence of $t$ actions from
+the start state is denoted $a_{1:t}$.) There is a search space variable for
+each ply in the game, and the values that a variable can take on are all
+possible legal moves.  Information set generation, then, is an all-solutions
+search through the space.  Whenever a node in the search tree is found to be
+inconsistent with the actual observation given in the input, the subtree rooted
+at that node is pruned (i.e., search is discontinued along that path).  There
+are three types of observations that must match in order for a state $s$ to be
+included in the information set $I$ at depth $t$: check status, pawn tries, and
+illegal moves available/attempted.
+
 \begin{figure}[htpb]
@@
 \EndFunction
-\label{codelisting1}
 \end{algorithmic}
@@
 \caption{Algorithm}
+\label{codelisting1}
 \end{figure}
@@
 \end{tabular}
-\vspace{-0.1in}
 
@@
 \label{speedups}
-\vspace{-0.3in}
 \end{table}
@@
 \includegraphics[width=0.9\columnwidth]{plots/3schemes.pdf}
-\vspace{-0.1in}
-
-\caption{(Top) Log-log plot of running time for three different problem sizes
-as the number of cores used increases.  (Bottom) Same data with a linear y-axis
-for the running time (to help see the isoefficiency.)}
 
@@
 \label{3schemes}
-\vspace{-0.2in}
 \end{figure}
@@
 } 
-\vspace{-0.1in}
 \caption{Grainsize Control for different problem sizes}
@@
 \label{fig:dep}
-\vspace{-0.1in}
 \end{figure*}
+%
+\begin{figure}[ht]
+\centering
+\begin{minipage}{0.7\linewidth}
+\includegraphics[width=\columnwidth]{plots/3depths.pdf}
+\end{minipage}
+\centering
+\begin{minipage}{0.7\linewidth}
+\includegraphics[width=\columnwidth]{plots/3depthsnolog.pdf}
+\end{minipage}
+\caption{(Top) Log-log plot of running time for three different sizes of problem as the number of cores used increases.
+(Bottom) Same data with a linear y-axis for the running time, to help see the isoefficiency.}
+\label{3depths}
+\end{figure}
 
@@
 \label{prob4} 
-\vspace{-0.2in}
 \end{figure}
@@
 
-\vspace{-0.2in}
 \label{tab:prob4}
@@
 Optimal grainsize for each number of cores is shown in bold.}
-
-\vspace{-0.2in}
 \label{tab:prob5}
@@
 We also evaluated our information set generation routine on 40 positions from
-actual kriespiel games, including human vs. human and computer vs. computer
+actual Kriespiel games, including human vs. human and computer vs. computer
 instances.  Many of these were trivial to solve in a few seconds or less on a
@@
 The running times and speedups for two of the more challenging instances on
-Ranger are shown in Tables~\ref{prob4}--\ref{tab:prob5speedups}.   These tables
+Ranger are shown in Tables~\ref{tab:prob4}--\ref{tab:prob5}.   These tables
 show the results for runs on one to $1,024$ cores.  Each column corresponds to
@@
 speedup of $443$ in the game between humans and up to $536$ in the game between
-computers.  Figures~\ref{prob4} and~\ref{prob5} show the plots of the running
-times for the best grainsize for each core setting.  These plots show that the
+computers.  Figure~\ref{prob4} shows the plots of the running
+times on these two problems with the best grainsize for each core setting.  These plots show that the
 running time continues to decrease as the number of cores increases.
@@
 \caption{Best Speedups reported for game tree applications.}
-\vspace{-0.2in}
 \label{bestspeedups}
@@
 A technical explanation of the concept of information sets can be found
-in~\cite{gilpin07algorithms}.  Koller \etal, showed that a Nash equilibrium can
+in~\cite{gilpin09algorithms}.  Koller \etal, showed that a Nash equilibrium can
 be found in time that is polynomial in the number of nodes in the game

### 2012-07-15 sbac/paper.tex
table captions - removed centering
Updates and error fixes to ISG section; a few caption updates

@@
 each player.  Figure~\ref{fig:figure2} shows White's perspective of the game at
-that poin. At move 9, Black attempts to block a
+that poin. At move 9, (Figure~\ref{filteredlisting}), Black attempts to block a
 potential threat along the diagonal by moving a pawn to g6, then attempts to
@@
 \end{mytinylisting}
+\vspace{-0.2in}
 \caption{Abbreviated output from example game.
@@
 \end{tabular}
+\vspace{-0.1in}
 
@@
 \label{speedups}
+\vspace{-0.3in}
 \end{table}
@@
 \includegraphics[width=0.9\columnwidth]{plots/3schemes.pdf}
+\vspace{-0.1in}
 
@@
 \label{3schemes}
+\vspace{-0.2in}
 \end{figure}
@@
 } 
+\vspace{-0.1in}
 \caption{Grainsize Control for different problem sizes}
@@
 \label{fig:dep}
+\vspace{-0.1in}
 \end{figure*}
@@
 \end{minipage}
+\vspace{-0.1in}
 \caption{(Top) Log-log plot of running time for three different sizes of problem as the number of cores used increases.
 (Bottom) Same data with a linear y-axis for the running time, to help see the isoefficiency.}
+\vspace{-0.2in}
 \label{3depths}
@@
 \label{prob4} 
+\vspace{-0.2in}
 \end{figure}
@@
 
+\vspace{-0.2in}
 \label{tab:prob4}
@@
 Optimal grainsize for each number of cores is shown in bold.}
+
+\vspace{-0.2in}
 \label{tab:prob5}

### 2012-07-15 sbac/paper.tex
spell check

--- 2012-07-15 sbac/paper.tex
@@
 \usepackage{amsmath,amsfonts,latexsym,graphicx}
+\usepackage[tableposition=top,font=bf]{caption}
+
 \lstset{language=C,numbers=left,basicstyle=\footnotesize,numberstyle=\footnotesize}
@@
 critical to the performance of game-playing agents.  ISG represent a class of
-combinatorial search problems which is computationaly intensive but difficult
+combinatorial search problems which is computationally intensive but difficult
 to efficiently parallelize.
@@
 each player.  Figure~\ref{fig:figure2} shows White's perspective of the game at
-that poin. At move 9, Black attempts to block a
+that coin. At move 9, Black attempts to block a
 potential threat along the diagonal by moving a pawn to g6, then attempts to
@@
 %
-%  if (whitePerspective == whiteMove) { // move from perspective plyaer 
+%  if (whitePerspective == whiteMove) { // move from perspective player 
 %    SetMove& failures = failedMoves[depth];
@@
 120 POWER5+ nodes, each having 16 cores and 64GB of memory. The system was
-decomissioned in June 2011 and is no longer in operation.
+decommissioned in June 2011 and is no longer in operation.
 
@@
 \begin{table}[ht]
+
+\caption{Speedups for a randomly selected problem instance on 1--1,024
+processors.  The input observations were from the 10th ply of the game tree and
+the size of the information set in this instance is 4,495,121.}
 \centering
@@
 \end{tabular}
-
-\caption{Speedups for a randomly selected problem instance on 1--1,024
-processors.  The input observations were from the 10th ply of the game tree and
-the size of the information set in this instance is 4,495,121.}
 
@@
 \begin{table}[ht]
+\caption{Running times in seconds and speedups (after $|$), to find full
+information set of size 764,209 at position 10 of Virgil-David 1969. Optimal
+grainsize (G) for each number of cores (PE) is shown in bold.}
+
 \centering
@@
 
-\caption{Running times in seconds and speedups (after $|$), to find full
-information set of size 764,209 at position 10 of Virgil-David 1969. Optimal
-grainsize (G) for each number of cores (PE) is shown in bold.}
-
 \label{tab:prob4}
@@
 \begin{table}[th]
+
+\caption{Running times, in seconds and speedups (after $|$), to find full
+information set of size 5,099,257 at position 10 of Game 6 Kbot-Darkboard 2006.
+Optimal grainsize for each number of cores is shown in bold.}
 \centering
@@
 
-\caption{Running times, in seconds and speedups (after $|$), to find full
-information set of size 5,099,257 at position 10 of Game 6 Kbot-Darkboard 2006.
-Optimal grainsize for each number of cores is shown in bold.}
 \label{tab:prob5}
@@
 We also evaluated our information set generation routine on 40 positions from
-actual Kriespiel games, including human vs. human and computer vs. computer
+actual Kriegspiel games, including human vs. human and computer vs. computer
 instances.  Many of these were trivial to solve in a few seconds or less on a
@@
 \begin{table}[ht]
+\caption{Best Speedups reported for game tree applications.}
 \centering
@@
 \end{tabular}
-\caption{Best Speedups reported for game tree applications.}
 \label{bestspeedups}

### 2012-07-15 sbac/paper.tex
Removed superfluous figure reference
Edited abstract

@@
 
-\author{Mark Richards, Abhishek Gupta, Osman Sarood and Laxmikant V. Kale\\
-Department of Computer Science\\ 
-University of Illinois\\ 
-Urbana, IL\\ {mdrichar, gupta59, sarood1, kale}@illinois.edu\\
+\author{
+\IEEEauthorblockN{Mark Richards, Abhishek Gupta, Osman Sarood and Laxmikant V. Kal\'e\\}
+\IEEEauthorblockA{University of Illinois at Urbana-Champaign\\
+Urbana, IL 61801, USA\\
+(mdrichar, gupta59, sarood1, kale)@illinois.edu\\}
 % For a paper whose authors are all at the same institution, 
@@
 \begin{abstract}
-Information Set Generation is the identification of the set of paths in an
+Information Set Generation (ISG) is the identification of the set of paths in an
 imperfect information game tree that are consistent with a player's
 observations.  The ability to reason about the possible game history is
-critical to the performance of game-playing agents.  ISG represent a class of
+critical to the performance of game-playing agents.  ISG represents a class of
 combinatorial search problems which is computationally intensive but difficult
@@
 algorithm on top of a general purpose combinatorial search engine and discuss
-its performance under different load balancing strategies and with different
-grainsize parameters.  We achieve speedups of over 500 on 1,024 processors, far
+its performance using datasets from real game instances in addition to benchmarks.
+Further, we demonstrate the effect of load balancing strategies, problem sizes and computational granularity (grainsize parameters) on performance.
+We achieve speedups of over 500 on 1,024 processors, far
 exceeding previous scalability results for game tree search applications.

### 2012-07-16 sbac/paper.tex
revise intro

@@
 critical to the performance of game-playing agents.  ISG represents a class of
-combinatorial search problems which is computationally intensive but difficult
+combinatorial search problems which is computationally intensive but challenging
 to efficiently parallelize.
@@
 Poker~\cite{billings02challenge} or Bridge~\cite{ginsberg96partition}, hidden
-tiles in Scrabble~\cite{richards07opponent}, or hidden pieces 
+tiles in Scrabble~\cite{richards07opponent}, or hidden pieces in 
 Kriegspiel (partially observable chess)~\cite{li94chess}. The game tree nodes
@@
 can use to identify nodes in the game tree that are consistent with their
-observations~\cite{richards12information}.  Because the operation is
+observations~\cite{richards12information}.  Since the operation is
 computationally intensive and may be repeatedly applied by a decision-making
@@
 game tree denotes not only the current state of the game, but also uniquely
-defines a path from the initial state or root node.  Thus, a game tree node
+defines a path from the initial state or root node.  Hence, a game tree node
 implicitly encodes not only the current state of the game but also the complete
@@
 
-In the game of Kriegspiel, information set generation is much more complicated.
+However, in the game of Kriegspiel, information set generation is much more complicated.
 After the first move by the opponent, a player faces uncertainty about the
@@
 In this work, we show that ISG for Kriegspiel can be efficiently parallelized.
-The primary contributions of this work are the following. We describe an
+The primary contributions of this work are the following:
+\begin{itemize}
+\item
+ We describe an
 approach to efficient parallelization of an application which is not typical of
-high performance computing domain. We present scalability results for our
+high performance computing domain.
+\item
+ We present scalability results for our
 parallel implementation of an information set generation algorithm for the game
-of Kriegspiel using problem instances from real games. Further, we demonstrate
-the effects of load balancing strategy and grainsize on parallel performance.
-
-The paper is organized as follows.  In Section~\ref{Kriegspiel}, we review
+of Kriegspiel using problem instances from real games. We achieve results which are unprecendented in game tree search application. 
+\item
+We demonstrate
+the effects of load balancing strategy and grainsize on the parallel performance of our implementation. 
+\end{itemize}
+
+The remainder of the paper is organized as follows.  In Section~\ref{Kriegspiel}, we review
 necessary background material, including the rules of Kriegspiel. Next, in
-Section~\ref{info} we describe the nature of the information set generation
+Section~\ref{info}, we describe the nature of the information set generation
 problem and the algorithm used.  In Section~\ref{ParSSSE}, we describe our
@@
 increases.  Our algorithm has performed well on problems with information sets
-with as many as 65 million nodes---much larger than any other treatment of the
-problem that we are aware of.  Because the algorithm can be efficiently
+with as many as 65 million nodes --- much larger than any other treatment of the
+problem that we are aware of.
+
+Since the algorithm can be efficiently
 parallelized, it would be reasonable to explore using it in a full Kriegspiel

### 2012-07-16 sbac/paper.tex
Changed a plot, muliple changes

@@
        }
+\title{Parallelizing  Information Set Generation for Game Tree Search Applications}
 
@@
 approach to efficient parallelization of an application which is not typical of
-high performance computing domain.
+high performance computing and scientific domain, yet important because of its usefulness in the field of artificial intelligence.
 \item
@@
 parallel implementation of an information set generation algorithm for the game
-of Kriegspiel using problem instances from real games. We achieve results which are unprecendented in game tree search application. 
+of Kriegspiel using problem instances from real games. We achieve results which are unprecedented in game tree search application. 
 \item
@@
 The concept of information sets is perhaps best explained through an example. 
-Figures ~\ref{listing} show a full game listing and the associated pictorial progression of an actual
+Figures~\ref{listing}, \ref{fig:figure1} and~\ref{fig:figure2} show a full game listing and the associated pictorial progression of an actual
 Kriegspiel game~\cite{li94chess}.
@@
 The actual move is shown
-first followed by a list of referee announcements.  Captures announcements are prefixed with an X and give the location
+first followed by a list of referee announcements.  Capture announcements are prefixed with an X and give the location
 of the captured piece.  Similarly, pawn try announcements are prefixed with T.  Attempted moves that were ultimately
@@
 each player.  Figure~\ref{fig:figure2} shows White's perspective of the game at
-that coin. At move 9, Black attempts to block a
+that point. At move 9, Black attempts to block a
 potential threat along the diagonal by moving a pawn to g6, then attempts to
@@
 of $t$ observations for one player $k$ (denoted $o_{1:t}^{(k)}$ in a Kriegspiel game.    
-
-The nodes in the search tree at depth $n$ correspond possible to possible move 
-sequences of $n$ plies from the start state.  (A sequence of $t$ actions from
-the start state is denoted $a_{1:t}$.) There is a search space variable for
-each ply in the game, and the values that a variable can take on are all
-possible legal moves.  Information set generation, then, is an all-solutions
-search through the space.  Whenever a node in the search tree is found to be
-inconsistent with the actual observation given in the input, the subtree rooted
-at that node is pruned (i.e., search is discontinued along that path).  There
-are three types of observations that must match in order for a state $s$ to be
-included in the information set $I$ at depth $t$: check status, pawn tries, and
-illegal moves available/attempted.
-
 \begin{figure}[htpb]
@@
 
+The nodes in the search tree at depth $n$ correspond to possible move 
+sequences of $n$ plies from the start state.  (A sequence of $t$ actions from
+the start state is denoted $a_{1:t}$.) There is a search space variable for
+each ply in the game, and the values that a variable can take on are all
+possible legal moves.  Information set generation, then, is an all-solutions
+search through the space.  Whenever a node in the search tree is found to be
+inconsistent with the actual observation given in the input, the subtree rooted
+at that node is pruned (i.e., search is discontinued along that path).  There
+are three types of observations that must match in order for a state $s$ to be
+included in the information set $I$ at depth $t$: check status, pawn tries, and
+illegal moves available/attempted.
+
 Figure~\ref{codelisting1} shows the function at the heart of our information set generation algorithm.  The actual
@@
 
-In contrast to other popular parallel programming frameworks such as {\sc MPI}
-and {\sc OpenMP}, the {\sc Charm++} language is based on object-oriented
+In contrast to other popular parallel programming frameworks such as {\sc MPI}~\cite{MPI}
+and {\sc OpenMP}~\cite{OpenMP}, {\sc Charm++} is based on object-oriented
 programming principles.  The programmer is responsible to decompose a problem
@@
 objects, which are executed according to a schedule set by the {\sc Charm++}
-runtime system.  The runtime is responsible to interleave computation and
+runtime system.  The runtime is responsible for interleaving computation and
 communication tasks in order to maximize efficiency.  To facilitate this, each
@@
 variety of graph-based and tree-based search problems~\cite{sun11adaptive}. In
-{\sc ParSSSE} the fundamental unit of work is the processing of a node.
+{\sc ParSSSE} the fundamental unit of work is the processing of a tree node.
 Generally, this means identifying whether the node corresponds to a solution
@@
 A {\em grainsize control} parameter specifies a bound on the size of a piece of
-work that can be passed along to the runtime system for scheduling.  Once a
+work that can be passed along to the runtime system for scheduling (in other words, computational granularity of a task which will be performed sequentially).  Once a
 task has been subdivided into an appropriate number of subtasks, those tasks
@@
 then the runtime system is flooded with a large number of trivial tasks, and
-the system gets bogged down by the overhead of creating and scheduling tasks
+application performance is degraded because of the overhead of creating and scheduling tasks
 instead of actually doing them.  {\sc ParSSSE} supports adaptive grainsize
@@
 during the execution of the program, depending on the properties of the problem
-and/or the realtime performance analysis of the runtime system.
-
-{\sc ParSSSE} currently supports two load balancing strategies.  The
+and/or the realtime performance analysis by the runtime system.
+
+Charm++ (and hence {\sc ParSSSE}) provides multiple load balancing strategies and we used three of them in our application. We describe these strategies next. The
 first is the randomized strategy.  In this approach, each newly created object
@@
 
+Another strategy we evaluated is the neighbor based load balancing which imposes a virtual topology on the processors and load exchange occur between neighbors only.
+In this strategy, the load balancing is initiated by overloaded processors.
+When a processor becomes overloaded, it sends work to its neighbors.
+
 In addition to these strategies, a user may define its own load balancing
@@
 \subsection{Analysis using Random Instance}
-The performance results reported in this section were obtained by running our
-program on Blueprint with up to 1,024 processors. This system was composed of
-120 POWER5+ nodes, each having 16 cores and 64GB of memory. The system was
-decommissioned in June 2011 and is no longer in operation.
-
-Table~\ref{speedups} shows the performance results for our algorithm on a
-randomly selected problem instance in which the size of the information set was
-about 4.5 million.  This instance is not necessarily representative of all
-games or states of the game, but it was not cherry-picked in any way. The table
-shows that the problem is indeed highly parallizable. The speedups are nearly
-linear for up to 32 processors, and the speedups continue to increase even up
-to 1,024 processors. In this particular instance, the sequential algorithm
-takes almost 25 minutes to run.  But when parallelized on 1,024
-processors---perfectly reasonable resource usage on modern hardware---the
-search takes less than three seconds.  This is a speedup of 576 times! In a
-tournament where each player is allotted an average of five minutes per move,
-an ISG operation that requires 25 minutes or more would be prohibitively
-expensive.  On the other hand, if the operation can be done in under three
-seconds, ISG can be used as a subroutine in the decision-making process
-(see~\cite{parker05game}).
 
@@
 \centering
-\begin{tabular}{rrr}
+\begin{tabular}{ccc}
 {\bf Procs}	&	{\bf Time (s)} 	&	{\bf Speedup}\\
@@
 
+The performance results reported in this section were obtained by running our
+program on Blueprint with up to 1,024 processors. This system was composed of
+120 POWER5+ nodes, each having 16 cores and 64GB of memory. The system was
+decommissioned in June 2011 and is no longer in operation.
+
+Table~\ref{speedups} shows the performance results for our algorithm on a
+randomly selected problem instance in which the size of the information set was
+about 4.5 million.  This instance is not necessarily representative of all
+games or states of the game, but it was not cherry-picked in any way. The table
+shows that our implementation achieves good parallel scaling. The speedups are nearly
+linear for up to 32 processors, and the speedups continue to increase even up
+to 1,024 processors. In this particular instance, the sequential algorithm
+takes almost 25 minutes to run.  However, when parallelized on 1,024
+processors --- perfectly reasonable resource usage on modern hardware --- the
+search takes less than three seconds.  This is a speedup of 576 times! In a
+tournament where each player is allotted an average of five minutes per move,
+an ISG operation that requires 25 minutes or more would be prohibitively
+expensive.  On the other hand, if the operation can be done in under three
+seconds, ISG can be used as a subroutine in the decision-making process
+(see~\cite{parker05game}).
+
 \subsubsection{Effect of Load Balancing}
 
-In our next round of experiments, we set out to compare load-balancing
-strategies for the problem.  We used the random, work stealing, and
+In our next round of experiments, we evaluated the impact of load-balancing
+strategies on our application's performance.  We used the random, work stealing, and
 neighborhood balancing strategies that are available in the general purpose
@@
 different strategies.  Across all of our experiments, the work stealing
-approach appeared to be consistently---if only marginally---better than the
+approach appeared to be consistently --- if only marginally --- better than the
 others.  Our hypothesis here is that the nodes are rarely starved for work.  In
@@
 naturally well-balanced and that the load-balancer rarely needs to do a
-significant amount of redistribution.  This is probably also related to the
+significant amount of redistribution.  This is also related to the
 fact that we are doing an all-solutions search. 
@@
 \subsubsection{Effect of Grainsize}
-
-\begin{figure*}[!ht]
+\label{grainsize}
+
+\begin{figure*}[ht]
 \centering
@@
 } 
-\caption{Grainsize Control for different problem sizes}
+\caption{Grainsize Control for different problem sizes.
+Sequential threshold denoted the depth in the search tree after which the remainder of the task (game tree search) is performed sequentially.
+Grain size = max game tree depth - sequential threshold.}
 %\vspace{-0.1in}
@@
 %
+
+Next, we evaluate the effect of grainsize on execution time. 
+Grainsize control is obtained in our implementation using a run-time argument called \emph{sequential threshold}, which specifies the depth in the search tree up to which parallel tasks would be created.
+With this terminology, grain size = max game tree depth - sequential threshold.
+Figures~\ref{dep8}--~\ref{dep12} illustrate the variation in application performance with sequential threshold for three different problem sizes.
+It is clear that there is a ``sweet
+spot'' for each problem size.  (At least it is clear for this particular
+instance.  It is possible that there would be different behavior on different
+problem instances.)  When the sequential threshold is large (grainsize is too small), the overall computation
+is hampered by excessive parallel overhead (communication, queueing of nodes to
+expand, load balancing operations).  On the other hand, when the sequential threshold is too small (grainsize is
+too large), it becomes too difficult to ensure that some processors are not
+idling with no work to do while a small number of processors are busy doing
+large chunks of work.
+
+In our experiments, the sequential threshold is clearly related to the size of the
+problem --- the depth of the search tree.  The optimal sequential threshold is $2-4$ less than the problem depth.  Since the depth of all solutions is
+the same for each problem instance, this means that the number of levels that
+should be searched sequentially (that is, optimal grain size) is fairly consistent across problem sizes.  We
+noted that the number of solutions found---in other words, the size of the
+information sets---is typically fairly large with respect to the total number
+of nodes expanded.  (We estimate that it was often $30-40\%$).  This high
+density of solutions limits the variability that can be present in the amount
+of work to be done at the nodes representing the last few levels of the tree.
+This means that the best grainsize is also likely to be fairly predictable,
+which is encouraging.
+
+\subsubsection{Effect of Problem Size}
+
+We next study the variation in the efficiency of the parallelization of the
+algorithm as size of the problem increases.  Here, we varied the number
+of plies in the test instance from 8 to 12. We used the work stealing
+load-balancing strategy.
+
+The results are shown in Figure~\ref{3depths}.  In the easiest case (depth 8),
+there is not enough additional parallel work to justify the jump from 256 to
+1,024 processors, and performance actually degrades.  Aside from this anomaly,
+however, we see consistent speedups as we increase the number of processors. 
+The plot on the bottom helps to illustrate the effect of problem size on parallel efficiency (defined as Speedup/P) and helps understand the iso-efficiency of the problem.
+Iso-efficiency~\cite{b:grama-isoefficiency} means how should the problem size be increased with increasing number of processors to maintain same level of parallel efficiency.
+As the size of the problem increases, the amount of parallelizable work available also increases, and hence parallel efficiency is higher for larger problem sizes and constant number of processors (Figure~\ref{3depths} bottom).
+As the number of processors increases, parallel efficiency decreases for constant problem size.
+Figure~\ref{3depths} suggests that in parallel information set generation, to maintain constant parallel efficiency, problem size (search depth) should be roughly increased by two when the number of processor in increased by 4 times.
 \begin{figure}[ht] 
@@
 \begin{minipage}{0.7\linewidth}
-\includegraphics[width=\columnwidth]{plots/3depthsnolog.pdf}
+\includegraphics[width=\columnwidth]{plots/3depthspe.pdf}
 \end{minipage}
-\caption{(Top) Log-log plot of running time for three different sizes of problem as the number of cores used increases.
-(Bottom) Same data with a linear y-axis for the running time, to help see the isoefficiency.}
+\caption{(Top) Execution time for three different sizes of problem as the number of cores used increases.
+(Bottom) Parallel Efficiency (wrt to 16 cores) vs number of cores, to help see the effect of problem size and iso-efficiency}
 \label{3depths}
 \end{figure}
-
-Next, we evaluate the effect of grainsize on execution time.  The results are
-shown in Tables~\ref{dep8}--~\ref{dep12}.  It is clear that there is a ``sweet
-spot'' for each problem size.  (At least it is clear for this particular
-instance.  It is possible that there would be different behavior on different
-problem instances.)  When the grainsize is too small, the overall computation
-is hampered by excessive parallel overhead (communication, queueing of nodes to
-expand, load balancing operations).  On the other hand, when the grainsize is
-too large, it becomes too difficult to ensure that some processors are not
-idling with no work to do while a small number of processors are busy doing
-large chunks of work.
-
-In our experiments, the optimal grainsize is clearly related to the size of the
-problem, the depth of the search tree.  The optimal sequential threshold seems
-to be $2-4$ less than the problem depth.  Since the depth of all solutions is
-the same for each problem instance, this means that the number of levels that
-should be searched sequentially is fairly consistent across problem sizes.  We
-noted that the number of solutions found---in other words, the size of the
-information sets---is typically fairly large with respect to the total number
-of nodes expanded.  (We estimate that it was often $30-40\%$).  This high
-density of solutions limits the variability that can be present in the amount
-of work to be done at the nodes representing the last few levels of the tree.
-This means that the best grainsize is also likely to be fairly predictable,
-which is encouraging.
-
-\subsubsection{Effect of Problem Size}
-
-We next study the variation in the efficiency of the parallelization of the
-algorithm as size of the problem increases.  To test this, we varied the number
-of plies in the test instance from 8 to 12. We used the work stealing
-load-balancing strategy.
-
-The results are shown in Figure~\ref{3depths}.  In the easiest case (depth 8),
-there is not enough additional parallel work to justify the jump from 256 to
-1,024 processors, and performance actually degrades.  Aside from this anomaly,
-however, we see consistent speedups as we increase the number of processors.
-The plot on the right helps to illustrate the good isoefficiency of the
-problem.  As the size of the problem increases, the amount of parallelizable
-work available also increases, and therefore the steepness of the speed curve
-is higher for larger problem sizes, as the number of processors increases.
-
 \subsection{Real Game Instance}
@@
 
-\begin{figure}[ht]
- \centering \includegraphics[scale=1]{plots/new.pdf} %
-\caption{Kriegspiel position from 2 games, position 10}
-\label{prob4} 
-\end{figure}
+We also evaluated our information set generation routine on 40 positions from
+actual Kriegspiel games, including human vs. human and computer vs. computer
+instances.  Many of these were trivial to solve in a few seconds or less on a
+desktop computer.  For these experiments, we used the Ranger supercomputer
+at TACC~\cite{ranger}. Ranger has 16-way SMP nodes containing AMD Opteron
+processors. The nodes are interconnected using Infiniband technology in a
+fully-CLOS topology providing a 1GBps point-to-point bandwidth.
+
+The running times and speedups for two of the more challenging instances are shown in Tables~\ref{tab:prob4}--\ref{tab:prob5}.   These tables
+show the results for runs on one to $1,024$ cores.  Each column corresponds to
+a different sequential threshold parameter - the number of
+levels of the search tree that are searched in parallel.  Up to that threshold,
+each node that is generated spawns a new piece of work, which the {\sc Charm++}
+runtime schedules to execute on one of the processing elements according to its
+load balancing strategy.  Beyond the threshold the entire subtree is searched to
+completion on a single processor, with no additional parallel overhead.
+
+These results confirm the observations made in Section~\ref{grainsize}. The optimal sequential threshold varies for the different numbers of cores and
+the pattern is consistent across problem instances.  The optimal
+sequential threshold increases as the number of cores increases.  This is expected,
+because as this parameter grows, the number of distinct pieces of work
+increases.  When the number of chares per processor is high, much of
+the overhead of parallelization is wasted.  But as the number of cores
+increases, the work can be efficiently broken up into smaller pieces and spread to all the cores.  With more pieces of work, it becomes easier to balance
+the load.
 
@@
 \centering
-\begin{tabular}{rrrrr}
+%\begin{tabular}{rrrrr}
+\begin{tabular}{ccccc}
 PE/G & 2 & 4 & 6 & 8 \\
@@
 \begin{table}[th]
-
 \caption{Running times, in seconds and speedups (after $|$), to find full
@@
 \end{table}
-
-We also evaluated our information set generation routine on 40 positions from
-actual Kriegspiel games, including human vs. human and computer vs. computer
-instances.  Many of these were trivial to solve in a few seconds or less on a
-desktop computer.  For these experiments, we used the Ranger supercomputer
-at TACC~\cite{ranger}. Ranger has 16-way SMP nodes containing AMD Opteron
-processors. The nodes are interconnected using Infiniband technology in a
-fully-CLOS topology providing a 1GBps point-to-point bandwidth.
-
-The running times and speedups for two of the more challenging instances on
-Ranger are shown in Tables~\ref{tab:prob4}--\ref{tab:prob5}.   These tables
-show the results for runs on one to $1,024$ cores.  Each column corresponds to
-a different grainsize parameter.  In this case, the grainsize is the number of
-levels of the search tree that are searched in parallel.  Up to that threshold,
-each node that is generated spawns a new piece of work, which the {\sc Charm++}
-runtime schedules to execute on one of the processing elements according to its
-load balancing strategy.  Above the threshold the entire subtree is searched to
-completion on a single processor, with no additional parallel overhead.
-
-Note that the optimal grainsize varies for the different numbers of cores and
-the pattern is consistent across problem instances.  The optimal
-grainsize increases as the number of cores increases.  This makes sense,
-because as this parameter grows, the number of distinct pieces of work
-increases.  When the number of chares per processor is high, much of
-the overhead of parallelization is wasted.  But as the number of cores
-increases, the work can be efficiently broken up into smaller pieces and doled
-out to all the cores.  With more pieces of work, it becomes easier to balance
-the load.
+\begin{figure}[ht]
+ \centering \includegraphics[scale=1]{plots/new.pdf} %
+\caption{Performance results for Information Set Generation for 2 games of Kriegspiel}
+\label{prob4} 
+\end{figure}
 
@@
 
+\section{Comparison with previous results, Discussion}
 Our results compare favorably to previous work in parallel game tree search,
@@
 
+For partially observable games, information set generation represents a
+different variety of tree search that does not suffer from speculative loss.
+All of the nodes that are visited in parallel also would have been visited in a
+sequential run.  The key challenge for parallelizing information set generation
+is load balancing: some parts of the tree will be pruned at a higher level than
+others, resulting in a wide variation in cost for each subtree.  The {\sc
+Charm++/ParSSSE} framework is perfectly suited to handle the load balancing
+issues that this problem naturally presents.
+
+Note also that information set generation is just one subroutine needed to
+reason about partially observable games.  Once plausible game histories have
+been found, an agent may simulate some sequences of moves up to previous moves
+made by opponents and then recursively call the ISG procedure from the
+perspective of the opponent.  This results in many parallel applications of
+ISG. Furthermore, given a node in an information set, it is common for an agent
+to perform forward search on that node (e.g., using MCTS).  All of these
+forward search operation can be performed independently in parallel. With these
+two additional levels of parallelism on top of the information set generation
+problem, it is likely that the problem of playing partially observable
+games could scale efficiently to millions of processors or more.
 \begin{table}[ht]
@@
 \end{table}
-
-For partially observable games, information set generation represents a
-different variety of tree search that does not suffer from speculative loss.
-All of the nodes that are visited in parallel also would have been visited in a
-sequential run.  The key challenge for parallelizing information set generation
-is load balancing: some parts of the tree will be pruned at a higher level than
-others, resulting in a wide variation in cost for each subtree.  The {\sc
-Charm++/ParSSSE} framework is perfectly suited to handle the load balancing
-issues that this problem naturally presents.
-
-Note also that information set generation is just one subroutine needed to
-reason about partially observable games.  Once plausible game histories have
-been found, an agent may simulate some sequences of moves up to previous moves
-made by opponents and then recursively call the ISG procedure from the
-perspective of the opponent.  This results in many parallel applications of
-ISG. Furthermore, given a node in an information set, it is common for an agent
-to perform forward search on that node (e.g., using MCTS).  All of these
-forward search operation can be performed independently in parallel. With these
-two additional levels of parallelism on top of the information set generation
-problem, it is likely that the problem of playing partially observable
-games could scale efficiently to millions of processors or more.
-
 \section{Related Work}
@@
 %of these would reduce the overall running time but would not necessarily be interesting from the standpoint of
-%evaluating the efficiency and isoefficiency of parallelization.  For example, in checking to see if a position leaves a
+%evaluating the efficiency and iso-efficiency of parallelization.  For example, in checking to see if a position leaves a
 %player in check, we currently utilize existing helper functions that generate possible moves for the opponent.  It would
@@
 \bibliographystyle{IEEEtran}
-\bibliography{paper}
+\bibliography{paper,cited}
 

### 2012-07-16 sbac/paper.tex
minor

@@
 \usepackage[tableposition=top,font=bf]{caption}
+%\usepackage[tableposition=top,font=bf]{caption}
 
@@
 Kriegspiel (partially observable chess)~\cite{li94chess}. The game tree nodes
-that are indistinguishable to a player because they differ only in the
+which are indistinguishable to a player since they differ only in the
 information that is hidden to the player by rule compose that player's {\em
@@
 cards is equally probable, as betting decisions made by the players up to that
-point would be affected by the quality of those players' cards.  But the set of
+point would be affected by the quality of those players' cards.  However, the set of
 {\em possible} hands for all of the opponents is easy to conceive and
@@
 Search Engine (ParSSSE) platform.  We then provide our experimental results in
-Section~\ref{Results}.  Related work is discussed in Section~\ref{Related}.  We
+Section~\ref{Results} followed by comparison with previous results and some discussion in Section~\ref{discuss}.  Related work is discussed in Section~\ref{Related}.  We
 conclude with a discussion of future work.
@@
 \label{listing}
+\vspace{-0.1in}
 \end{figure}
@@
 \end{boxedminipage}
-\caption{Algorithm}
+\caption{Core of Information Set Generation Algorithm}
 \label{codelisting1}
@@
 \section{Parallel Implementation using Charm++ and ParSSSE}\label{ParSSSE}
-We built our search engine on top of the {\sc Charm++} run-time
+We built our parallel ISG search engine on top of the {\sc Charm++} run-time
 system for parallel processing~\cite{kale93charm,kale09charm}.  {\sc Charm++}
@@
 
-Some chares may inherently require more computation than others, but the amount
+Some chares may inherently require more computation than others, however the amount
 of computation associated with each piece of work is not always known before
@@
 pruning~\cite{knuth75analysis,baudet78analysis}.  A natural decomposition is to
-assign different nodes to different processors for evaluation.  Because some
+assign different nodes to different processors for evaluation.  Since some
 nodes will be pruned at shallower levels of the tree than others, the amount of
@@
 then the runtime system is flooded with a large number of trivial tasks, and
-application performance is degraded because of the overhead of creating and scheduling tasks
+application performance is degraded due to the overhead of creating and scheduling tasks
 instead of actually doing them.  {\sc ParSSSE} supports adaptive grainsize
@@
 The advantage of the work stealing strategy is that the message-passing
-overhead is only incurred if at least one processor is idling.  Furthermore,
-because victims respond by sending the highest priority work, the most
+overhead is only incurred if at least one processor is idling.  Moreover,
+since victims respond by sending the highest priority work, the most
 important tasks are scheduled sooner.
@@
 
-\begin{table}[ht]
+The performance results reported in this section were obtained by running our
+program on Blueprint with up to 1,024 processors. This system was composed of
+120 POWER5+ nodes, each having 16 cores and 64GB of memory. The system was
+decommissioned in June 2011 and is no longer in operation.
+
+Table~\ref{speedups} shows the performance results for our algorithm on a
+randomly selected problem instance in which the size of the information set was
+about 4.5 million.  This instance is not necessarily representative of all
+games or states of the game, but it was not cherry-picked in any way. The table
+shows that our implementation achieves good parallel scaling. The speedups are nearly
+linear for up to 32 processors, and the speedups continue to increase even up
+to 1,024 processors. In this particular instance, the sequential algorithm
+takes almost 25 minutes to run.  However, when parallelized on 1,024
+processors --- perfectly reasonable resource usage on modern hardware --- the
+search takes less than three seconds.  This is a speedup of 576 times! In a
+tournament where each player is allotted an average of five minutes per move,
+an ISG operation that requires 25 minutes or more would be prohibitively
+expensive.  On the other hand, if the operation can be done in under three
+seconds, ISG can be used as a subroutine in the decision-making process
+(see~\cite{parker05game}).
+
+\subsubsection{Effect of Load Balancing}
+
+In our next round of experiments, we evaluated the impact of load-balancing
+strategies on our application's performance.  We used the random, work stealing, and
+neighborhood balancing strategies that are available in the general purpose
+search engine.  These results are shown in Figure~\ref{3schemes}.  Our results
+show that there was actually little variation in performance across the
+different strategies.  Across all of our experiments, the work stealing
+approach appeared to be consistently --- if only marginally --- better than the
+others.  Our hypothesis here is that the nodes are rarely starved for work.  In
+other words, perhaps the very nature of the problem means that the load is
+naturally well-balanced and that the load-balancer rarely needs to do a
+significant amount of redistribution.  This is also related to the
+fact that we are doing an all-solutions search.
+
+\subsubsection{Effect of Grainsize}
+
+Next, we evaluate the effect of grainsize on execution time. 
+Grainsize control is obtained in our implementation using a run-time argument called \emph{sequential threshold}, which specifies the depth in the search tree up to which parallel tasks would be created.
+With this terminology, grain size = max game tree depth - sequential threshold.
+Figures~\ref{dep8}--~\ref{dep12} illustrate the variation in application performance with sequential threshold for three different problem sizes.
+It is clear that there is a ``sweet
+spot'' for each problem size.  (At least it is clear for this particular
+instance.  It is possible that there would be different behavior on different
+problem instances.)  When the sequential threshold is large (grainsize is too small), the overall computation
+is hampered by excessive parallel overhead (communication, queueing of nodes to
+expand, load balancing operations).  On the other hand, when the sequential threshold is too small (grainsize is
+too large), it becomes too difficult to ensure that some processors are not
+idling with no work to do while a small number of processors are busy doing
+large chunks of work.
+
+\begin{table}[t]
 
@@
 
-The performance results reported in this section were obtained by running our
-program on Blueprint with up to 1,024 processors. This system was composed of
-120 POWER5+ nodes, each having 16 cores and 64GB of memory. The system was
-decommissioned in June 2011 and is no longer in operation.
-
-Table~\ref{speedups} shows the performance results for our algorithm on a
-randomly selected problem instance in which the size of the information set was
-about 4.5 million.  This instance is not necessarily representative of all
-games or states of the game, but it was not cherry-picked in any way. The table
-shows that our implementation achieves good parallel scaling. The speedups are nearly
-linear for up to 32 processors, and the speedups continue to increase even up
-to 1,024 processors. In this particular instance, the sequential algorithm
-takes almost 25 minutes to run.  However, when parallelized on 1,024
-processors --- perfectly reasonable resource usage on modern hardware --- the
-search takes less than three seconds.  This is a speedup of 576 times! In a
-tournament where each player is allotted an average of five minutes per move,
-an ISG operation that requires 25 minutes or more would be prohibitively
-expensive.  On the other hand, if the operation can be done in under three
-seconds, ISG can be used as a subroutine in the decision-making process
-(see~\cite{parker05game}).
-
-\subsubsection{Effect of Load Balancing}
-
-In our next round of experiments, we evaluated the impact of load-balancing
-strategies on our application's performance.  We used the random, work stealing, and
-neighborhood balancing strategies that are available in the general purpose
-search engine.  These results are shown in Figure~\ref{3schemes}.  Our results
-show that there was actually little variation in performance across the
-different strategies.  Across all of our experiments, the work stealing
-approach appeared to be consistently --- if only marginally --- better than the
-others.  Our hypothesis here is that the nodes are rarely starved for work.  In
-other words, perhaps the very nature of the problem means that the load is
-naturally well-balanced and that the load-balancer rarely needs to do a
-significant amount of redistribution.  This is also related to the
-fact that we are doing an all-solutions search. 
 \begin{figure}[ht]
@@
 \label{3schemes}
+\vspace{-0.2in}
 \end{figure}
 
-\subsubsection{Effect of Grainsize}
 \label{grainsize}
@@
 \caption{Grainsize Control for different problem sizes.
-Sequential threshold denoted the depth in the search tree after which the remainder of the task (game tree search) is performed sequentially.
+Sequential threshold denotes the depth in the search tree after which the remainder of the task (game tree search) is performed sequentially.
 Grain size = max game tree depth - sequential threshold.}
@@
 \label{fig:dep}
+\vspace{-0.2in}
 \end{figure*}
 %
-
-Next, we evaluate the effect of grainsize on execution time. 
-Grainsize control is obtained in our implementation using a run-time argument called \emph{sequential threshold}, which specifies the depth in the search tree up to which parallel tasks would be created.
-With this terminology, grain size = max game tree depth - sequential threshold.
-Figures~\ref{dep8}--~\ref{dep12} illustrate the variation in application performance with sequential threshold for three different problem sizes.
-It is clear that there is a ``sweet
-spot'' for each problem size.  (At least it is clear for this particular
-instance.  It is possible that there would be different behavior on different
-problem instances.)  When the sequential threshold is large (grainsize is too small), the overall computation
-is hampered by excessive parallel overhead (communication, queueing of nodes to
-expand, load balancing operations).  On the other hand, when the sequential threshold is too small (grainsize is
-too large), it becomes too difficult to ensure that some processors are not
-idling with no work to do while a small number of processors are busy doing
-large chunks of work.
 
@@
 load-balancing strategy.  
-
-The results are shown in Figure~\ref{3depths}.  In the easiest case (depth 8),
-there is not enough additional parallel work to justify the jump from 256 to
-1,024 processors, and performance actually degrades.  Aside from this anomaly,
-however, we see consistent speedups as we increase the number of processors. 
-The plot on the bottom helps to illustrate the effect of problem size on parallel efficiency (defined as Speedup/P) and helps understand the iso-efficiency of the problem.
-Iso-efficiency~\cite{b:grama-isoefficiency} means how should the problem size be increased with increasing number of processors to maintain same level of parallel efficiency.
-As the size of the problem increases, the amount of parallelizable work available also increases, and hence parallel efficiency is higher for larger problem sizes and constant number of processors (Figure~\ref{3depths} bottom).
-As the number of processors increases, parallel efficiency decreases for constant problem size.
-Figure~\ref{3depths} suggests that in parallel information set generation, to maintain constant parallel efficiency, problem size (search depth) should be roughly increased by two when the number of processor in increased by 4 times.
 \begin{figure}[ht] 
@@
 \label{3depths}
+\vspace{-0.3in}
 \end{figure}
-\subsection{Real Game Instance}
-%\begin{figure}[ht]
-% \centering \includegraphics[viewport= 280 200 300 550, scale=0.35]{images/KriegspielProblem4.pdf} %
-%
-%\caption{Kriegspiel position from Virgil-David 1969 with information set size
-%764,209.}
-%
-%\label{prob4} 
-%\vspace{-0.2in}
-%\end{figure}
-%
-%\begin{figure}[ht]
-% \centering \includegraphics[viewport= 280 200 300 550, scale=0.35]{images/KriegspielProblem5.pdf} %
-%
-%\caption{Kriegspiel position from Game 6 of Kbot-Darkboard at the 2006 Computer
-%Chess championship.  Information set size is 5,099,257.}
-%
-%\label{prob5} 
-%\vspace{-0.2in}
-%\end{figure}
-
-We also evaluated our information set generation routine on 40 positions from
-actual Kriegspiel games, including human vs. human and computer vs. computer
-instances.  Many of these were trivial to solve in a few seconds or less on a
-desktop computer.  For these experiments, we used the Ranger supercomputer
-at TACC~\cite{ranger}. Ranger has 16-way SMP nodes containing AMD Opteron
-processors. The nodes are interconnected using Infiniband technology in a
-fully-CLOS topology providing a 1GBps point-to-point bandwidth.
-
-The running times and speedups for two of the more challenging instances are shown in Tables~\ref{tab:prob4}--\ref{tab:prob5}.   These tables
-show the results for runs on one to $1,024$ cores.  Each column corresponds to
-a different sequential threshold parameter - the number of
-levels of the search tree that are searched in parallel.  Up to that threshold,
-each node that is generated spawns a new piece of work, which the {\sc Charm++}
-runtime schedules to execute on one of the processing elements according to its
-load balancing strategy.  Beyond the threshold the entire subtree is searched to
-completion on a single processor, with no additional parallel overhead.
-
-These results confirm the observations made in Section~\ref{grainsize}. The optimal sequential threshold varies for the different numbers of cores and
-the pattern is consistent across problem instances.  The optimal
-sequential threshold increases as the number of cores increases.  This is expected,
-because as this parameter grows, the number of distinct pieces of work
-increases.  When the number of chares per processor is high, much of
-the overhead of parallelization is wasted.  But as the number of cores
-increases, the work can be efficiently broken up into smaller pieces and spread to all the cores.  With more pieces of work, it becomes easier to balance
-the load.
+The results are shown in Figure~\ref{3depths}.  In the easiest case (depth 8),
+there is not enough additional parallel work to justify the jump from 256 to
+1,024 processors, and performance actually degrades.  Aside from this anomaly,
+however, we see consistent speedups as we increase the number of processors. 
+The plot on the bottom helps to illustrate the effect of problem size on parallel efficiency (defined as Speedup/P) and helps understand the iso-efficiency of the problem.
+Iso-efficiency~\cite{b:grama-isoefficiency} means how should the problem size be increased with increasing number of processors to maintain same level of parallel efficiency.
+As the size of the problem increases, the amount of parallelizable work available also increases, and hence parallel efficiency is higher for larger problem sizes and constant number of processors (Figure~\ref{3depths} bottom).
+As the number of processors increases, parallel efficiency decreases for constant problem size.
+Figure~\ref{3depths} suggests that in parallel information set generation, to maintain constant parallel efficiency, problem size (search depth) should be roughly increased by two when the number of processor in increased by 4 times.
 
@@
 \label{tab:prob4}
+\vspace{-0.1in}
 \end{table}
@@
 \label{tab:prob5}
+\vspace{-0.1in}
 \end{table}
-\begin{figure}[ht]
- \centering \includegraphics[scale=1]{plots/new.pdf} %
-\caption{Performance results for Information Set Generation for 2 games of Kriegspiel}
-\label{prob4} 
-\end{figure}
+
+\subsection{Real Game Instance}
+%\begin{figure}[ht]
+% \centering \includegraphics[viewport= 280 200 300 550, scale=0.35]{images/KriegspielProblem4.pdf} %
+%
+%\caption{Kriegspiel position from Virgil-David 1969 with information set size
+%764,209.}
+%
+%\label{prob4} 
+%\vspace{-0.2in}
+%\end{figure}
+%
+%\begin{figure}[ht]
+% \centering \includegraphics[viewport= 280 200 300 550, scale=0.35]{images/KriegspielProblem5.pdf} %
+%
+%\caption{Kriegspiel position from Game 6 of Kbot-Darkboard at the 2006 Computer
+%Chess championship.  Information set size is 5,099,257.}
+%
+%\label{prob5} 
+%\vspace{-0.2in}
+%\end{figure}
+
+We also evaluated our information set generation routine on 40 positions from
+actual Kriegspiel games, including human vs. human and computer vs. computer
+instances.  Many of these were trivial to solve in a few seconds or less on a
+desktop computer.  For these experiments, we used the Ranger supercomputer
+at TACC~\cite{ranger}. Ranger has 16-way SMP nodes containing AMD Opteron
+processors. The nodes are interconnected using Infiniband technology in a
+fully-CLOS topology providing a 1GBps point-to-point bandwidth. We used Charm++'s \texttt{net-ibverbs} machine layer build and $-O3$ optimization for these experiments.
+
+The running times and speedups for two of the more challenging instances are shown in Tables~\ref{tab:prob4}--\ref{tab:prob5}.   These tables
+show the results for runs on one to $1,024$ cores.  Each column corresponds to
+a different sequential threshold parameter - the number of
+levels of the search tree that are searched in parallel.  Up to that threshold,
+each node that is generated spawns a new piece of work, which the {\sc Charm++}
+runtime schedules to execute on one of the processing elements according to its
+load balancing strategy.  Beyond the threshold the entire subtree is searched to
+completion on a single processor, with no additional parallel overhead.
+
+These results confirm the observations made in Section~\ref{grainsize}. The optimal sequential threshold varies for the different numbers of cores and
+the pattern is consistent across problem instances.  The optimal
+sequential threshold increases as the number of cores increases.  This is expected,
+because as this parameter grows, the number of distinct pieces of work
+increases.  When the number of chares per processor is high, much of
+the overhead of parallelization is wasted.  However, as the number of cores
+increases, the work can be efficiently broken up into smaller pieces and spread to all the cores.  With more pieces of work, it becomes easier to balance
+the load.
 
@@
 speedup of $443$ in the game between humans and up to $536$ in the game between
-computers.  Figure~\ref{prob4} shows the plots of the running
+computers.  Figure~\ref{prob4} shows the plot of the running
 times on these two problems with the best grainsize for each core setting.  These plots show that the
@@
 \section{Comparison with previous results, Discussion}
+\label{discuss}
 Our results compare favorably to previous work in parallel game tree search,
@@
 issues that this problem naturally presents.
+
+\begin{figure}[t]
+ \centering \includegraphics[scale=1]{plots/new.pdf} %
+\caption{Performance results for Information Set Generation for two games of Kriegspiel, position ten}
+\label{prob4} 
+\end{figure}
 

### 2012-07-16 sbac/paper.tex
more

@@
 \label{filteredlisting}
-%\vspace{-0.2in}
+\vspace{-0.2in}
 \end{figure}
@@
 \label{abbrevoutput}
+\vspace{-0.1in}
 \end{figure}
@@
 \label{codelisting1}
+\vspace{-0.2in}
 \end{figure}
@@
 its accompanying runtime system have been ported to many shared-memory and
-distributed-memory platforms.
+distributed-memory platforms. {\sc Charm++} is used by many highly scalable scientific applications such as NAMD~\cite{NAMDIPDPS08} and ChaNGa~\cite{2007_ChaNGaScaling}
 
@@
 
-Charm++ (and hence {\sc ParSSSE}) provides multiple load balancing strategies and we used three of them in our application. We describe these strategies next. The
+{\sc Charm++} (and hence {\sc ParSSSE}) provides multiple load balancing strategies and we used three of them in our application~\cite{GengbinThesis}. We describe these strategies next. The
 first is the randomized strategy.  In this approach, each newly created object
@@
 processors. The nodes are interconnected using Infiniband technology in a
-fully-CLOS topology providing a 1GBps point-to-point bandwidth. We used Charm++'s \texttt{net-ibverbs} machine layer build and $-O3$ optimization for these experiments.
+fully-CLOS topology providing a 1GBps point-to-point bandwidth. We used {\sc Charm++}'s \texttt{net-ibverbs} machine layer build and $-O3$ optimization for these experiments.
 
@@
 
+The effect of load balancing and computational granularity on combinatorial search problems has been studied before~\cite{sun11adaptive,WeatherAmpiSBAC10}.
+In this work, we addressed the parallelization of an information set generation algorithm and studied the impact of these factors applied to game tree applications.
+%We explored an application  which has largely been unexplored by the HPC community
+
 %\Section{Conclusion Future Work}
@@
 \bibliographystyle{IEEEtran}
-\bibliography{paper,cited}
+\bibliography{paper,cited,group}
 

### 2012-07-16 sbac/paper.tex
more conclusion

@@
 
-However, in the game of Kriegspiel, information set generation is much more complicated.
+However, for many games (such as Kriegspiel), information set generation is much more complicated.
 After the first move by the opponent, a player faces uncertainty about the
@@
  We describe an
-approach to efficient parallelization of an application which is not typical of
+approach to efficient parallelization of an application - information set generation,  which is not typical of
 high performance computing and scientific domain, yet important because of its usefulness in the field of artificial intelligence.
@@
 
-\section{Kriegspiel}\label{Kriegspiel}
+\section{Background}\label{Kriegspiel}
 Information Set Generation is the process of identifying the nodes in an
@@
 can be applied to ISG in other imperfect information games.
+%\subsection{Kriegspiel}
 
@@
 \section{Conclusion and Future Work}
-We have shown that information set generation for Kriegspiel can be efficiently
-parallelized.  The opportunities for parallelism increase as the problem size
+In this paper, we showed that information set generation for Kriegspiel can be efficiently
+parallelized.
+Through extensive experimentation, we demonstrated that the choice of a suitable load balancing strategy and optimal grain size control is crucial to achieve good scalability for parallel ISG.
+The opportunities for parallelism increase as the problem size
 increases.  Our algorithm has performed well on problems with information sets

### 2012-09-07 paper/paper.tex
Adding all pending

@@
-\documentclass[11pt]{article}
+%\documentclass[11pt]{article}
+\documentclass[conference]{IEEEtran}
+
 \usepackage{amsmath,amsfonts,latexsym,graphicx}
@@
 engine and discuss its performance under different load balancing strategies and with different grainsize parameters.
-We discuss some variations in the search strategy and their effect on performance.  \end{abstract}
+Our results suggest that our algorithm can be efficiently parallelized.  \end{abstract}
 
@@
 
-Information set generation is not the same problem as solving a game
-tree, although it is arguably a necessary subroutine for quality game play in imperfect information games. The standard
-game tree problem is to find an equilibrium (i.e., minimax) strategy for the game. This can be done using alpha-beta
-pruning for perfect information games or linear programming for imperfect information games. Information set generation,
-by contrast, is the problem of finding sequences of moves in the game tree that are consistent with a player's
-observations.
+Information set generation is not the same problem as solving a game tree, although it is arguably a necessary
+subroutine for quality game play in imperfect information games.  In this work, we describe an approach to generating
+information sets for the game of kriegspiel.
 
@@
 
-The concept of information sets is perhaps best explained through an example.  Figures~\ref{listing} and~\ref{sequence}
-show a full listing and the associated pictorial simulation of an actual kriegspiel game from ~\cite{li94chess}.  We
-have adapted the notation used by Wolfe for Berkeley Kriegspiel.  The moves are numbered in pairs: a move by white on
-one line followed by black's move on the next line.  The actual move is shown first followed by a list of referee
-announcements.  Captures announcements are prefixed with an X and give the location of the captured piece.  Similarly,
-pawn tries announcements are prefixed with T.  Attempted moves that were ultimately declared illegal are shown in a list
-following the ':'.  For example at move 9, (Figure~\ref{fig:9W}--\ref{fig:9B}), black attempts to block a potential
-threat along the diagonal by moving a pawn to f6, then attempts to escape or capture a threat at f7.  When both of these
-attempts fail, black can infer that his king is being threatened by a protected bishop or queen at f7.  The attempt to
-move to e7 is black's way of finding out whether the threatening piece is a queen or bishop.  When this move fails,
-black knows it is white's queen and ultimately retreats to d8.  Note that the failed moves by white on turns 3 and 4 are
-a result of being blocked by the pawn at e2.
-
-Note that the listing in Figure~\ref{listing} corresponds to the referee's view of the game and would be useful for
+The concept of information sets is perhaps best explained through an example.  (See also Appendix A.)
+Figures~\ref{listing} and~\ref{sequence} show a full listing and the associated pictorial progression of an actual
+kriegspiel game from ~\cite{li94chess}.  We have adapted the notation used by Wolfe for Berkeley Kriegspiel.  The moves
+are numbered in pairs: a move by white on one line followed by black's move on the next line.  The actual move is shown
+first followed by a list of referee announcements.  Captures announcements are prefixed with an X and give the location
+of the captured piece.  Similarly, pawn tries announcements are prefixed with T.  Attempted moves that were ultimately
+declared illegal are shown in a list following the `:'.  For example at move 9, (Figure~\ref{fig:9W}--\ref{fig:9B}),
+black attempts to block a potential threat along the diagonal by moving a pawn to f6, then attempts to escape or capture
+a threat at f7.  When both of these attempts fail, black can infer that his king is being threatened by a protected
+bishop or queen at f7.  The attempt to move to e7 is black's way of finding out whether the threatening piece is a queen
+or bishop.  When this move fails, black knows it is white's queen and ultimately retreats to d8.  Note that the failed
+moves by white on turns 3 and 4 are a result of being blocked by the pawn at e2.
+
+The listing in Figure~\ref{listing} corresponds to the referee's view of the game and would be useful for
 post-game analysis or commentary.  During the game, players would not have access to the full information.  The
@@
 
-Figure~\ref{3schemes}
-\begin{figure}
+\begin{figure}[h]
 \centering
@@
 
-Figure~\ref{3depths}
-\begin{figure}
+\begin{figure}[h]
 \begin{minipage}{0.5\linewidth}
@@
 \end{minipage}
-\caption{3depths}
+\caption{(Left) Log-log plot of running time for three different sizes of problem as the number of cores used increases.
+(Right) Same data with a linear y-axis for the running time (to help see the isoefficiency.)}
 \label{3depths}
@@
 
-Figure~\ref{dep8}
-\begin{figure}
+\begin{figure}[h]
 \centering
 \includegraphics[width=0.6\textwidth]{plots/dep8.pdf}
-\caption{dep8}
-\label{Grainsize control data for depth 8 problems.}
+\caption{Grainsize control data for depth 8 problems.}
+\label{dep8}
 \end{figure}
 
-Figure~\ref{dep10}
-\begin{figure}
+\begin{figure}[h]
 \centering
 \includegraphics[width=0.6\textwidth]{plots/dep10.pdf}
-\caption{dep10}
-\label{Grainsize control data for depth 10 problems.}
+\caption{Grainsize control data for depth 10 problems.}
+\label{dep10}
 \end{figure}
 
-Figure~\ref{dep12}
-\begin{figure}
+\begin{figure}[h]
 \centering
 \includegraphics[width=0.6\textwidth]{plots/dep12.pdf}
-\caption{dep12}
-\label{Grainsize control data for depth 12 problems.}
+\caption{Grainsize control data for depth 12 problems.}
+\label{dep12}
 \end{figure}
@@
 \section{Related Work}
-An overview of the theory of imperfect information games and other basic topics in game theory can be found in
-~\cite{kuhn03lectures} and~\cite{kuhn97classics}.  A technical explanation of the concept of information sets can be
+An overview of the theory of imperfect information games and other basic topics in game theory can be found
+in~\cite{kuhn03lectures} and~\cite{kuhn97classics}.  A technical explanation of the concept of information sets can be
 found in~\cite{gilpin07algorithms}.  Algorithms for ``solving'' imperfect information games---in the sense of finding a
 Nash equilibrium---can be found in~\cite{koller94fast} and~\cite{koller97representations}.  Koller \etal, showed that an
-equilibrium can be found in time that is polynomial in the size of the game tree.  And in fact, they found that it took
-about as much time to generate the full game tree as it did to actually solve it.  Unfortunately, this is feasible only
-for small games.  In larger games where theoretically optimal strategies cannot be computed efficiently, computer
-systems tend to utilize the concept of infomration sets in some form or another.  A common strategy is to estimate the
-value of a move by averaging the estimated value of the associated descendant for each node in the current information
-set (or a random sample from the set).  While this sort of ``averaging over clairvoyance'' {\em can} lead to very poor
-decisions, it is nevertheless a reasonable strategy in some games (e.g., Scrabble~\cite{sheppard02world}).
+equilibrium can be found in time that is polynomial in the number of nodes in the game tree.  And in fact, they found
+that it took about as much time to generate the full game tree as it did to actually solve it.  Unfortunately, this is
+feasible only for small games.  In larger games where theoretically optimal strategies cannot be computed efficiently,
+computer systems tend to utilize the concept of infomration sets or belief states in some form or another.  A common
+strategy is to estimate the value of a move by averaging the estimated value of the associated descendant for each node
+in the current information set (or a random sample from the set).  While this sort of ``averaging over clairvoyance''
+{\em can} lead to very poor decisions, it is nevertheless a reasonable strategy in some games (e.g.,
+Scrabble~\cite{sheppard02world}).
 
 In this work, we have argued for an information set generation approach to kriegspiel, but other authors have approached
-the problem differently.  Parker \etal, consider the problem of sampling from belief states in a variety of large
-game trees, including kriegspiel~\cite{parker05game}.  They show that some performance gains can be achieved by only
+the problem differently.  Parker \etal, consider the problem of sampling from belief states in a variety of large game
+trees, including kriegspiel~\cite{parker05game}.  They show that some performance gains can be achieved by only
 approximate sampling from the belief state.  Instead of generating the full information set, or even sampling from it,
@@
 applied naively using a belief state sampling algorithm (i.e., an algorithm that samples from the distribution of
-possible worlds in the current state without respect to the overall history mof moves).  However, it has been shown that
+possible worlds in the current state without respect to the overall history of moves).  However, it has been shown that
 significant improvement in play can be achieved by estimating the value of future moves from the perspective of the
@@
 some games ourselves.  Currently, the manual encoding of games is not difficult but is tedious.  Still, it would be nice
-to have some games that go on for 40 or 50 moves and have positions where the information sets are on the order of a few
-hundred thousand or a few million in size.
-
-The information set generation algorithm that we have implemented here is one the most straightforward one and is
-alluded to in~\cite{parker05game} and~\cite{russell05efficient}.  It is analogous to the ``expand-at-the-tip'' strategy
-for the Hamiltonian Circuit problem.  An alternative information set generation algorithm is described in
-~\cite{richards09information} and could potentially improve the search efficiency greatly by exploiting variable
-ordering.  This would be analgous to the disconnected edge-pairs\footnote{This is what we referred to as ``algorithm 2'' in
-class.} heuristic for the Hamiltonian Circuit problem.  Rather than search from the root node going forward, this
-algorithm would seek opportunities for pruning based on the propagation of logical consequences from each move both
-forward and backward in time.  This, incidentally, is more consistent with the way human players would analyze the game.
-For example, if a player's proposed bishop move is rejected and the bishop is sufficiently removed from its own king,
-then the player may infer that one of his opponent's pieces is along that rejected path.  At least one of those spaces
-must have been the destination of a prior move by the opponent.  If a recent move by that bishop crossed the same path,
-then that would be hard evidence that those squares were empty at the time.  And therefore, the arrival of the
-opponent's piece must have been between the previous (successful) bishop move and the current (unsuccessful) bishop
-move.  To implement this algorithm, we envision the use of planning graph type data structures ~\cite{blum97fast}.
-Planning graphs have a sequence of alternating levels of state constraints and action constraints.  State constraints
-would be of the form $at(bishop,d4) \vee at(bishop,e6)$ or $!occupied(e4)$.  Action constraints would be of the form
-$move(knight,a4,c5) \vee move(knight,e6)$.  At a minimum, each level would require a data structure with a bit for each
-possible action/state.  Propagating the constraints would require much of the same machinery used in STRIPS-like
-planning algorithms~\cite{chen05solving}, and one of the many knobs that would need to be tuned would be the depth of the
-propagation for each constraint.  (Should the consequences of a known action or non-action be propagated one move into
-the future and one move into the past?  More?  Less?)  Additionally, there would be a significant and interesting
-tradeoff in the effectiveness of pruning and the and the size of the data structures that would have to be encoded in
-each chare.  The best algorithm might be a hybrid strategy that does forward propagation from the root at the parallel
-level and constraint propagation at the sequential level.\footnote{This work is also in progress.}
+to have some games that go on for 40 or 50 moves and have positions where the information sets are of moderate size (on
+the order of a few hundred thousand to a few million).
+
+The information set generation algorithm that we have implemented here is the most straightforward one and is alluded to
+in~\cite{parker05game} and~\cite{russell05efficient}.  It is analogous to the ``expand-at-the-tip'' strategy for the
+Hamiltonian Circuit problem.  An alternative information set generation algorithm is described
+in~\cite{richards09information} (but not implemented in parallel) and could potentially improve the search efficiency
+greatly by exploiting variable ordering.  This would be analgous to the disconnected edge-pairs\footnote{This is what we
+referred to as ``algorithm 2'' in class.} heuristic for the Hamiltonian Circuit problem.  Rather than search from the
+root node going forward, this algorithm would seek opportunities for pruning based on the propagation of logical
+consequences from each move both forward and backward in time.  This, incidentally, is more consistent with the way
+human players would analyze the game.  For example, if a player's proposed bishop move is rejected and the bishop is
+sufficiently removed from its own king, then the player may infer that one of his opponent's pieces is along that
+rejected path.  At least one of those spaces must have been the destination of a prior move by the opponent.  If a
+recent move by that bishop crossed the same path, then that would be hard evidence that those squares were empty at the
+time.  And therefore, the arrival of the opponent's piece must have been between the previous (successful) bishop move
+and the current (unsuccessful) bishop move.  To implement this algorithm, we envision the use of planning graph type
+data structures~\cite{blum97fast}.  Planning graphs have a sequence of alternating levels of state constraints and
+action constraints.  State constraints would be of the form $at(bishop,d4) \vee at(bishop,e6)$ or $!occupied(e4)$.
+Action constraints would be of the form $move(knight,a4,c5) \vee move(knight,e6)$.  At a minimum, each level would
+require a data structure with a bit for each possible action/state.  Propagating the constraints would require much of
+the same machinery used in STRIPS-like planning algorithms~\cite{chen05solving}, and one of the many knobs that would
+need to be tuned would be the depth of the propagation for each constraint.  (Should the consequences of a known action
+or non-action be propagated one move into the future and one move into the past?  More?  Less?)  Additionally, there
+would be a significant and interesting tradeoff in the effectiveness of pruning and the size of the data
+structures that would have to be encoded in each chare.  The best algorithm might be a hybrid strategy that does forward
+propagation from the root at the parallel level and constraint propagation at the sequential level.\footnote{This work
+is also in progress.}
 
@@
 allocation issues that could be improved.
+
+\section{Conclusion}
+We have shown that information set generation for kriegspiel can be efficiently parallelized.  The opportunities for
+parallelism increase as the problem size increases.  Our algorithm has performed well on problems with information sets
+with as many as 65 million nodes---much larger than any other treatment of the problem that we are aware of.  Because
+the algorithm can be efficiently parallelized, it would be reasonable to explore using it in a full kriegspiel player.
+It remains to be seen whether these scalability properties hold over a wide variety of playing styles and problem
+instances.  We suspect that fruitful research remains with respect to alternative algorithms that implement variable
+ordering heuristics.
 
@@
 \section{Appendix B}
+Verbose output from the information set generation algorithm.  The listing shows possible sequences of moves that are
+consistent with the black's observations after ten plies.  The resulting position after each consistent sequence of
+moves is shown below.  Based on these results, black can know with certainty the location of white's pawn at d6 and
+white's bishop at g5.
+
 \begin{verbatim}
@@
 
-\bibliographystyle{acm}
+%\bibliographystyle{acm}
+\bibliographystyle{IEEEtran}
+
 \bibliography{paper}

### 2012-09-15 sbac/paper.tex
ack
Addressed reviewer comments; fixed errors in figures.

@@
 \usepackage{listings}
-\usepackage{amsmath,amsfonts,latexsym,graphicx}
+\usepackage[cmex10]{amsmath}
+\usepackage{amsfonts,latexsym,graphicx}
 \usepackage[tableposition=top,font=bf]{caption}
@@
 \newcommand{\eps}{\epsilon}
+%\renewcommand{\thefootnote}{\fnsymbol{footnote}}
 %\documentstyle[times,art10,twocolumn,latex8]{article}
@@
    c5 {(:)}
-4. d5 {(:Bd8;Tc5)}
+4. d5 {(Tc5:Bd8)}
    Na6 {(:)}
 5. d6 {(:)}
-   f6 {(:Td6)}
+   f6 {(Td6:)}
 6. e4 {(:)}
-   fxg5 {(Xg5:Td6;Tg5)}
+   fxg5 {(Td6,Tg5,Xg5:)}
 7. Bc4 {(:)}
-   exd6 {(Xd6:Td6)}
+   exd6 {(Td6,Xd6:)}
 8. Qd5 {(:)}
@@
 queen and ultimately retreats to d8.  Note that the failed moves by White on
-turns 3 and 4 are a result of being blocked by the pawn at e2.
+turns 3 and 4 are a result of being blocked by the pawn at e7.
 
@@
    ??  {(:0)}
-4. d5 {(:Bd8;Tc5)}
+4. d5 {(Tc5:Bd8)}
    ??  {(:0)}
 5. d6 {(:)}
-   ?? {(:0,Td6)}
+   ?? {(Td6:0)}
 6. e4 {(:)}
-   fxg5 {(Xg5:Td6;Tg5)}
+   fxg5 {(Td6,Tg5,Xg5:0)}
 7. Bc4 {(:)}
-   ?? {(Xd6:Td6)}
+   ?? {(Td6,Xd6:0)}
 8. Qd5 {(:)}
@@
 9. Qf7+ {(:)} 
-   ?? {:0}
+   ?? {:3}
 10. Qxf8++ {(1-0:)}
@@
 randomly selected problem instance in which the size of the information set was
-about 4.5 million.  This instance is not necessarily representative of all
+about 4.5 million sequences.  This instance is not necessarily representative of all
 games or states of the game, but it was not cherry-picked in any way. The table
@@
 processors.  The input observations were from the 10th ply of the game tree and
-the size of the information set in this instance is 4,495,121.}
+the size of the information set in this instance is 4,495,121 sequences.}
 \centering
@@
 \caption{(Top) Execution time for three different sizes of problem as the number of cores used increases.
-(Bottom) Parallel Efficiency (wrt to 16 cores) vs number of cores, to help see the effect of problem size and iso-efficiency}
+(Bottom) Parallel Efficiency (wrt to 16 cores) vs number of cores, to help see the effect of problem size and iso-efficiency.}
 \label{3depths}
@@
 PE/G & 2 & 4 & 6 & 8 \\
-1 & {\bf 177$|$1} & 178$|$0.99 & 1844$|$0.96 & 493$|$0.36 \\
+1 & {\bf 177$|$1} & 178$|$0.99 & 184$|$0.96 & 493$|$0.36 \\
 4 & 71$|$2.48 & 53$|$3.34 & {\bf 48$|$3.63} & 75$|$2.35 \\
@@
 \label{tab:prob5}
-\vspace{-0.1in}
+\vspace{-0.2in}
 \end{table}
@@
 \label{prob4}
+\vspace{-0.2in} 
 \end{figure}
@@
 \label{bestspeedups}
+\vspace{-0.2in} 
 \end{table}
@@
 
+\section*{Acknowledgments}
+
+Authors were supported by grants 
+NSF-0904844, DOE DE-SC0001845. We used machine resources on the
+Ranger cluster (TACC), under TeraGrid allocation grant TG-ASC050039N supported
+by NSF. First and second author contributed equally. 
 %\Section{Appendix A}

### 2012-09-15 sbac/paper.tex
Several minor typos and mechanics issues addressed.

@@
 Kriegspiel (partially observable chess)~\cite{li94chess}. The game tree nodes
-which are indistinguishable to a player since they differ only in the
+that are indistinguishable to a player since they differ only in the
 information that is hidden to the player by rule compose that player's {\em
@@
 
-The listing of a sequence of moves from the perspective of one player lists all of that player's moves exactly and
+The listing of a sequence of moves from the perspective of one player shows all of that player's moves exactly and
 includes the declarations from the referee.  We refer to such a list as a player's observations.  Given a list of
@@
 the same sequence of observations for Black is in Black's information set.  Figure~\ref{abbrevoutput} shows a portion of the
-information set for Black after both player's have made five moves.  From this list of possibilities, Black can infer
+information set for Black after both players have made five moves.  From this list of possibilities, Black can infer
 that White must have a bishop at g5 and a pawn at d6.  There are 25 possible sequences of moves that are consistent with
@@
 (ISG) operation can be formulated as a combinatorial search problem.  The input to the algorithm is a list
-of $t$ observations for one player $k$ (denoted $o_{1:t}^{(k)}$ in a Kriegspiel game.    
+of $t$ observations for one player $k$ (denoted $o_{1:t}^{(k)}$) in a Kriegspiel game.    
 \begin{figure}[htpb]
@@
 its accompanying runtime system have been ported to many shared-memory and
-distributed-memory platforms. {\sc Charm++} is used by many highly scalable scientific applications such as NAMD~\cite{NAMDIPDPS08} and ChaNGa~\cite{2007_ChaNGaScaling}
+distributed-memory platforms. {\sc Charm++} is used by many highly scalable scientific applications such as NAMD~\cite{NAMDIPDPS08} and ChaNGa~\cite{2007_ChaNGaScaling}.
 
@@
 
-Another strategy we evaluated is the neighbor based load balancing which imposes a virtual topology on the processors and load exchange occur between neighbors only.
+Another strategy we evaluated is the neighbor-based load balancing which imposes a virtual topology on the processors and load exchange occur between neighbors only.
 In this strategy, the load balancing is initiated by overloaded processors.
@@
 The plot on the bottom helps to illustrate the effect of problem size on parallel efficiency (defined as Speedup/P) and helps understand the iso-efficiency of the problem.
-Iso-efficiency~\cite{b:grama-isoefficiency} means how should the problem size be increased with increasing number of processors to maintain same level of parallel efficiency.
+Iso-efficiency~\cite{b:grama-isoefficiency} quantifies how the problem size should be increased with increasing number of processors to maintain same level of parallel efficiency.
 As the size of the problem increases, the amount of parallelizable work available also increases, and hence parallel efficiency is higher for larger problem sizes and constant number of processors (Figure~\ref{3depths} bottom).
 As the number of processors increases, parallel efficiency decreases for constant problem size.
-Figure~\ref{3depths} suggests that in parallel information set generation, to maintain constant parallel efficiency, problem size (search depth) should be roughly increased by two when the number of processor in increased by 4 times.
+Figure~\ref{3depths} suggests that in parallel information set generation, to maintain constant parallel efficiency, problem size (search depth) should be roughly increased by two when the number of processors is increased by 4 times.
 
@@
  \centering \includegraphics[scale=1]{plots/new.pdf} %
-\caption{Performance results for Information Set Generation for two games of Kriegspiel, position ten}
+\caption{Performance results for Information Set Generation for two games of Kriegspiel, position ten.}
 \label{prob4}
